---
title: "Latin text classification: Data Processing"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of christian vs. non-christian documents


## Data loading and cleaning

### Loading in and filtering the data (this code chunk can be skipped if corpus.csv already exists in the script directory)

```{r}
library(tidyverse)
library(stringi)

data_path <- "../../data/processed_texts/600ad_unlemmatized"

meta_csv <- file.path("../../data/processed_texts/metadata/latinise_metadata_2026.csv")
texts_dir <- "../../data/non_lemmatized_texts_2026"
corpus_rds  <- file.path(data_path,"corpus_600ad_unlemmatized.rds")
corpus_csv  <- file.path(data_path,"corpus_600ad_unlemmatized.csv")

meta <- readr::read_csv(meta_csv, show_col_types = FALSE) %>%
  mutate(file = basename(as.character(file)))
```


### Declaring text-processing functions 

```{r}

read_txt <- function(path){
  x <- readr::read_file(path) 
  x <- stri_trans_nfc(x)      
  x <- str_replace_all(x, "\\r\\n?", "\n") 
  x <- str_squish(x) 
  x
}

strip_hash_nums <- function(x) str_replace_all(x, "#\\d+", "")
```


### Processing the text files with corresponding metadata using the functions we just declared

```{r}

texts <- tibble(file_full = dir(texts_dir, pattern = "\\.txt$", full.names = TRUE, recursive = TRUE)) %>%
  mutate(file = basename(file_full),
         text_raw = map_chr(file_full, read_txt),
         text     = strip_hash_nums(text_raw))
```

### Combining metadata with texts into a data frame and counting characters + words for downstream tasks

```{r}
df <- meta %>%
  left_join(select(texts, file, text_raw, text), by = "file") %>% 
  mutate(
    n_chars_raw = nchar(text_raw, allowNA = TRUE), 
    n_chars     = nchar(text,     allowNA = TRUE),
    n_words_raw = if_else(is.na(text_raw), NA_integer_, str_count(text_raw, "(?u)\\p{L}+")), 
    n_words     = if_else(is.na(text),     NA_integer_, str_count(text,     "(?u)\\p{L}+"))
  )
```


### Saving the processed files 

```{r}
df <- dplyr::filter(df, date_range_end <= 605)
readr::write_csv(df, file.path(data_path,"corpus_600ad_unlemmatized.csv"))
saveRDS(df, file.path(data_path,"corpus_600ad_unlemmatized.rds"))
print(head(select(df, file, n_chars, n_words)))
```


## Sampling from the corpus for manual labeling

### Stratifying the text corpus

```{r}
library(dplyr)

get_latin_era <- function(year) {
  case_when(
    is.na(year) ~ "Unknown",
    year < -80  ~ "Archaic",
    year < 14   ~ "Classical",
    year < 180  ~ "Silver",
    TRUE        ~ "Late"
  )
}

df <- df %>%
  mutate(start_era = get_latin_era(date_range_start)) %>%
  mutate(end_era = get_latin_era(date_range_end))
head(df)

df <- df %>%
  mutate(stratum_start = paste(start_era, type, sep = "_")) %>%
  mutate(stratum_end = paste(end_era, type, sep = "_"))

```

### Quick sanity check on the stratal distribution of the corpus
```{r}
print("Distribution of Strata in Full Corpus:")
print(table(df$stratum_start,df$stratum_end))
```

### Creating a sample corpus

```{r}

set.seed(42) # Setting seed for reproducibility

sample_df <- df %>%
  group_by(stratum_start,stratum_end) %>%
  slice_sample(prop = 100/nrow(df)) %>% 
  ungroup()

print(paste(nrow(sample_df)))
```

### Ensuring that the sample is a good representative of the entire corpus

#### Extracting top words for analysis of similarity (inspired by Burrows delta)
```{r}
library(tidytext)

get_top_words <- function(df, n = 50) {
  df %>%
    unnest_tokens(word, text) %>%
    count(word, sort = TRUE) %>%
    slice_max(n = n, order_by = n) %>% 
    pull(word)
}

target_words <- get_top_words(df)

```


#### Extracting statistics on top words for the sample and entire corpus

```{r}
get_fingerprint <- function(data, text_col, target_words) {
  
  all_text <- paste(data[[text_col]], collapse = " ")
  all_text <- str_to_lower(all_text)
  all_words <- unlist(str_split(all_text, "\\W+"))
  total_count <- length(all_words)
  results <- numeric(length(target_words))
  names(results) <- target_words
  
  for (word in target_words) {
    word_count <- sum(all_words == word)
    results[word] <- (word_count / total_count) * 100
  }
  
  return(results)
}

```


```{r}
corpus_stats <- get_fingerprint(df, "text_raw", target_words)
sample_stats <- get_fingerprint(sample_df, "text_raw", target_words)
```


#### Comparing the sample and entire corpus

##### Preparing the data for statistical tests

```{r}
library(knitr)
library(tidytext)

word_comp <- data.frame(
  Word = names(corpus_stats),
  Corpus_Pct = as.numeric(corpus_stats),
  Sample_Pct = as.numeric(sample_stats)
)

era_comp <- data.frame(
  Era = names(prop.table(table(df$start_era))),
  Corpus_Pct = as.numeric(prop.table(table(df$start_era)) * 100),
  Sample_Pct = as.numeric(prop.table(table(sample_df$start_era)) * 100)
)

validation_df <- rbind(
  data.frame(Feature = paste0("Word: ", word_comp$Word), 
             Corpus_Pct = word_comp$Corpus_Pct, 
             Sample_Pct = word_comp$Sample_Pct),
  data.frame(Feature = paste0("Era: ", era_comp$Era), 
             Corpus_Pct = era_comp$Corpus_Pct, 
             Sample_Pct = era_comp$Sample_Pct)
) %>%
  mutate(Diff = abs(Corpus_Pct - Sample_Pct))


kable(head(validation_df, 20))
```

##### Pearson's correlation test for word percentages

```{r}
library(ggplot2)
library(ggrepel) 

word_cor_test <- cor.test(word_comp$Corpus_Pct, word_comp$Sample_Pct, method = "pearson")


p <- ggplot(word_comp, aes(x = Corpus_Pct, y = Sample_Pct)) +
  geom_point(alpha = 0.6, color = "#FFD100", size = 3) +
  geom_smooth(method = "lm", color = "#2774AE", fill = "#8BB8E8", alpha = 0.2) +
  
  annotate("text", x = min(word_comp$Corpus_Pct), y = max(word_comp$Sample_Pct), 
           label = sprintf("Pearson's r = %.3f\np < .001", word_cor_test$estimate),
           hjust = 0, vjust = 1, size = 5, fontface = "italic", color = "black") +
  
  labs(
    title = "Comparison of Word Frequencies",
    subtitle = paste("Top", nrow(word_comp), "Words: Full Corpus vs. Sample Stratified Subset"),
    x = "Full Corpus Frequency (%)",
    y = "Sample Frequency (%)"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

print(p)


```

##### Kolmogorov-Smirnov (KS) Test to see if the sample's distribution of document lengths matches the full corpus
```{r}

library(ggplot2)
library(dplyr)
ks_len_test <- ks.test(df$n_words, sample_df$n_words)
plot_data <- bind_rows(
  data.frame(n_words = df$n_words, Group = "Full Corpus"),
  data.frame(n_words = sample_df$n_words, Group = "Sample")
)

p_ks <- ggplot(plot_data, aes(x = n_words, color = Group)) +
  stat_ecdf(geom = "step", linewidth = 1.2) +
  scale_color_manual(values = c("Full Corpus" = "#2774AE", "Sample" = "#FFD100")) +
  scale_x_log10() + # Log scale is often better for word counts (skewed data)
  labs(
    title = "Cumulative Distribution of Document Lengths",
    subtitle = paste("Kolmogorov-Smirnov Test Comparison"),
    x = "Number of Words (Log Scale)",
    y = "Cumulative Probability (ECDF)",
    color = "Dataset"
  ) +
  
  annotate("text", x = min(plot_data$n_words), y = 0.8, 
           label = sprintf("KS Statistic (D) = %.3f\np-value = %.3f", 
                           ks_len_test$statistic, 
                           ks_len_test$p.value),
           hjust = 0, size = 5, fontface = "italic", color = "black") +
  
  theme_classic() +
  theme(
    legend.position = c(0.8, 0.2), # Place legend inside plot
    plot.title = element_text(face = "bold", size = 14)
  )

print(p_ks)

```

#### Saving the sample corpus 

```{r}
library(tidyverse)
labeling_sheet <- sample_df %>%
  select(id, title, creator, date_range_start,date_range_end, start_era, end_era, stratum_start,stratum_end, type)

write_excel_csv(labeling_sheet, file.path(data_path,"latin_sample_600ad_unlemmatized_metadata_only.csv"), na = "")


```

### Creating additional subsets that exclude the works already sampled
```{r}

sample_df <- read.csv(file.path(data_path,"latin_sample_600ad_unlemmatized_metadata_only.csv"))
remaining_pool <- df %>%
  anti_join(sample_df, by = "id") 


target_size <- 100

set.seed(42) 

sample_df_2 <- remaining_pool %>%
  group_by(stratum_start) %>%
  slice_sample(prop = target_size / nrow(remaining_pool)) %>%
  ungroup()

overlap_check <- intersect(sample_df$id, sample_df_2$id)
print(paste("Number of overlapping documents:", length(overlap_check)))
```

#### Making sure our second sample also fits the empirical distributions of the corpus 
```{r}
library(knitr)
library(ggplot2)
library(tidymodels)

sample_2_stats <- get_fingerprint(sample_df_2, "text_raw", target_words)

comparison_2 <- data.frame(
  Word = names(corpus_stats),
  Corpus_Pct = as.numeric(corpus_stats),
  Sample_2_Pct = as.numeric(sample_2_stats)
) %>%
  mutate(Diff = abs(Corpus_Pct - Sample_2_Pct)) %>%
  arrange(desc(Diff)) 

cor_test_2 <- cor.test(comparison_2$Corpus_Pct, comparison_2$Sample_2_Pct)

cat("\n=== SAMPLE 2 VALIDATION SUMMARY ===\n")
cat(sprintf("Correlation to Corpus: r = %.4f (p < .001)\n", cor_test_2$estimate))
cat(sprintf("Mean Deviation:        %.2f%%\n", mean(comparison_2$Diff)))
cat("-----------------------------------\n")

print(kable(comparison_2, digits = 2, row.names = FALSE,
            caption = "Word Frequency Comparison (Sorted by Deviation)",
            col.names = c("Word", "Corpus %", "Sample 2 %", "Diff")))

```

#### Saving the new corpus 

```{r}

labeling_sheet_2 <- sample_df_2 %>%
  select(id, title, creator, date_range_start,date_range_end,start_era,end_era, stratum_start,stratum_end, type)

write_excel_csv(labeling_sheet_2, file.path(data_path,"latin_sample_2_600ad_unlemmatized_metadata_only.csv"), na = "")


```