---
title: "multimodel_feature_drivers_600ad"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(stringi)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(ggrepel)
library(forcats)
corpus_path <- "../../data/processed_texts/600ad"
data_path <- "../../results/predictions/non_transformer_lemmatized/600ad"
fig_dir <- "../../results/figures/600ad/multimodel_feature_drivers"
dir.create(fig_dir, recursive = TRUE, showWarnings = FALSE)

`%||%` <- function(x, y) if (is.null(x)) y else x

min_docfreq <- 3L
alpha_nb <- 1
min_present_group <- 2L

col_christian     <- "#0072B2"
col_non_christian <- "#D55E00"
col_mid           <- "#F0E442"
col_grid          <- "grey88"
col_segment       <- "grey72"
col_zero_line     <- "grey50"
col_text          <- "grey20"
```

## Data

```{r}
df <- readr::read_csv(file.path(corpus_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels_raw <- readr::read_csv(file.path(corpus_path, "corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  distinct(id, label)

df_labeled <- df %>%
  inner_join(labels_raw %>% transmute(id, manual_label = label), by = "id") %>%
  arrange(row_id)

hyper_full <- readr::read_csv(file.path(data_path, "model_hyperparameters_full.csv"), show_col_types = FALSE)
```

## Stopwords

```{r}
custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href",
  "fridericus", "heinricus", "conradus", "meus", "tuus"
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]
```

## Helpers

```{r}
safe_prob <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)

make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat <- as(x_counts, "dgCMatrix")
  x_vals <- x_mat@x
  x_mat@x <- 1 + log(x_vals)
  idf_vec <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}




fit_enet_fixed <- function(x, y, alpha, lambda) {
  glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    lambda = lambda,
    standardize = TRUE,
    maxit = 1e6
  )
}

svm_to_matrix <- function(x) {
  if (inherits(x, "dgCMatrix") || inherits(x, "dsCMatrix") || inherits(x, "dgRMatrix")) return(x)
  if (inherits(x, "matrix")) return(x)
  as.matrix(x)
}

fit_svm_fixed <- function(x, y, cost) {
  x_use <- svm_to_matrix(x)
  fit <- LiblineaR::LiblineaR(data = x_use, target = y, type = 7, cost = cost)
  pred_tr <- predict(fit, x_use, decisionValues = TRUE)
  raw <- as.numeric(pred_tr$decisionValues[, 1])
  sgn <- if (mean(raw[y == "christian"]) >= mean(raw[y == "non_christian"])) 1 else -1
  list(model = fit, sign = sgn, cost = cost)
}

make_xgb_params <- function(row, scale_pos_weight = 1) {
  list(
    booster = "gbtree",
    objective = "binary:logistic",
    eval_metric = "logloss",
    tree_method = "hist",
    eta = as.numeric(row$eta),
    max_depth = as.integer(row$max_depth),
    min_child_weight = as.numeric(row$min_child_weight),
    subsample = as.numeric(row$subsample),
    colsample_bytree = as.numeric(row$colsample_bytree),
    gamma = as.numeric(row$gamma),
    scale_pos_weight = scale_pos_weight
  )
}

fit_xgb_fixed <- function(x, y, params, nrounds, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  xgboost::xgb.train(
    params = params,
    data = xgboost::xgb.DMatrix(x, label = as.integer(y == "christian")),
    nrounds = as.integer(nrounds),
    verbose = 0
  )
}

make_lgb_params <- function(row, seed = 123) {
  list(
    objective = "binary",
    metric = "binary_logloss",
    boosting = "gbdt",
    learning_rate = as.numeric(row$learning_rate),
    num_leaves = as.integer(row$num_leaves),
    min_data_in_leaf = as.integer(row$min_data_in_leaf),
    feature_fraction = as.numeric(row$feature_fraction),
    bagging_fraction = as.numeric(row$bagging_fraction),
    bagging_freq = 1L,
    is_unbalance = TRUE,
    first_metric_only = TRUE,
    seed = seed,
    feature_fraction_seed = seed,
    bagging_seed = seed,
    verbosity = -1
  )
}

make_lgb_dataset_params <- function() {
  list(feature_pre_filter = FALSE)
}

fit_lgb_fixed <- function(x, y, params, nrounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x,
    label = as.integer(y == "christian"),
    params = make_lgb_dataset_params(),
    free_raw_data = TRUE
  )
  lightgbm::lgb.train(
    params = params,
    data = dtrain,
    nrounds = as.integer(nrounds),
    verbose = -1
  )
}

zscore <- function(x) {
  s <- stats::sd(x, na.rm = TRUE)
  m <- mean(x, na.rm = TRUE)
  if (!is.finite(s) || s == 0) return(rep(0, length(x)))
  (x - m) / s
}

safe_divide <- function(num, den) {
  num <- as.numeric(num)
  den <- as.numeric(den)
  out <- num / den
  out[den == 0 | !is.finite(out)] <- NA_real_
  out
}


sparse_weight_contrib <- function(x, weights) {
  if (length(weights) != ncol(x)) stop("weights must have length ncol(x)")
  out <- x %*% Matrix::Diagonal(x = as.numeric(weights))
  Matrix::drop0(out)
}

mask_dense_contrib_to_present <- function(contrib, x_ref) {
  present <- if (inherits(x_ref, "Matrix")) as.matrix(x_ref != 0) else (x_ref != 0)
  if (!all(dim(contrib) == dim(present))) stop("contrib and x_ref must have the same dimensions")
  contrib * present
}



extract_enet_weights <- function(fit, feature_names) {

  beta <- as.matrix(stats::coef(fit))
  beta <- beta[, 1, drop = FALSE]
  out <- tibble(term = rownames(beta), weight = as.numeric(beta[, 1])) %>%
    filter(term != "(Intercept)") %>%
    right_join(tibble(term = feature_names), by = "term") %>%
    mutate(weight = replace_na(weight, 0))
  out$weight
}

extract_svm_weights <- function(obj, feature_names) {
  p <- length(feature_names)
  w_raw <- obj$model$W
  if (is.null(w_raw)) stop("LiblineaR model does not expose a weight vector in $W.")
  w <- as.numeric(unlist(w_raw, recursive = TRUE, use.names = FALSE))

  if (length(w) == p + 1L) {
    w <- w[seq_len(p)]
  } else if (length(w) %% p == 0L) {
    k <- as.integer(length(w) / p)
    w_mat <- matrix(w, nrow = p, ncol = k)
    if (k == 1L) {
      w <- w_mat[, 1]
    } else if (k == 2L && stats::cor(w_mat[, 1], -w_mat[, 2], use = "complete.obs") > 0.999) {
      w <- w_mat[, 1]
    } else {
      sds <- apply(w_mat, 2, stats::sd)
      w <- w_mat[, which.max(sds)]
    }
  }

  if (length(w) != p) {
    stop(sprintf(
      "SVM weight vector length %d does not match the feature count %d.",
      length(w), p
    ))
  }

  obj$sign * w
}

extract_nb_weights <- function(x_counts, y, feature_names, alpha = alpha_nb) {
  y_ch <- y == "christian"
  y_nc <- y == "non_christian"
  cnt_ch <- Matrix::colSums(x_counts[y_ch, , drop = FALSE])
  cnt_nc <- Matrix::colSums(x_counts[y_nc, , drop = FALSE])
  V <- length(feature_names)
  p_ch <- (cnt_ch + alpha) / (sum(cnt_ch) + alpha * V)
  p_nc <- (cnt_nc + alpha) / (sum(cnt_nc) + alpha * V)
  tibble(term = feature_names, weight = as.numeric(log(p_ch / p_nc)))
}

predict_xgb_contrib <- function(fit, x, feature_names) {
  out <- predict(
    fit,
    xgboost::xgb.DMatrix(x),
    predcontrib = TRUE,
    approxcontrib = FALSE
  )
  out <- as.matrix(out)
  if (ncol(out) != length(feature_names) + 1) {
    stop("XGBoost contribution output does not have p + 1 columns.")
  }
  colnames(out) <- c(feature_names, "BIAS")
  list(shap = out[, feature_names, drop = FALSE], bias = out[, "BIAS"])
}

predict_lgb_contrib <- function(model, x, feature_names, num_iteration = NULL) {
  args <- list(model, x, type = "contrib")
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, args),
    error = function(e) tryCatch(
      do.call(predict, c(args, list(predcontrib = TRUE))),
      error = function(e2) stop(e2)
    )
  )
  out <- as.matrix(out)
  if (ncol(out) != length(feature_names) + 1) {
    stop("LightGBM contribution output does not have p + 1 columns.")
  }
  colnames(out) <- c(feature_names, "BIAS")
  list(shap = out[, feature_names, drop = FALSE], bias = out[, "BIAS"])
}
```

```{r}
summarize_driver_matrix <- function(contrib, x_ref, y, feature_names, model_name) {
  if (ncol(contrib) != length(feature_names)) stop("contrib has wrong number of columns")
  if (ncol(x_ref) != length(feature_names)) stop("x_ref has wrong number of columns")

  y_ch <- y == "christian"
  y_nc <- y == "non_christian"
  present <- x_ref != 0

  n_present    <- Matrix::colSums(present)
  n_present_ch <- Matrix::colSums(present[y_ch, , drop = FALSE])
  n_present_nc <- Matrix::colSums(present[y_nc, , drop = FALSE])


  if (inherits(contrib, "Matrix")) {
    contrib <- contrib * present
    contrib <- Matrix::drop0(contrib)
    sum_all <- Matrix::colSums(contrib)
    sum_ch  <- Matrix::colSums(contrib[y_ch, , drop = FALSE])
    sum_nc  <- Matrix::colSums(contrib[y_nc, , drop = FALSE])
    sum_abs <- Matrix::colSums(abs(contrib))
  } else {
    present_dense <- if (inherits(present, "Matrix")) as.matrix(present) else present
    contrib <- contrib * present_dense
    sum_all <- colSums(contrib)
    sum_ch  <- colSums(contrib[y_ch, , drop = FALSE])
    sum_nc  <- colSums(contrib[y_nc, , drop = FALSE])
    sum_abs <- colSums(abs(contrib))
  }

  mu_all   <- safe_divide(sum_all, n_present)
  mu_ch    <- safe_divide(sum_ch, n_present_ch)
  mu_nc    <- safe_divide(sum_nc, n_present_nc)
  mean_abs <- safe_divide(sum_abs, n_present)

  tibble(
    model            = model_name,
    term             = feature_names,
    mu_all           = as.numeric(mu_all),
    mu_ch            = as.numeric(mu_ch),
    mu_nc            = as.numeric(mu_nc),
    mean_abs_contrib = as.numeric(mean_abs),
    n_present        = as.integer(n_present),
    n_present_ch     = as.integer(n_present_ch),
    n_present_nc     = as.integer(n_present_nc),
    present_rate     = as.numeric(safe_divide(n_present, nrow(x_ref)))
  ) %>%
    mutate(
      direction = case_when(
        is.na(mu_all) ~ NA_character_,
        mu_all >= 0   ~ "christian",
        TRUE          ~ "non_christian"
      )
    )
}


driver_theme <- function(base_size = 16) {
  theme_minimal(base_size = base_size) +
    theme(
      text               = element_text(colour = col_text),
      panel.grid.minor   = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.major.x = element_line(colour = col_grid, linewidth = 0.3),
      axis.title.y       = element_blank(),
      axis.title.x       = element_text(size = rel(1.0), margin = margin(t = 8)),
      axis.text.y        = element_text(size = rel(0.92)),
      axis.text.x        = element_text(size = rel(0.88)),
      strip.text          = element_text(face = "bold", size = rel(1.1)),
      legend.position     = "bottom",
      legend.box          = "horizontal",
      legend.text         = element_text(size = rel(0.88)),
      legend.title        = element_text(size = rel(0.92)),
      plot.title          = element_text(face = "bold", size = rel(1.2)),
      plot.subtitle       = element_text(colour = "grey40", size = rel(0.95)),
      plot.margin         = margin(10, 14, 10, 14)
    )
}

top_driver_plot <- function(tbl, top_n = 20, min_present_group = 2L) {
  plot_tbl <- bind_rows(
    tbl %>%
      filter(
        is.finite(mu_all),
        mu_all > 0,
        n_present_ch >= min_present_group,
        n_present_nc >= min_present_group
      ) %>%
      slice_max(mu_all, n = top_n, with_ties = FALSE),
    tbl %>%
      filter(
        is.finite(mu_all),
        mu_all < 0,
        n_present_ch >= min_present_group,
        n_present_nc >= min_present_group
      ) %>%
      slice_min(mu_all, n = top_n, with_ties = FALSE)
  ) %>%
    mutate(direction = factor(direction, levels = c("christian", "non_christian"))) %>%
    group_by(direction) %>%
    arrange(mu_all, .by_group = TRUE) %>%
    mutate(term_plot = paste(direction, sprintf("%03d", row_number()), term, sep = "|||")) %>%
    ungroup() %>%
    mutate(term_plot = factor(term_plot, levels = term_plot))

  seg_tbl <- plot_tbl %>%
    transmute(
      direction,
      term_plot,
      x    = pmin(mu_ch, mu_nc),
      xend = pmax(mu_ch, mu_nc)
    )

  point_tbl <- plot_tbl %>%
    select(direction, term_plot, present_rate, mu_ch, mu_nc) %>%
    pivot_longer(
      cols      = c(mu_ch, mu_nc),
      names_to  = "group",
      values_to = "mean_contrib"
    ) %>%
    mutate(
      group = recode(
        group,
        mu_ch = "present in christian docs",
        mu_nc = "present in non-christian docs"
      )
    )

  ggplot() +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.45,
               color = col_zero_line) +
    geom_segment(
      data = seg_tbl,
      aes(x = x, xend = xend, y = term_plot, yend = term_plot),
      linewidth = 0.85, color = col_segment
    ) +
    geom_point(
      data = point_tbl,
      aes(x = mean_contrib, y = term_plot, fill = group, size = present_rate),
      shape = 21, color = col_text, stroke = 0.3, alpha = 0.92
    ) +
    facet_wrap(~ direction, scales = "free_y", ncol = 2) +
    scale_y_discrete(labels = function(x) sub("^.*\\|\\|\\|\\d+\\|\\|\\|", "", x)) +
    scale_fill_manual(
      values = c(
        "present in christian docs"     = col_christian,
        "present in non-christian docs" = col_non_christian
      )
    ) +
    scale_size_continuous(
      range  = c(2.8, 6.5),
      labels = scales::percent_format(accuracy = 1)
    ) +
    labs(
      x    = "mean contribution when term is present",
      fill = NULL,
      size = "document prevalence"
    ) +
    driver_theme()
}


agreement_plot <- function(tbl, direction_keep, top_n = 30) {
  plot_tbl <- tbl %>%
    filter(direction == direction_keep, is.finite(mean_z)) %>%
    {
      if (direction_keep == "christian") {
        slice_max(., mean_z, n = top_n, with_ties = FALSE)
      } else {
        slice_min(., mean_z, n = top_n, with_ties = FALSE)
      }
    } %>%
    arrange(mean_z) %>%
    mutate(term = factor(term, levels = term))

  fill_col <- if (direction_keep == "christian") col_christian else col_non_christian

  ggplot(plot_tbl, aes(x = mean_z, y = term)) +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.45,
               color = col_zero_line) +
    geom_segment(
      aes(x = 0, xend = mean_z, yend = term),
      linewidth = 0.85, color = col_segment
    ) +
    geom_point(
      aes(size = mean_present_rate),
      shape = 21, fill = fill_col, color = col_text,
      stroke = 0.3, alpha = 0.92
    ) +
    scale_size_continuous(
      range  = c(2.8, 6.5),
      labels = scales::percent_format(accuracy = 1)
    ) +
    labs(
      x    = "mean standardised present-only driver score (z)",
      size = "mean document prevalence"
    ) +
    driver_theme()
}


consensus_both_plot <- function(tbl, top_n_each = 25, min_present_group = 2L) {
  plot_tbl <- bind_rows(
    tbl %>%
      filter(
        direction == "christian",
        is.finite(mean_z),
        mean_n_present_ch >= min_present_group,
        mean_n_present_nc >= min_present_group
      ) %>%
      slice_max(mean_z, n = top_n_each, with_ties = FALSE),
    tbl %>%
      filter(
        direction == "non_christian",
        is.finite(mean_z),
        mean_n_present_ch >= min_present_group,
        mean_n_present_nc >= min_present_group
      ) %>%
      slice_min(mean_z, n = top_n_each, with_ties = FALSE)
  ) %>%
    arrange(mean_z) %>%
    mutate(
      term      = factor(term, levels = term),
      direction = factor(direction, levels = c("non_christian", "christian"))
    )

  ggplot(plot_tbl, aes(x = mean_z, y = term)) +
    geom_vline(xintercept = 0, linetype = "dashed", linewidth = 0.45,
               color = col_zero_line) +
    geom_segment(
      aes(x = 0, xend = mean_z, yend = term),
      linewidth = 0.85, color = col_segment
    ) +
    geom_point(
      aes(fill = direction, size = mean_present_rate),
      shape = 21, color = col_text, stroke = 0.3, alpha = 0.92
    ) +
    scale_fill_manual(
      values = c("christian" = col_christian, "non_christian" = col_non_christian)
    ) +
    scale_size_continuous(
      range  = c(2.8, 6.5),
      labels = scales::percent_format(accuracy = 1)
    ) +
    labs(
      x    = "mean standardised present-only driver score (z)",
      y    = NULL,
      fill = NULL,
      size = "mean document prevalence"
    ) +
    driver_theme() +
    theme(panel.grid.major.y = element_blank())
}
```

## Vectorization shared across models

```{r}
y <- factor(df_labeled$manual_label, levels = c("non_christian", "christian"))

vec_full <- fit_vectorizer(df_labeled$text, latin_stops)
x_counts_dfm <- transform_counts_dfm(df_labeled$text, vec_full, latin_stops)
x_counts <- as(x_counts_dfm, "dgCMatrix")
x_tfidf <- transform_tfidf_sparse(df_labeled$text, vec_full, latin_stops)

feature_names <- colnames(x_tfidf)
if (is.null(feature_names)) feature_names <- vec_full$features
colnames(x_counts) <- feature_names
colnames(x_tfidf) <- feature_names
```

## Full-data model refits under saved hyperparameters

```{r}
get_model_row <- function(model_name) {
  hyper_full %>% filter(model == model_name) %>% slice(1)
}

n_pos <- sum(y == "christian")
n_neg <- sum(y == "non_christian")
scale_pos_weight_full <- n_neg / max(n_pos, 1)

nb_row    <- get_model_row("naive_bayes")
ridge_row <- get_model_row("ridge")
svm_row   <- get_model_row("svm_platt")
xgb_row   <- get_model_row("xgboost")
lgb_row   <- get_model_row("lightgbm")

nb_fit <- fit_nb_model(x_counts_dfm, y)

ridge_alpha <- if ("alpha" %in% names(ridge_row) && !is.na(ridge_row$alpha[[1]])) {
  as.numeric(ridge_row$alpha[[1]])
} else {
  message("NOTE: 'alpha' column not found in model_hyperparameters_full.csv — using alpha = 0 (pure ridge).")
  0
}

message(sprintf("Ridge/elastic-net model: alpha = %s, lambda = %s",
                ridge_alpha, ridge_row$lambda_1se[[1]]))

ridge_fit <- fit_enet_fixed(
  x_tfidf,
  y,
  alpha  = ridge_alpha,
  lambda = as.numeric(ridge_row$lambda_1se[[1]])
)

svm_fit <- fit_svm_fixed(
  x_tfidf,
  y,
  cost = as.numeric(svm_row$cost[[1]])
)

xgb_fit <- fit_xgb_fixed(
  x_tfidf,
  y,
  params  = make_xgb_params(xgb_row, scale_pos_weight = scale_pos_weight_full),
  nrounds = as.integer(xgb_row$best_nrounds[[1]]),
  seed    = 6204
)

lgb_fit <- fit_lgb_fixed(
  x      = x_tfidf,
  y      = y,
  params  = make_lgb_params(lgb_row, seed = 6205),
  nrounds = as.integer(lgb_row$best_nrounds[[1]])
)
```

## Model-specific contribution matrices

```{r}
nb_weights    <- extract_nb_weights(x_counts, y, feature_names, alpha = alpha_nb)


ridge_weights <- tibble(term = feature_names, weight = extract_enet_weights(ridge_fit, feature_names))
svm_weights   <- tibble(term = feature_names, weight = extract_svm_weights(svm_fit, feature_names))

nb_contrib    <- sparse_weight_contrib(x_counts, nb_weights$weight)
ridge_contrib <- sparse_weight_contrib(x_tfidf,  ridge_weights$weight)
svm_contrib   <- sparse_weight_contrib(x_tfidf,  svm_weights$weight)

xgb_contrib_obj <- predict_xgb_contrib(xgb_fit, x_tfidf, feature_names)
xgb_contrib     <- mask_dense_contrib_to_present(xgb_contrib_obj$shap, x_tfidf)

lgb_contrib_obj <- predict_lgb_contrib(
  lgb_fit,
  x_tfidf,
  feature_names,
  num_iteration = as.integer(lgb_row$best_nrounds[[1]])
)
lgb_contrib <- mask_dense_contrib_to_present(lgb_contrib_obj$shap, x_tfidf)


n_nonzero_ridge <- sum(ridge_weights$weight != 0)
message(sprintf("Elastic net: %d / %d features have non-zero coefficients (%.1f%%)",
                n_nonzero_ridge, length(feature_names),
                100 * n_nonzero_ridge / length(feature_names)))
```

## Per-model driver tables

```{r}
driver_nb    <- summarize_driver_matrix(nb_contrib,    x_counts, y, feature_names, "naive_bayes")
driver_ridge <- summarize_driver_matrix(ridge_contrib, x_tfidf,  y, feature_names, "ridge")
driver_svm   <- summarize_driver_matrix(svm_contrib,   x_tfidf,  y, feature_names, "svm_platt")
driver_xgb   <- summarize_driver_matrix(xgb_contrib,   x_tfidf,  y, feature_names, "xgboost")
driver_lgb   <- summarize_driver_matrix(lgb_contrib,   x_tfidf,  y, feature_names, "lightgbm")

driver_tbl <- bind_rows(
  driver_nb,
  driver_ridge,
  driver_svm,
  driver_xgb,
  driver_lgb
) %>%
  group_by(model) %>%
  mutate(
    z_mu_all = zscore(mu_all),
    z_abs    = zscore(mean_abs_contrib),
    rank_abs = min_rank(desc(mean_abs_contrib)),
    rank_pos = min_rank(desc(mu_all)),
    rank_neg = min_rank(mu_all)
  ) %>%
  ungroup()

readr::write_csv(driver_tbl, file.path(fig_dir, "feature_driver_table_all_models.csv"))
```

## Individual model plots and tables

```{r}
plot_nb    <- top_driver_plot(driver_nb,    top_n = 20, min_present_group = min_present_group)
plot_ridge <- top_driver_plot(driver_ridge, top_n = 20, min_present_group = min_present_group)
plot_svm   <- top_driver_plot(driver_svm,   top_n = 20, min_present_group = min_present_group)
plot_xgb   <- top_driver_plot(driver_xgb,   top_n = 20, min_present_group = min_present_group)
plot_lgb   <- top_driver_plot(driver_lgb,   top_n = 20, min_present_group = min_present_group)

ggsave(file.path(fig_dir, "top_drivers_naive_bayes.png"),  plot_nb,    width = 12, height = 8.5, dpi = 320)
ggsave(file.path(fig_dir, "top_drivers_ridge.png"),        plot_ridge, width = 12, height = 8.5, dpi = 320)
ggsave(file.path(fig_dir, "top_drivers_svm_platt.png"),    plot_svm,   width = 12, height = 8.5, dpi = 320)
ggsave(file.path(fig_dir, "top_drivers_xgboost.png"),      plot_xgb,   width = 12, height = 8.5, dpi = 320)
ggsave(file.path(fig_dir, "top_drivers_lightgbm.png"),     plot_lgb,   width = 12, height = 8.5, dpi = 320)

top_driver_table <- driver_tbl %>%
  filter(
    is.finite(mu_all),
    n_present_ch >= min_present_group,
    n_present_nc >= min_present_group
  ) %>%
  group_by(model) %>%
  group_modify(
    ~ bind_rows(
      slice_max(.x, mu_all, n = 30, with_ties = FALSE) %>% mutate(side = "christian"),
      slice_min(.x, mu_all, n = 30, with_ties = FALSE) %>% mutate(side = "non_christian")
    )
  ) %>%
  ungroup() %>%
  arrange(model, side, desc(abs(mu_all)), desc(present_rate))

readr::write_csv(top_driver_table, file.path(fig_dir, "top_driver_table_by_model.csv"))
```

## Correlation of feature-driver profiles across models

```{r}
driver_wide <- driver_tbl %>%
  select(model, term, z_mu_all) %>%
  pivot_wider(names_from = model, values_from = z_mu_all)

driver_cor <- cor(as.matrix(driver_wide %>% select(-term)), use = "pairwise.complete.obs")

model_order <- c("naive_bayes", "ridge", "svm_platt", "xgboost", "lightgbm")

driver_cor_long <- as.data.frame(as.table(driver_cor), stringsAsFactors = FALSE) %>%
  setNames(c("model_1", "model_2", "correlation")) %>%
  tibble::as_tibble() %>%
  mutate(
    model_1 = factor(model_1, levels = model_order),
    model_2 = factor(model_2, levels = rev(model_order))
  )

readr::write_csv(driver_cor_long, file.path(fig_dir, "feature_driver_correlations.csv"))

p_cor <- ggplot(driver_cor_long, aes(x = model_1, y = model_2, fill = correlation)) +
  geom_tile(color = "white", linewidth = 0.7) +
  geom_text(aes(label = sprintf("%.2f", correlation)), size = 5, colour = col_text) +
  scale_fill_gradient2(
    low      = col_non_christian,
    mid      = "white",
    high     = col_christian,
    midpoint = 0,
    limits   = c(-1, 1)
  ) +
  coord_equal() +
  labs(x = NULL, y = NULL, fill = "correlation") +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid      = element_blank(),
    axis.text.x     = element_text(angle = 45, hjust = 1, colour = col_text, size = 14),
    axis.text.y     = element_text(colour = col_text, size = 14),
    legend.position = "bottom",
    legend.text     = element_text(size = 12),
    legend.title    = element_text(size = 13),
    plot.margin     = margin(10, 14, 10, 14)
  )

ggsave(file.path(fig_dir, "feature_driver_model_correlation_heatmap.png"),
       p_cor, width = 8.8, height = 7.4, dpi = 320)

p_cor
```

## Cross-model agreement on Christian and non-Christian drivers

```{r}
n_models <- n_distinct(driver_tbl$model)

consensus_tbl <- driver_tbl %>%
  filter(!is.na(mu_all)) %>%
  group_by(term) %>%
  summarise(
    n_models            = n(),
    christian_votes     = sum(mu_all > 0, na.rm = TRUE),
    non_christian_votes = sum(mu_all < 0, na.rm = TRUE),
    mean_z              = mean(z_mu_all, na.rm = TRUE),
    median_z            = median(z_mu_all, na.rm = TRUE),
    mean_abs_z          = mean(abs(z_mu_all), na.rm = TRUE),
    mean_abs_contrib    = mean(mean_abs_contrib, na.rm = TRUE),
    mean_rank_abs       = mean(rank_abs, na.rm = TRUE),
    mean_present_rate   = mean(present_rate, na.rm = TRUE),
    mean_n_present      = mean(n_present, na.rm = TRUE),
    mean_n_present_ch   = mean(n_present_ch, na.rm = TRUE),
    mean_n_present_nc   = mean(n_present_nc, na.rm = TRUE),
    agreement_prop      = pmax(christian_votes, non_christian_votes) / n_models,
    all_agree           = agreement_prop == 1,
    .groups             = "drop"
  ) %>%
  mutate(
    direction = case_when(
      christian_votes == n_models     ~ "christian",
      non_christian_votes == n_models ~ "non_christian",
      mean_z > 0                      ~ "christian",
      mean_z < 0                      ~ "non_christian",
      TRUE                            ~ "mixed"
    )
  ) %>%
  arrange(desc(agreement_prop), desc(abs(mean_z)))

readr::write_csv(consensus_tbl, file.path(fig_dir, "feature_driver_consensus_all_models.csv"))

consensus_all_agree_ch <- consensus_tbl %>%
  filter(
    christian_votes == n_models,
    mean_n_present_ch >= min_present_group,
    mean_n_present_nc >= min_present_group
  ) %>%
  arrange(desc(mean_z), desc(mean_abs_z))

consensus_all_agree_nc <- consensus_tbl %>%
  filter(
    non_christian_votes == n_models,
    mean_n_present_ch >= min_present_group,
    mean_n_present_nc >= min_present_group
  ) %>%
  arrange(mean_z, desc(mean_abs_z))

consensus_top_both <- bind_rows(
  consensus_tbl %>%
    filter(
      direction == "christian",
      mean_n_present_ch >= min_present_group,
      mean_n_present_nc >= min_present_group
    ) %>%
    slice_max(mean_z, n = 30, with_ties = FALSE),
  consensus_tbl %>%
    filter(
      direction == "non_christian",
      mean_n_present_ch >= min_present_group,
      mean_n_present_nc >= min_present_group
    ) %>%
    slice_min(mean_z, n = 30, with_ties = FALSE)
) %>%
  arrange(mean_z)

readr::write_csv(consensus_all_agree_ch, file.path(fig_dir, "feature_driver_consensus_christian_unanimous.csv"))
readr::write_csv(consensus_all_agree_nc, file.path(fig_dir, "feature_driver_consensus_non_christian_unanimous.csv"))
readr::write_csv(consensus_top_both,     file.path(fig_dir, "feature_driver_consensus_top_both.csv"))

p_cons_both <- consensus_both_plot(consensus_tbl, top_n_each = 25, min_present_group = min_present_group)
p_cons_ch   <- agreement_plot(consensus_all_agree_ch, "christian",     top_n = 30)
p_cons_nc   <- agreement_plot(consensus_all_agree_nc, "non_christian", top_n = 30)

ggsave(file.path(fig_dir, "feature_driver_consensus_both.png"),
       p_cons_both, width = 11.2, height = 9.2, dpi = 320)
ggsave(file.path(fig_dir, "feature_driver_consensus_christian_unanimous.png"),
       p_cons_ch, width = 10.5, height = 8.2, dpi = 320)
ggsave(file.path(fig_dir, "feature_driver_consensus_non_christian_unanimous.png"),
       p_cons_nc, width = 10.5, height = 8.2, dpi = 320)

p_cons_both
p_cons_ch
p_cons_nc
```

## Agreement summary by model and term

```{r}
agreement_detail <- driver_tbl %>%
  select(model, term, mu_all, z_mu_all, mean_abs_contrib, rank_abs, direction,
         n_present, n_present_ch, n_present_nc, present_rate) %>%
  left_join(
    consensus_tbl %>%
      select(
        term,
        christian_votes,
        non_christian_votes,
        agreement_prop,
        all_agree,
        mean_z,
        mean_present_rate,
        mean_n_present_ch,
        mean_n_present_nc
      ),
    by = "term"
  ) %>%
  arrange(desc(all_agree), desc(agreement_prop), desc(abs(z_mu_all)))

readr::write_csv(agreement_detail, file.path(fig_dir, "feature_driver_agreement_detail.csv"))

agreement_detail %>%
  filter(all_agree) %>%
  arrange(desc(abs(mean_z))) %>%
  slice_head(n = 20)
```
