---
title: "Latin text classification"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of christian vs. non-christian documents


## Data loading and cleaning

### Loading in and filtering the data (this code chunk can be skipped if corpus.csv already exists in the script directory)

```{r}
library(tidyverse)
library(stringi)

data_path <- "../../data/processed_texts/600ad"
results_path <- "../../results/predictions/non_transformer_lemmatized/600ad"
```


```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels <- readr::read_csv(file.path(data_path,"corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  transmute(id, manual_label = label) %>%
  distinct(id, .keep_all = TRUE)

df_labeled <- df %>%
  inner_join(labels, by = "id") 

```


### Loading in the stopwowrds

```{r}

custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href", 
  "fridericus", "heinricus", "conradus", "meus", "tuus" 
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")

stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- c(stops_v1, stops_v2, custom_additions)

latin_stops <- latin_stops[latin_stops != "*"]

latin_stops <- unique(latin_stops)

latin_stops <- sort(latin_stops)

head(latin_stops, 200)

```



### Defining a tokenization function to count relative frequencies per token for the NB classifier


```{r}
library(quanteda)
library(quanteda.textmodels)
library(tidyverse)
library(knitr)


get_standard_dfm <- function(text_data, stop_list) {

  toks <- tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_remove(pattern = stop_list)
  

  dfm(toks)
}


dfm_train <- get_standard_dfm(df_labeled$text, latin_stops)

dim(dfm_train)

```


## Training classification models

### Setting up model parameters


```{r}

library(tidyverse)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(pROC)

set.seed(123)

checkpoint_path <- "../../results/predictions/non_transformer_lemmatized/600ad"

k_folds_outer <- 5L
k_folds_inner <- 5L
xgb_max_rounds <- 4000L
lgb_max_rounds <- 4000L
early_stop_rounds <- 100L
min_docfreq <- 3L

checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
output_dir <- file.path(checkpoint_path, "model_outputs_checkpointed")
progress_path <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

append_progress_row <- function(row, path) {
  if (!file.exists(path)) {
    write_csv_atomic(row, path)
  } else {
    old <- readr::read_csv(path, show_col_types = FALSE)
    write_csv_atomic(bind_rows(old, row), path)
  }
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step, if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

safe_prob <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb))) {
    stop(paste0("Both classes are required", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
  if (any(tb < 2)) {
    stop(paste0("Each class needs at least 2 observations", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for cross-validation.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

make_foldid <- function(y, k, seed = 123) {
  folds <- make_folds(y, k, seed)
  foldid <- integer(length(y))
  for (i in seq_along(folds)) foldid[folds[[i]]] <- i
  foldid
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = s,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1)
  tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1)
  fn <- sum(y01 == 1 & p01 == 0)
  precision <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1 <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(
    accuracy = (tp + tn) / length(y),
    balanced_accuracy = bal_acc,
    precision = precision,
    recall = recall,
    specificity = specificity,
    f1 = f1,
    mcc = mcc,
    tp = tp,
    tn = tn,
    fp = fp,
    fn = fn
  )
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(
          response = y,
          predictor = p,
          levels = c("non_christian", "christian"),
          direction = "<",
          quiet = TRUE
        ),
        x = "best",
        input = "threshold",
        best.method = "youden",
        ret = "threshold",
        transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01 <- as.integer(y == "christian")
    cand <- sort(unique(p))
    cand <- unique(c(0, cand, 0.5, 1))
    f1s <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    tp_youden = hy$tp,
    tn_youden = hy$tn,
    fp_youden = hy$fp,
    fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc,
    tp_f1 = hf$tp,
    tn_f1 = hf$tn,
    fp_f1 = hf$fp,
    fn_f1 = hf$fn
  )
}

summarize_model_fold <- function(y, p, pred_youden, pred_f1, model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold,
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc
  )
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]), by = keys)
  bind_rows(keep, new_rows)
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(), control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores), type = "response"))
  }
}

make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat <- as(x_counts, "dgCMatrix")
  x_vals <- x_mat@x
  x_mat@x <- 1 + log(x_vals)
  idf_vec <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}

nb_prob_christian <- function(pred_obj) {
  if ("christian" %in% colnames(pred_obj)) safe_prob(pred_obj[, "christian"]) else safe_prob(pred_obj[, ncol(pred_obj)])
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}

predict_nb_score <- function(fit, newx_dfm) {
  safe_logit(nb_prob_christian(predict(fit, newdata = newx_dfm, type = "probability")))
}

enet_alpha_grid <- c(0.05, 0.1, 0.2, 0.3, 0.4, 0.5)

fit_enet_fixed <- function(x, y, alpha, lambda) {
  glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    lambda = lambda,
    standardize = TRUE,
    maxit = 1e6
  )
}

enet_has_nonzero <- function(fit) {
  cf <- tryCatch(as.matrix(coef(fit)), error = function(e) NULL)
  if (is.null(cf) || nrow(cf) <= 1) return(FALSE)
  any(abs(cf[-1, 1]) > 0)
}

predict_enet_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_enet_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

enet_lambda_grid <- function(x, y, alpha, nlambda = 60, lambda.min.ratio = 1e-3) {
  base_fit <- glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    standardize = TRUE,
    nlambda = nlambda,
    lambda.min.ratio = lambda.min.ratio,
    maxit = 1e6
  )
  as.numeric(base_fit$lambda)
}

tune_enet_lambda_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir, nlambda = 60, lambda.min.ratio = 1e-3) {
  ensure_dir(ckpt_dir)
  folds <- make_folds(y, inner_k, seed)
  alpha_rows <- vector("list", length(enet_alpha_grid))
  for (a_idx in seq_along(enet_alpha_grid)) {
    alpha <- enet_alpha_grid[a_idx]
    lambda_grid <- enet_lambda_grid(x, y, alpha = alpha, nlambda = nlambda, lambda.min.ratio = lambda.min.ratio)
    tuning_rows <- vector("list", length(lambda_grid))
    for (g in seq_along(lambda_grid)) {
      lam <- lambda_grid[g]
      fold_losses <- rep(NA_real_, length(folds))
      for (j in seq_along(folds)) {
        log_progress("enet_tuning_fold", paste(prefix, "alpha", alpha, "lambda", g, "of", length(lambda_grid), "fold", j, "of", length(folds)))
        va <- folds[[j]]
        tr <- setdiff(seq_along(y), va)
        fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lam)
        if (!enet_has_nonzero(fit)) {
          fold_losses[j] <- Inf
        } else {
          p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
          fold_losses[j] <- log_loss_safe(y[va], p_va)
        }
      }
      cvm <- mean(fold_losses, na.rm = TRUE)
      cvsd <- stats::sd(fold_losses, na.rm = TRUE) / sqrt(sum(is.finite(fold_losses)))
      tuning_rows[[g]] <- tibble(alpha = alpha, lambda = lam, cvm = cvm, cvsd = cvsd)
    }
    alpha_tuning <- bind_rows(tuning_rows) %>% arrange(desc(lambda))
    idx_min <- which.min(alpha_tuning$cvm)
    lambda_min <- alpha_tuning$lambda[idx_min]
    cutoff <- alpha_tuning$cvm[idx_min] + alpha_tuning$cvsd[idx_min]
    lambda_1se <- max(alpha_tuning$lambda[alpha_tuning$cvm <= cutoff], na.rm = TRUE)
    alpha_rows[[a_idx]] <- tibble(alpha = alpha, lambda_1se = as.numeric(lambda_1se), lambda_min = as.numeric(lambda_min), cvm_min = alpha_tuning$cvm[idx_min])
  }
  alpha_summary <- bind_rows(alpha_rows) %>% arrange(cvm_min, alpha)
  best <- alpha_summary[1, , drop = FALSE]
  list(alpha = as.numeric(best$alpha), lambda = as.numeric(best$lambda_1se), lambda_min = as.numeric(best$lambda_min), tuning = alpha_summary)
}

cv_enet_scores_ckpt <- function(x, y, alpha, lambda, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("enet_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    state$scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

cv_svm_scores_ckpt <- function(x, y, cost, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("svm_calibration_fold", paste(prefix, "cost", cost, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    state$scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_svm_cost_ckpt <- function(x, y, cost_grid = c(0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 8, 16), inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  rows_path <- file.path(ckpt_dir, "summary.rds")
  state <- load_rds_or_null(rows_path)
  if (is.null(state) || length(state$done) != length(cost_grid)) {
    state <- list(done = rep(FALSE, length(cost_grid)), rows = vector("list", length(cost_grid)))
  }
  for (i in seq_along(cost_grid)) {
    if (isTRUE(state$done[i])) next
    scores_i <- cv_svm_scores_ckpt(
      x = x,
      y = y,
      cost = cost_grid[i],
      k = inner_k,
      seed = seed + i,
      prefix = prefix,
      ckpt_dir = file.path(ckpt_dir, sprintf("cost_%03d", i))
    )
    state$rows[[i]] <- tibble(cost = cost_grid[i], auc_margin = auc_margin_safe(y, scores_i))
    state$done[i] <- TRUE
    save_rds_atomic(state, rows_path)
  }
  tuning <- bind_rows(state$rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}

tune_xgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- xgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  n_pos <- sum(y == "christian")
  n_neg <- sum(y == "non_christian")
  spw <- n_neg / max(n_pos, 1)

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("xgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "christian")
      n_neg_tr <- sum(y[tr] == "non_christian")
      spw_tr <- n_neg_tr / max(n_pos_tr, 1)
      dtr <- xgboost::xgb.DMatrix(x[tr, , drop = FALSE], label = as.integer(y[tr] == "christian"))
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE], label = as.integer(y[va] == "christian"))
      fit <- xgboost::xgb.train(
        params = make_xgb_params(pars_row, scale_pos_weight = spw_tr),
        data = dtr,
        nrounds = xgb_max_rounds,
        evals = list(valid = dva),
        early_stopping_rounds = early_stop_rounds,
        verbose = 0
      )
      cand_state$oof_prob[va] <- safe_prob(as.numeric(predict(fit, dva)))
      cand_state$best_iters[j] <- best_iter_xgb(fit, xgb_max_rounds)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_xgb_params(best, scale_pos_weight = spw), tuning = tuning)
}

cv_xgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("xgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "christian")
    n_neg_tr <- sum(y[tr] == "non_christian")
    fold_params <- modifyList(params, list(scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr], params = fold_params, nrounds = nrounds, seed = seed + j)
    state$scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_lgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- lgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("lgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      lgb_params_fold <- make_lgb_params(pars_row, seed = seed + 1000L * g + j)
      result <- fit_lgb_early_stop(
        x_tr = x[tr, , drop = FALSE],
        y_tr = y[tr],
        x_va = x[va, , drop = FALSE],
        y_va = y[va],
        params = lgb_params_fold,
        max_rounds = lgb_max_rounds,
        early_rounds = early_stop_rounds
      )
      cand_state$oof_prob[va] <- predict_lgb_prob(result$model, x[va, , drop = FALSE], num_iteration = result$best_iter)
      cand_state$best_iters[j] <- result$best_iter
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_lgb_params(best, seed = seed + 9999L), tuning = tuning)
}

cv_lgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("lgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE],
      y = y[tr],
      params = modifyList(
        params,
        list(
          seed = seed + j,
          feature_fraction_seed = seed + j,
          bagging_seed = seed + j
        )
      ),
      nrounds = nrounds
    )
    state$scores[va] <- predict_lgb_score(fit, x[va, , drop = FALSE], num_iteration = nrounds)
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

run_lgb_preflight <- function(text_vec, y, stop_list, seed = 999L) {
  folds <- make_folds(y, 2L, seed)
  va <- folds[[1]]
  tr <- setdiff(seq_along(y), va)
  vec <- fit_vectorizer(text_vec[tr], stop_list)
  x_tr <- transform_tfidf_sparse(text_vec[tr], vec, stop_list)
  x_va <- transform_tfidf_sparse(text_vec[va], vec, stop_list)
  pars_row <- lgb_grid()[1, , drop = FALSE]
  pars <- make_lgb_params(pars_row, seed = seed)
  result <- fit_lgb_early_stop(
    x_tr = x_tr,
    y_tr = y[tr],
    x_va = x_va,
    y_va = y[va],
    params = pars,
    max_rounds = 100L,
    early_rounds = 20L
  )
  if (!is.finite(result$best_iter)) stop("LightGBM preflight failed to select a valid num_iteration.")
  invisible(result)
}

save_model_object <- function(model, path, engine) {
  ensure_dir(dirname(path))
  if (engine == "xgboost") {
    xgboost::xgb.save(model, path)
  } else if (engine == "lightgbm") {
    lightgbm::lgb.save(model, path)
  } else {
    save_rds_atomic(model, path)
  }
  invisible(path)
}

init_run_state <- function(n_lab) {
  list(
    preflight_done = FALSE,
    outer_fold_id = integer(n_lab),

    nb_oof = rep(NA_real_, n_lab),
    enet_oof = rep(NA_real_, n_lab),
    svm_oof = rep(NA_real_, n_lab),
    xgb_oof = rep(NA_real_, n_lab),
    lgb_oof = rep(NA_real_, n_lab),

    nb_score_oof = rep(NA_real_, n_lab),
    enet_score_oof = rep(NA_real_, n_lab),
    svm_score_oof = rep(NA_real_, n_lab),
    xgb_score_oof = rep(NA_real_, n_lab),
    lgb_score_oof = rep(NA_real_, n_lab),

    nb_pred_youden_oof = rep(NA_character_, n_lab),
    enet_pred_youden_oof = rep(NA_character_, n_lab),
    svm_pred_youden_oof = rep(NA_character_, n_lab),
    xgb_pred_youden_oof = rep(NA_character_, n_lab),
    lgb_pred_youden_oof = rep(NA_character_, n_lab),

    nb_pred_f1_oof = rep(NA_character_, n_lab),
    enet_pred_f1_oof = rep(NA_character_, n_lab),
    svm_pred_f1_oof = rep(NA_character_, n_lab),
    xgb_pred_f1_oof = rep(NA_character_, n_lab),
    lgb_pred_f1_oof = rep(NA_character_, n_lab),

    outer_threshold_df = tibble(),
    fold_metric_df = tibble(),
    selected_outer_df = tibble(),
    model_done = list(),
    full_done = list()
  )
}

mark_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  state[[bucket]][[key]] <- TRUE
  state
}

is_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  isTRUE(state[[bucket]][[key]])
}

state_path <- file.path(checkpoint_root, "run_state.rds")

log_progress("load_data", "reading corpus and labels")

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

dup_corpus <- df %>% count(id, name = "n") %>% filter(n > 1)
if (nrow(dup_corpus) > 0) {
  stop(paste0("Duplicate ids in corpus_600ad.csv: ", paste(head(dup_corpus$id, 20), collapse = ", ")))
}

labels_raw <- readr::read_csv(file.path(data_path, "corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  select(id, label)

label_conflicts <- labels_raw %>%
  distinct(id, label) %>%
  count(id, name = "n_labels") %>%
  filter(n_labels > 1)

if (nrow(label_conflicts) > 0) {
  stop(paste0("Conflicting labels in corpus_600ad_labeled.csv: ", paste(head(label_conflicts$id, 20), collapse = ", ")))
}

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
check_binary_y(df_labeled$manual_label, "labeled data")

custom_additions <- c("m", "c", "l", "d", "gt", "lt", "br", "href", "fridericus", "heinricus", "conradus", "meus", "tuus")
stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")
latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]

lab <- df_labeled %>% arrange(row_id)
out <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos <- match(lab$row_id, out$row_id)

outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)
n_lab <- nrow(lab)

state <- load_rds_or_null(state_path)
if (is.null(state) || length(state$nb_oof) != n_lab) {
  state <- init_run_state(n_lab)
  save_rds_atomic(state, state_path)
}

if (!isTRUE(state$preflight_done)) {
  log_progress("lgb_preflight_start", "testing LightGBM before the full run")
  invisible(run_lgb_preflight(lab$text, lab$manual_label, latin_stops, seed = 999L))
  state$preflight_done <- TRUE
  save_rds_atomic(state, state_path)
  log_progress("lgb_preflight_done", "LightGBM preflight passed")
}

for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start", paste("fold", i, "of", length(outer_folds)))
  va <- outer_folds[[i]]
  tr <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]

  outer_dir <- file.path(checkpoint_root, sprintf("outer_fold_%02d", i))
  ensure_dir(outer_dir)

  vec_fold <- fit_vectorizer(lab$text[tr], latin_stops)
  x_tr_dfm <- transform_counts_dfm(lab$text[tr], vec_fold, latin_stops)
  x_va_dfm <- transform_counts_dfm(lab$text[va], vec_fold, latin_stops)
  x_tr <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
  x_va <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)

  state$outer_fold_id[va] <- i
  save_rds_atomic(state, state_path)

  key <- paste0("outer_", i, "_lightgbm")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_lgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "lightgbm")
    lgb_tuned <- tune_lgb_model_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 5000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    lgb_train_scores <- cv_lgb_scores_ckpt(
      x_tr, y_tr,
      params = lgb_tuned$params,
      nrounds = as.integer(lgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner,
      seed = 5100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    lgb_platt <- fit_platt(lgb_train_scores, y_tr)
    lgb_fit <- fit_lgb_fixed(
      x = x_tr,
      y = y_tr,
      params = lgb_tuned$params,
      nrounds = as.integer(lgb_tuned$best$best_nrounds[1])
    )
    lgb_va_scores <- predict_lgb_score(lgb_fit, x_va, num_iteration = as.integer(lgb_tuned$best$best_nrounds[1]))
    lgb_train_prob <- apply_platt(lgb_platt, lgb_train_scores)
    lgb_va_prob <- apply_platt(lgb_platt, lgb_va_scores)
    lgb_th_y <- choose_threshold(y_tr, lgb_train_prob, "youden")
    lgb_th_f <- choose_threshold(y_tr, lgb_train_prob, "f1")

    state$lgb_oof[va] <- lgb_va_prob
    state$lgb_score_oof[va] <- lgb_va_scores
    state$lgb_pred_youden_oof[va] <- ifelse(lgb_va_prob >= lgb_th_y, "christian", "non_christian")
    state$lgb_pred_f1_oof[va] <- ifelse(lgb_va_prob >= lgb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "lightgbm", threshold_youden = lgb_th_y, threshold_f1 = lgb_th_f)
    metric_row <- summarize_model_fold(y_va, lgb_va_prob, state$lgb_pred_youden_oof[va], state$lgb_pred_f1_oof[va], "lightgbm", i)
    selected_row <- lgb_tuned$best %>%
      mutate(
        outer_fold = i,
        model = "lightgbm",
        threshold_youden = lgb_th_y,
        threshold_f1 = lgb_th_f
      )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(best = lgb_tuned$best, tuning = lgb_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_naive_bayes")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_nb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "naive_bayes")
    nb_train_scores <- cv_nb_scores_ckpt(
      x_tr_dfm, y_tr,
      k = k_folds_inner,
      seed = 1000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    nb_platt <- fit_platt(nb_train_scores, y_tr)
    nb_fit <- fit_nb_model(x_tr_dfm, y_tr)
    nb_va_scores <- predict_nb_score(nb_fit, x_va_dfm)
    nb_train_prob <- apply_platt(nb_platt, nb_train_scores)
    nb_va_prob <- apply_platt(nb_platt, nb_va_scores)
    nb_th_y <- choose_threshold(y_tr, nb_train_prob, "youden")
    nb_th_f <- choose_threshold(y_tr, nb_train_prob, "f1")

    state$nb_oof[va] <- nb_va_prob
    state$nb_score_oof[va] <- nb_va_scores
    state$nb_pred_youden_oof[va] <- ifelse(nb_va_prob >= nb_th_y, "christian", "non_christian")
    state$nb_pred_f1_oof[va] <- ifelse(nb_va_prob >= nb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "naive_bayes", threshold_youden = nb_th_y, threshold_f1 = nb_th_f)
    metric_row <- summarize_model_fold(y_va, nb_va_prob, state$nb_pred_youden_oof[va], state$nb_pred_f1_oof[va], "naive_bayes", i)
    selected_row <- tibble(outer_fold = i, model = "naive_bayes", threshold_youden = nb_th_y, threshold_f1 = nb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_enet")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_enet_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "enet")
    enet_tuned <- tune_enet_lambda_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 2000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    enet_train_scores <- cv_enet_scores_ckpt(
      x_tr, y_tr,
      alpha = enet_tuned$alpha,
      lambda = enet_tuned$lambda,
      k = k_folds_inner,
      seed = 2100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    enet_platt <- fit_platt(enet_train_scores, y_tr)
    enet_fit <- fit_enet_fixed(x_tr, y_tr, alpha = enet_tuned$alpha, lambda = enet_tuned$lambda)
    enet_va_scores <- predict_enet_score(enet_fit, x_va)
    enet_train_prob <- apply_platt(enet_platt, enet_train_scores)
    enet_va_prob <- apply_platt(enet_platt, enet_va_scores)
    enet_th_y <- choose_threshold(y_tr, enet_train_prob, "youden")
    enet_th_f <- choose_threshold(y_tr, enet_train_prob, "f1")

    state$enet_oof[va] <- enet_va_prob
    state$enet_score_oof[va] <- enet_va_scores
    state$enet_pred_youden_oof[va] <- ifelse(enet_va_prob >= enet_th_y, "christian", "non_christian")
    state$enet_pred_f1_oof[va] <- ifelse(enet_va_prob >= enet_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "enet", threshold_youden = enet_th_y, threshold_f1 = enet_th_f)
    metric_row <- summarize_model_fold(y_va, enet_va_prob, state$enet_pred_youden_oof[va], state$enet_pred_f1_oof[va], "enet", i)
    selected_row <- tibble(
      outer_fold = i,
      model = "enet",
      alpha = enet_tuned$alpha,
      lambda_1se = enet_tuned$lambda,
      lambda_min = enet_tuned$lambda_min,
      threshold_youden = enet_th_y,
      threshold_f1 = enet_th_f
    )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(tuning = enet_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_svm_platt")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_svm_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "svm_platt")
    svm_tuned <- tune_svm_cost_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 3000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    svm_train_scores <- cv_svm_scores_ckpt(
      x_tr, y_tr,
      cost = svm_tuned$cost,
      k = k_folds_inner,
      seed = 3100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    svm_platt <- fit_platt(svm_train_scores, y_tr)
    svm_fit <- fit_svm_fixed(x_tr, y_tr, cost = svm_tuned$cost)
    svm_va_scores <- predict_svm_score(svm_fit, x_va)
    svm_train_prob <- apply_platt(svm_platt, svm_train_scores)
    svm_va_prob <- apply_platt(svm_platt, svm_va_scores)
    svm_th_y <- choose_threshold(y_tr, svm_train_prob, "youden")
    svm_th_f <- choose_threshold(y_tr, svm_train_prob, "f1")

    state$svm_oof[va] <- svm_va_prob
    state$svm_score_oof[va] <- svm_va_scores
    state$svm_pred_youden_oof[va] <- ifelse(svm_va_prob >= svm_th_y, "christian", "non_christian")
    state$svm_pred_f1_oof[va] <- ifelse(svm_va_prob >= svm_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "svm_platt", threshold_youden = svm_th_y, threshold_f1 = svm_th_f)
    metric_row <- summarize_model_fold(y_va, svm_va_prob, state$svm_pred_youden_oof[va], state$svm_pred_f1_oof[va], "svm_platt", i)
    selected_row <- tibble(
      outer_fold = i,
      model = "svm_platt",
      cost = svm_tuned$cost,
      threshold_youden = svm_th_y,
      threshold_f1 = svm_th_f
    )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(tuning = svm_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_xgboost")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_xgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "xgboost")
    xgb_tuned <- tune_xgb_model_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 4000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    xgb_train_scores <- cv_xgb_scores_ckpt(
      x_tr, y_tr,
      params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner,
      seed = 4100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    xgb_platt <- fit_platt(xgb_train_scores, y_tr)
    xgb_fit <- fit_xgb_fixed(
      x_tr, y_tr,
      params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      seed = 4200 + i
    )
    xgb_va_scores <- predict_xgb_score(xgb_fit, x_va)
    xgb_train_prob <- apply_platt(xgb_platt, xgb_train_scores)
    xgb_va_prob <- apply_platt(xgb_platt, xgb_va_scores)
    xgb_th_y <- choose_threshold(y_tr, xgb_train_prob, "youden")
    xgb_th_f <- choose_threshold(y_tr, xgb_train_prob, "f1")

    state$xgb_oof[va] <- xgb_va_prob
    state$xgb_score_oof[va] <- xgb_va_scores
    state$xgb_pred_youden_oof[va] <- ifelse(xgb_va_prob >= xgb_th_y, "christian", "non_christian")
    state$xgb_pred_f1_oof[va] <- ifelse(xgb_va_prob >= xgb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "xgboost", threshold_youden = xgb_th_y, threshold_f1 = xgb_th_f)
    metric_row <- summarize_model_fold(y_va, xgb_va_prob, state$xgb_pred_youden_oof[va], state$xgb_pred_f1_oof[va], "xgboost", i)
    selected_row <- xgb_tuned$best %>%
      mutate(
        outer_fold = i,
        model = "xgboost",
        threshold_youden = xgb_th_y,
        threshold_f1 = xgb_th_f
      )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(best = xgb_tuned$best, tuning = xgb_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  log_progress("outer_fold_done", paste("fold", i, "of", length(outer_folds)))
}

oof_metrics <- bind_rows(
  summarize_model_oof(lab$manual_label, state$nb_oof, state$nb_pred_youden_oof, state$nb_pred_f1_oof, "naive_bayes"),
  summarize_model_oof(lab$manual_label, state$enet_oof, state$enet_pred_youden_oof, state$enet_pred_f1_oof, "enet"),
  summarize_model_oof(lab$manual_label, state$svm_oof, state$svm_pred_youden_oof, state$svm_pred_f1_oof, "svm_platt"),
  summarize_model_oof(lab$manual_label, state$xgb_oof, state$xgb_pred_youden_oof, state$xgb_pred_f1_oof, "xgboost"),
  summarize_model_oof(lab$manual_label, state$lgb_oof, state$lgb_pred_youden_oof, state$lgb_pred_f1_oof, "lightgbm")
)

oof_predictions <- lab %>%
  transmute(
    row_id,
    id,
    manual_label,
    outer_fold = state$outer_fold_id,
    nb_oof = state$nb_oof,
    nb_score_oof = state$nb_score_oof,
    nb_pred_youden_oof = state$nb_pred_youden_oof,
    nb_pred_f1_oof = state$nb_pred_f1_oof,
    prob_enet = state$enet_oof,
    score_enet_logit = state$enet_score_oof,
    label_enet_youden = state$enet_pred_youden_oof,
    label_enet_f1 = state$enet_pred_f1_oof,
    svm_oof = state$svm_oof,
    svm_score_oof = state$svm_score_oof,
    svm_pred_youden_oof = state$svm_pred_youden_oof,
    svm_pred_f1_oof = state$svm_pred_f1_oof,
    xgb_oof = state$xgb_oof,
    xgb_score_oof = state$xgb_score_oof,
    xgb_pred_youden_oof = state$xgb_pred_youden_oof,
    xgb_pred_f1_oof = state$xgb_pred_f1_oof,
    lgb_oof = state$lgb_oof,
    lgb_score_oof = state$lgb_score_oof,
    lgb_pred_youden_oof = state$lgb_pred_youden_oof,
    lgb_pred_f1_oof = state$lgb_pred_f1_oof
  )

write_csv_atomic(arrange(state$outer_threshold_df, outer_fold, model), file.path(output_dir, "outer_thresholds.csv"))
write_csv_atomic(arrange(state$fold_metric_df, outer_fold, model), file.path(output_dir, "outer_fold_metrics.csv"))
write_csv_atomic(arrange(state$selected_outer_df, outer_fold, model), file.path(output_dir, "selected_outer_hyperparameters.csv"))
write_csv_atomic(oof_metrics, file.path(output_dir, "oof_model_metrics.csv"))
write_csv_atomic(oof_predictions, file.path(output_dir, "oof_predictions_labeled.csv"))

full_dir <- file.path(checkpoint_root, "full_fit")
ensure_dir(full_dir)

vec_full_path <- file.path(full_dir, "vectorizer_full.rds")
vec_full <- load_rds_or_null(vec_full_path)
if (is.null(vec_full)) {
  vec_full <- fit_vectorizer(lab$text, latin_stops)
  save_rds_atomic(vec_full, vec_full_path)
}

x_lab_dfm_full <- transform_counts_dfm(lab$text, vec_full, latin_stops)
x_all_dfm_full <- transform_counts_dfm(out$text, vec_full, latin_stops)
x_lab_full <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all_full <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

full_predictions <- tibble(row_id = out$row_id, id = out$id, manual_label = out$manual_label)

key <- "full_naive_bayes"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_nb_full", "full-data fit")
  model_dir <- file.path(full_dir, "naive_bayes")
  nb_cv_scores <- cv_nb_scores_ckpt(
    x_lab_dfm_full, lab$manual_label,
    k = k_folds_inner,
    seed = 8001,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  nb_platt <- fit_platt(nb_cv_scores, lab$manual_label)
  nb_fit <- fit_nb_model(x_lab_dfm_full, lab$manual_label)
  save_model_object(nb_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(nb_platt, file.path(model_dir, "platt.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

nb_fit <- readRDS(file.path(full_dir, "naive_bayes", "model.rds"))
nb_platt <- load_rds_or_null(file.path(full_dir, "naive_bayes", "platt.rds"))
nb_all_scores <- predict_nb_score(nb_fit, x_all_dfm_full)
full_predictions <- full_predictions %>%
  mutate(
    nb_score = nb_all_scores,
    nb_prob = apply_platt(nb_platt, nb_all_scores)
  )

key <- "full_enet"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_enet_full", "full-data fit")
  model_dir <- file.path(full_dir, "enet")
  enet_tuned <- tune_enet_lambda_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8002,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  enet_cv_scores <- cv_enet_scores_ckpt(
    x_lab_full, lab$manual_label,
    alpha = enet_tuned$alpha,
    lambda = enet_tuned$lambda,
    k = k_folds_inner,
    seed = 8102,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  enet_platt <- fit_platt(enet_cv_scores, lab$manual_label)
  enet_fit <- fit_enet_fixed(x_lab_full, lab$manual_label, alpha = enet_tuned$alpha, lambda = enet_tuned$lambda)
  save_model_object(enet_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(enet_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(enet_tuned, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

enet_fit <- readRDS(file.path(full_dir, "enet", "model.rds"))
enet_platt <- load_rds_or_null(file.path(full_dir, "enet", "platt.rds"))
enet_all_scores <- predict_enet_score(enet_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    score_enet_logit = enet_all_scores,
    prob_enet = apply_platt(enet_platt, enet_all_scores)
  )

key <- "full_svm_platt"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_svm_full", "full-data fit")
  model_dir <- file.path(full_dir, "svm_platt")
  svm_tuned <- tune_svm_cost_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8003,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  svm_cv_scores <- cv_svm_scores_ckpt(
    x_lab_full, lab$manual_label,
    cost = svm_tuned$cost,
    k = k_folds_inner,
    seed = 8103,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  svm_platt <- fit_platt(svm_cv_scores, lab$manual_label)
  svm_fit <- fit_svm_fixed(x_lab_full, lab$manual_label, cost = svm_tuned$cost)
  save_model_object(svm_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(svm_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(svm_tuned, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

svm_fit <- readRDS(file.path(full_dir, "svm_platt", "model.rds"))
svm_platt <- load_rds_or_null(file.path(full_dir, "svm_platt", "platt.rds"))
svm_all_scores <- predict_svm_score(svm_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    svm_score = svm_all_scores,
    svm_prob = apply_platt(svm_platt, svm_all_scores)
  )

key <- "full_xgboost"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_xgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "xgboost")
  xgb_tuned <- tune_xgb_model_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8004,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  xgb_cv_scores <- cv_xgb_scores_ckpt(
    x_lab_full, lab$manual_label,
    params = xgb_tuned$params,
    nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
    k = k_folds_inner,
    seed = 8104,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  xgb_platt <- fit_platt(xgb_cv_scores, lab$manual_label)
  xgb_fit <- fit_xgb_fixed(
    x_lab_full, lab$manual_label,
    params = xgb_tuned$params,
    nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
    seed = 8204
  )
  save_model_object(xgb_fit, file.path(model_dir, "model.xgb"), "xgboost")
  save_rds_atomic(xgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = xgb_tuned$best, tuning = xgb_tuned$tuning), file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

xgb_fit <- xgboost::xgb.load(file.path(full_dir, "xgboost", "model.xgb"))
xgb_platt <- load_rds_or_null(file.path(full_dir, "xgboost", "platt.rds"))
xgb_all_scores <- predict_xgb_score(xgb_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    xgb_score = xgb_all_scores,
    xgb_prob = apply_platt(xgb_platt, xgb_all_scores)
  )

key <- "full_lightgbm"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_lgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "lightgbm")
  lgb_tuned <- tune_lgb_model_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8005,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  lgb_cv_scores <- cv_lgb_scores_ckpt(
    x_lab_full, lab$manual_label,
    params = lgb_tuned$params,
    nrounds = as.integer(lgb_tuned$best$best_nrounds[1]),
    k = k_folds_inner,
    seed = 8105,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  lgb_platt <- fit_platt(lgb_cv_scores, lab$manual_label)
  lgb_fit <- fit_lgb_fixed(
    x = x_lab_full,
    y = lab$manual_label,
    params = lgb_tuned$params,
    nrounds = as.integer(lgb_tuned$best$best_nrounds[1])
  )
  save_model_object(lgb_fit, file.path(model_dir, "model.txt"), "lightgbm")
  save_rds_atomic(lgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = lgb_tuned$best, tuning = lgb_tuned$tuning), file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

lgb_fit <- lightgbm::lgb.load(file.path(full_dir, "lightgbm", "model.txt"))
lgb_platt <- load_rds_or_null(file.path(full_dir, "lightgbm", "platt.rds"))
lgb_tuning_result <- readRDS(file.path(full_dir, "lightgbm", "tuning_result.rds"))
lgb_all_scores <- predict_lgb_score(lgb_fit, x_all_full, num_iteration = as.integer(lgb_tuning_result$best$best_nrounds[1]))
full_predictions <- full_predictions %>%
  mutate(
    lgb_score = lgb_all_scores,
    lgb_prob = apply_platt(lgb_platt, lgb_all_scores)
  )

write_csv_atomic(full_predictions, file.path(output_dir, "full_corpus_model_scores.csv"))

log_progress("done", "checkpointed run completed")

```




