---
title: "Latin text classification"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of christian vs. non-christian documents


## Data loading and cleaning

### Loading in and filtering the data (this code chunk can be skipped if corpus.csv already exists in the script directory)

```{r}
library(tidyverse)
library(stringi)

data_path <- "../../data/processed_texts/600ad"
results_path <- "../../results/predictions/non_transformer_lemmatized/600ad"
```


```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels <- readr::read_csv(file.path(data_path,"corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  transmute(id, manual_label = label) %>%
  distinct(id, .keep_all = TRUE)

df_labeled <- df %>%
  inner_join(labels, by = "id") 

```


### Loading in the stopwowrds

```{r}

custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href", 
  "fridericus", "heinricus", "conradus", "meus", "tuus" 
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")

stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- c(stops_v1, stops_v2, custom_additions)

latin_stops <- latin_stops[latin_stops != "*"]

latin_stops <- unique(latin_stops)

latin_stops <- sort(latin_stops)

head(latin_stops, 200)

```



### Defining a tokenization function to count relative frequencies per token for the NB classifier


```{r}
library(quanteda)
library(quanteda.textmodels)
library(tidyverse)
library(knitr)


get_standard_dfm <- function(text_data, stop_list) {

  toks <- tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_remove(pattern = stop_list)
  

  dfm(toks)
}


dfm_train <- get_standard_dfm(df_labeled$text, latin_stops)

dim(dfm_train)

```


## Training classification models

### Setting up model parameters


```{r}

library(tidyverse)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(pROC)
if (!requireNamespace("kernlab", quietly = TRUE)) {
  install.packages("kernlab", repos = "https://cloud.r-project.org")
}
library(kernlab)

set.seed(123)

checkpoint_path <- "../../results/predictions/non_transformer_lemmatized/600ad"

k_folds_outer <- 5L
k_folds_inner <- 5L
xgb_max_rounds <- 4000L
lgb_max_rounds <- 4000L
early_stop_rounds <- 100L
min_docfreq <- 3L

checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
output_dir <- file.path(checkpoint_path, "model_outputs_checkpointed")
progress_path <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

append_progress_row <- function(row, path) {
  row <- row %>% mutate(across(everything(), as.character))
  if (!file.exists(path)) {
    write_csv_atomic(row, path)
  } else {
    old <- readr::read_csv(
      path,
      show_col_types = FALSE,
      col_types = readr::cols(.default = readr::col_character())
    ) %>%
      mutate(across(everything(), as.character))
    write_csv_atomic(bind_rows(old, row), path)
  }
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step, if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

safe_prob <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb))) {
    stop(paste0("Both classes are required", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
  if (any(tb < 2)) {
    stop(paste0("Each class needs at least 2 observations", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for cross-validation.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

make_foldid <- function(y, k, seed = 123) {
  folds <- make_folds(y, k, seed)
  foldid <- integer(length(y))
  for (i in seq_along(folds)) foldid[folds[[i]]] <- i
  foldid
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = s,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1)
  tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1)
  fn <- sum(y01 == 1 & p01 == 0)
  precision <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1 <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(
    accuracy = (tp + tn) / length(y),
    balanced_accuracy = bal_acc,
    precision = precision,
    recall = recall,
    specificity = specificity,
    f1 = f1,
    mcc = mcc,
    tp = tp,
    tn = tn,
    fp = fp,
    fn = fn
  )
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(
          response = y,
          predictor = p,
          levels = c("non_christian", "christian"),
          direction = "<",
          quiet = TRUE
        ),
        x = "best",
        input = "threshold",
        best.method = "youden",
        ret = "threshold",
        transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01 <- as.integer(y == "christian")
    cand <- sort(unique(p))
    cand <- unique(c(0, cand, 0.5, 1))
    f1s <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    tp_youden = hy$tp,
    tn_youden = hy$tn,
    fp_youden = hy$fp,
    fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc,
    tp_f1 = hf$tp,
    tn_f1 = hf$tn,
    fp_f1 = hf$fp,
    fn_f1 = hf$fn
  )
}

summarize_model_fold <- function(y, p, pred_youden, pred_f1, model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold,
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc
  )
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]), by = keys)
  bind_rows(keep, new_rows)
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(), control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores), type = "response"))
  }
}

make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat <- as(x_counts, "dgCMatrix")
  x_vals <- x_mat@x
  x_mat@x <- 1 + log(x_vals)
  idf_vec <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- Matrix::Diagonal(x = 1 / row_norms) %*% x_tfidf
  Matrix::drop0(x_tfidf)
}

nb_prob_christian <- function(pred_obj) {
  if ("christian" %in% colnames(pred_obj)) safe_prob(pred_obj[, "christian"]) else safe_prob(pred_obj[, ncol(pred_obj)])
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}

predict_nb_score <- function(fit, newx_dfm) {
  safe_logit(nb_prob_christian(predict(fit, newdata = newx_dfm, type = "probability")))
}

enet_alpha_grid <- c(0.05, 0.1, 0.2, 0.3, 0.4, 0.5)

fit_enet_fixed <- function(x, y, alpha, lambda) {
  glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    lambda = lambda,
    standardize = TRUE,
    maxit = 1e6
  )
}

enet_has_nonzero <- function(fit) {
  cf <- tryCatch(as.matrix(coef(fit)), error = function(e) NULL)
  if (is.null(cf) || nrow(cf) <= 1) return(FALSE)
  any(abs(cf[-1, 1]) > 0)
}

predict_enet_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_enet_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

enet_lambda_grid <- function(x, y, alpha, nlambda = 60, lambda.min.ratio = 1e-3) {
  base_fit <- glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    standardize = TRUE,
    nlambda = nlambda,
    lambda.min.ratio = lambda.min.ratio,
    maxit = 1e6
  )
  as.numeric(base_fit$lambda)
}

tune_enet_lambda_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir, nlambda = 60, lambda.min.ratio = 1e-3) {
  ensure_dir(ckpt_dir)
  folds <- make_folds(y, inner_k, seed)
  alpha_rows <- vector("list", length(enet_alpha_grid))
  for (a_idx in seq_along(enet_alpha_grid)) {
    alpha <- enet_alpha_grid[a_idx]
    lambda_grid <- enet_lambda_grid(x, y, alpha = alpha, nlambda = nlambda, lambda.min.ratio = lambda.min.ratio)
    tuning_rows <- vector("list", length(lambda_grid))
    for (g in seq_along(lambda_grid)) {
      lam <- lambda_grid[g]
      fold_losses <- rep(NA_real_, length(folds))
      for (j in seq_along(folds)) {
        log_progress("enet_tuning_fold", paste(prefix, "alpha", alpha, "lambda", g, "of", length(lambda_grid), "fold", j, "of", length(folds)))
        va <- folds[[j]]
        tr <- setdiff(seq_along(y), va)
        fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lam)
        if (!enet_has_nonzero(fit)) {
          fold_losses[j] <- Inf
        } else {
          p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
          fold_losses[j] <- log_loss_safe(y[va], p_va)
        }
      }
      cvm <- mean(fold_losses, na.rm = TRUE)
      cvsd <- stats::sd(fold_losses, na.rm = TRUE) / sqrt(sum(is.finite(fold_losses)))
      tuning_rows[[g]] <- tibble(alpha = alpha, lambda = lam, cvm = cvm, cvsd = cvsd)
    }
    alpha_tuning <- bind_rows(tuning_rows) %>% arrange(desc(lambda))
    idx_min <- which.min(alpha_tuning$cvm)
    lambda_min <- alpha_tuning$lambda[idx_min]
    cutoff <- alpha_tuning$cvm[idx_min] + alpha_tuning$cvsd[idx_min]
    lambda_1se <- max(alpha_tuning$lambda[alpha_tuning$cvm <= cutoff], na.rm = TRUE)
    alpha_rows[[a_idx]] <- tibble(alpha = alpha, lambda_1se = as.numeric(lambda_1se), lambda_min = as.numeric(lambda_min), cvm_min = alpha_tuning$cvm[idx_min])
  }
  alpha_summary <- bind_rows(alpha_rows) %>% arrange(cvm_min, alpha)
  best <- alpha_summary[1, , drop = FALSE]
  list(alpha = as.numeric(best$alpha), lambda = as.numeric(best$lambda_1se), lambda_min = as.numeric(best$lambda_min), tuning = alpha_summary)
}

cv_enet_scores_ckpt <- function(x, y, alpha, lambda, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("enet_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    state$scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

cv_svm_scores_ckpt <- function(x, y, cost, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("svm_calibration_fold", paste(prefix, "cost", cost, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    state$scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_svm_cost_ckpt <- function(x, y, cost_grid = c(0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 8, 16), inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  rows_path <- file.path(ckpt_dir, "summary.rds")
  state <- load_rds_or_null(rows_path)
  if (is.null(state) || length(state$done) != length(cost_grid)) {
    state <- list(done = rep(FALSE, length(cost_grid)), rows = vector("list", length(cost_grid)))
  }
  for (i in seq_along(cost_grid)) {
    if (isTRUE(state$done[i])) next
    scores_i <- cv_svm_scores_ckpt(
      x = x,
      y = y,
      cost = cost_grid[i],
      k = inner_k,
      seed = seed + i,
      prefix = prefix,
      ckpt_dir = file.path(ckpt_dir, sprintf("cost_%03d", i))
    )
    state$rows[[i]] <- tibble(cost = cost_grid[i], auc_margin = auc_margin_safe(y, scores_i))
    state$done[i] <- TRUE
    save_rds_atomic(state, rows_path)
  }
  tuning <- bind_rows(state$rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}

tune_xgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- xgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  n_pos <- sum(y == "christian")
  n_neg <- sum(y == "non_christian")
  spw <- n_neg / max(n_pos, 1)

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("xgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "christian")
      n_neg_tr <- sum(y[tr] == "non_christian")
      spw_tr <- n_neg_tr / max(n_pos_tr, 1)
      dtr <- xgboost::xgb.DMatrix(x[tr, , drop = FALSE], label = as.integer(y[tr] == "christian"))
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE], label = as.integer(y[va] == "christian"))
      fit <- xgboost::xgb.train(
        params = make_xgb_params(pars_row, scale_pos_weight = spw_tr),
        data = dtr,
        nrounds = xgb_max_rounds,
        evals = list(valid = dva),
        early_stopping_rounds = early_stop_rounds,
        verbose = 0
      )
      cand_state$oof_prob[va] <- safe_prob(as.numeric(predict(fit, dva)))
      cand_state$best_iters[j] <- best_iter_xgb(fit, xgb_max_rounds)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_xgb_params(best, scale_pos_weight = spw), tuning = tuning)
}

cv_xgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("xgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "christian")
    n_neg_tr <- sum(y[tr] == "non_christian")
    fold_params <- modifyList(params, list(scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr], params = fold_params, nrounds = nrounds, seed = seed + j)
    state$scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_lgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- lgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("lgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      lgb_params_fold <- make_lgb_params(pars_row, seed = seed + 1000L * g + j)
      result <- fit_lgb_early_stop(
        x_tr = x[tr, , drop = FALSE],
        y_tr = y[tr],
        x_va = x[va, , drop = FALSE],
        y_va = y[va],
        params = lgb_params_fold,
        max_rounds = lgb_max_rounds,
        early_rounds = early_stop_rounds
      )
      cand_state$oof_prob[va] <- predict_lgb_prob(result$model, x[va, , drop = FALSE], num_iteration = result$best_iter)
      cand_state$best_iters[j] <- result$best_iter
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_lgb_params(best, seed = seed + 9999L), tuning = tuning)
}

cv_lgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("lgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE],
      y = y[tr],
      params = modifyList(
        params,
        list(
          seed = seed + j,
          feature_fraction_seed = seed + j,
          bagging_seed = seed + j
        )
      ),
      nrounds = nrounds
    )
    state$scores[va] <- predict_lgb_score(fit, x[va, , drop = FALSE], num_iteration = nrounds)
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

run_lgb_preflight <- function(text_vec, y, stop_list, seed = 999L) {
  folds <- make_folds(y, 2L, seed)
  va <- folds[[1]]
  tr <- setdiff(seq_along(y), va)
  vec <- fit_vectorizer(text_vec[tr], stop_list)
  x_tr <- transform_tfidf_sparse(text_vec[tr], vec, stop_list)
  x_va <- transform_tfidf_sparse(text_vec[va], vec, stop_list)
  pars_row <- lgb_grid()[1, , drop = FALSE]
  pars <- make_lgb_params(pars_row, seed = seed)
  result <- fit_lgb_early_stop(
    x_tr = x_tr,
    y_tr = y[tr],
    x_va = x_va,
    y_va = y[va],
    params = pars,
    max_rounds = 100L,
    early_rounds = 20L
  )
  if (!is.finite(result$best_iter)) stop("LightGBM preflight failed to select a valid num_iteration.")
  invisible(result)
}

save_model_object <- function(model, path, engine) {
  ensure_dir(dirname(path))
  if (engine == "xgboost") {
    xgboost::xgb.save(model, path)
  } else if (engine == "lightgbm") {
    lightgbm::lgb.save(model, path)
  } else {
    save_rds_atomic(model, path)
  }
  invisible(path)
}

init_run_state <- function(n_lab) {
  list(
    preflight_done = FALSE,
    outer_fold_id = integer(n_lab),

    nb_oof = rep(NA_real_, n_lab),
    enet_oof = rep(NA_real_, n_lab),
    svm_oof = rep(NA_real_, n_lab),
    xgb_oof = rep(NA_real_, n_lab),
    lgb_oof = rep(NA_real_, n_lab),

    nb_score_oof = rep(NA_real_, n_lab),
    enet_score_oof = rep(NA_real_, n_lab),
    svm_score_oof = rep(NA_real_, n_lab),
    xgb_score_oof = rep(NA_real_, n_lab),
    lgb_score_oof = rep(NA_real_, n_lab),

    nb_pred_youden_oof = rep(NA_character_, n_lab),
    enet_pred_youden_oof = rep(NA_character_, n_lab),
    svm_pred_youden_oof = rep(NA_character_, n_lab),
    xgb_pred_youden_oof = rep(NA_character_, n_lab),
    lgb_pred_youden_oof = rep(NA_character_, n_lab),

    nb_pred_f1_oof = rep(NA_character_, n_lab),
    enet_pred_f1_oof = rep(NA_character_, n_lab),
    svm_pred_f1_oof = rep(NA_character_, n_lab),
    xgb_pred_f1_oof = rep(NA_character_, n_lab),
    lgb_pred_f1_oof = rep(NA_character_, n_lab),

    outer_threshold_df = tibble(),
    fold_metric_df = tibble(),
    selected_outer_df = tibble(),
    model_done = list(),
    full_done = list()
  )
}

mark_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  state[[bucket]][[key]] <- TRUE
  state
}

is_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  isTRUE(state[[bucket]][[key]])
}

prediction_only <- TRUE

pick_new_corpus_file <- function(path) {
  candidates <- list.files(
    path,
    pattern = "cltk_latin_onefile_merged_with_text_and_lemmas.*\\.csv$",
    full.names = TRUE
  )
  if (!length(candidates)) {
    stop("Could not find a CSV named like cltk_latin_onefile_merged_with_text_and_lemmas*.csv in data_path.")
  }
  info <- file.info(candidates)
  candidates[order(info$mtime, decreasing = TRUE)][1]
}

pick_existing_file <- function(paths, label = "file") {
  paths <- unique(paths)
  hit <- paths[file.exists(paths)]
  if (!length(hit)) return(NULL)
  message(label, ": using ", hit[1])
  hit[1]
}

normalize_model_name <- function(x) {
  x <- tolower(as.character(x))
  dplyr::case_when(
    x %in% c("naive_bayes", "nb") ~ "naive_bayes",
    x %in% c("enet", "elastic_net", "elastic-net", "ridge") ~ "enet",
    x %in% c("svm", "svm_platt", "linear_svm") ~ "svm_platt",
    x %in% c("xgb", "xgboost") ~ "xgboost",
    x %in% c("lgb", "lightgbm") ~ "lightgbm",
    TRUE ~ x
  )
}

first_present_col <- function(df, candidates) {
  hit <- intersect(candidates, names(df))
  if (length(hit)) hit[1] else NULL
}

as_numeric_maybe <- function(x) suppressWarnings(as.numeric(x))

standardize_thresholds <- function(df) {
  nm <- names(df)
  model_col <- first_present_col(df, c("model", "model_name", "classifier"))
  youden_col <- first_present_col(df, c("threshold_youden", "youden_threshold", "youden", "thresh_youden"))
  f1_col <- first_present_col(df, c("threshold_f1", "f1_threshold", "f1", "thresh_f1"))
  if (is.null(model_col) || is.null(youden_col) || is.null(f1_col)) return(NULL)
  df %>%
    transmute(
      model = normalize_model_name(.data[[model_col]]),
      threshold_youden = as_numeric_maybe(.data[[youden_col]]),
      threshold_f1 = as_numeric_maybe(.data[[f1_col]])
    ) %>%
    group_by(model) %>%
    summarise(
      threshold_youden = median(threshold_youden, na.rm = TRUE),
      threshold_f1 = median(threshold_f1, na.rm = TRUE),
      .groups = "drop"
    )
}

collapse_outer_hyperparameters <- function(df) {
  model_col <- first_present_col(df, c("model", "model_name", "classifier"))
  if (is.null(model_col)) return(NULL)
  df <- df %>% mutate(model = normalize_model_name(.data[[model_col]]))
  drop_cols <- c(
    "outer_fold", "fold", "model", "model_name", "classifier",
    "threshold_youden", "threshold_f1",
    "auc", "log_loss", "brier", "cvm", "cvsd", "cvm_min"
  )
  keep_cols <- setdiff(names(df), drop_cols)
  out <- vector("list", length(unique(df$model)))
  mods <- sort(unique(df$model))
  for (i in seq_along(mods)) {
    m <- mods[i]
    sub <- df %>% filter(model == m)
    row <- list(model = m)
    for (col in keep_cols) {
      vals <- sub[[col]]
      vals2 <- vals[!is.na(vals)]
      if (!length(vals2)) next
      num <- suppressWarnings(as.numeric(vals2))
      if (all(!is.na(num))) {
        value <- median(num)
        if (col %in% c("best_nrounds", "max_depth", "num_leaves", "bagging_freq", "min_data_in_leaf")) {
          value <- as.integer(round(value))
        }
        row[[col]] <- value
      } else {
        tab <- sort(table(as.character(vals2)), decreasing = TRUE)
        row[[col]] <- names(tab)[1]
      }
    }
    out[[i]] <- tibble::as_tibble(row)
  }
  bind_rows(out)
}

read_threshold_table <- function() {
  candidates <- c(
    file.path(output_dir, "outer_thresholds.csv"),
    file.path(results_path, "model_thresholds_oof.csv"),
    file.path(checkpoint_path, "model_thresholds_oof.csv")
  )
  path <- pick_existing_file(candidates, "threshold file")
  if (is.null(path)) return(NULL)
  standardize_thresholds(readr::read_csv(path, show_col_types = FALSE))
}

read_hyper_table <- function() {
  full_candidates <- c(
    file.path(output_dir, "selected_full_hyperparameters.csv"),
    file.path(results_path, "model_hyperparameters_full.csv"),
    file.path(checkpoint_path, "model_hyperparameters_full.csv")
  )
  outer_candidates <- c(
    file.path(output_dir, "selected_outer_hyperparameters.csv"),
    file.path(results_path, "model_hyperparameters_outer.csv"),
    file.path(checkpoint_path, "model_hyperparameters_outer.csv")
  )
  full_path <- pick_existing_file(full_candidates, "full hyperparameter file")
  if (!is.null(full_path)) {
    df <- readr::read_csv(full_path, show_col_types = FALSE)
    model_col <- first_present_col(df, c("model", "model_name", "classifier"))
    if (!is.null(model_col)) {
      df <- df %>% mutate(model = normalize_model_name(.data[[model_col]]))
      return(df)
    }
  }
  outer_path <- pick_existing_file(outer_candidates, "outer hyperparameter file")
  if (is.null(outer_path)) return(NULL)
  collapse_outer_hyperparameters(readr::read_csv(outer_path, show_col_types = FALSE))
}

pick_num_from_row <- function(row, candidates, default = NA_real_) {
  for (nm in candidates) {
    if (nm %in% names(row)) {
      val <- suppressWarnings(as.numeric(row[[nm]][1]))
      if (length(val) && is.finite(val)) return(val)
    }
  }
  default
}

threshold_value_or <- function(threshold_df, model_name, which = c("youden", "f1"), default = 0.5) {
  which <- match.arg(which)
  if (is.null(threshold_df) || !nrow(threshold_df)) return(default)
  x <- threshold_df %>% filter(model == model_name)
  if (!nrow(x)) return(default)
  val <- if (which == "youden") x$threshold_youden[[1]] else x$threshold_f1[[1]]
  if (is.finite(val)) val else default
}

get_hyper_row <- function(hyper_df, model_name) {
  if (is.null(hyper_df) || !nrow(hyper_df)) return(tibble(model = model_name))
  x <- hyper_df %>% filter(model == model_name)
  if (!nrow(x)) tibble(model = model_name) else x[1, , drop = FALSE]
}


extract_svm_decision_values <- function(pred, pred_lab = NULL) {
  s <- if (is.matrix(pred) || is.data.frame(pred)) {
    as.numeric(pred[, 1])
  } else {
    as.numeric(pred)
  }
  if (!is.null(pred_lab)) {
    pred_lab <- as.character(pred_lab)
    if (length(unique(pred_lab)) > 1) {
      m_chr <- suppressWarnings(mean(s[pred_lab == "christian"], na.rm = TRUE))
      m_non <- suppressWarnings(mean(s[pred_lab == "non_christian"], na.rm = TRUE))
      if (is.finite(m_chr) && is.finite(m_non) && m_chr < m_non) s <- -s
    }
  }
  s
}

svm_to_matrix <- function(x) {
  if (inherits(x, "Matrix")) return(as.matrix(x))
  if (is.matrix(x)) return(x)
  as.matrix(x)
}

fit_svm_linear <- function(x, y, cost = 1) {
  x_in <- svm_to_matrix(x)
  y_fac <- factor(y, levels = c("christian", "non_christian"))
  if (nrow(x_in) != length(y_fac)) {
    stop("SVM fit failed: number of rows in x does not match length(y)")
  }
  fit <- kernlab::ksvm(
    x = x_in,
    y = y_fac,
    type = "C-svc",
    kernel = "vanilladot",
    C = cost,
    scaled = FALSE,
    prob.model = FALSE
  )
  tr_pred_lab <- as.character(predict(fit, x_in, type = "response"))
  tr_scores <- extract_svm_decision_values(
    predict(fit, x_in, type = "decision"),
    pred_lab = tr_pred_lab
  )
  if (!(length(tr_scores) == length(y_fac) && all(is.finite(tr_scores)) && length(unique(signif(tr_scores, 12))) > 1L)) {
    stop("SVM fit failed: kernlab produced constant or invalid training margins")
  }
  structure(list(model = fit), class = "safe_svm_fit")
}

predict_svm_margin <- function(fit, newx) {
  model <- if (inherits(fit, "safe_svm_fit")) fit$model else fit
  newx_in <- svm_to_matrix(newx)
  pred_lab <- as.character(predict(model, newx_in, type = "response"))
  extract_svm_decision_values(
    predict(model, newx_in, type = "decision"),
    pred_lab = pred_lab
  )
}

make_xgb_params_fixed <- function(row, scale_pos_weight = 1, seed = 123) {
  list(
    objective = "binary:logistic",
    eval_metric = "logloss",
    eta = pick_num_from_row(row, c("eta", "learning_rate"), 0.05),
    max_depth = as.integer(round(pick_num_from_row(row, c("max_depth"), 6))),
    min_child_weight = pick_num_from_row(row, c("min_child_weight"), 1),
    subsample = pick_num_from_row(row, c("subsample"), 0.8),
    colsample_bytree = pick_num_from_row(row, c("colsample_bytree", "feature_fraction"), 0.8),
    gamma = pick_num_from_row(row, c("gamma", "min_split_loss"), 0),
    lambda = pick_num_from_row(row, c("lambda", "reg_lambda"), 1),
    alpha = pick_num_from_row(row, c("alpha", "reg_alpha"), 0),
    scale_pos_weight = scale_pos_weight,
    seed = seed,
    nthread = 1
  )
}

fit_xgb_fixed_fast <- function(x, y, row, nrounds, seed = 123) {
  spw <- sum(y == "non_christian") / max(sum(y == "christian"), 1)
  dtrain <- xgboost::xgb.DMatrix(x, label = as.integer(y == "christian"))
  xgboost::xgb.train(
    params = make_xgb_params_fixed(row, scale_pos_weight = spw, seed = seed),
    data = dtrain,
    nrounds = as.integer(nrounds),
    verbose = 0
  )
}

predict_xgb_score_fast <- function(fit, newx) {
  safe_logit(as.numeric(predict(fit, xgboost::xgb.DMatrix(newx))))
}

make_lgb_params_fixed <- function(row, seed = 123) {
  list(
    objective = "binary",
    metric = "binary_logloss",
    learning_rate = pick_num_from_row(row, c("learning_rate", "eta"), 0.05),
    num_leaves = as.integer(round(pick_num_from_row(row, c("num_leaves"), 31))),
    min_data_in_leaf = as.integer(round(pick_num_from_row(row, c("min_data_in_leaf"), 20))),
    feature_fraction = pick_num_from_row(row, c("feature_fraction", "colsample_bytree"), 0.8),
    bagging_fraction = pick_num_from_row(row, c("bagging_fraction", "subsample"), 0.8),
    bagging_freq = as.integer(round(pick_num_from_row(row, c("bagging_freq"), 1))),
    min_gain_to_split = pick_num_from_row(row, c("min_gain_to_split", "gamma"), 0),
    lambda_l1 = pick_num_from_row(row, c("lambda_l1", "alpha", "reg_alpha"), 0),
    lambda_l2 = pick_num_from_row(row, c("lambda_l2", "lambda", "reg_lambda"), 0),
    verbosity = -1,
    num_threads = 1,
    seed = seed,
    feature_fraction_seed = seed,
    bagging_seed = seed,
    deterministic = TRUE,
    force_col_wise = TRUE
  )
}

fit_lgb_fixed_fast <- function(x, y, row, nrounds, seed = 123) {
  dtrain <- lightgbm::lgb.Dataset(x, label = as.integer(y == "christian"))
  lightgbm::lgb.train(
    params = make_lgb_params_fixed(row, seed = seed),
    data = dtrain,
    nrounds = as.integer(nrounds),
    verbose = -1
  )
}

predict_lgb_prob_fast <- function(fit, newx, num_iteration = NULL) {
  safe_prob(as.numeric(predict(fit, newx, num_iteration = num_iteration)))
}

predict_lgb_score_fast <- function(fit, newx, num_iteration = NULL) {
  safe_logit(predict_lgb_prob_fast(fit, newx, num_iteration = num_iteration))
}

cv_nb_scores_fixed <- function(x_dfm, y, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_nb_model(x_dfm[tr, ], y[tr])
    scores[va] <- predict_nb_score(fit, x_dfm[va, ])
  }
  scores
}

cv_enet_scores_fixed <- function(x, y, alpha, lambda, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
  }
  scores
}

cv_svm_scores_fixed <- function(x, y, cost, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_linear(x[tr, , drop = FALSE], y[tr], cost = cost)
    scores[va] <- predict_svm_margin(fit, x[va, , drop = FALSE])
  }
  scores
}

tune_svm_cost_inline <- function(x, y,
                                 cost_grid = c(0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 8, 16),
                                 inner_k = 5, seed = 8003) {
  rows <- vector("list", length(cost_grid))
  for (i in seq_along(cost_grid)) {
    s <- cv_svm_scores_fixed(x, y, cost = cost_grid[i], k = inner_k, seed = seed + i)
    rows[[i]] <- tibble(cost = cost_grid[i], auc_margin = auc_margin_safe(y, s))
  }
  tuning <- bind_rows(rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}

cv_xgb_scores_fixed <- function(x, y, row, nrounds, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_xgb_fixed_fast(x[tr, , drop = FALSE], y[tr], row = row, nrounds = nrounds, seed = seed + j)
    scores[va] <- predict_xgb_score_fast(fit, x[va, , drop = FALSE])
  }
  scores
}

cv_lgb_scores_fixed <- function(x, y, row, nrounds, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed_fast(x[tr, , drop = FALSE], y[tr], row = row, nrounds = nrounds, seed = seed + j)
    scores[va] <- predict_lgb_score_fast(fit, x[va, , drop = FALSE], num_iteration = nrounds)
  }
  scores
}

log_progress("prediction_only", "refitting once on labeled corpus with saved hyperparameters; no nested tuning")

new_csv_path <- pick_new_corpus_file(data_path)
log_progress("new_corpus", new_csv_path)

thresholds <- read_threshold_table()
hyper_df <- read_hyper_table()

new_raw <- readr::read_csv(new_csv_path, show_col_types = FALSE)

if (!"text_lemma" %in% names(new_raw)) {
  stop("The new CSV must contain a text_lemma column.")
}

if (!"id" %in% names(new_raw)) {
  new_raw$id <- NA_character_
}

if (!"manual_label" %in% names(new_raw)) {
  new_raw$manual_label <- NA_character_
}

new_raw <- new_raw %>%
  mutate(
    source_row_id = row_number(),
    id = as.character(id),
    manual_label = as.character(manual_label),
    text_lemma = if_else(is.na(text_lemma), NA_character_, as.character(text_lemma)),
    text_for_model = stringi::stri_trim_both(coalesce(text_lemma, "")),
    classify_status = if_else(text_for_model == "", "skipped_missing_text_lemma", "scored")
  )

new_skip <- new_raw %>%
  filter(classify_status != "scored") %>%
  select(-text_for_model)

new_ok <- new_raw %>%
  filter(classify_status == "scored")

score_output_path <- file.path(output_dir, "new_corpus_model_scores.csv")
skip_output_path <- file.path(output_dir, "new_corpus_rows_skipped_missing_text_lemma.csv")
params_output_path <- file.path(output_dir, "prediction_only_hyperparameters_used.csv")

if (nrow(new_ok) == 0) {
  results_no_text <- new_raw %>%
    select(-any_of(c("text", "text_lemma", "text_for_model")))

  new_skip_no_text <- new_skip %>%
    select(-any_of(c("text", "text_lemma", "text_for_model")))

  write_csv_atomic(results_no_text, score_output_path)
  write_csv_atomic(new_skip_no_text, skip_output_path)
  log_progress("done", "no scorable rows found in the new CSV")
} else {
  train_df <- df_labeled %>%
    mutate(
      text = as.character(text),
      text = if_else(is.na(text), "", text),
      text = stringi::stri_trim_both(text)
    ) %>%
    filter(text != "")

  log_progress("vectorizer", "fitting full-corpus vectorizer")
  vec_full <- fit_vectorizer(train_df$text, latin_stops)
  x_lab_dfm_full <- transform_counts_dfm(train_df$text, vec_full, latin_stops)
  x_lab_full <- transform_tfidf_sparse(train_df$text, vec_full, latin_stops)

  keep_train <- Matrix::rowSums(abs(x_lab_full)) > 0
  if (any(!keep_train)) {
    log_progress("train_filter", paste("dropping", sum(!keep_train), "blank/zero-feature training rows"))
    train_df <- train_df[keep_train, , drop = FALSE]
    x_lab_dfm_full <- x_lab_dfm_full[keep_train, ]
    x_lab_full <- x_lab_full[keep_train, , drop = FALSE]
  }

  y_train <- train_df$manual_label
  calibration_k <- min(5L, min(table(y_train)))
  if (calibration_k < 2L) stop("Not enough labeled data for calibration after filtering.")

  x_new_dfm <- transform_counts_dfm(new_ok$text_for_model, vec_full, latin_stops)
  x_new <- transform_tfidf_sparse(new_ok$text_for_model, vec_full, latin_stops)

  keep_new <- Matrix::rowSums(abs(x_new)) > 0
  if (any(!keep_new)) {
    zero_skip <- new_ok[!keep_new, , drop = FALSE] %>%
      mutate(classify_status = "skipped_zero_feature_text") %>%
      select(-text_for_model)
    new_skip <- bind_rows(new_skip, zero_skip)
    new_ok <- new_ok[keep_new, , drop = FALSE]
    x_new_dfm <- x_new_dfm[keep_new, ]
    x_new <- x_new[keep_new, , drop = FALSE]
    new_raw <- new_raw %>%
      mutate(classify_status = if_else(source_row_id %in% zero_skip$source_row_id, "skipped_zero_feature_text", classify_status))
    log_progress("new_filter", paste("dropping", nrow(zero_skip), "zero-feature new rows"))
  }

  params_used <- list()

  if (nrow(new_ok) == 0) {
    results_no_text <- new_raw %>%
      select(-any_of(c("text", "text_lemma", "text_for_model")))

    new_skip_no_text <- new_skip %>%
      select(-any_of(c("text", "text_lemma", "text_for_model")))

    write_csv_atomic(results_no_text, score_output_path)
    write_csv_atomic(new_skip_no_text, skip_output_path)
    log_progress("done", "all candidate rows became zero-feature after vectorization")
  } else {
    log_progress("fit_nb_full", "naive bayes")
    nb_cv_scores <- cv_nb_scores_fixed(x_lab_dfm_full, y_train, k = calibration_k, seed = 8001)
    nb_platt <- fit_platt(nb_cv_scores, y_train)
    nb_train_prob <- apply_platt(nb_platt, nb_cv_scores)
    nb_fit <- fit_nb_model(x_lab_dfm_full, y_train)
    nb_scores <- predict_nb_score(nb_fit, x_new_dfm)
    nb_probs <- apply_platt(nb_platt, nb_scores)
    nb_th_y <- threshold_value_or(thresholds, "naive_bayes", "youden", choose_threshold(y_train, nb_train_prob, "youden"))
    nb_th_f <- threshold_value_or(thresholds, "naive_bayes", "f1", choose_threshold(y_train, nb_train_prob, "f1"))
    params_used[[length(params_used) + 1]] <- tibble(model = "naive_bayes")

    enet_row <- get_hyper_row(hyper_df, "enet")
    enet_alpha <- pick_num_from_row(enet_row, c("alpha"), 0.1)
    enet_lambda <- pick_num_from_row(enet_row, c("lambda_1se", "lambda", "lambda_min"), 0.01)
    log_progress("fit_enet_full", paste("alpha", signif(enet_alpha, 4), "lambda", signif(enet_lambda, 4)))
    enet_cv_scores <- cv_enet_scores_fixed(x_lab_full, y_train, alpha = enet_alpha, lambda = enet_lambda, k = calibration_k, seed = 8002)
    enet_platt <- fit_platt(enet_cv_scores, y_train)
    enet_train_prob <- apply_platt(enet_platt, enet_cv_scores)
    enet_fit <- fit_enet_fixed(x_lab_full, y_train, alpha = enet_alpha, lambda = enet_lambda)
    enet_scores <- predict_enet_score(enet_fit, x_new)
    enet_probs <- apply_platt(enet_platt, enet_scores)
    enet_th_y <- threshold_value_or(thresholds, "enet", "youden", choose_threshold(y_train, enet_train_prob, "youden"))
    enet_th_f <- threshold_value_or(thresholds, "enet", "f1", choose_threshold(y_train, enet_train_prob, "f1"))
    params_used[[length(params_used) + 1]] <- tibble(model = "enet", alpha = enet_alpha, lambda = enet_lambda)

    svm_row <- get_hyper_row(hyper_df, "svm_platt")
    svm_cost <- pick_num_from_row(svm_row, c("cost"), 1)
    log_progress("fit_svm_full", paste("using saved cost", signif(svm_cost, 4)))
    svm_scores <- rep(NA_real_, nrow(new_ok))
    svm_probs <- rep(NA_real_, nrow(new_ok))
    svm_th_y <- NA_real_
    svm_th_f <- NA_real_
    svm_status <- "ok"
    svm_detail <- ""
    tryCatch({
      svm_cv_scores <- tryCatch(
        cv_svm_scores_fixed(x_lab_full, y_train, cost = svm_cost, k = calibration_k, seed = 8003),
        error = function(e) {
          log_progress("fit_svm_full", paste("cv failed; falling back to in-sample margins:", conditionMessage(e)))
          rep(NA_real_, length(y_train))
        }
      )
      if (!all(is.finite(svm_cv_scores)) || length(unique(signif(stats::na.omit(svm_cv_scores), 12))) < 2L) {
        svm_fit_cal <- fit_svm_linear(x_lab_full, y_train, cost = svm_cost)
        svm_cv_scores <- predict_svm_margin(svm_fit_cal, x_lab_full)
      }
      svm_platt <- fit_platt(svm_cv_scores, y_train)
      svm_train_prob <- apply_platt(svm_platt, svm_cv_scores)
      svm_fit <- fit_svm_linear(x_lab_full, y_train, cost = svm_cost)
      svm_scores <- predict_svm_margin(svm_fit, x_new)
      svm_probs <- apply_platt(svm_platt, svm_scores)
      svm_th_y <- threshold_value_or(thresholds, "svm_platt", "youden", choose_threshold(y_train, svm_train_prob, "youden"))
      svm_th_f <- threshold_value_or(thresholds, "svm_platt", "f1", choose_threshold(y_train, svm_train_prob, "f1"))
    }, error = function(e) {
      svm_status <<- "failed"
      svm_detail <<- conditionMessage(e)
      log_progress("fit_svm_full", paste("SVM failed; writing NA SVM outputs:", svm_detail))
    })
    params_used[[length(params_used) + 1]] <- tibble(model = "svm_platt", cost = svm_cost, status = svm_status, detail = svm_detail)

    xgb_row <- get_hyper_row(hyper_df, "xgboost")
    xgb_nrounds <- as.integer(round(pick_num_from_row(xgb_row, c("best_nrounds", "nrounds"), 200)))
    log_progress("fit_xgb_full", paste("nrounds", xgb_nrounds))
    xgb_cv_scores <- cv_xgb_scores_fixed(x_lab_full, y_train, row = xgb_row, nrounds = xgb_nrounds, k = calibration_k, seed = 8004)
    xgb_platt <- fit_platt(xgb_cv_scores, y_train)
    xgb_train_prob <- apply_platt(xgb_platt, xgb_cv_scores)
    xgb_fit <- fit_xgb_fixed_fast(x_lab_full, y_train, row = xgb_row, nrounds = xgb_nrounds, seed = 8104)
    xgb_scores <- predict_xgb_score_fast(xgb_fit, x_new)
    xgb_probs <- apply_platt(xgb_platt, xgb_scores)
    xgb_th_y <- threshold_value_or(thresholds, "xgboost", "youden", choose_threshold(y_train, xgb_train_prob, "youden"))
    xgb_th_f <- threshold_value_or(thresholds, "xgboost", "f1", choose_threshold(y_train, xgb_train_prob, "f1"))
    params_used[[length(params_used) + 1]] <- tibble(
      model = "xgboost",
      eta = pick_num_from_row(xgb_row, c("eta", "learning_rate"), 0.05),
      max_depth = as.integer(round(pick_num_from_row(xgb_row, c("max_depth"), 6))),
      min_child_weight = pick_num_from_row(xgb_row, c("min_child_weight"), 1),
      subsample = pick_num_from_row(xgb_row, c("subsample"), 0.8),
      colsample_bytree = pick_num_from_row(xgb_row, c("colsample_bytree", "feature_fraction"), 0.8),
      gamma = pick_num_from_row(xgb_row, c("gamma", "min_split_loss"), 0),
      lambda = pick_num_from_row(xgb_row, c("lambda", "reg_lambda"), 1),
      alpha = pick_num_from_row(xgb_row, c("alpha", "reg_alpha"), 0),
      best_nrounds = xgb_nrounds
    )

    lgb_row <- get_hyper_row(hyper_df, "lightgbm")
    lgb_nrounds <- as.integer(round(pick_num_from_row(lgb_row, c("best_nrounds", "nrounds"), 200)))
    log_progress("fit_lgb_full", paste("nrounds", lgb_nrounds))
    lgb_cv_scores <- cv_lgb_scores_fixed(x_lab_full, y_train, row = lgb_row, nrounds = lgb_nrounds, k = calibration_k, seed = 8005)
    lgb_platt <- fit_platt(lgb_cv_scores, y_train)
    lgb_train_prob <- apply_platt(lgb_platt, lgb_cv_scores)
    lgb_fit <- fit_lgb_fixed_fast(x_lab_full, y_train, row = lgb_row, nrounds = lgb_nrounds, seed = 8105)
    lgb_scores <- predict_lgb_score_fast(lgb_fit, x_new, num_iteration = lgb_nrounds)
    lgb_probs <- apply_platt(lgb_platt, lgb_scores)
    lgb_th_y <- threshold_value_or(thresholds, "lightgbm", "youden", choose_threshold(y_train, lgb_train_prob, "youden"))
    lgb_th_f <- threshold_value_or(thresholds, "lightgbm", "f1", choose_threshold(y_train, lgb_train_prob, "f1"))
    params_used[[length(params_used) + 1]] <- tibble(
      model = "lightgbm",
      learning_rate = pick_num_from_row(lgb_row, c("learning_rate", "eta"), 0.05),
      num_leaves = as.integer(round(pick_num_from_row(lgb_row, c("num_leaves"), 31))),
      min_data_in_leaf = as.integer(round(pick_num_from_row(lgb_row, c("min_data_in_leaf"), 20))),
      feature_fraction = pick_num_from_row(lgb_row, c("feature_fraction", "colsample_bytree"), 0.8),
      bagging_fraction = pick_num_from_row(lgb_row, c("bagging_fraction", "subsample"), 0.8),
      bagging_freq = as.integer(round(pick_num_from_row(lgb_row, c("bagging_freq"), 1))),
      min_gain_to_split = pick_num_from_row(lgb_row, c("min_gain_to_split", "gamma"), 0),
      lambda_l1 = pick_num_from_row(lgb_row, c("lambda_l1", "alpha", "reg_alpha"), 0),
      lambda_l2 = pick_num_from_row(lgb_row, c("lambda_l2", "lambda", "reg_lambda"), 0),
      best_nrounds = lgb_nrounds
    )

    params_used_df <- bind_rows(params_used)

scored_preds <- tibble(
    source_row_id = new_ok$source_row_id,
    nb_score = nb_scores,
    nb_prob = nb_probs,
    label_nb_youden = if_else(nb_probs >= nb_th_y, "christian", "non_christian"),
    label_nb_f1 = if_else(nb_probs >= nb_th_f, "christian", "non_christian"),
    score_enet_logit = enet_scores,
    prob_enet = enet_probs,
    label_enet_youden = if_else(enet_probs >= enet_th_y, "christian", "non_christian"),
    label_enet_f1 = if_else(enet_probs >= enet_th_f, "christian", "non_christian"),
    svm_score = svm_scores,
    svm_prob = svm_probs,
    label_svm_youden = if_else(svm_probs >= svm_th_y, "christian", "non_christian"),
    label_svm_f1 = if_else(svm_probs >= svm_th_f, "christian", "non_christian"),
    xgb_score = xgb_scores,
    xgb_prob = xgb_probs,
    label_xgb_youden = if_else(xgb_probs >= xgb_th_y, "christian", "non_christian"),
    label_xgb_f1 = if_else(xgb_probs >= xgb_th_f, "christian", "non_christian"),
    lgb_score = lgb_scores,
    lgb_prob = lgb_probs,
    label_lgb_youden = if_else(lgb_probs >= lgb_th_y, "christian", "non_christian"),
    label_lgb_f1 = if_else(lgb_probs >= lgb_th_f, "christian", "non_christian")
  )

  results <- new_raw %>%
    select(-text_for_model) %>%
    left_join(scored_preds, by = "source_row_id")

  results_no_text <- results %>%
    select(-any_of(c("text", "text_lemma", "text_for_model")))

  new_skip_no_text <- new_skip %>%
    select(-any_of(c("text", "text_lemma", "text_for_model")))

  write_csv_atomic(results_no_text, score_output_path)
  write_csv_atomic(new_skip_no_text, skip_output_path)
  write_csv_atomic(params_used_df, params_output_path)

  log_progress("saved_scores", score_output_path)
  log_progress("saved_skipped_rows", skip_output_path)
  log_progress("saved_params", params_output_path)
  log_progress(
    "done",
    paste(
      "rows total:", nrow(results),
      "| rows scored:", sum(results$classify_status == "scored"),
      "| rows skipped:", sum(results$classify_status != "scored")
    )
  )
  }
}


```
