---
title: "Latin text classification"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of christian vs. non-christian documents


## Data loading and cleaning

### Loading in and filtering the data (this code chunk can be skipped if corpus.csv already exists in the script directory)

```{r}
library(tidyverse)
library(stringi)

data_path <- "../../data/processed_texts/600ad"
results_path <- "../../results/predictions/non_transformer_lemmatized/600ad/temporal_sweep"
```


```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels <- readr::read_csv(file.path(data_path,"corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  transmute(id, manual_label = label) %>%
  distinct(id, .keep_all = TRUE)

df_labeled <- df %>%
  inner_join(labels, by = "id") 

```


### Loading in the stopwowrds

```{r}

custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href", 
  "fridericus", "heinricus", "conradus", "meus", "tuus" 
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")

stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- c(stops_v1, stops_v2, custom_additions)

latin_stops <- latin_stops[latin_stops != "*"]

latin_stops <- unique(latin_stops)

latin_stops <- sort(latin_stops)

head(latin_stops, 200)

```


### Defining a tokenization function to count relative frequencies per token for the NB classifier


```{r}
library(quanteda)
library(quanteda.textmodels)
library(tidyverse)
library(knitr)


get_standard_dfm <- function(text_data, stop_list) {

  toks <- tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_remove(pattern = stop_list)
  

  dfm(toks)
}


dfm_train <- get_standard_dfm(df_labeled$text, latin_stops)

dim(dfm_train)

```


### Getting rid of highly-date-correlated features

```{r}
library(Matrix)
library(ggplot2)

date_df <- df %>%
  filter(!is.na(date_range_start), !is.na(date_range_end), text != "") %>%
  mutate(
    date_range_start = as.numeric(date_range_start),
    date_range_end = as.numeric(date_range_end)
  )

dfm_date <- get_standard_dfm(date_df$text, latin_stops)
keep_docs <- ntoken(dfm_date) > 0
date_df <- date_df[keep_docs, ]
dfm_date <- dfm_date[keep_docs, ] %>%
  dfm_trim(min_docfreq = max(5, ceiling(nrow(date_df) * 0.02))) %>%
  dfm_weight(scheme = "prop")

X <- as(dfm_date, "dgCMatrix")

safe_spearman <- function(x, y) {
  if (sd(x) == 0 || sd(y) == 0) return(NA_real_)
  suppressWarnings(cor(x, y, method = "spearman"))
}

rho_start <- vapply(seq_len(ncol(X)), function(j) safe_spearman(as.numeric(X[, j]), date_df$date_range_start), numeric(1))
rho_end   <- vapply(seq_len(ncol(X)), function(j) safe_spearman(as.numeric(X[, j]), date_df$date_range_end), numeric(1))

cor_summary <- tibble(
  term = featnames(dfm_date),
  docfreq = docfreq(dfm_date),
  rho_start = rho_start,
  rho_end = rho_end,
  rho_mean = (rho_start + rho_end) / 2,
  mean_abs_rho = (abs(rho_start) + abs(rho_end)) / 2
) %>%
  drop_na(rho_start, rho_end) %>%
  arrange(desc(mean_abs_rho))

top_date_terms <- bind_rows(
  cor_summary %>% slice_max(rho_mean, n = 50),
  cor_summary %>% slice_min(rho_mean, n = 50)
) %>%
  distinct(term, .keep_all = TRUE) %>%
  arrange(rho_mean) %>%
  mutate(term = factor(term, levels = term))

plot_df <- top_date_terms %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(-term, names_to = "bound", values_to = "rho") %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))
```


```{r}
library(forcats)

out_dir <- "../../results/alternative_analyses/word_correlation"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

top_date_terms <- cor_summary %>%
  slice_head(n = 100) %>%
  arrange(rho_mean) %>%
  mutate(term = fct_inorder(term))

plot_df <- top_date_terms %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(-term, names_to = "bound", values_to = "rho") %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))

p <- ggplot(top_date_terms, aes(y = term)) +
  geom_segment(
    aes(x = rho_start, xend = rho_end, yend = term, color = rho_mean > 0),
    linewidth = 1.2, alpha = 0.7, show.legend = FALSE
  ) +
  geom_point(
    data = plot_df,
    aes(x = rho, shape = bound, fill = rho > 0),
    size = 2.6, color = "white", stroke = 0.4
  ) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5, color = "grey60") +
  scale_shape_manual(values = c("Start date" = 21, "End date" = 24), name = NULL) +
  scale_fill_manual(values = c(`TRUE` = "#2b6cb0", `FALSE` = "#c53030"), guide = "none") +
  scale_color_manual(values = c(`TRUE` = "#2b6cb0", `FALSE` = "#c53030")) +
  labs(x = "Spearman correlation with document date", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top",
    axis.text.y = element_text(size = 6.5),
    plot.margin = margin(10, 20, 10, 10)
  )

plot_height <- max(16, 0.22 * nrow(top_date_terms))

ggsave(
  filename = file.path(out_dir, "top100_date_correlated_terms.png"),
  plot = p,
  width = 10,
  height = plot_height,
  dpi = 400,
  bg = "white"
)

ggsave(
  filename = file.path(out_dir, "top100_date_correlated_terms.pdf"),
  plot = p,
  width = 10,
  height = plot_height,
  bg = "white"
)

p
```

```{r}
out_dir <- "../../results/alternative_analyses/word_correlation"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

dist_df <- cor_summary %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(
    cols = c(rho_start, rho_end),
    names_to = "bound",
    values_to = "rho"
  ) %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))

p_dist <- ggplot(dist_df, aes(x = rho, fill = bound, color = bound)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 60, alpha = 0.30, position = "identity") +
  geom_density(linewidth = 0.9) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5, color = "grey50") +
  scale_fill_manual(values = c("Start date" = "#2b6cb0", "End date" = "#c53030")) +
  scale_color_manual(values = c("Start date" = "#2b6cb0", "End date" = "#c53030")) +
  labs(x = "Spearman correlation", y = "Density", fill = NULL, color = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

ggsave(
  file.path(out_dir, "correlation_distribution_all_terms.png"),
  p_dist,
  width = 9,
  height = 5.5,
  dpi = 400,
  bg = "white"
)

ggsave(
  file.path(out_dir, "correlation_distribution_all_terms.pdf"),
  p_dist,
  width = 9,
  height = 5.5,
  bg = "white"
)

p_dist

```

## Training classification models with temporal-confounding correction

This pipeline evaluates how much of each classifier's signal is driven by lexical drift over time
rather than by genuinely Christian / non-Christian vocabulary. In every outer training fold we:

1. Use **only the non-Christian training texts** to rank the vocabulary by absolute Spearman
   correlation between token relative frequency and document midpoint date.
2. For each cutoff $n \in \{0, 25, 50, 100, 200\}$, remove the top-$n$ most date-correlated words
   from the fold's vocabulary. $n = 0$ is the unrestricted baseline.
3. Fit the classifier on the full training fold (both classes) using that reduced vocabulary, and
   score the held-out test fold with the same vocabulary.

The removed-word list is recomputed inside every fold so nothing leaks from the held-out texts.
If a classifier still performs well at $n = 200$, that's evidence the signal isn't purely a proxy
for when the text was written.

### Setting up model parameters

```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(pROC)

set.seed(123)

checkpoint_path <- "../../results/predictions/non_transformer_lemmatized/600ad/temporal_sweep"

k_folds_outer     <- 5L
k_folds_inner     <- 5L
min_docfreq       <- 3L

# n-values for the top-n date-word removal sweep. n = 0 is the no-removal baseline.
n_remove_grid <- c(0L, 25L, 50L, 100L, 200L)

checkpoint_root <- file.path(checkpoint_path, "checkpoints_temporal_sweep")
output_dir      <- file.path(checkpoint_path, "model_outputs_temporal_sweep")
progress_path   <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir,      recursive = TRUE, showWarnings = FALSE)

`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) { file.copy(tmp, path, overwrite = TRUE); unlink(tmp) }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) { file.copy(tmp, path, overwrite = TRUE); unlink(tmp) }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

```

```{r}

append_progress_row <- function(row, path) {
  if (!file.exists(path)) {
    write_csv_atomic(row, path)
  } else {
    old <- readr::read_csv(path, show_col_types = FALSE) %>%
      mutate(time = as.character(time))
    row <- row %>%
      mutate(time = as.character(time))
    write_csv_atomic(bind_rows(old, row), path)
  }
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time   = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step   = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step,
                 if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

safe_prob  <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb)))
    stop(paste0("Both classes are required",
                if (nzchar(where)) paste0(" in ", where) else "", "."))
  if (any(tb < 2))
    stop(paste0("Each class needs at least 2 observations",
                if (nzchar(where)) paste0(" in ", where) else "", "."))
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for cross-validation.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = s,
      levels = c("non_christian", "christian"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1); tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1); fn <- sum(y01 == 1 & p01 == 0)
  precision   <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall      <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1  <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(
    accuracy = (tp + tn) / length(y), balanced_accuracy = bal_acc,
    precision = precision, recall = recall, specificity = specificity,
    f1 = f1, mcc = mcc, tp = tp, tn = tn, fp = fp, fn = fn
  )
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(response = y, predictor = p,
                  levels = c("non_christian", "christian"),
                  direction = "<", quiet = TRUE),
        x = "best", input = "threshold", best.method = "youden",
        ret = "threshold", transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01  <- as.integer(y == "christian")
    cand <- sort(unique(c(0, p, 0.5, 1)))
    f1s  <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(), control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores), type = "response"))
  }
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]), by = keys)
  bind_rows(keep, new_rows)
}
```

### Vectorizer and date-word ranking helpers

```{r}
make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

# Fit a vectorizer (feature list + IDF) on a given text sample, then optionally
# intersect with an allowed-feature whitelist. This is how we apply the
# fold-specific "reduced vocabulary" after removing date-loaded words.
fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq,
                           allowed_features = NULL) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (!is.null(allowed_features)) {
    keep <- intersect(quanteda::featnames(x), allowed_features)
    if (length(keep) == 0) stop("No features remain after applying allowed_features whitelist.")
    x <- quanteda::dfm_select(x, pattern = keep, selection = "keep", valuetype = "fixed")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N   <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat    <- as(x_counts, "dgCMatrix")
  if (length(x_mat@x) > 0) x_mat@x <- 1 + log(x_mat@x)
  idf_vec  <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf  <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}

safe_spearman <- function(x, y) {
  if (length(x) < 3) return(NA_real_)
  if (stats::sd(x, na.rm = TRUE) == 0 || stats::sd(y, na.rm = TRUE) == 0) return(NA_real_)
  suppressWarnings(stats::cor(x, y, method = "spearman", use = "pairwise.complete.obs"))
}

# Core: rank words by absolute Spearman correlation with date, computed on a
# specified subset of documents. Returns a tibble sorted most-to-least
# date-associated. We use the midpoint of the document date range so each doc
# contributes a single date value.
rank_date_words <- function(text_vec, date_mid, stop_list,
                            min_docfreq_rank = 5L) {
  dfm_raw <- make_counts_dfm(text_vec, stop_list)
  keep_docs <- quanteda::ntoken(dfm_raw) > 0 & is.finite(date_mid)
  dfm_raw   <- dfm_raw[keep_docs, ]
  date_mid  <- date_mid[keep_docs]

  if (quanteda::ndoc(dfm_raw) < 5) {
    warning("Too few documents with valid dates to rank date-correlated words; returning empty ranking.")
    return(tibble(term = character(0), rho = numeric(0), abs_rho = numeric(0),
                  docfreq = integer(0)))
  }

  min_df_use <- max(min_docfreq_rank,
                    ceiling(quanteda::ndoc(dfm_raw) * 0.02))
  dfm_trim <- quanteda::dfm_trim(dfm_raw,
                                 min_docfreq = min_df_use,
                                 docfreq_type = "count")
  if (quanteda::nfeat(dfm_trim) == 0) {
    return(tibble(term = character(0), rho = numeric(0), abs_rho = numeric(0),
                  docfreq = integer(0)))
  }

  dfm_prop <- quanteda::dfm_weight(dfm_trim, scheme = "prop")
  X <- as(dfm_prop, "dgCMatrix")

  rho <- vapply(
    seq_len(ncol(X)),
    function(j) safe_spearman(as.numeric(X[, j]), date_mid),
    numeric(1)
  )

  tibble(
    term    = quanteda::featnames(dfm_trim),
    rho     = rho,
    abs_rho = abs(rho),
    docfreq = as.integer(quanteda::docfreq(dfm_trim, scheme = "count"))
  ) %>%
    dplyr::filter(!is.na(rho)) %>%
    dplyr::arrange(dplyr::desc(abs_rho))
}

# Given the date-word ranking computed on non-Christian training texts, return
# the top n words to remove. n = 0 returns an empty character vector (baseline).
top_date_words <- function(ranking_df, n) {
  if (n <= 0 || nrow(ranking_df) == 0) return(character(0))
  head(ranking_df$term, as.integer(n))
}
```

### Model fit / predict helpers

```{r}
nb_prob_christian <- function(pred_obj) {
  if ("christian" %in% colnames(pred_obj)) safe_prob(pred_obj[, "christian"])
  else safe_prob(pred_obj[, ncol(pred_obj)])
}
fit_nb_model       <- function(x_dfm, y) quanteda.textmodels::textmodel_nb(x_dfm, y)
predict_nb_score   <- function(fit, newx_dfm) safe_logit(nb_prob_christian(predict(fit, newdata = newx_dfm, type = "probability")))

# --- Elastic net (replaces ridge) ---
fit_enet_fixed <- function(x, y, alpha, lambda) {
  suppressWarnings(glmnet::glmnet(
    x = x, y = as.integer(y == "christian"),
    family = "binomial", alpha = alpha, lambda = lambda,
    standardize = TRUE, maxit = 1e6
  ))
}
enet_has_nonzero   <- function(fit) sum(as.numeric(coef(fit)[-1, 1]) != 0) > 0
predict_enet_score <- function(fit, newx) as.numeric(predict(fit, newx = newx, type = "link"))
predict_enet_prob  <- function(fit, newx) safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))

# SVM
svm_to_matrix <- function(x) {
  if (inherits(x, "dgCMatrix") || inherits(x, "dsCMatrix") || inherits(x, "dgRMatrix")) return(x)
  if (inherits(x, "matrix")) return(x)
  as.matrix(x)
}
fit_svm_fixed <- function(x, y, cost) {
  x_use <- svm_to_matrix(x)
  fit <- LiblineaR::LiblineaR(data = x_use, target = y, type = 7, cost = cost)
  pred_tr <- predict(fit, x_use, decisionValues = TRUE)
  raw <- as.numeric(pred_tr$decisionValues[, 1])
  sgn <- if (mean(raw[y == "christian"]) >= mean(raw[y == "non_christian"])) 1 else -1
  list(model = fit, sign = sgn, cost = cost)
}
predict_svm_score <- function(obj, x) {
  pred <- predict(obj$model, svm_to_matrix(x), decisionValues = TRUE)
  obj$sign * as.numeric(pred$decisionValues[, 1])
}

# XGBoost
make_xgb_params <- function(row, scale_pos_weight = 1) {
  list(
    booster = "gbtree", objective = "binary:logistic", eval_metric = "logloss",
    tree_method = "hist",
    eta = as.numeric(row$eta),
    max_depth = as.integer(row$max_depth),
    min_child_weight = as.numeric(row$min_child_weight),
    subsample = as.numeric(row$subsample),
    colsample_bytree = as.numeric(row$colsample_bytree),
    gamma = as.numeric(row$gamma),
    scale_pos_weight = scale_pos_weight
  )
}
fit_xgb_fixed <- function(x, y, params, nrounds, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  xgboost::xgb.train(
    params = params,
    data = xgboost::xgb.DMatrix(x, label = as.integer(y == "christian")),
    nrounds = as.integer(nrounds), verbose = 0
  )
}
predict_xgb_score <- function(fit, x) as.numeric(predict(fit, xgboost::xgb.DMatrix(x), outputmargin = TRUE))

# LightGBM
make_lgb_params <- function(row, seed = 123) {
  list(
    objective = "binary", metric = "binary_logloss", boosting = "gbdt",
    learning_rate = as.numeric(row$learning_rate),
    num_leaves = as.integer(row$num_leaves),
    min_data_in_leaf = as.integer(row$min_data_in_leaf),
    feature_fraction = as.numeric(row$feature_fraction),
    bagging_fraction = as.numeric(row$bagging_fraction),
    bagging_freq = 1L, is_unbalance = TRUE, first_metric_only = TRUE,
    seed = seed, feature_fraction_seed = seed, bagging_seed = seed,
    verbosity = -1
  )
}
make_lgb_dataset_params <- function() list(feature_pre_filter = FALSE)
predict_lgb_prob <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "response"))),
    error = function(e) tryCatch(
      do.call(predict, args),
      error = function(e2) rep(0.5, nrow(x))
    )
  )
  safe_prob(as.numeric(out))
}
predict_lgb_score <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "raw"))),
    error = function(e) tryCatch(
      do.call(predict, c(args, list(rawscore = TRUE))),
      error = function(e2) qlogis(safe_prob(predict_lgb_prob(model, x, num_iteration)))
    )
  )
  as.numeric(out)
}
fit_lgb_fixed <- function(x, y, params, nrounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x, label = as.integer(y == "christian"),
    params = make_lgb_dataset_params(), free_raw_data = TRUE
  )
  lightgbm::lgb.train(params = params, data = dtrain,
                     nrounds = as.integer(nrounds), verbose = -1)
}
```

### Seeded hyperparameters and minimal local search

Rather than rerun the full nested tuning grid inside every (fold, n) combination, we **load the
pre-tuned hyperparameters from `model_hyperparameters_full.csv`** (produced by the earlier full
pipeline) and only do a very small local refinement in each fold. The tree-shape hyperparameters
(eta, max_depth, leaves, etc.) and the elastic-net alpha are held fixed at the seeded best — we
only sweep a handful of values around lambda (elastic net), cost (SVM), and nrounds (boosters).

```{r}
# Locate the seeded hyperparameter CSV. The user stated the path is
# ../../results/predictions/non_transformer_lemmatized/600ad relative to this
# notebook, which is what `results_path` already points to. We fall back to a
# hardcoded relative path in case this chunk is run from a different working dir.
candidate_paths <- c(
  file.path(results_path, "model_hyperparameters_full.csv"),
  file.path("../../results/predictions/non_transformer_lemmatized/600ad",
            "model_hyperparameters_full.csv")
)
hyper_seed_path <- candidate_paths[file.exists(candidate_paths)][1]
if (is.na(hyper_seed_path)) {
  stop("Could not locate model_hyperparameters_full.csv. Tried: ",
       paste(candidate_paths, collapse = " ; "))
}

hyper_seed <- readr::read_csv(hyper_seed_path, show_col_types = FALSE)
cat("\n===== Seeded hyperparameters loaded from", hyper_seed_path, "=====\n")
print(as.data.frame(hyper_seed))

pick_seed <- function(model_name) {
  row <- hyper_seed %>% filter(model == model_name)
  if (nrow(row) == 0) stop("No seeded row for model = ", model_name)
  as.list(row[1, , drop = FALSE])
}

enet_seed <- pick_seed("elastic_net")
svm_seed  <- pick_seed("svm_platt")
xgb_seed  <- pick_seed("xgboost")
lgb_seed  <- pick_seed("lightgbm")

# --- Minimal search grids centered on the seeded best values.

# Elastic net: fix alpha at the seeded value, sweep lambda on a small log-grid around lambda_1se.
enet_fixed_alpha   <- as.numeric(enet_seed$alpha)
enet_lambda_center <- as.numeric(enet_seed$lambda_1se)
enet_lambda_local  <- enet_lambda_center * c(0.5, 0.75, 1.0, 1.33, 2.0)

# SVM: sweep cost locally around the seeded best.
svm_cost_center <- as.numeric(svm_seed$cost)
svm_cost_local  <- unique(c(svm_cost_center / 2, svm_cost_center, svm_cost_center * 2))

# XGBoost: fix every tree-shape parameter at the seeded best, sweep only nrounds.
xgb_fixed_pars_row <- tibble(
  eta              = as.numeric(xgb_seed$eta),
  max_depth        = as.integer(xgb_seed$max_depth),
  min_child_weight = as.numeric(xgb_seed$min_child_weight),
  subsample        = as.numeric(xgb_seed$subsample),
  colsample_bytree = as.numeric(xgb_seed$colsample_bytree),
  gamma            = as.numeric(xgb_seed$gamma)
)
xgb_nrounds_center <- as.integer(xgb_seed$best_nrounds)
xgb_nrounds_local  <- unique(pmax(50L, as.integer(round(xgb_nrounds_center * c(0.5, 1.0, 1.5)))))

# LightGBM: fix every tree-shape parameter at the seeded best, sweep only nrounds.
lgb_fixed_pars_row <- tibble(
  learning_rate    = as.numeric(lgb_seed$learning_rate),
  num_leaves       = as.integer(lgb_seed$num_leaves),
  min_data_in_leaf = as.integer(lgb_seed$min_data_in_leaf),
  feature_fraction = as.numeric(lgb_seed$feature_fraction),
  bagging_fraction = as.numeric(lgb_seed$bagging_fraction)
)
lgb_nrounds_center <- as.integer(lgb_seed$best_nrounds)
lgb_nrounds_local  <- unique(pmax(20L, as.integer(round(lgb_nrounds_center * c(0.5, 1.0, 1.5)))))

cat(sprintf("\nEN:   alpha = %s (fixed), lambda grid = {%s}\n",
            enet_fixed_alpha,
            paste(signif(enet_lambda_local, 4), collapse = ", ")))
cat(sprintf("SVM:  cost grid = {%s}\n",
            paste(signif(svm_cost_local, 4), collapse = ", ")))
cat(sprintf("XGB:  tree pars fixed, nrounds grid = {%s}\n",
            paste(xgb_nrounds_local, collapse = ", ")))
cat(sprintf("LGB:  tree pars fixed, nrounds grid = {%s}\n",
            paste(lgb_nrounds_local, collapse = ", ")))
```

### Minimal-tuning CV helpers

These are the per-model helpers used inside the outer-fold loop. Each one does a single short
inner-CV pass over a tiny local grid (or no grid at all, for NB), and returns the selected
hyperparameter + the inner-CV scores needed for Platt calibration.

```{r}
# Naive Bayes: no hyperparameters. Return inner-CV scores for Platt calibration.
cv_nb_scores <- function(x_dfm, y, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_nb_model(x_dfm[tr, ], y[tr])
    scores[va] <- predict_nb_score(fit, x_dfm[va, ])
  }
  scores
}

# Elastic net: fix alpha, choose lambda from a short local grid by inner-CV log loss.
tune_enet_local <- function(x, y, alpha, lambda_grid, inner_k = 5, seed = 123, prefix = "") {
  folds <- make_folds(y, inner_k, seed)
  rows <- vector("list", length(lambda_grid))
  for (g in seq_along(lambda_grid)) {
    lam <- lambda_grid[g]
    fold_losses <- rep(NA_real_, length(folds))
    fold_ok <- TRUE
    for (j in seq_along(folds)) {
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lam)
      if (!enet_has_nonzero(fit)) { fold_ok <- FALSE; break }
      p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
      fold_losses[j] <- log_loss_safe(y[va], p_va)
    }
    rows[[g]] <- tibble(
      alpha  = alpha,
      lambda = lam,
      cvm    = if (fold_ok) mean(fold_losses, na.rm = TRUE) else Inf
    )
  }
  tuning <- bind_rows(rows) %>% filter(is.finite(cvm))
  if (nrow(tuning) == 0) {
    warning(prefix, ": all lambda candidates produced all-zero coefficients; using seeded lambda.")
    return(list(alpha = alpha, lambda = lambda_grid[which.min(abs(lambda_grid - enet_lambda_center))],
                tuning = bind_rows(rows)))
  }
  best <- tuning %>% arrange(cvm) %>% slice(1)
  list(alpha = best$alpha, lambda = best$lambda, tuning = bind_rows(rows))
}

cv_enet_scores <- function(x, y, alpha, lambda, k = 5, seed = 123) {
  folds  <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
  }
  scores
}

# SVM: choose cost from a small local grid by inner-CV margin AUC.
cv_svm_scores <- function(x, y, cost, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
  }
  scores
}

tune_svm_local <- function(x, y, cost_grid, inner_k = 5, seed = 123, prefix = "") {
  rows <- vector("list", length(cost_grid))
  for (i in seq_along(cost_grid)) {
    scores_i <- cv_svm_scores(x, y, cost = cost_grid[i], k = inner_k, seed = seed + i)
    rows[[i]] <- tibble(cost = cost_grid[i], auc_margin = auc_margin_safe(y, scores_i))
  }
  tuning <- bind_rows(rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = bind_rows(rows))
}

# XGBoost: tree-shape params fixed; choose nrounds from a tiny local grid by inner-CV log loss.
tune_xgb_local <- function(x, y, fixed_pars_row, nrounds_grid,
                           inner_k = 5, seed = 123, prefix = "") {
  folds <- make_folds(y, inner_k, seed)
  n_pos <- sum(y == "christian"); n_neg <- sum(y == "non_christian")
  spw <- n_neg / max(n_pos, 1)

  rows <- vector("list", length(nrounds_grid))
  for (g in seq_along(nrounds_grid)) {
    nr <- nrounds_grid[g]
    oof_prob <- rep(NA_real_, length(y))
    for (j in seq_along(folds)) {
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "christian"); n_neg_tr <- sum(y[tr] == "non_christian")
      spw_tr <- n_neg_tr / max(n_pos_tr, 1)
      fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr],
                           params = make_xgb_params(fixed_pars_row,
                                                    scale_pos_weight = spw_tr),
                           nrounds = nr, seed = seed + j)
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE], label = as.integer(y[va] == "christian"))
      oof_prob[va] <- safe_prob(as.numeric(predict(fit, dva)))
    }
    rows[[g]] <- fixed_pars_row %>%
      mutate(nrounds  = as.integer(nr),
             log_loss = log_loss_safe(y, oof_prob),
             auc      = auc_safe(y, oof_prob))
  }
  tuning <- bind_rows(rows) %>% arrange(log_loss, desc(auc), nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best         = best %>% mutate(best_nrounds = nrounds),
       params       = make_xgb_params(fixed_pars_row, scale_pos_weight = spw),
       best_nrounds = as.integer(best$nrounds),
       tuning       = tuning)
}

cv_xgb_scores <- function(x, y, params, nrounds, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "christian"); n_neg_tr <- sum(y[tr] == "non_christian")
    fold_params <- modifyList(params, list(scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr],
                         params = fold_params, nrounds = nrounds, seed = seed + j)
    scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
  }
  scores
}

# LightGBM: tree-shape params fixed; choose nrounds from a tiny local grid by inner-CV log loss.
tune_lgb_local <- function(x, y, fixed_pars_row, nrounds_grid,
                           inner_k = 5, seed = 123, prefix = "") {
  folds <- make_folds(y, inner_k, seed)
  rows <- vector("list", length(nrounds_grid))
  for (g in seq_along(nrounds_grid)) {
    nr <- nrounds_grid[g]
    oof_prob <- rep(NA_real_, length(y))
    for (j in seq_along(folds)) {
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      params_j <- make_lgb_params(fixed_pars_row, seed = seed + 1000L * g + j)
      fit <- fit_lgb_fixed(x = x[tr, , drop = FALSE], y = y[tr],
                           params = params_j, nrounds = nr)
      oof_prob[va] <- predict_lgb_prob(fit, x[va, , drop = FALSE], num_iteration = nr)
    }
    rows[[g]] <- fixed_pars_row %>%
      mutate(nrounds  = as.integer(nr),
             log_loss = log_loss_safe(y, oof_prob),
             auc      = auc_safe(y, oof_prob))
  }
  tuning <- bind_rows(rows) %>% arrange(log_loss, desc(auc), nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best         = best %>% mutate(best_nrounds = nrounds),
       params       = make_lgb_params(fixed_pars_row, seed = seed + 9999L),
       best_nrounds = as.integer(best$nrounds),
       tuning       = tuning)
}

cv_lgb_scores <- function(x, y, params, nrounds, k = 5, seed = 123) {
  folds <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE], y = y[tr],
      params = modifyList(params, list(seed = seed + j,
                                        feature_fraction_seed = seed + j,
                                        bagging_seed = seed + j)),
      nrounds = nrounds
    )
    scores[va] <- predict_lgb_score(fit, x[va, , drop = FALSE], num_iteration = nrounds)
  }
  scores
}
```

### Data loading

```{r}
log_progress("load_data", "reading corpus and labels")

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

if (!all(c("date_range_start", "date_range_end") %in% names(df))) {
  stop("corpus_600ad.csv must contain date_range_start and date_range_end columns.")
}

df <- df %>%
  mutate(
    date_range_start = suppressWarnings(as.numeric(date_range_start)),
    date_range_end   = suppressWarnings(as.numeric(date_range_end)),
    date_mid         = (date_range_start + date_range_end) / 2
  )

dup_corpus <- df %>% count(id, name = "n") %>% filter(n > 1)
if (nrow(dup_corpus) > 0) {
  stop(paste0("Duplicate ids in corpus_600ad.csv: ",
              paste(head(dup_corpus$id, 20), collapse = ", ")))
}

labels_raw <- readr::read_csv(file.path(data_path, "corpus_600ad_labeled.csv"),
                              show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  select(id, label)

label_conflicts <- labels_raw %>%
  distinct(id, label) %>%
  count(id, name = "n_labels") %>%
  filter(n_labels > 1)
if (nrow(label_conflicts) > 0) {
  stop(paste0("Conflicting labels in corpus_600ad_labeled.csv: ",
              paste(head(label_conflicts$id, 20), collapse = ", ")))
}

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
check_binary_y(df_labeled$manual_label, "labeled data")

custom_additions <- c("m", "c", "l", "d", "gt", "lt", "br", "href",
                      "fridericus", "heinricus", "conradus", "meus", "tuus")
stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt",
                 what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt",
                 what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")
latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]

lab     <- df_labeled %>% arrange(row_id)
out     <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos <- match(lab$row_id, out$row_id)
n_lab   <- nrow(lab)

# Report how many labeled texts actually have valid dates — these are the only
# documents that can contribute to the date-word ranking step. Christian texts
# are still trained on and tested on regardless of their date coverage.
n_nc       <- sum(lab$manual_label == "non_christian")
n_nc_dated <- sum(lab$manual_label == "non_christian" & is.finite(lab$date_mid))
log_progress("date_coverage",
             sprintf("non-Christian labeled = %d, with valid date_mid = %d",
                     n_nc, n_nc_dated))
```

### Nested cross-validation with date-word-removal sweep

The outer loop is the standard 5-fold CV. The inner per-fold work does:

1. Compute the fold-specific date-word ranking from **non-Christian training texts only**.
2. For each $n$ in `n_remove_grid`, define a reduced vocabulary (fold vocab minus top-$n$), fit a
   vectorizer against that vocabulary, and train every classifier on the full training fold.
3. Score the held-out test fold and record OOF predictions indexed by $(\text{model}, n)$.

```{r}
outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)

models     <- c("naive_bayes", "enet", "svm_platt", "xgboost", "lightgbm")
n_grid_str <- as.character(n_remove_grid)

# Storage: OOF predictions indexed by model x n
make_oof_store <- function() {
  list(
    prob         = rep(NA_real_, n_lab),
    score        = rep(NA_real_, n_lab),
    pred_youden  = rep(NA_character_, n_lab),
    pred_f1      = rep(NA_character_, n_lab),
    outer_fold   = integer(n_lab)
  )
}
oof_store <- list()
for (m in models) {
  oof_store[[m]] <- list()
  for (nstr in n_grid_str) oof_store[[m]][[nstr]] <- make_oof_store()
}

removed_word_log  <- list()   # one row per (outer_fold, n, rank) with the removed word
fold_metric_rows  <- list()
fold_hyper_rows   <- list()
fold_threshold_rows <- list()

for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start", paste("fold", i, "of", length(outer_folds)))
  va   <- outer_folds[[i]]
  tr   <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]

  outer_dir <- file.path(checkpoint_root, sprintf("outer_fold_%02d", i))
  ensure_dir(outer_dir)

  # --- Step 1: fold-specific date-word ranking from non-Christian training texts only.
  nc_mask  <- y_tr == "non_christian"
  nc_idx   <- tr[nc_mask]
  nc_texts <- lab$text[nc_idx]
  nc_dates <- lab$date_mid[nc_idx]

  rank_path <- file.path(outer_dir, "date_word_ranking.rds")
  ranking <- load_rds_or_null(rank_path)
  if (is.null(ranking)) {
    ranking <- rank_date_words(nc_texts, nc_dates, latin_stops,
                               min_docfreq_rank = 5L)
    save_rds_atomic(ranking, rank_path)
  }
  log_progress("date_rank_done",
               sprintf("outer %d | ranked %d non-Christian terms", i, nrow(ranking)))

  # --- Step 2: for each n, build reduced-vocab matrices and train every model.
  for (n_drop in n_remove_grid) {
    nstr <- as.character(n_drop)
    words_to_drop <- top_date_words(ranking, n_drop)

    # Record the words removed for this fold/n.
    if (length(words_to_drop) > 0) {
      removed_word_log[[length(removed_word_log) + 1L]] <- tibble(
        outer_fold = i,
        n_removed  = n_drop,
        rank       = seq_along(words_to_drop),
        term       = words_to_drop,
        abs_rho    = head(ranking$abs_rho, n_drop),
        rho        = head(ranking$rho, n_drop)
      )
    }

    # Fit vectorizer on the full training fold, but exclude the date-loaded words.
    full_vec <- fit_vectorizer(lab$text[tr], latin_stops, min_df = min_docfreq)
    allowed  <- setdiff(full_vec$features, words_to_drop)
    vec_fold <- fit_vectorizer(lab$text[tr], latin_stops,
                               min_df = min_docfreq,
                               allowed_features = allowed)

    x_tr_dfm <- transform_counts_dfm(lab$text[tr], vec_fold, latin_stops)
    x_va_dfm <- transform_counts_dfm(lab$text[va], vec_fold, latin_stops)
    x_tr     <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
    x_va     <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)

    log_progress("fold_matrix_built",
                 sprintf("outer %d | n_remove = %d | %d features", i, n_drop, length(vec_fold$features)))

    # Convenience: fold model results into the OOF store.
    store_predictions <- function(m, va_prob, va_score, th_y, th_f) {
      oof_store[[m]][[nstr]]$prob[va]        <<- va_prob
      oof_store[[m]][[nstr]]$score[va]       <<- va_score
      oof_store[[m]][[nstr]]$pred_youden[va] <<- ifelse(va_prob >= th_y, "christian", "non_christian")
      oof_store[[m]][[nstr]]$pred_f1[va]     <<- ifelse(va_prob >= th_f, "christian", "non_christian")
      oof_store[[m]][[nstr]]$outer_fold[va]  <<- i
    }
    record_fold_rows <- function(m, va_prob, th_y, th_f, hyper = NULL) {
      metric_row <- tibble(
        outer_fold = i, model = m, n_removed = n_drop,
        auc       = auc_safe(y_va, va_prob),
        log_loss  = log_loss_safe(y_va, va_prob),
        brier     = brier_safe(y_va, va_prob)
      ) %>%
        bind_cols(hard_metrics(y_va, oof_store[[m]][[nstr]]$pred_youden[va]) %>%
                    rename_with(~ paste0(.x, "_youden"))) %>%
        bind_cols(hard_metrics(y_va, oof_store[[m]][[nstr]]$pred_f1[va]) %>%
                    rename_with(~ paste0(.x, "_f1")))
      fold_metric_rows[[length(fold_metric_rows) + 1L]]    <<- metric_row
      fold_threshold_rows[[length(fold_threshold_rows) + 1L]] <<- tibble(
        outer_fold = i, model = m, n_removed = n_drop,
        threshold_youden = th_y, threshold_f1 = th_f
      )
      if (!is.null(hyper)) {
        fold_hyper_rows[[length(fold_hyper_rows) + 1L]] <<- hyper %>%
          mutate(outer_fold = i, model = m, n_removed = n_drop)
      }
    }

    # ---- Naive Bayes ----
    log_progress("fit_nb", paste("outer", i, "| n =", n_drop))
    nb_train_scores <- cv_nb_scores(x_tr_dfm, y_tr, k = k_folds_inner, seed = 1000 + 10 * i + n_drop)
    nb_platt <- fit_platt(nb_train_scores, y_tr)
    nb_fit   <- fit_nb_model(x_tr_dfm, y_tr)
    nb_va_sc <- predict_nb_score(nb_fit, x_va_dfm)
    nb_tr_pr <- apply_platt(nb_platt, nb_train_scores)
    nb_va_pr <- apply_platt(nb_platt, nb_va_sc)
    nb_th_y  <- choose_threshold(y_tr, nb_tr_pr, "youden")
    nb_th_f  <- choose_threshold(y_tr, nb_tr_pr, "f1")
    store_predictions("naive_bayes", nb_va_pr, nb_va_sc, nb_th_y, nb_th_f)
    record_fold_rows("naive_bayes", nb_va_pr, nb_th_y, nb_th_f)

    # ---- Elastic net ----
    log_progress("fit_enet", paste("outer", i, "| n =", n_drop))
    enet_tuned <- tune_enet_local(x_tr, y_tr,
                                  alpha = enet_fixed_alpha,
                                  lambda_grid = enet_lambda_local,
                                  inner_k = k_folds_inner,
                                  seed = 2000 + 100 * i + n_drop,
                                  prefix = paste("outer", i, "n", n_drop))
    enet_train_scores <- cv_enet_scores(x_tr, y_tr,
                                        alpha = enet_tuned$alpha,
                                        lambda = enet_tuned$lambda,
                                        k = k_folds_inner,
                                        seed = 2100 + 100 * i + n_drop)
    enet_platt <- fit_platt(enet_train_scores, y_tr)
    enet_fit   <- fit_enet_fixed(x_tr, y_tr, alpha = enet_tuned$alpha, lambda = enet_tuned$lambda)
    enet_va_sc <- predict_enet_score(enet_fit, x_va)
    enet_tr_pr <- apply_platt(enet_platt, enet_train_scores)
    enet_va_pr <- apply_platt(enet_platt, enet_va_sc)
    enet_th_y  <- choose_threshold(y_tr, enet_tr_pr, "youden")
    enet_th_f  <- choose_threshold(y_tr, enet_tr_pr, "f1")
    store_predictions("enet", enet_va_pr, enet_va_sc, enet_th_y, enet_th_f)
    record_fold_rows("enet", enet_va_pr, enet_th_y, enet_th_f,
                     hyper = tibble(alpha = enet_tuned$alpha,
                                    lambda = enet_tuned$lambda))

    # ---- SVM ----
    log_progress("fit_svm", paste("outer", i, "| n =", n_drop))
    svm_tuned <- tune_svm_local(x_tr, y_tr,
                                cost_grid = svm_cost_local,
                                inner_k = k_folds_inner,
                                seed = 3000 + 100 * i + n_drop,
                                prefix = paste("outer", i, "n", n_drop))
    svm_train_scores <- cv_svm_scores(x_tr, y_tr, cost = svm_tuned$cost,
                                      k = k_folds_inner,
                                      seed = 3100 + 100 * i + n_drop)
    svm_platt <- fit_platt(svm_train_scores, y_tr)
    svm_fit   <- fit_svm_fixed(x_tr, y_tr, cost = svm_tuned$cost)
    svm_va_sc <- predict_svm_score(svm_fit, x_va)
    svm_tr_pr <- apply_platt(svm_platt, svm_train_scores)
    svm_va_pr <- apply_platt(svm_platt, svm_va_sc)
    svm_th_y  <- choose_threshold(y_tr, svm_tr_pr, "youden")
    svm_th_f  <- choose_threshold(y_tr, svm_tr_pr, "f1")
    store_predictions("svm_platt", svm_va_pr, svm_va_sc, svm_th_y, svm_th_f)
    record_fold_rows("svm_platt", svm_va_pr, svm_th_y, svm_th_f,
                     hyper = tibble(cost = svm_tuned$cost))

    # ---- XGBoost ----
    log_progress("fit_xgb", paste("outer", i, "| n =", n_drop))
    xgb_tuned <- tune_xgb_local(x_tr, y_tr,
                                fixed_pars_row = xgb_fixed_pars_row,
                                nrounds_grid = xgb_nrounds_local,
                                inner_k = k_folds_inner,
                                seed = 4000 + 100 * i + n_drop,
                                prefix = paste("outer", i, "n", n_drop))
    xgb_train_scores <- cv_xgb_scores(x_tr, y_tr, params = xgb_tuned$params,
                                      nrounds = xgb_tuned$best_nrounds,
                                      k = k_folds_inner,
                                      seed = 4100 + 100 * i + n_drop)
    xgb_platt <- fit_platt(xgb_train_scores, y_tr)
    xgb_fit   <- fit_xgb_fixed(x_tr, y_tr, params = xgb_tuned$params,
                               nrounds = xgb_tuned$best_nrounds,
                               seed = 4200 + 100 * i + n_drop)
    xgb_va_sc <- predict_xgb_score(xgb_fit, x_va)
    xgb_tr_pr <- apply_platt(xgb_platt, xgb_train_scores)
    xgb_va_pr <- apply_platt(xgb_platt, xgb_va_sc)
    xgb_th_y  <- choose_threshold(y_tr, xgb_tr_pr, "youden")
    xgb_th_f  <- choose_threshold(y_tr, xgb_tr_pr, "f1")
    store_predictions("xgboost", xgb_va_pr, xgb_va_sc, xgb_th_y, xgb_th_f)
    record_fold_rows("xgboost", xgb_va_pr, xgb_th_y, xgb_th_f,
                     hyper = xgb_tuned$best %>% select(-any_of("candidate_id")))

    # ---- LightGBM ----
    log_progress("fit_lgb", paste("outer", i, "| n =", n_drop))
    lgb_tuned <- tune_lgb_local(x_tr, y_tr,
                                fixed_pars_row = lgb_fixed_pars_row,
                                nrounds_grid = lgb_nrounds_local,
                                inner_k = k_folds_inner,
                                seed = 5000 + 100 * i + n_drop,
                                prefix = paste("outer", i, "n", n_drop))
    lgb_train_scores <- cv_lgb_scores(x_tr, y_tr, params = lgb_tuned$params,
                                      nrounds = lgb_tuned$best_nrounds,
                                      k = k_folds_inner,
                                      seed = 5100 + 100 * i + n_drop)
    lgb_platt <- fit_platt(lgb_train_scores, y_tr)
    lgb_fit   <- fit_lgb_fixed(x = x_tr, y = y_tr, params = lgb_tuned$params,
                               nrounds = lgb_tuned$best_nrounds)
    lgb_va_sc <- predict_lgb_score(lgb_fit, x_va,
                                   num_iteration = lgb_tuned$best_nrounds)
    lgb_tr_pr <- apply_platt(lgb_platt, lgb_train_scores)
    lgb_va_pr <- apply_platt(lgb_platt, lgb_va_sc)
    lgb_th_y  <- choose_threshold(y_tr, lgb_tr_pr, "youden")
    lgb_th_f  <- choose_threshold(y_tr, lgb_tr_pr, "f1")
    store_predictions("lightgbm", lgb_va_pr, lgb_va_sc, lgb_th_y, lgb_th_f)
    record_fold_rows("lightgbm", lgb_va_pr, lgb_th_y, lgb_th_f,
                     hyper = lgb_tuned$best %>% select(-any_of("candidate_id")))
  } # end n_drop loop

  log_progress("outer_fold_done", paste("fold", i, "of", length(outer_folds)))
}
```

### OOF metrics aggregated by (model, n)

```{r}
summarize_oof <- function(y, p, pred_y, pred_f, model_name, n_removed) {
  hy <- hard_metrics(y, pred_y)
  hf <- hard_metrics(y, pred_f)
  tibble(
    model = model_name, n_removed = n_removed,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p), brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    tp_youden = hy$tp, tn_youden = hy$tn, fp_youden = hy$fp, fn_youden = hy$fn,
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc,
    tp_f1 = hf$tp, tn_f1 = hf$tn, fp_f1 = hf$fp, fn_f1 = hf$fn
  )
}

oof_rows <- list()
for (m in models) for (nstr in n_grid_str) {
  s <- oof_store[[m]][[nstr]]
  oof_rows[[length(oof_rows) + 1L]] <- summarize_oof(
    lab$manual_label, s$prob, s$pred_youden, s$pred_f1, m, as.integer(nstr)
  )
}
oof_metrics_by_n <- bind_rows(oof_rows) %>% arrange(model, n_removed)

cat("\n===== OOF metrics by model and n_removed =====\n")
print(
  oof_metrics_by_n %>%
    select(model, n_removed, auc, log_loss, f1_youden, mcc_youden),
  n = nrow(oof_metrics_by_n)
)

fold_metric_df    <- bind_rows(fold_metric_rows)    %>% arrange(model, n_removed, outer_fold)
fold_threshold_df <- bind_rows(fold_threshold_rows) %>% arrange(model, n_removed, outer_fold)
fold_hyper_df     <- bind_rows(fold_hyper_rows)     %>% arrange(model, n_removed, outer_fold)
removed_words_df  <- bind_rows(removed_word_log)    %>% arrange(outer_fold, n_removed, rank)

write_csv_atomic(oof_metrics_by_n,   file.path(output_dir, "oof_metrics_by_model_and_n.csv"))
write_csv_atomic(fold_metric_df,     file.path(output_dir, "fold_metrics_by_model_and_n.csv"))
write_csv_atomic(fold_threshold_df,  file.path(output_dir, "fold_thresholds_by_model_and_n.csv"))
write_csv_atomic(fold_hyper_df,      file.path(output_dir, "fold_hyperparameters_by_model_and_n.csv"))
write_csv_atomic(removed_words_df,   file.path(output_dir, "removed_date_words_by_fold.csv"))

# Per-document OOF predictions, one row per labeled doc per (model, n).
oof_predictions_long <- purrr::map_dfr(models, function(m) {
  purrr::map_dfr(n_grid_str, function(nstr) {
    s <- oof_store[[m]][[nstr]]
    tibble(
      row_id         = lab$row_id,
      id             = lab$id,
      manual_label   = lab$manual_label,
      model          = m,
      n_removed      = as.integer(nstr),
      outer_fold     = s$outer_fold,
      prob           = s$prob,
      score          = s$score,
      pred_youden    = s$pred_youden,
      pred_f1        = s$pred_f1
    )
  })
})
write_csv_atomic(oof_predictions_long,
                 file.path(output_dir, "oof_predictions_by_model_and_n.csv"))

log_progress("oof_summary_done", "nested CV with temporal sweep completed")
```

### Visualizing performance vs. n

The point of this sweep: each curve is one classifier. A classifier whose AUC/F1 collapses as $n$
grows was leaning heavily on time-correlated vocabulary. A classifier that stays flat is picking
up genuine Christian / non-Christian signal independent of date.

```{r}
library(ggplot2)

plot_df_long <- oof_metrics_by_n %>%
  select(model, n_removed, auc, f1_youden, mcc_youden) %>%
  pivot_longer(c(auc, f1_youden, mcc_youden),
               names_to = "metric", values_to = "value") %>%
  mutate(metric = recode(metric,
                         auc = "AUC",
                         f1_youden = "F1 (Youden threshold)",
                         mcc_youden = "MCC (Youden threshold)"))

p_sweep <- ggplot(plot_df_long,
                  aes(x = n_removed, y = value, color = model, group = model)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.2) +
  facet_wrap(~ metric, scales = "free_y") +
  labs(
    x = "Number of top date-correlated words removed (n)",
    y = NULL,
    color = "Model"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank())

sweep_dir <- file.path(checkpoint_path, "temporal_sweep_plots")
dir.create(sweep_dir, recursive = TRUE, showWarnings = FALSE)
ggsave(file.path(sweep_dir, "metrics_vs_n_removed.png"),
       p_sweep, width = 10, height = 5.5, dpi = 400, bg = "white")
ggsave(file.path(sweep_dir, "metrics_vs_n_removed.pdf"),
       p_sweep, width = 10, height = 5.5, bg = "white")
p_sweep
```

### Full-data deployment models, one per (model, n)

For each $n$ we identify the date-correlated words from the non-Christian **labeled** texts (no
held-out set here since this is the deployment fit), remove them from the full vocabulary, and
retrain each classifier. This gives a deployable model per $n$ that downstream scripts can pick
from depending on how aggressively they want to strip time-loaded vocabulary.

```{r}
full_dir <- file.path(checkpoint_root, "full_fit")
ensure_dir(full_dir)

# One date-word ranking from all non-Christian labeled texts.
nc_full_idx   <- which(lab$manual_label == "non_christian")
full_ranking  <- rank_date_words(lab$text[nc_full_idx],
                                 lab$date_mid[nc_full_idx],
                                 latin_stops, min_docfreq_rank = 5L)
save_rds_atomic(full_ranking, file.path(full_dir, "date_word_ranking_full.rds"))

full_predictions <- tibble(row_id = out$row_id, id = out$id, manual_label = out$manual_label)
deployment_thresholds <- list()
deployment_hypers     <- list()
full_removed_words    <- list()

for (n_drop in n_remove_grid) {
  nstr <- as.character(n_drop)
  log_progress("full_fit_start", paste("n_remove =", n_drop))

  words_to_drop <- top_date_words(full_ranking, n_drop)
  if (length(words_to_drop) > 0) {
    full_removed_words[[length(full_removed_words) + 1L]] <- tibble(
      n_removed = n_drop,
      rank      = seq_along(words_to_drop),
      term      = words_to_drop,
      abs_rho   = head(full_ranking$abs_rho, n_drop),
      rho       = head(full_ranking$rho, n_drop)
    )
  }

  base_vec <- fit_vectorizer(lab$text, latin_stops, min_df = min_docfreq)
  allowed  <- setdiff(base_vec$features, words_to_drop)
  vec_full <- fit_vectorizer(lab$text, latin_stops, min_df = min_docfreq,
                             allowed_features = allowed)

  x_lab_dfm <- transform_counts_dfm(lab$text, vec_full, latin_stops)
  x_all_dfm <- transform_counts_dfm(out$text, vec_full, latin_stops)
  x_lab     <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
  x_all     <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

  # --- NB ---
  nb_cv_sc  <- cv_nb_scores(x_lab_dfm, lab$manual_label, k = k_folds_outer, seed = 8001 + n_drop)
  nb_platt  <- fit_platt(nb_cv_sc, lab$manual_label)
  nb_fit    <- fit_nb_model(x_lab_dfm, lab$manual_label)
  nb_all_sc <- predict_nb_score(nb_fit, x_all_dfm)
  nb_all_pr <- apply_platt(nb_platt, nb_all_sc)
  nb_cf_pr  <- apply_platt(nb_platt, nb_cv_sc)
  nb_th_y   <- choose_threshold(lab$manual_label, nb_cf_pr, "youden")
  nb_th_f   <- choose_threshold(lab$manual_label, nb_cf_pr, "f1")

  # --- Elastic net ---
  enet_tuned_f <- tune_enet_local(x_lab, lab$manual_label,
                                  alpha = enet_fixed_alpha,
                                  lambda_grid = enet_lambda_local,
                                  inner_k = k_folds_inner,
                                  seed = 8002 + n_drop, prefix = paste("full n", n_drop))
  enet_cv_sc   <- cv_enet_scores(x_lab, lab$manual_label,
                                 alpha = enet_tuned_f$alpha,
                                 lambda = enet_tuned_f$lambda,
                                 k = k_folds_outer, seed = 8102 + n_drop)
  enet_platt   <- fit_platt(enet_cv_sc, lab$manual_label)
  enet_fit     <- fit_enet_fixed(x_lab, lab$manual_label,
                                 alpha = enet_tuned_f$alpha, lambda = enet_tuned_f$lambda)
  enet_all_sc  <- predict_enet_score(enet_fit, x_all)
  enet_all_pr  <- apply_platt(enet_platt, enet_all_sc)
  enet_cf_pr   <- apply_platt(enet_platt, enet_cv_sc)
  enet_th_y    <- choose_threshold(lab$manual_label, enet_cf_pr, "youden")
  enet_th_f    <- choose_threshold(lab$manual_label, enet_cf_pr, "f1")

  # --- SVM ---
  svm_tuned_f  <- tune_svm_local(x_lab, lab$manual_label,
                                 cost_grid = svm_cost_local,
                                 inner_k = k_folds_inner,
                                 seed = 8003 + n_drop, prefix = paste("full n", n_drop))
  svm_cv_sc    <- cv_svm_scores(x_lab, lab$manual_label, cost = svm_tuned_f$cost,
                                k = k_folds_outer, seed = 8103 + n_drop)
  svm_platt    <- fit_platt(svm_cv_sc, lab$manual_label)
  svm_fit      <- fit_svm_fixed(x_lab, lab$manual_label, cost = svm_tuned_f$cost)
  svm_all_sc   <- predict_svm_score(svm_fit, x_all)
  svm_all_pr   <- apply_platt(svm_platt, svm_all_sc)
  svm_cf_pr    <- apply_platt(svm_platt, svm_cv_sc)
  svm_th_y     <- choose_threshold(lab$manual_label, svm_cf_pr, "youden")
  svm_th_f     <- choose_threshold(lab$manual_label, svm_cf_pr, "f1")

  # --- XGBoost ---
  xgb_tuned_f  <- tune_xgb_local(x_lab, lab$manual_label,
                                 fixed_pars_row = xgb_fixed_pars_row,
                                 nrounds_grid = xgb_nrounds_local,
                                 inner_k = k_folds_inner,
                                 seed = 8004 + n_drop, prefix = paste("full n", n_drop))
  xgb_cv_sc    <- cv_xgb_scores(x_lab, lab$manual_label, params = xgb_tuned_f$params,
                                nrounds = xgb_tuned_f$best_nrounds,
                                k = k_folds_outer, seed = 8104 + n_drop)
  xgb_platt    <- fit_platt(xgb_cv_sc, lab$manual_label)
  xgb_fit      <- fit_xgb_fixed(x_lab, lab$manual_label, params = xgb_tuned_f$params,
                                nrounds = xgb_tuned_f$best_nrounds,
                                seed = 8204 + n_drop)
  xgb_all_sc   <- predict_xgb_score(xgb_fit, x_all)
  xgb_all_pr   <- apply_platt(xgb_platt, xgb_all_sc)
  xgb_cf_pr    <- apply_platt(xgb_platt, xgb_cv_sc)
  xgb_th_y     <- choose_threshold(lab$manual_label, xgb_cf_pr, "youden")
  xgb_th_f     <- choose_threshold(lab$manual_label, xgb_cf_pr, "f1")

  # --- LightGBM ---
  lgb_tuned_f  <- tune_lgb_local(x_lab, lab$manual_label,
                                 fixed_pars_row = lgb_fixed_pars_row,
                                 nrounds_grid = lgb_nrounds_local,
                                 inner_k = k_folds_inner,
                                 seed = 8005 + n_drop, prefix = paste("full n", n_drop))
  lgb_cv_sc    <- cv_lgb_scores(x_lab, lab$manual_label, params = lgb_tuned_f$params,
                                nrounds = lgb_tuned_f$best_nrounds,
                                k = k_folds_outer, seed = 8105 + n_drop)
  lgb_platt    <- fit_platt(lgb_cv_sc, lab$manual_label)
  lgb_fit      <- fit_lgb_fixed(x = x_lab, y = lab$manual_label,
                                params = lgb_tuned_f$params,
                                nrounds = lgb_tuned_f$best_nrounds)
  lgb_all_sc   <- predict_lgb_score(lgb_fit, x_all,
                                    num_iteration = lgb_tuned_f$best_nrounds)
  lgb_all_pr   <- apply_platt(lgb_platt, lgb_all_sc)
  lgb_cf_pr    <- apply_platt(lgb_platt, lgb_cv_sc)
  lgb_th_y     <- choose_threshold(lab$manual_label, lgb_cf_pr, "youden")
  lgb_th_f     <- choose_threshold(lab$manual_label, lgb_cf_pr, "f1")

  # Add prediction columns to the full-corpus output, tagged by n_removed.
  pred_cols <- tibble(
    !!paste0("prob_nb_n",    n_drop) := nb_all_pr,
    !!paste0("prob_enet_n",  n_drop) := enet_all_pr,
    !!paste0("prob_svm_n",   n_drop) := svm_all_pr,
    !!paste0("prob_xgb_n",   n_drop) := xgb_all_pr,
    !!paste0("prob_lgb_n",   n_drop) := lgb_all_pr,

    !!paste0("score_nb_n",   n_drop) := nb_all_sc,
    !!paste0("score_enet_n", n_drop) := enet_all_sc,
    !!paste0("score_svm_n",  n_drop) := svm_all_sc,
    !!paste0("score_xgb_n",  n_drop) := xgb_all_sc,
    !!paste0("score_lgb_n",  n_drop) := lgb_all_sc,

    !!paste0("label_nb_youden_n",   n_drop) := ifelse(nb_all_pr   >= nb_th_y,   "christian", "non_christian"),
    !!paste0("label_enet_youden_n", n_drop) := ifelse(enet_all_pr >= enet_th_y, "christian", "non_christian"),
    !!paste0("label_svm_youden_n",  n_drop) := ifelse(svm_all_pr  >= svm_th_y,  "christian", "non_christian"),
    !!paste0("label_xgb_youden_n",  n_drop) := ifelse(xgb_all_pr  >= xgb_th_y,  "christian", "non_christian"),
    !!paste0("label_lgb_youden_n",  n_drop) := ifelse(lgb_all_pr  >= lgb_th_y,  "christian", "non_christian"),

    !!paste0("label_nb_f1_n",   n_drop) := ifelse(nb_all_pr   >= nb_th_f,   "christian", "non_christian"),
    !!paste0("label_enet_f1_n", n_drop) := ifelse(enet_all_pr >= enet_th_f, "christian", "non_christian"),
    !!paste0("label_svm_f1_n",  n_drop) := ifelse(svm_all_pr  >= svm_th_f,  "christian", "non_christian"),
    !!paste0("label_xgb_f1_n",  n_drop) := ifelse(xgb_all_pr  >= xgb_th_f,  "christian", "non_christian"),
    !!paste0("label_lgb_f1_n",  n_drop) := ifelse(lgb_all_pr  >= lgb_th_f,  "christian", "non_christian")
  )
  full_predictions <- bind_cols(full_predictions, pred_cols)

  deployment_thresholds[[length(deployment_thresholds) + 1L]] <- tibble(
    n_removed = n_drop,
    model = c("naive_bayes", "enet", "svm_platt", "xgboost", "lightgbm"),
    threshold_youden = c(nb_th_y, enet_th_y, svm_th_y, xgb_th_y, lgb_th_y),
    threshold_f1     = c(nb_th_f, enet_th_f, svm_th_f, xgb_th_f, lgb_th_f),
    threshold_scope  = "full_labeled_cross_fitted_calibrated"
  )

  deployment_hypers[[length(deployment_hypers) + 1L]] <- bind_rows(
    tibble(n_removed = n_drop, model = "naive_bayes"),
    tibble(n_removed = n_drop, model = "enet",
           alpha = enet_tuned_f$alpha,
           lambda = enet_tuned_f$lambda),
    tibble(n_removed = n_drop, model = "svm_platt", cost = svm_tuned_f$cost),
    xgb_tuned_f$best %>%
      select(-any_of("candidate_id")) %>%
      mutate(n_removed = n_drop, model = "xgboost"),
    lgb_tuned_f$best %>%
      select(-any_of("candidate_id")) %>%
      mutate(n_removed = n_drop, model = "lightgbm")
  )

  log_progress("full_fit_done_n", paste("n_remove =", n_drop))
}
library(reticulate)

# 1) run a .py script
source_python("my_script.py")

# 2) run Python code stored as a string
py_run_string("
x = 2 + 2
print(x)
")

# 3) import Python packages directly
pd <- import("pandas")
deployment_thresholds_df <- bind_rows(deployment_thresholds) %>%
  arrange(model, n_removed)
deployment_hypers_df     <- bind_rows(deployment_hypers) %>%
  arrange(model, n_removed)
full_removed_words_df    <- bind_rows(full_removed_words) %>%
  arrange(n_removed, rank)

write_csv_atomic(full_predictions,
                 file.path(output_dir, "full_corpus_predictions_by_n.csv"))
write_csv_atomic(deployment_thresholds_df,
                 file.path(output_dir, "deployment_thresholds_by_model_and_n.csv"))
write_csv_atomic(deployment_hypers_df,
                 file.path(output_dir, "deployment_hyperparameters_by_model_and_n.csv"))
write_csv_atomic(full_removed_words_df,
                 file.path(output_dir, "removed_date_words_full_fit.csv"))

log_progress("all_done", "temporal-sweep pipeline finished")
```

```{r}
Sys.setenv(RETICULATE_PYTHON = "~/.virtualenvs/bert-env/bin/python")
library(reticulate)
py_config()
```

```{r}
library(reticulate)
```

```{r}
Sys.setenv(RETICULATE_PYTHON = path.expand("~/.virtualenvs/bert-env/bin/python"))

py_config()
```


```{r}
f <- "../600ad_unlemmatized/bert_temporal_sweep.py"
x <- readLines(f, warn = FALSE)

x <- gsub(
  "^[[:space:]]*tokenizer = AutoTokenizer\\.from_pretrained\\(cfg\\['model_name'\\], use_fast=False\\)$",
  "",
  x
)

x <- gsub(
  "^[[:space:]]*ttokenizer_probe = AutoTokenizer\\.from_pretrained\\(cfg\\['model_name'\\], use_fast=False\\)$",
  "",
  x
)

x <- gsub(
  "mask_token = MASK_WITH_TOKEN or \\(tokenizer_probe\\.mask_token if tokenizer_probe\\.mask_token is not None else '\\[MASK\\]'\\)",
  "mask_token = MASK_WITH_TOKEN or \"[MASK]\"",
  x
)

writeLines(x, f)
```

```{r}

system2(
  command = path.expand("~/.virtualenvs/bert-env/bin/python"),
  args = "../600ad_unlemmatized/bert_temporal_sweep_fixed.py"
)
```

 