---
title: "Latin text classification"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of christian vs. non-christian documents


## Data loading and cleaning

### Loading in and filtering the data (this code chunk can be skipped if corpus.csv already exists in the script directory)

```{r}
library(tidyverse)
library(stringi)

data_path <- "../../data/processed_texts/600ad"
results_path <- "../../results/predictions/non_transformer_lemmatized/600ad"
```


```{r}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels <- readr::read_csv(file.path(data_path,"corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  transmute(id, manual_label = label) %>%
  distinct(id, .keep_all = TRUE)

df_labeled <- df %>%
  inner_join(labels, by = "id") 

```


### Loading in the stopwowrds

```{r}

custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href", 
  "fridericus", "heinricus", "conradus", "meus", "tuus" 
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")

stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- c(stops_v1, stops_v2, custom_additions)

latin_stops <- latin_stops[latin_stops != "*"]

latin_stops <- unique(latin_stops)

latin_stops <- sort(latin_stops)

head(latin_stops, 200)

```


### Defining a tokenization function to count relative frequencies per token for the NB classifier


```{r}
library(quanteda)
library(quanteda.textmodels)
library(tidyverse)
library(knitr)


get_standard_dfm <- function(text_data, stop_list) {

  toks <- tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    tokens_remove(pattern = stop_list)
  

  dfm(toks)
}


dfm_train <- get_standard_dfm(df_labeled$text, latin_stops)

dim(dfm_train)

```


### Getting rid of highly-date-correlated features

```{r}
library(Matrix)
library(ggplot2)

date_df <- df %>%
  filter(!is.na(date_range_start), !is.na(date_range_end), text != "") %>%
  mutate(
    date_range_start = as.numeric(date_range_start),
    date_range_end = as.numeric(date_range_end)
  )

dfm_date <- get_standard_dfm(date_df$text, latin_stops)
keep_docs <- ntoken(dfm_date) > 0
date_df <- date_df[keep_docs, ]
dfm_date <- dfm_date[keep_docs, ] %>%
  dfm_trim(min_docfreq = max(5, ceiling(nrow(date_df) * 0.02))) %>%
  dfm_weight(scheme = "prop")

X <- as(dfm_date, "dgCMatrix")

safe_spearman <- function(x, y) {
  if (sd(x) == 0 || sd(y) == 0) return(NA_real_)
  suppressWarnings(cor(x, y, method = "spearman"))
}

rho_start <- vapply(seq_len(ncol(X)), function(j) safe_spearman(as.numeric(X[, j]), date_df$date_range_start), numeric(1))
rho_end   <- vapply(seq_len(ncol(X)), function(j) safe_spearman(as.numeric(X[, j]), date_df$date_range_end), numeric(1))

cor_summary <- tibble(
  term = featnames(dfm_date),
  docfreq = docfreq(dfm_date),
  rho_start = rho_start,
  rho_end = rho_end,
  rho_mean = (rho_start + rho_end) / 2,
  mean_abs_rho = (abs(rho_start) + abs(rho_end)) / 2
) %>%
  drop_na(rho_start, rho_end) %>%
  arrange(desc(mean_abs_rho))

top_date_terms <- bind_rows(
  cor_summary %>% slice_max(rho_mean, n = 50),
  cor_summary %>% slice_min(rho_mean, n = 50)
) %>%
  distinct(term, .keep_all = TRUE) %>%
  arrange(rho_mean) %>%
  mutate(term = factor(term, levels = term))

plot_df <- top_date_terms %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(-term, names_to = "bound", values_to = "rho") %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))
```


```{r}
library(forcats)

out_dir <- "../../results/alternative_analyses/word_correlation"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

top_date_terms <- cor_summary %>%
  slice_head(n = 100) %>%
  arrange(rho_mean) %>%
  mutate(term = fct_inorder(term))

plot_df <- top_date_terms %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(-term, names_to = "bound", values_to = "rho") %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))

p <- ggplot(top_date_terms, aes(y = term)) +
  geom_segment(
    aes(x = rho_start, xend = rho_end, yend = term, color = rho_mean > 0),
    linewidth = 1.2, alpha = 0.7, show.legend = FALSE
  ) +
  geom_point(
    data = plot_df,
    aes(x = rho, shape = bound, fill = rho > 0),
    size = 2.6, color = "white", stroke = 0.4
  ) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5, color = "grey60") +
  scale_shape_manual(values = c("Start date" = 21, "End date" = 24), name = NULL) +
  scale_fill_manual(values = c(`TRUE` = "#2b6cb0", `FALSE` = "#c53030"), guide = "none") +
  scale_color_manual(values = c(`TRUE` = "#2b6cb0", `FALSE` = "#c53030")) +
  labs(x = "Spearman correlation with document date", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "top",
    axis.text.y = element_text(size = 6.5),
    plot.margin = margin(10, 20, 10, 10)
  )

plot_height <- max(16, 0.22 * nrow(top_date_terms))

ggsave(
  filename = file.path(out_dir, "top100_date_correlated_terms.png"),
  plot = p,
  width = 10,
  height = plot_height,
  dpi = 400,
  bg = "white"
)

ggsave(
  filename = file.path(out_dir, "top100_date_correlated_terms.pdf"),
  plot = p,
  width = 10,
  height = plot_height,
  bg = "white"
)

p
```
```{r}
out_dir <- "../../results/alternative_analyses/word_correlation"
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

dist_df <- cor_summary %>%
  select(term, rho_start, rho_end) %>%
  pivot_longer(
    cols = c(rho_start, rho_end),
    names_to = "bound",
    values_to = "rho"
  ) %>%
  mutate(bound = recode(bound, rho_start = "Start date", rho_end = "End date"))

p_dist <- ggplot(dist_df, aes(x = rho, fill = bound, color = bound)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 60, alpha = 0.30, position = "identity") +
  geom_density(linewidth = 0.9) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 0.5, color = "grey50") +
  scale_fill_manual(values = c("Start date" = "#2b6cb0", "End date" = "#c53030")) +
  scale_color_manual(values = c("Start date" = "#2b6cb0", "End date" = "#c53030")) +
  labs(x = "Spearman correlation", y = "Density", fill = NULL, color = NULL) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "top"
  )

ggsave(
  file.path(out_dir, "correlation_distribution_all_terms.png"),
  p_dist,
  width = 9,
  height = 5.5,
  dpi = 400,
  bg = "white"
)

ggsave(
  file.path(out_dir, "correlation_distribution_all_terms.pdf"),
  p_dist,
  width = 9,
  height = 5.5,
  bg = "white"
)

p_dist

```

## Training classification models

### Setting up model parameters


```{r}

library(tidyverse)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(pROC)

set.seed(123)

checkpoint_path <- "../../results/predictions/non_transformer_lemmatized/600ad"

k_folds_outer <- 5L
k_folds_inner <- 5L
xgb_max_rounds <- 4000L
lgb_max_rounds <- 4000L
early_stop_rounds <- 100L
min_docfreq <- 3L

checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
output_dir <- file.path(checkpoint_path, "model_outputs_checkpointed")
progress_path <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) {
    file.copy(tmp, path, overwrite = TRUE)
    unlink(tmp)
  }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

append_progress_row <- function(row, path) {
  if (!file.exists(path)) {
    write_csv_atomic(row, path)
  } else {
    old <- readr::read_csv(path, show_col_types = FALSE)
    write_csv_atomic(bind_rows(old, row), path)
  }
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step, if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

safe_prob <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb))) {
    stop(paste0("Both classes are required", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
  if (any(tb < 2)) {
    stop(paste0("Each class needs at least 2 observations", if (nzchar(where)) paste0(" in ", where) else "", "."))
  }
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for cross-validation.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

make_foldid <- function(y, k, seed = 123) {
  folds <- make_folds(y, k, seed)
  foldid <- integer(length(y))
  for (i in seq_along(folds)) foldid[folds[[i]]] <- i
  foldid
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y,
      predictor = s,
      levels = c("non_christian", "christian"),
      direction = "<",
      quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1)
  tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1)
  fn <- sum(y01 == 1 & p01 == 0)
  precision <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1 <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(
    accuracy = (tp + tn) / length(y),
    balanced_accuracy = bal_acc,
    precision = precision,
    recall = recall,
    specificity = specificity,
    f1 = f1,
    mcc = mcc,
    tp = tp,
    tn = tn,
    fp = fp,
    fn = fn
  )
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(
          response = y,
          predictor = p,
          levels = c("non_christian", "christian"),
          direction = "<",
          quiet = TRUE
        ),
        x = "best",
        input = "threshold",
        best.method = "youden",
        ret = "threshold",
        transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01 <- as.integer(y == "christian")
    cand <- sort(unique(p))
    cand <- unique(c(0, cand, 0.5, 1))
    f1s <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    tp_youden = hy$tp,
    tn_youden = hy$tn,
    fp_youden = hy$fp,
    fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc,
    tp_f1 = hf$tp,
    tn_f1 = hf$tn,
    fp_f1 = hf$fp,
    fn_f1 = hf$fn
  )
}

summarize_model_fold <- function(y, p, pred_youden, pred_f1, model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold,
    model = model_name,
    auc = auc_safe(y, p),
    log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy,
    balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision,
    recall_youden = hy$recall,
    specificity_youden = hy$specificity,
    f1_youden = hy$f1,
    mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy,
    balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision,
    recall_f1 = hf$recall,
    specificity_f1 = hf$specificity,
    f1_f1 = hf$f1,
    mcc_f1 = hf$mcc
  )
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]), by = keys)
  bind_rows(keep, new_rows)
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(), control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores), type = "response"))
  }
}

make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat <- as(x_counts, "dgCMatrix")
  x_vals <- x_mat@x
  x_mat@x <- 1 + log(x_vals)
  idf_vec <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}

nb_prob_christian <- function(pred_obj) {
  if ("christian" %in% colnames(pred_obj)) safe_prob(pred_obj[, "christian"]) else safe_prob(pred_obj[, ncol(pred_obj)])
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}

predict_nb_score <- function(fit, newx_dfm) {
  safe_logit(nb_prob_christian(predict(fit, newdata = newx_dfm, type = "probability")))
}

fit_ridge_fixed <- function(x, y, lambda) {
  glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = 0,
    lambda = lambda,
    standardize = TRUE,
    maxit = 1e6
  )
}

predict_ridge_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_ridge_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

svm_to_matrix <- function(x) {
  if (inherits(x, "dgCMatrix") || inherits(x, "dsCMatrix") || inherits(x, "dgRMatrix")) return(x)
  if (inherits(x, "matrix")) return(x)
  as.matrix(x)
}

fit_svm_fixed <- function(x, y, cost) {
  x_use <- svm_to_matrix(x)
  fit <- LiblineaR::LiblineaR(data = x_use, target = y, type = 7, cost = cost)
  pred_tr <- predict(fit, x_use, decisionValues = TRUE)
  raw <- as.numeric(pred_tr$decisionValues[, 1])
  sgn <- if (mean(raw[y == "christian"]) >= mean(raw[y == "non_christian"])) 1 else -1
  list(model = fit, sign = sgn, cost = cost)
}

predict_svm_score <- function(obj, x) {
  pred <- predict(obj$model, svm_to_matrix(x), decisionValues = TRUE)
  obj$sign * as.numeric(pred$decisionValues[, 1])
}

xgb_grid <- function() {
  tidyr::crossing(
    eta = c(0.01, 0.03, 0.05),
    max_depth = c(3L, 4L, 6L),
    min_child_weight = c(1, 3, 5),
    subsample = c(0.7, 0.85),
    colsample_bytree = c(0.6, 0.8),
    gamma = c(0, 0.1, 0.5)
  ) %>%
    mutate(candidate_id = row_number()) %>%
    select(candidate_id, everything())
}

best_iter_xgb <- function(fit, fallback) {
  bi <- fit$best_iteration
  if (is.null(bi) || length(bi) == 0 || !is.finite(bi)) as.integer(fallback) else as.integer(bi)
}

make_xgb_params <- function(row, scale_pos_weight = 1) {
  list(
    booster = "gbtree",
    objective = "binary:logistic",
    eval_metric = "logloss",
    tree_method = "hist",
    eta = as.numeric(row$eta),
    max_depth = as.integer(row$max_depth),
    min_child_weight = as.numeric(row$min_child_weight),
    subsample = as.numeric(row$subsample),
    colsample_bytree = as.numeric(row$colsample_bytree),
    gamma = as.numeric(row$gamma),
    scale_pos_weight = scale_pos_weight
  )
}

fit_xgb_fixed <- function(x, y, params, nrounds, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  xgboost::xgb.train(
    params = params,
    data = xgboost::xgb.DMatrix(x, label = as.integer(y == "christian")),
    nrounds = as.integer(nrounds),
    verbose = 0
  )
}

predict_xgb_score <- function(fit, x) {
  as.numeric(predict(fit, xgboost::xgb.DMatrix(x), outputmargin = TRUE))
}

lgb_grid <- function() {
  tidyr::crossing(
    learning_rate = c(0.01, 0.03, 0.05),
    num_leaves = c(15L, 31L, 63L),
    min_data_in_leaf = c(5L, 10L, 20L),
    feature_fraction = c(0.6, 0.8),
    bagging_fraction = c(0.7, 0.85)
  ) %>%
    mutate(candidate_id = row_number()) %>%
    select(candidate_id, everything())
}

make_lgb_params <- function(row, seed = 123) {
  list(
    objective = "binary",
    metric = "binary_logloss",
    boosting = "gbdt",
    learning_rate = as.numeric(row$learning_rate),
    num_leaves = as.integer(row$num_leaves),
    min_data_in_leaf = as.integer(row$min_data_in_leaf),
    feature_fraction = as.numeric(row$feature_fraction),
    bagging_fraction = as.numeric(row$bagging_fraction),
    bagging_freq = 1L,
    is_unbalance = TRUE,
    first_metric_only = TRUE,
    seed = seed,
    feature_fraction_seed = seed,
    bagging_seed = seed,
    verbosity = -1
  )
}

make_lgb_dataset_params <- function() {
  list(feature_pre_filter = FALSE)
}

predict_lgb_prob <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "response"))),
    error = function(e) tryCatch(
      do.call(predict, args),
      error = function(e2) rep(0.5, nrow(x))
    )
  )
  safe_prob(as.numeric(out))
}

predict_lgb_score <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "raw"))),
    error = function(e) tryCatch(
      do.call(predict, c(args, list(rawscore = TRUE))),
      error = function(e2) {
        p <- predict_lgb_prob(model, x, num_iteration)
        qlogis(safe_prob(p))
      }
    )
  )
  as.numeric(out)
}

fit_lgb_fixed <- function(x, y, params, nrounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x,
    label = as.integer(y == "christian"),
    params = make_lgb_dataset_params(),
    free_raw_data = TRUE
  )
  lightgbm::lgb.train(
    params = params,
    data = dtrain,
    nrounds = as.integer(nrounds),
    verbose = -1
  )
}

fit_lgb_early_stop <- function(x_tr, y_tr, x_va, y_va, params, max_rounds, early_rounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x_tr,
    label = as.integer(y_tr == "christian"),
    params = make_lgb_dataset_params(),
    free_raw_data = FALSE
  )
  dval <- lightgbm::lgb.Dataset(
    data = x_va,
    label = as.integer(y_va == "christian"),
    params = make_lgb_dataset_params(),
    reference = dtrain,
    free_raw_data = FALSE
  )
  fit <- lightgbm::lgb.train(
    params = params,
    data = dtrain,
    nrounds = as.integer(max_rounds),
    valids = list(valid = dval),
    early_stopping_rounds = as.integer(early_rounds),
    verbose = -1
  )
  best_iter <- fit$best_iter
  if (is.null(best_iter) || !is.finite(best_iter) || best_iter < 1) best_iter <- as.integer(max_rounds)
  list(model = fit, best_iter = as.integer(best_iter))
}

cv_nb_scores_ckpt <- function(x_dfm, y, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("nb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_nb_model(x_dfm[tr, ], y[tr])
    state$scores[va] <- predict_nb_score(fit, x_dfm[va, ])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

ridge_lambda_grid <- function(x, y, nlambda = 60, lambda.min.ratio = 1e-3) {
  base_fit <- glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = 0,
    standardize = TRUE,
    nlambda = nlambda,
    lambda.min.ratio = lambda.min.ratio,
    maxit = 1e6
  )
  as.numeric(base_fit$lambda)
}

tune_ridge_lambda_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir, nlambda = 60, lambda.min.ratio = 1e-3) {
  ensure_dir(ckpt_dir)
  folds <- make_folds(y, inner_k, seed)
  lambda_grid <- ridge_lambda_grid(x, y, nlambda = nlambda, lambda.min.ratio = lambda.min.ratio)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != length(lambda_grid)) {
    summary_state <- list(done = rep(FALSE, length(lambda_grid)))
  }

  for (g in seq_along(lambda_grid)) {
    if (isTRUE(summary_state$done[g])) next
    lam <- lambda_grid[g]
    cand_path <- file.path(ckpt_dir, sprintf("lambda_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$fold_losses) != length(folds)) {
      cand_state <- list(fold_losses = rep(NA_real_, length(folds)), done = rep(FALSE, length(folds)))
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("ridge_tuning_fold", paste(prefix, "lambda", g, "of", length(lambda_grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      fit <- fit_ridge_fixed(x[tr, , drop = FALSE], y[tr], lambda = lam)
      p_va <- predict_ridge_prob(fit, x[va, , drop = FALSE])
      cand_state$fold_losses[j] <- log_loss_safe(y[va], p_va)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", length(lambda_grid))
  for (g in seq_along(lambda_grid)) {
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("lambda_%03d.rds", g)))
    cvm <- mean(cand_state$fold_losses, na.rm = TRUE)
    cvsd <- stats::sd(cand_state$fold_losses, na.rm = TRUE) / sqrt(sum(is.finite(cand_state$fold_losses)))
    tuning_rows[[g]] <- tibble(lambda = lambda_grid[g], cvm = cvm, cvsd = cvsd)
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(desc(lambda))
  idx_min <- which.min(tuning$cvm)
  lambda_min <- tuning$lambda[idx_min]
  cutoff <- tuning$cvm[idx_min] + tuning$cvsd[idx_min]
  lambda_1se <- max(tuning$lambda[tuning$cvm <= cutoff], na.rm = TRUE)
  list(
    lambda = as.numeric(lambda_1se),
    cvfit = list(lambda.min = as.numeric(lambda_min)),
    tuning = tuning
  )
}

cv_ridge_scores_ckpt <- function(x, y, lambda, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("ridge_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_ridge_fixed(x[tr, , drop = FALSE], y[tr], lambda = lambda)
    state$scores[va] <- predict_ridge_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

cv_svm_scores_ckpt <- function(x, y, cost, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("svm_calibration_fold", paste(prefix, "cost", cost, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    state$scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_svm_cost_ckpt <- function(x, y, cost_grid = c(0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 8, 16), inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  rows_path <- file.path(ckpt_dir, "summary.rds")
  state <- load_rds_or_null(rows_path)
  if (is.null(state) || length(state$done) != length(cost_grid)) {
    state <- list(done = rep(FALSE, length(cost_grid)), rows = vector("list", length(cost_grid)))
  }
  for (i in seq_along(cost_grid)) {
    if (isTRUE(state$done[i])) next
    scores_i <- cv_svm_scores_ckpt(
      x = x,
      y = y,
      cost = cost_grid[i],
      k = inner_k,
      seed = seed + i,
      prefix = prefix,
      ckpt_dir = file.path(ckpt_dir, sprintf("cost_%03d", i))
    )
    state$rows[[i]] <- tibble(cost = cost_grid[i], auc_margin = auc_margin_safe(y, scores_i))
    state$done[i] <- TRUE
    save_rds_atomic(state, rows_path)
  }
  tuning <- bind_rows(state$rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}

tune_xgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- xgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  n_pos <- sum(y == "christian")
  n_neg <- sum(y == "non_christian")
  spw <- n_neg / max(n_pos, 1)

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("xgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "christian")
      n_neg_tr <- sum(y[tr] == "non_christian")
      spw_tr <- n_neg_tr / max(n_pos_tr, 1)
      dtr <- xgboost::xgb.DMatrix(x[tr, , drop = FALSE], label = as.integer(y[tr] == "christian"))
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE], label = as.integer(y[va] == "christian"))
      fit <- xgboost::xgb.train(
        params = make_xgb_params(pars_row, scale_pos_weight = spw_tr),
        data = dtr,
        nrounds = xgb_max_rounds,
        evals = list(valid = dva),
        early_stopping_rounds = early_stop_rounds,
        verbose = 0
      )
      cand_state$oof_prob[va] <- safe_prob(as.numeric(predict(fit, dva)))
      cand_state$best_iters[j] <- best_iter_xgb(fit, xgb_max_rounds)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_xgb_params(best, scale_pos_weight = spw), tuning = tuning)
}

cv_xgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("xgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "christian")
    n_neg_tr <- sum(y[tr] == "non_christian")
    fold_params <- modifyList(params, list(scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr], params = fold_params, nrounds = nrounds, seed = seed + j)
    state$scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_lgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- lgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) || length(cand_state$done) != length(folds)) {
      cand_state <- list(
        oof_prob = rep(NA_real_, length(y)),
        best_iters = rep(NA_integer_, length(folds)),
        done = rep(FALSE, length(folds))
      )
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("lgb_tuning_fold", paste(prefix, "candidate", g, "of", nrow(grid), "fold", j, "of", length(folds)))
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      lgb_params_fold <- make_lgb_params(pars_row, seed = seed + 1000L * g + j)
      result <- fit_lgb_early_stop(
        x_tr = x[tr, , drop = FALSE],
        y_tr = y[tr],
        x_va = x[va, , drop = FALSE],
        y_va = y[va],
        params = lgb_params_fold,
        max_rounds = lgb_max_rounds,
        early_rounds = early_stop_rounds
      )
      cand_state$oof_prob[va] <- predict_lgb_prob(result$model, x[va, , drop = FALSE], num_iteration = result$best_iter)
      cand_state$best_iters[j] <- result$best_iter
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss = log_loss_safe(y, cand_state$oof_prob),
        auc = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_lgb_params(best, seed = seed + 9999L), tuning = tuning)
}

cv_lgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) || length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)), done = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("lgb_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE],
      y = y[tr],
      params = modifyList(
        params,
        list(
          seed = seed + j,
          feature_fraction_seed = seed + j,
          bagging_seed = seed + j
        )
      ),
      nrounds = nrounds
    )
    state$scores[va] <- predict_lgb_score(fit, x[va, , drop = FALSE], num_iteration = nrounds)
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

run_lgb_preflight <- function(text_vec, y, stop_list, seed = 999L) {
  folds <- make_folds(y, 2L, seed)
  va <- folds[[1]]
  tr <- setdiff(seq_along(y), va)
  vec <- fit_vectorizer(text_vec[tr], stop_list)
  x_tr <- transform_tfidf_sparse(text_vec[tr], vec, stop_list)
  x_va <- transform_tfidf_sparse(text_vec[va], vec, stop_list)
  pars_row <- lgb_grid()[1, , drop = FALSE]
  pars <- make_lgb_params(pars_row, seed = seed)
  result <- fit_lgb_early_stop(
    x_tr = x_tr,
    y_tr = y[tr],
    x_va = x_va,
    y_va = y[va],
    params = pars,
    max_rounds = 100L,
    early_rounds = 20L
  )
  if (!is.finite(result$best_iter)) stop("LightGBM preflight failed to select a valid num_iteration.")
  invisible(result)
}

save_model_object <- function(model, path, engine) {
  ensure_dir(dirname(path))
  if (engine == "xgboost") {
    xgboost::xgb.save(model, path)
  } else if (engine == "lightgbm") {
    lightgbm::lgb.save(model, path)
  } else {
    save_rds_atomic(model, path)
  }
  invisible(path)
}

init_run_state <- function(n_lab) {
  list(
    preflight_done = FALSE,
    outer_fold_id = integer(n_lab),

    nb_oof = rep(NA_real_, n_lab),
    ridge_oof = rep(NA_real_, n_lab),
    svm_oof = rep(NA_real_, n_lab),
    xgb_oof = rep(NA_real_, n_lab),
    lgb_oof = rep(NA_real_, n_lab),

    nb_score_oof = rep(NA_real_, n_lab),
    ridge_score_oof = rep(NA_real_, n_lab),
    svm_score_oof = rep(NA_real_, n_lab),
    xgb_score_oof = rep(NA_real_, n_lab),
    lgb_score_oof = rep(NA_real_, n_lab),

    nb_pred_youden_oof = rep(NA_character_, n_lab),
    ridge_pred_youden_oof = rep(NA_character_, n_lab),
    svm_pred_youden_oof = rep(NA_character_, n_lab),
    xgb_pred_youden_oof = rep(NA_character_, n_lab),
    lgb_pred_youden_oof = rep(NA_character_, n_lab),

    nb_pred_f1_oof = rep(NA_character_, n_lab),
    ridge_pred_f1_oof = rep(NA_character_, n_lab),
    svm_pred_f1_oof = rep(NA_character_, n_lab),
    xgb_pred_f1_oof = rep(NA_character_, n_lab),
    lgb_pred_f1_oof = rep(NA_character_, n_lab),

    outer_threshold_df = tibble(),
    fold_metric_df = tibble(),
    selected_outer_df = tibble(),
    model_done = list(),
    full_done = list()
  )
}

mark_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  state[[bucket]][[key]] <- TRUE
  state
}

is_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  isTRUE(state[[bucket]][[key]])
}

state_path <- file.path(checkpoint_root, "run_state.rds")

log_progress("load_data", "reading corpus and labels")

df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

dup_corpus <- df %>% count(id, name = "n") %>% filter(n > 1)
if (nrow(dup_corpus) > 0) {
  stop(paste0("Duplicate ids in corpus_600ad.csv: ", paste(head(dup_corpus$id, 20), collapse = ", ")))
}

labels_raw <- readr::read_csv(file.path(data_path, "corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  select(id, label)

label_conflicts <- labels_raw %>%
  distinct(id, label) %>%
  count(id, name = "n_labels") %>%
  filter(n_labels > 1)

if (nrow(label_conflicts) > 0) {
  stop(paste0("Conflicting labels in corpus_600ad_labeled.csv: ", paste(head(label_conflicts$id, 20), collapse = ", ")))
}

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
check_binary_y(df_labeled$manual_label, "labeled data")

custom_additions <- c("m", "c", "l", "d", "gt", "lt", "br", "href", "fridericus", "heinricus", "conradus", "meus", "tuus")
stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")
latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]

lab <- df_labeled %>% arrange(row_id)
out <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos <- match(lab$row_id, out$row_id)

outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)
n_lab <- nrow(lab)

state <- load_rds_or_null(state_path)
if (is.null(state) || length(state$nb_oof) != n_lab) {
  state <- init_run_state(n_lab)
  save_rds_atomic(state, state_path)
}

if (!isTRUE(state$preflight_done)) {
  log_progress("lgb_preflight_start", "testing LightGBM before the full run")
  invisible(run_lgb_preflight(lab$text, lab$manual_label, latin_stops, seed = 999L))
  state$preflight_done <- TRUE
  save_rds_atomic(state, state_path)
  log_progress("lgb_preflight_done", "LightGBM preflight passed")
}

for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start", paste("fold", i, "of", length(outer_folds)))
  va <- outer_folds[[i]]
  tr <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]

  outer_dir <- file.path(checkpoint_root, sprintf("outer_fold_%02d", i))
  ensure_dir(outer_dir)

  vec_fold <- fit_vectorizer(lab$text[tr], latin_stops)
  x_tr_dfm <- transform_counts_dfm(lab$text[tr], vec_fold, latin_stops)
  x_va_dfm <- transform_counts_dfm(lab$text[va], vec_fold, latin_stops)
  x_tr <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
  x_va <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)

  state$outer_fold_id[va] <- i
  save_rds_atomic(state, state_path)

  key <- paste0("outer_", i, "_lightgbm")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_lgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "lightgbm")
    lgb_tuned <- tune_lgb_model_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 5000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    lgb_train_scores <- cv_lgb_scores_ckpt(
      x_tr, y_tr,
      params = lgb_tuned$params,
      nrounds = as.integer(lgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner,
      seed = 5100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    lgb_platt <- fit_platt(lgb_train_scores, y_tr)
    lgb_fit <- fit_lgb_fixed(
      x = x_tr,
      y = y_tr,
      params = lgb_tuned$params,
      nrounds = as.integer(lgb_tuned$best$best_nrounds[1])
    )
    lgb_va_scores <- predict_lgb_score(lgb_fit, x_va, num_iteration = as.integer(lgb_tuned$best$best_nrounds[1]))
    lgb_train_prob <- apply_platt(lgb_platt, lgb_train_scores)
    lgb_va_prob <- apply_platt(lgb_platt, lgb_va_scores)
    lgb_th_y <- choose_threshold(y_tr, lgb_train_prob, "youden")
    lgb_th_f <- choose_threshold(y_tr, lgb_train_prob, "f1")

    state$lgb_oof[va] <- lgb_va_prob
    state$lgb_score_oof[va] <- lgb_va_scores
    state$lgb_pred_youden_oof[va] <- ifelse(lgb_va_prob >= lgb_th_y, "christian", "non_christian")
    state$lgb_pred_f1_oof[va] <- ifelse(lgb_va_prob >= lgb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "lightgbm", threshold_youden = lgb_th_y, threshold_f1 = lgb_th_f)
    metric_row <- summarize_model_fold(y_va, lgb_va_prob, state$lgb_pred_youden_oof[va], state$lgb_pred_f1_oof[va], "lightgbm", i)
    selected_row <- lgb_tuned$best %>%
      mutate(
        outer_fold = i,
        model = "lightgbm",
        threshold_youden = lgb_th_y,
        threshold_f1 = lgb_th_f
      )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(best = lgb_tuned$best, tuning = lgb_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_naive_bayes")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_nb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "naive_bayes")
    nb_train_scores <- cv_nb_scores_ckpt(
      x_tr_dfm, y_tr,
      k = k_folds_inner,
      seed = 1000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    nb_platt <- fit_platt(nb_train_scores, y_tr)
    nb_fit <- fit_nb_model(x_tr_dfm, y_tr)
    nb_va_scores <- predict_nb_score(nb_fit, x_va_dfm)
    nb_train_prob <- apply_platt(nb_platt, nb_train_scores)
    nb_va_prob <- apply_platt(nb_platt, nb_va_scores)
    nb_th_y <- choose_threshold(y_tr, nb_train_prob, "youden")
    nb_th_f <- choose_threshold(y_tr, nb_train_prob, "f1")

    state$nb_oof[va] <- nb_va_prob
    state$nb_score_oof[va] <- nb_va_scores
    state$nb_pred_youden_oof[va] <- ifelse(nb_va_prob >= nb_th_y, "christian", "non_christian")
    state$nb_pred_f1_oof[va] <- ifelse(nb_va_prob >= nb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "naive_bayes", threshold_youden = nb_th_y, threshold_f1 = nb_th_f)
    metric_row <- summarize_model_fold(y_va, nb_va_prob, state$nb_pred_youden_oof[va], state$nb_pred_f1_oof[va], "naive_bayes", i)
    selected_row <- tibble(outer_fold = i, model = "naive_bayes", threshold_youden = nb_th_y, threshold_f1 = nb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_ridge")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_ridge_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "ridge")
    ridge_tuned <- tune_ridge_lambda_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 2000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    ridge_train_scores <- cv_ridge_scores_ckpt(
      x_tr, y_tr,
      lambda = ridge_tuned$lambda,
      k = k_folds_inner,
      seed = 2100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    ridge_platt <- fit_platt(ridge_train_scores, y_tr)
    ridge_fit <- fit_ridge_fixed(x_tr, y_tr, lambda = ridge_tuned$lambda)
    ridge_va_scores <- predict_ridge_score(ridge_fit, x_va)
    ridge_train_prob <- apply_platt(ridge_platt, ridge_train_scores)
    ridge_va_prob <- apply_platt(ridge_platt, ridge_va_scores)
    ridge_th_y <- choose_threshold(y_tr, ridge_train_prob, "youden")
    ridge_th_f <- choose_threshold(y_tr, ridge_train_prob, "f1")

    state$ridge_oof[va] <- ridge_va_prob
    state$ridge_score_oof[va] <- ridge_va_scores
    state$ridge_pred_youden_oof[va] <- ifelse(ridge_va_prob >= ridge_th_y, "christian", "non_christian")
    state$ridge_pred_f1_oof[va] <- ifelse(ridge_va_prob >= ridge_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "ridge", threshold_youden = ridge_th_y, threshold_f1 = ridge_th_f)
    metric_row <- summarize_model_fold(y_va, ridge_va_prob, state$ridge_pred_youden_oof[va], state$ridge_pred_f1_oof[va], "ridge", i)
    selected_row <- tibble(
      outer_fold = i,
      model = "ridge",
      lambda_1se = ridge_tuned$lambda,
      lambda_min = ridge_tuned$cvfit$lambda.min,
      threshold_youden = ridge_th_y,
      threshold_f1 = ridge_th_f
    )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(tuning = ridge_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_svm_platt")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_svm_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "svm_platt")
    svm_tuned <- tune_svm_cost_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 3000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    svm_train_scores <- cv_svm_scores_ckpt(
      x_tr, y_tr,
      cost = svm_tuned$cost,
      k = k_folds_inner,
      seed = 3100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    svm_platt <- fit_platt(svm_train_scores, y_tr)
    svm_fit <- fit_svm_fixed(x_tr, y_tr, cost = svm_tuned$cost)
    svm_va_scores <- predict_svm_score(svm_fit, x_va)
    svm_train_prob <- apply_platt(svm_platt, svm_train_scores)
    svm_va_prob <- apply_platt(svm_platt, svm_va_scores)
    svm_th_y <- choose_threshold(y_tr, svm_train_prob, "youden")
    svm_th_f <- choose_threshold(y_tr, svm_train_prob, "f1")

    state$svm_oof[va] <- svm_va_prob
    state$svm_score_oof[va] <- svm_va_scores
    state$svm_pred_youden_oof[va] <- ifelse(svm_va_prob >= svm_th_y, "christian", "non_christian")
    state$svm_pred_f1_oof[va] <- ifelse(svm_va_prob >= svm_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "svm_platt", threshold_youden = svm_th_y, threshold_f1 = svm_th_f)
    metric_row <- summarize_model_fold(y_va, svm_va_prob, state$svm_pred_youden_oof[va], state$svm_pred_f1_oof[va], "svm_platt", i)
    selected_row <- tibble(
      outer_fold = i,
      model = "svm_platt",
      cost = svm_tuned$cost,
      threshold_youden = svm_th_y,
      threshold_f1 = svm_th_f
    )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(tuning = svm_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  key <- paste0("outer_", i, "_xgboost")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_xgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "xgboost")
    xgb_tuned <- tune_xgb_model_ckpt(
      x_tr, y_tr,
      inner_k = k_folds_inner,
      seed = 4000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    xgb_train_scores <- cv_xgb_scores_ckpt(
      x_tr, y_tr,
      params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner,
      seed = 4100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    xgb_platt <- fit_platt(xgb_train_scores, y_tr)
    xgb_fit <- fit_xgb_fixed(
      x_tr, y_tr,
      params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      seed = 4200 + i
    )
    xgb_va_scores <- predict_xgb_score(xgb_fit, x_va)
    xgb_train_prob <- apply_platt(xgb_platt, xgb_train_scores)
    xgb_va_prob <- apply_platt(xgb_platt, xgb_va_scores)
    xgb_th_y <- choose_threshold(y_tr, xgb_train_prob, "youden")
    xgb_th_f <- choose_threshold(y_tr, xgb_train_prob, "f1")

    state$xgb_oof[va] <- xgb_va_prob
    state$xgb_score_oof[va] <- xgb_va_scores
    state$xgb_pred_youden_oof[va] <- ifelse(xgb_va_prob >= xgb_th_y, "christian", "non_christian")
    state$xgb_pred_f1_oof[va] <- ifelse(xgb_va_prob >= xgb_th_f, "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "xgboost", threshold_youden = xgb_th_y, threshold_f1 = xgb_th_f)
    metric_row <- summarize_model_fold(y_va, xgb_va_prob, state$xgb_pred_youden_oof[va], state$xgb_pred_f1_oof[va], "xgboost", i)
    selected_row <- xgb_tuned$best %>%
      mutate(
        outer_fold = i,
        model = "xgboost",
        threshold_youden = xgb_th_y,
        threshold_f1 = xgb_th_f
      )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df, threshold_row, c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row, c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df, selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(best = xgb_tuned$best, tuning = xgb_tuned$tuning, thresholds = threshold_row, metrics = metric_row), file.path(model_dir, "result.rds"))
  }

  log_progress("outer_fold_done", paste("fold", i, "of", length(outer_folds)))
}

oof_metrics <- bind_rows(
  summarize_model_oof(lab$manual_label, state$nb_oof, state$nb_pred_youden_oof, state$nb_pred_f1_oof, "naive_bayes"),
  summarize_model_oof(lab$manual_label, state$ridge_oof, state$ridge_pred_youden_oof, state$ridge_pred_f1_oof, "ridge"),
  summarize_model_oof(lab$manual_label, state$svm_oof, state$svm_pred_youden_oof, state$svm_pred_f1_oof, "svm_platt"),
  summarize_model_oof(lab$manual_label, state$xgb_oof, state$xgb_pred_youden_oof, state$xgb_pred_f1_oof, "xgboost"),
  summarize_model_oof(lab$manual_label, state$lgb_oof, state$lgb_pred_youden_oof, state$lgb_pred_f1_oof, "lightgbm")
)

oof_predictions <- lab %>%
  transmute(
    row_id,
    id,
    manual_label,
    outer_fold = state$outer_fold_id,
    nb_oof = state$nb_oof,
    nb_score_oof = state$nb_score_oof,
    nb_pred_youden_oof = state$nb_pred_youden_oof,
    nb_pred_f1_oof = state$nb_pred_f1_oof,
    ridge_oof = state$ridge_oof,
    ridge_score_oof = state$ridge_score_oof,
    ridge_pred_youden_oof = state$ridge_pred_youden_oof,
    ridge_pred_f1_oof = state$ridge_pred_f1_oof,
    svm_oof = state$svm_oof,
    svm_score_oof = state$svm_score_oof,
    svm_pred_youden_oof = state$svm_pred_youden_oof,
    svm_pred_f1_oof = state$svm_pred_f1_oof,
    xgb_oof = state$xgb_oof,
    xgb_score_oof = state$xgb_score_oof,
    xgb_pred_youden_oof = state$xgb_pred_youden_oof,
    xgb_pred_f1_oof = state$xgb_pred_f1_oof,
    lgb_oof = state$lgb_oof,
    lgb_score_oof = state$lgb_score_oof,
    lgb_pred_youden_oof = state$lgb_pred_youden_oof,
    lgb_pred_f1_oof = state$lgb_pred_f1_oof
  )

write_csv_atomic(arrange(state$outer_threshold_df, outer_fold, model), file.path(output_dir, "outer_thresholds.csv"))
write_csv_atomic(arrange(state$fold_metric_df, outer_fold, model), file.path(output_dir, "outer_fold_metrics.csv"))
write_csv_atomic(arrange(state$selected_outer_df, outer_fold, model), file.path(output_dir, "selected_outer_hyperparameters.csv"))
write_csv_atomic(oof_metrics, file.path(output_dir, "oof_model_metrics.csv"))
write_csv_atomic(oof_predictions, file.path(output_dir, "oof_predictions_labeled.csv"))

full_dir <- file.path(checkpoint_root, "full_fit")
ensure_dir(full_dir)

vec_full_path <- file.path(full_dir, "vectorizer_full.rds")
vec_full <- load_rds_or_null(vec_full_path)
if (is.null(vec_full)) {
  vec_full <- fit_vectorizer(lab$text, latin_stops)
  save_rds_atomic(vec_full, vec_full_path)
}

x_lab_dfm_full <- transform_counts_dfm(lab$text, vec_full, latin_stops)
x_all_dfm_full <- transform_counts_dfm(out$text, vec_full, latin_stops)
x_lab_full <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all_full <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

full_predictions <- tibble(row_id = out$row_id, id = out$id, manual_label = out$manual_label)

key <- "full_naive_bayes"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_nb_full", "full-data fit")
  model_dir <- file.path(full_dir, "naive_bayes")
  nb_cv_scores <- cv_nb_scores_ckpt(
    x_lab_dfm_full, lab$manual_label,
    k = k_folds_inner,
    seed = 8001,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  nb_platt <- fit_platt(nb_cv_scores, lab$manual_label)
  nb_fit <- fit_nb_model(x_lab_dfm_full, lab$manual_label)
  save_model_object(nb_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(nb_platt, file.path(model_dir, "platt.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

nb_fit <- readRDS(file.path(full_dir, "naive_bayes", "model.rds"))
nb_platt <- load_rds_or_null(file.path(full_dir, "naive_bayes", "platt.rds"))
nb_all_scores <- predict_nb_score(nb_fit, x_all_dfm_full)
full_predictions <- full_predictions %>%
  mutate(
    nb_score = nb_all_scores,
    nb_prob = apply_platt(nb_platt, nb_all_scores)
  )

key <- "full_ridge"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_ridge_full", "full-data fit")
  model_dir <- file.path(full_dir, "ridge")
  ridge_tuned <- tune_ridge_lambda_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8002,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  ridge_cv_scores <- cv_ridge_scores_ckpt(
    x_lab_full, lab$manual_label,
    lambda = ridge_tuned$lambda,
    k = k_folds_inner,
    seed = 8102,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  ridge_platt <- fit_platt(ridge_cv_scores, lab$manual_label)
  ridge_fit <- fit_ridge_fixed(x_lab_full, lab$manual_label, lambda = ridge_tuned$lambda)
  save_model_object(ridge_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(ridge_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(ridge_tuned, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

ridge_fit <- readRDS(file.path(full_dir, "ridge", "model.rds"))
ridge_platt <- load_rds_or_null(file.path(full_dir, "ridge", "platt.rds"))
ridge_all_scores <- predict_ridge_score(ridge_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    ridge_score = ridge_all_scores,
    ridge_prob = apply_platt(ridge_platt, ridge_all_scores)
  )

key <- "full_svm_platt"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_svm_full", "full-data fit")
  model_dir <- file.path(full_dir, "svm_platt")
  svm_tuned <- tune_svm_cost_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8003,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  svm_cv_scores <- cv_svm_scores_ckpt(
    x_lab_full, lab$manual_label,
    cost = svm_tuned$cost,
    k = k_folds_inner,
    seed = 8103,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  svm_platt <- fit_platt(svm_cv_scores, lab$manual_label)
  svm_fit <- fit_svm_fixed(x_lab_full, lab$manual_label, cost = svm_tuned$cost)
  save_model_object(svm_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(svm_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(svm_tuned, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

svm_fit <- readRDS(file.path(full_dir, "svm_platt", "model.rds"))
svm_platt <- load_rds_or_null(file.path(full_dir, "svm_platt", "platt.rds"))
svm_all_scores <- predict_svm_score(svm_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    svm_score = svm_all_scores,
    svm_prob = apply_platt(svm_platt, svm_all_scores)
  )

key <- "full_xgboost"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_xgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "xgboost")
  xgb_tuned <- tune_xgb_model_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8004,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  xgb_cv_scores <- cv_xgb_scores_ckpt(
    x_lab_full, lab$manual_label,
    params = xgb_tuned$params,
    nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
    k = k_folds_inner,
    seed = 8104,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  xgb_platt <- fit_platt(xgb_cv_scores, lab$manual_label)
  xgb_fit <- fit_xgb_fixed(
    x_lab_full, lab$manual_label,
    params = xgb_tuned$params,
    nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
    seed = 8204
  )
  save_model_object(xgb_fit, file.path(model_dir, "model.xgb"), "xgboost")
  save_rds_atomic(xgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = xgb_tuned$best, tuning = xgb_tuned$tuning), file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

xgb_fit <- xgboost::xgb.load(file.path(full_dir, "xgboost", "model.xgb"))
xgb_platt <- load_rds_or_null(file.path(full_dir, "xgboost", "platt.rds"))
xgb_all_scores <- predict_xgb_score(xgb_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(
    xgb_score = xgb_all_scores,
    xgb_prob = apply_platt(xgb_platt, xgb_all_scores)
  )

key <- "full_lightgbm"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_lgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "lightgbm")
  lgb_tuned <- tune_lgb_model_ckpt(
    x_lab_full, lab$manual_label,
    inner_k = k_folds_inner,
    seed = 8005,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "tuning")
  )
  lgb_cv_scores <- cv_lgb_scores_ckpt(
    x_lab_full, lab$manual_label,
    params = lgb_tuned$params,
    nrounds = as.integer(lgb_tuned$best$best_nrounds[1]),
    k = k_folds_inner,
    seed = 8105,
    prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration")
  )
  lgb_platt <- fit_platt(lgb_cv_scores, lab$manual_label)
  lgb_fit <- fit_lgb_fixed(
    x = x_lab_full,
    y = lab$manual_label,
    params = lgb_tuned$params,
    nrounds = as.integer(lgb_tuned$best$best_nrounds[1])
  )
  save_model_object(lgb_fit, file.path(model_dir, "model.txt"), "lightgbm")
  save_rds_atomic(lgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = lgb_tuned$best, tuning = lgb_tuned$tuning), file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

lgb_fit <- lightgbm::lgb.load(file.path(full_dir, "lightgbm", "model.txt"))
lgb_platt <- load_rds_or_null(file.path(full_dir, "lightgbm", "platt.rds"))
lgb_tuning_result <- readRDS(file.path(full_dir, "lightgbm", "tuning_result.rds"))
lgb_all_scores <- predict_lgb_score(lgb_fit, x_all_full, num_iteration = as.integer(lgb_tuning_result$best$best_nrounds[1]))
full_predictions <- full_predictions %>%
  mutate(
    lgb_score = lgb_all_scores,
    lgb_prob = apply_platt(lgb_platt, lgb_all_scores)
  )

write_csv_atomic(full_predictions, file.path(output_dir, "full_corpus_model_scores.csv"))

log_progress("done", "checkpointed run completed")

```



```{r}

model_metrics_oof <- bind_rows(
  summarize_model_oof(lab$manual_label, nb_oof, nb_pred_youden_oof, nb_pred_f1_oof, "naive_bayes"),
  summarize_model_oof(lab$manual_label, ridge_oof, ridge_pred_youden_oof, ridge_pred_f1_oof, "ridge"),
  summarize_model_oof(lab$manual_label, svm_oof, svm_pred_youden_oof, svm_pred_f1_oof, "svm_platt"),
  summarize_model_oof(lab$manual_label, xgb_oof, xgb_pred_youden_oof, xgb_pred_f1_oof, "xgboost"),
  summarize_model_oof(lab$manual_label, lgb_oof, lgb_pred_youden_oof, lgb_pred_f1_oof, "lightgbm")
)

model_metrics_oof_by_fold <- bind_rows(fold_metric_rows)
outer_thresholds <- bind_rows(outer_threshold_rows)
selected_outer <- bind_rows(selected_outer)

log_progress("fit_full_models", "training full-data deployment models")

vec_full <- fit_vectorizer(lab$text, latin_stops)
x_train_full_dfm <- transform_counts_dfm(lab$text, vec_full, latin_stops)
x_all_dfm <- transform_counts_dfm(out$text, vec_full, latin_stops)
x_train_full <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

nb_lab_score_cf <- cv_nb_scores_fixed(x_train_full_dfm, lab$manual_label, k = k_folds_outer, seed = 6001, prefix = "full")
nb_platt_full <- fit_platt(nb_lab_score_cf, lab$manual_label)
nb_lab_prob_cf <- apply_platt(nb_platt_full, nb_lab_score_cf)
nb_model_full <- fit_nb_model(x_train_full_dfm, lab$manual_label)
nb_all_score <- predict_nb_score(nb_model_full, x_all_dfm)
nb_all_prob <- apply_platt(nb_platt_full, nb_all_score)
nb_deploy_th_y <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "youden")
nb_deploy_th_f <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "f1")

ridge_tuned_full <- tune_ridge_lambda(x_train_full, lab$manual_label, inner_k = k_folds_inner, seed = 6002)
ridge_lab_score_cf <- cv_ridge_scores_fixed(x_train_full, lab$manual_label, lambda = ridge_tuned_full$lambda, k = k_folds_outer, seed = 6102, prefix = "full")
ridge_platt_full <- fit_platt(ridge_lab_score_cf, lab$manual_label)
ridge_lab_prob_cf <- apply_platt(ridge_platt_full, ridge_lab_score_cf)
ridge_model_full <- fit_ridge_fixed(x_train_full, lab$manual_label, lambda = ridge_tuned_full$lambda)
ridge_all_score <- predict_ridge_score(ridge_model_full, x_all)
ridge_all_prob <- apply_platt(ridge_platt_full, ridge_all_score)
ridge_deploy_th_y <- choose_threshold(lab$manual_label, ridge_lab_prob_cf, "youden")
ridge_deploy_th_f <- choose_threshold(lab$manual_label, ridge_lab_prob_cf, "f1")

svm_tuned_full <- tune_svm_cost(x_train_full, lab$manual_label, inner_k = k_folds_inner, seed = 6003, prefix = "full")
svm_lab_score_cf <- cv_svm_scores_fixed(x_train_full, lab$manual_label, cost = svm_tuned_full$cost, k = k_folds_outer, seed = 6103, prefix = "full")
svm_platt_full <- fit_platt(svm_lab_score_cf, lab$manual_label)
svm_lab_prob_cf <- apply_platt(svm_platt_full, svm_lab_score_cf)
svm_model_full <- fit_svm_fixed(x_train_full, lab$manual_label, cost = svm_tuned_full$cost)
svm_all_score <- predict_svm_score(svm_model_full, x_all)
svm_all_prob <- apply_platt(svm_platt_full, svm_all_score)
svm_deploy_th_y <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "youden")
svm_deploy_th_f <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "f1")

xgb_tuned_full <- tune_xgb_model(x_train_full, lab$manual_label, inner_k = k_folds_inner, seed = 6004, prefix = "full")
xgb_lab_score_cf <- cv_xgb_scores_fixed(
  x_train_full, lab$manual_label,
  params = xgb_tuned_full$params,
  nrounds = as.integer(xgb_tuned_full$best$best_nrounds[1]),
  k = k_folds_outer,
  seed = 6104,
  prefix = "full"
)
xgb_platt_full <- fit_platt(xgb_lab_score_cf, lab$manual_label)
xgb_lab_prob_cf <- apply_platt(xgb_platt_full, xgb_lab_score_cf)
xgb_model_full <- fit_xgb_fixed(
  x_train_full, lab$manual_label,
  params = xgb_tuned_full$params,
  nrounds = as.integer(xgb_tuned_full$best$best_nrounds[1]),
  seed = 6204
)
xgb_all_score <- predict_xgb_score(xgb_model_full, x_all)
xgb_all_prob <- apply_platt(xgb_platt_full, xgb_all_score)
xgb_deploy_th_y <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "youden")
xgb_deploy_th_f <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "f1")

lgb_tuned_full <- tune_lgb_model(x_train_full, lab$manual_label, inner_k = k_folds_inner, seed = 6005, prefix = "full")
lgb_lab_score_cf <- cv_lgb_scores_fixed(
  x_train_full, lab$manual_label,
  params = lgb_tuned_full$params,
  nrounds = as.integer(lgb_tuned_full$best$best_nrounds[1]),
  k = k_folds_outer,
  seed = 6105,
  prefix = "full"
)
lgb_platt_full <- fit_platt(lgb_lab_score_cf, lab$manual_label)
lgb_lab_prob_cf <- apply_platt(lgb_platt_full, lgb_lab_score_cf)
lgb_model_full <- fit_lgb_fixed(
  x_train_full, lab$manual_label,
  params = lgb_tuned_full$params,
  nrounds = as.integer(lgb_tuned_full$best$best_nrounds[1])
)
lgb_all_score <- predict_lgb_score(lgb_model_full, x_all)
lgb_all_prob <- apply_platt(lgb_platt_full, lgb_all_score)
lgb_deploy_th_y <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "youden")
lgb_deploy_th_f <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "f1")

prob_nb <- nb_all_prob
prob_ridge <- ridge_all_prob
prob_svm <- svm_all_prob
prob_xgb <- xgb_all_prob
prob_lgb <- lgb_all_prob

score_nb_logit <- nb_all_score
score_ridge_logit <- ridge_all_score
score_svm_margin <- svm_all_score
score_xgb_margin <- xgb_all_score
score_lgb_margin <- lgb_all_score

prob_nb[lab_pos] <- nb_lab_prob_cf
prob_ridge[lab_pos] <- ridge_lab_prob_cf
prob_svm[lab_pos] <- svm_lab_prob_cf
prob_xgb[lab_pos] <- xgb_lab_prob_cf
prob_lgb[lab_pos] <- lgb_lab_prob_cf

score_nb_logit[lab_pos] <- nb_lab_score_cf
score_ridge_logit[lab_pos] <- ridge_lab_score_cf
score_svm_margin[lab_pos] <- svm_lab_score_cf
score_xgb_margin[lab_pos] <- xgb_lab_score_cf
score_lgb_margin[lab_pos] <- lgb_lab_score_cf

deployment_thresholds <- tibble(
  model = c("naive_bayes", "ridge", "svm_platt", "xgboost", "lightgbm"),
  threshold_youden = c(nb_deploy_th_y, ridge_deploy_th_y, svm_deploy_th_y, xgb_deploy_th_y, lgb_deploy_th_y),
  threshold_f1 = c(nb_deploy_th_f, ridge_deploy_th_f, svm_deploy_th_f, xgb_deploy_th_f, lgb_deploy_th_f),
  threshold_scope = "full_labeled_cross_fitted_calibrated"
)

get_threshold <- function(model_name, kind = c("youden", "f1")) {
  kind <- match.arg(kind)
  x <- deployment_thresholds %>%
    filter(model == model_name) %>%
    pull(if (kind == "youden") threshold_youden else threshold_f1)
  x[[1]]
}

out <- out %>%
  mutate(
    prediction_source = if_else(!is.na(manual_label), "cross_fitted_labeled", "full_fit_unlabeled"),

    prob_nb = safe_prob(prob_nb),
    score_nb_logit = score_nb_logit,
    label_nb_youden = if_else(prob_nb >= get_threshold("naive_bayes", "youden"), "christian", "non_christian"),
    label_nb_f1 = if_else(prob_nb >= get_threshold("naive_bayes", "f1"), "christian", "non_christian"),

    prob_ridge = safe_prob(prob_ridge),
    score_ridge_logit = score_ridge_logit,
    label_ridge_youden = if_else(prob_ridge >= get_threshold("ridge", "youden"), "christian", "non_christian"),
    label_ridge_f1 = if_else(prob_ridge >= get_threshold("ridge", "f1"), "christian", "non_christian"),

    prob_svm = safe_prob(prob_svm),
    score_svm_margin = score_svm_margin,
    score_svm_logit = safe_logit(prob_svm),
    label_svm_youden = if_else(prob_svm >= get_threshold("svm_platt", "youden"), "christian", "non_christian"),
    label_svm_f1 = if_else(prob_svm >= get_threshold("svm_platt", "f1"), "christian", "non_christian"),

    prob_xgb = safe_prob(prob_xgb),
    score_xgb_margin = score_xgb_margin,
    score_xgb_logit = safe_logit(prob_xgb),
    label_xgb_youden = if_else(prob_xgb >= get_threshold("xgboost", "youden"), "christian", "non_christian"),
    label_xgb_f1 = if_else(prob_xgb >= get_threshold("xgboost", "f1"), "christian", "non_christian"),

    prob_lgb = safe_prob(prob_lgb),
    score_lgb_margin = score_lgb_margin,
    score_lgb_logit = safe_logit(prob_lgb),
    label_lgb_youden = if_else(prob_lgb >= get_threshold("lightgbm", "youden"), "christian", "non_christian"),
    label_lgb_f1 = if_else(prob_lgb >= get_threshold("lightgbm", "f1"), "christian", "non_christian")
  )

selected_full <- bind_rows(
  tibble(model = "naive_bayes"),
  tibble(model = "ridge", lambda_1se = ridge_tuned_full$lambda, lambda_min = ridge_tuned_full$cvfit$lambda.min),
  tibble(model = "svm_platt", cost = svm_tuned_full$cost),
  xgb_tuned_full$best %>% mutate(model = "xgboost"),
  lgb_tuned_full$best %>% mutate(model = "lightgbm")
)

tuning_full <- bind_rows(
  tibble(model = "ridge", lambda_1se = ridge_tuned_full$lambda, lambda_min = ridge_tuned_full$cvfit$lambda.min),
  svm_tuned_full$tuning %>% mutate(model = "svm_platt"),
  xgb_tuned_full$tuning %>% mutate(model = "xgboost"),
  lgb_tuned_full$tuning %>% mutate(model = "lightgbm")
)

oof_eval_predictions <- tibble(
  id = lab$id,
  row_id = lab$row_id,
  outer_fold = outer_fold_id,
  manual_label = lab$manual_label,

  prob_nb = nb_oof,
  score_nb_logit = nb_score_oof,
  label_nb_youden = nb_pred_youden_oof,
  label_nb_f1 = nb_pred_f1_oof,

  prob_ridge = ridge_oof,
  score_ridge_logit = ridge_score_oof,
  label_ridge_youden = ridge_pred_youden_oof,
  label_ridge_f1 = ridge_pred_f1_oof,

  prob_svm = svm_oof,
  score_svm_margin = svm_score_oof,
  label_svm_youden = svm_pred_youden_oof,
  label_svm_f1 = svm_pred_f1_oof,

  prob_xgb = xgb_oof,
  score_xgb_margin = xgb_score_oof,
  label_xgb_youden = xgb_pred_youden_oof,
  label_xgb_f1 = xgb_pred_f1_oof,

  prob_lgb = lgb_oof,
  score_lgb_margin = lgb_score_oof,
  label_lgb_youden = lgb_pred_youden_oof,
  label_lgb_f1 = lgb_pred_f1_oof
)

oof_confusion_counts <- bind_rows(
  hard_metrics(lab$manual_label, nb_pred_youden_oof) %>% mutate(model = "naive_bayes", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, ridge_pred_youden_oof) %>% mutate(model = "ridge", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, svm_pred_youden_oof) %>% mutate(model = "svm_platt", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, xgb_pred_youden_oof) %>% mutate(model = "xgboost", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, lgb_pred_youden_oof) %>% mutate(model = "lightgbm", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, nb_pred_f1_oof) %>% mutate(model = "naive_bayes", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, ridge_pred_f1_oof) %>% mutate(model = "ridge", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, svm_pred_f1_oof) %>% mutate(model = "svm_platt", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, xgb_pred_f1_oof) %>% mutate(model = "xgboost", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, lgb_pred_f1_oof) %>% mutate(model = "lightgbm", threshold_kind = "f1")
)

audit_summary <- tibble(
  corpus_rows = nrow(df),
  labeled_rows = nrow(df_labeled),
  labels_raw_rows = nrow(labels_raw),
  duplicate_corpus_ids = nrow(dup_corpus),
  conflicting_label_ids = nrow(label_conflicts),
  repeated_identical_labels_collapsed = nrow(label_duplicates_collapsed),
  feature_weighting = "sublinear_tfidf_l2norm",
  min_docfreq_filter = min_docfreq,
  probability_policy = "all_models_platt_calibrated",
  oof_probability_policy = "outer_validation_predictions_with_training_only_calibration",
  oof_threshold_policy = "nested_outer_fold_train_only",
  deployment_threshold_policy = "full_labeled_cross_fitted_calibrated"
)

readr::write_csv(
  out %>% select(-any_of(c("text", "text_raw"))),
  file.path(results_path, "corpus_600ad_model_outputs_oof.csv")
)

readr::write_csv(
  deployment_thresholds,
  file.path(results_path, "model_thresholds_oof.csv")
)

readr::write_csv(
  deployment_thresholds,
  file.path(results_path, "model_thresholds_deployment.csv")
)

readr::write_csv(
  outer_thresholds,
  file.path(results_path, "model_thresholds_outer.csv")
)

readr::write_csv(
  model_metrics_oof,
  file.path(results_path, "model_metrics_oof.csv")
)

readr::write_csv(
  model_metrics_oof_by_fold,
  file.path(results_path, "model_metrics_oof_by_fold.csv")
)

readr::write_csv(
  oof_eval_predictions,
  file.path(results_path, "oof_eval_predictions.csv")
)

readr::write_csv(
  oof_confusion_counts,
  file.path(results_path, "oof_confusion_counts.csv")
)

readr::write_csv(
  selected_outer,
  file.path(results_path, "model_hyperparameters_outer.csv")
)

readr::write_csv(
  selected_full,
  file.path(results_path, "model_hyperparameters_full.csv")
)

readr::write_csv(
  tuning_full,
  file.path(results_path, "model_tuning_full.csv")
)

readr::write_csv(
  audit_summary,
  file.path(results_path, "data_audit_summary.csv")
)

if (nrow(label_duplicates_collapsed) > 0) {
  readr::write_csv(
    label_duplicates_collapsed,
    file.path(data_path, "label_duplicates_collapsed.csv")
  )
}

log_progress("done", "all outputs written")

df <- out
```


```{r}


suppressPackageStartupMessages({
  library(tidyverse)
  library(quanteda)
  library(quanteda.textmodels)
  library(glmnet)
  library(Matrix)
  library(pROC)
})

set.seed(123)


data_path       <- "../../data/processed_texts/600ad"
results_path    <- "../../results/predictions/non_transformer_lemmatized/600ad"
checkpoint_path <- results_path
checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
enet_root       <- file.path(checkpoint_path, "checkpoints_elasticnet")

dir.create(enet_root, recursive = TRUE, showWarnings = FALSE)


k_folds_outer   <- 5L
k_folds_inner   <- 5L
min_docfreq     <- 3L



`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_", format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) { file.copy(tmp, path, overwrite = TRUE); unlink(tmp) }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

safe_prob  <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb)))
    stop(paste0("Both classes required", if (nzchar(where)) paste0(" in ", where) else "", "."))
  if (any(tb < 2))
    stop(paste0("Each class needs >=2 obs", if (nzchar(where)) paste0(" in ", where) else "", "."))
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for CV.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg)) folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1); tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1); fn <- sum(y01 == 1 & p01 == 0)
  precision   <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall      <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1  <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(accuracy = (tp + tn) / length(y), balanced_accuracy = bal_acc,
         precision = precision, recall = recall, specificity = specificity,
         f1 = f1, mcc = mcc, tp = tp, tn = tn, fp = fp, fn = fn)
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(response = y, predictor = p,
                  levels = c("non_christian", "christian"),
                  direction = "<", quiet = TRUE),
        x = "best", input = "threshold", best.method = "youden",
        ret = "threshold", transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01  <- as.integer(y == "christian")
    cand <- sort(unique(c(0, p, 0.5, 1)))
    f1s  <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(), control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores), type = "response"))
  }
}

log_progress <- function(step, detail = "") {
  message(paste0("[", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "] ", step,
                 if (nzchar(detail)) paste0(" | ", detail) else ""))
}



make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N   <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat    <- as(x_counts, "dgCMatrix")
  x_mat@x  <- 1 + log(x_mat@x)
  idf_vec  <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf  <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}






enet_alpha_grid <- c(0.05, 0.1, 0.2, 0.3, 0.4, 0.5)

fit_enet_fixed <- function(x, y, alpha, lambda) {



  suppressWarnings(
    glmnet::glmnet(
      x = x,
      y = as.integer(y == "christian"),
      family = "binomial",
      alpha = alpha,
      lambda = lambda,
      standardize = TRUE,
      maxit = 1e6
    )
  )
}


enet_has_nonzero <- function(fit) {
  beta <- as.numeric(coef(fit)[-1, 1])
  sum(beta != 0) > 0
}

predict_enet_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_enet_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

enet_lambda_grid <- function(x, y, alpha, nlambda = 60, lambda.min.ratio = 1e-3) {
  base_fit <- suppressWarnings(glmnet::glmnet(
    x = x,
    y = as.integer(y == "christian"),
    family = "binomial",
    alpha = alpha,
    standardize = TRUE,
    nlambda = nlambda,
    lambda.min.ratio = lambda.min.ratio,
    maxit = 1e6
  ))
  lams <- as.numeric(base_fit$lambda)
  if (length(lams) == 0) {
    warning("glmnet returned no lambda path for alpha = ", alpha, "; using fallback grid.")
    lams <- 10^seq(log10(0.001), log10(100), length.out = nlambda)
  }
  lams
}





tune_enet <- function(x, y, inner_k = 5, seed = 123, prefix = "",
                      nlambda = 60, lambda.min.ratio = 1e-3) {
  folds <- make_folds(y, inner_k, seed)
  all_results <- list()

  for (a_idx in seq_along(enet_alpha_grid)) {
    alpha <- enet_alpha_grid[a_idx]
    log_progress("enet_tuning_alpha", paste(prefix, "alpha =", alpha))

    lambda_grid <- enet_lambda_grid(x, y, alpha = alpha,
                                    nlambda = nlambda,
                                    lambda.min.ratio = lambda.min.ratio)


    lambda_results <- vector("list", length(lambda_grid))
    for (g in seq_along(lambda_grid)) {
      lam <- lambda_grid[g]
      fold_losses <- rep(NA_real_, length(folds))
      fold_ok     <- TRUE
      for (j in seq_along(folds)) {
        va <- folds[[j]]
        tr <- setdiff(seq_along(y), va)
        fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lam)

        if (!enet_has_nonzero(fit)) {
          fold_ok <- FALSE
          break
        }
        p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
        fold_losses[j] <- log_loss_safe(y[va], p_va)
      }
      if (!fold_ok) {

        lambda_results[[g]] <- tibble(alpha = alpha, lambda = lam, cvm = Inf, cvsd = NA_real_)
      } else {
        cvm  <- mean(fold_losses, na.rm = TRUE)
        cvsd <- sd(fold_losses, na.rm = TRUE) / sqrt(sum(is.finite(fold_losses)))
        lambda_results[[g]] <- tibble(alpha = alpha, lambda = lam, cvm = cvm, cvsd = cvsd)
      }
    }
    all_results[[a_idx]] <- bind_rows(lambda_results)
  }

  tuning <- bind_rows(all_results)


  tuning <- tuning %>% filter(is.finite(cvm))

  if (nrow(tuning) == 0) {
    stop("All alpha/lambda combinations failed to converge. Check your data.")
  }


  best_per_alpha <- tuning %>%
    group_by(alpha) %>%
    summarise(
      idx_min    = which.min(cvm),
      lambda_min = lambda[idx_min],
      cvm_min    = cvm[idx_min],
      cvsd_min   = cvsd[idx_min],
      .groups    = "drop"
    ) %>%
    mutate(
      cutoff = cvm_min + cvsd_min
    )


  best_1se <- tuning %>%
    inner_join(best_per_alpha %>% select(alpha, cutoff), by = "alpha") %>%
    filter(cvm <= cutoff) %>%
    group_by(alpha) %>%
    summarise(lambda_1se = max(lambda), .groups = "drop")


  global_best <- best_per_alpha %>%
    inner_join(best_1se, by = "alpha") %>%
    arrange(cvm_min) %>%
    slice(1)

  list(
    alpha      = global_best$alpha,
    lambda     = global_best$lambda_1se,
    lambda_min = global_best$lambda_min,
    tuning     = tuning
  )
}



cv_enet_scores <- function(x, y, alpha, lambda, k = 5, seed = 123, prefix = "") {
  folds  <- make_folds(y, k, seed)
  scores <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    log_progress("enet_calibration_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
  }
  scores
}



cv_enet_probs <- function(x, y, alpha, lambda, k = 5, seed = 123, prefix = "") {
  folds <- make_folds(y, k, seed)
  probs <- rep(NA_real_, length(y))
  for (j in seq_along(folds)) {
    log_progress("enet_prob_fold", paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr], alpha = alpha, lambda = lambda)
    probs[va] <- predict_enet_prob(fit, x[va, , drop = FALSE])
  }
  probs
}



summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p), brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    tp_youden = hy$tp, tn_youden = hy$tn, fp_youden = hy$fp, fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc,
    tp_f1 = hf$tp, tn_f1 = hf$tn, fp_f1 = hf$fp, fn_f1 = hf$fn
  )
}





log_progress("start", "elastic net replacement script")



df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels_raw <- readr::read_csv(file.path(data_path, "corpus_600ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  select(id, label)

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
check_binary_y(df_labeled$manual_label, "labeled data")

custom_additions <- c("m", "c", "l", "d", "gt", "lt", "br", "href",
                       "fridericus", "heinricus", "conradus", "meus", "tuus")
stops_v1    <- scan("../../data/stopwords/stopwords_latin1.txt",
                    what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2    <- scan("../../data/stopwords/stopwords_latin2.txt",
                    what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")
latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]

lab <- df_labeled %>% arrange(row_id)
out <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos <- match(lab$row_id, out$row_id)

n_lab <- nrow(lab)



outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)



enet_oof              <- rep(NA_real_, n_lab)
enet_score_oof        <- rep(NA_real_, n_lab)
enet_pred_youden_oof  <- rep(NA_character_, n_lab)
enet_pred_f1_oof      <- rep(NA_character_, n_lab)
enet_outer_fold_id    <- integer(n_lab)


fold_threshold_rows <- vector("list", length(outer_folds))
fold_metric_rows    <- vector("list", length(outer_folds))
fold_hyper_rows     <- vector("list", length(outer_folds))

summarize_model_fold <- function(y, p, pred_youden, pred_f1, model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold, model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p), brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc
  )
}

for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start", paste("fold", i, "of", length(outer_folds)))
  va <- outer_folds[[i]]
  tr <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]


  vec_fold <- fit_vectorizer(lab$text[tr], latin_stops)
  x_tr     <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
  x_va     <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)



  enet_tuned <- tune_enet(
    x_tr, y_tr,
    inner_k = k_folds_inner,
    seed    = 2000 + i,
    prefix  = paste("outer", i)
  )

  log_progress("enet_outer_tuned", paste("outer", i,
               "| alpha =", enet_tuned$alpha,
               "| lambda_1se =", round(enet_tuned$lambda, 6)))




  enet_train_prob <- cv_enet_probs(
    x_tr, y_tr,
    alpha  = enet_tuned$alpha,
    lambda = enet_tuned$lambda,
    k      = k_folds_inner,
    seed   = 2100 + i,
    prefix = paste("outer", i)
  )


  enet_fit       <- fit_enet_fixed(x_tr, y_tr,
                                   alpha  = enet_tuned$alpha,
                                   lambda = enet_tuned$lambda)
  enet_va_scores <- predict_enet_score(enet_fit, x_va)
  enet_va_prob   <- predict_enet_prob(enet_fit, x_va)


  enet_th_y <- choose_threshold(y_tr, enet_train_prob, "youden")
  enet_th_f <- choose_threshold(y_tr, enet_train_prob, "f1")


  enet_oof[va]             <- enet_va_prob
  enet_score_oof[va]       <- enet_va_scores
  enet_pred_youden_oof[va] <- ifelse(enet_va_prob >= enet_th_y, "christian", "non_christian")
  enet_pred_f1_oof[va]     <- ifelse(enet_va_prob >= enet_th_f, "christian", "non_christian")
  enet_outer_fold_id[va]   <- i


  fold_threshold_rows[[i]] <- tibble(
    outer_fold = i, model = "ridge",
    threshold_youden = enet_th_y, threshold_f1 = enet_th_f
  )
  fold_metric_rows[[i]] <- summarize_model_fold(
    y_va, enet_va_prob,
    enet_pred_youden_oof[va], enet_pred_f1_oof[va],
    "ridge", i
  )
  fold_hyper_rows[[i]] <- tibble(
    outer_fold = i, model = "ridge",
    threshold_youden = enet_th_y, threshold_f1 = enet_th_f,
    alpha = enet_tuned$alpha,
    lambda_1se = enet_tuned$lambda,
    lambda_min = enet_tuned$lambda_min
  )

  log_progress("outer_fold_done", paste("fold", i, "| th_youden =",
               round(enet_th_y, 4), "| th_f1 =", round(enet_th_f, 4)))
}



oof_metrics_enet <- summarize_model_oof(
  lab$manual_label, enet_oof, enet_pred_youden_oof, enet_pred_f1_oof, "ridge"
)

cat("\n===== Elastic net OOF metrics (replacing ridge) =====\n")
print(as.data.frame(oof_metrics_enet %>% select(model, auc, log_loss, brier,
                                                 f1_youden, mcc_youden,
                                                 f1_f1, mcc_f1)))



log_progress("full_fit", "tuning elastic net on all labeled data")

vec_full      <- fit_vectorizer(lab$text, latin_stops)
x_lab_full    <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all_full    <- transform_tfidf_sparse(out$text, vec_full, latin_stops)


enet_tuned_full <- tune_enet(
  x_lab_full, lab$manual_label,
  inner_k = k_folds_inner,
  seed    = 8002,
  prefix  = "full"
)

log_progress("full_fit_tuned", paste("alpha =", enet_tuned_full$alpha,
             "| lambda =", round(enet_tuned_full$lambda, 6)))


enet_lab_prob_cf <- cv_enet_probs(
  x_lab_full, lab$manual_label,
  alpha  = enet_tuned_full$alpha,
  lambda = enet_tuned_full$lambda,
  k      = k_folds_inner,
  seed   = 8102,
  prefix = "full"
)


enet_lab_score_cf <- cv_enet_scores(
  x_lab_full, lab$manual_label,
  alpha  = enet_tuned_full$alpha,
  lambda = enet_tuned_full$lambda,
  k      = k_folds_inner,
  seed   = 8102,
  prefix = "full"
)


enet_fit_full <- fit_enet_fixed(
  x_lab_full, lab$manual_label,
  alpha  = enet_tuned_full$alpha,
  lambda = enet_tuned_full$lambda
)


enet_all_scores <- predict_enet_score(enet_fit_full, x_all_full)
enet_all_prob   <- predict_enet_prob(enet_fit_full, x_all_full)


enet_deploy_th_y <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "youden")
enet_deploy_th_f <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "f1")

log_progress("full_fit_done", paste("deploy th_youden =", round(enet_deploy_th_y, 4),
             "| deploy th_f1 =", round(enet_deploy_th_f, 4)))





prob_ridge         <- enet_all_prob
score_ridge_logit  <- enet_all_scores

prob_ridge[lab_pos]        <- enet_lab_prob_cf
score_ridge_logit[lab_pos] <- enet_lab_score_cf

label_ridge_youden <- ifelse(prob_ridge >= enet_deploy_th_y, "christian", "non_christian")
label_ridge_f1     <- ifelse(prob_ridge >= enet_deploy_th_f, "christian", "non_christian")



output_csv_path <- file.path(results_path, "corpus_600ad_model_outputs_oof.csv")
existing <- readr::read_csv(output_csv_path, show_col_types = FALSE)

stopifnot(nrow(existing) == length(prob_ridge))

existing$prob_ridge         <- safe_prob(prob_ridge)
existing$score_ridge_logit  <- score_ridge_logit
existing$label_ridge_youden <- label_ridge_youden
existing$label_ridge_f1     <- label_ridge_f1


readr::write_csv(existing, output_csv_path)
log_progress("csv_updated", paste("wrote", output_csv_path))





hyper_csv_path <- file.path(results_path, "model_hyperparameters_full.csv")
if (file.exists(hyper_csv_path)) {
  hyper_full <- readr::read_csv(hyper_csv_path, show_col_types = FALSE)


  hyper_full <- hyper_full %>% filter(model != "ridge")


  new_ridge_row <- tibble(
    model      = "ridge",
    alpha      = enet_tuned_full$alpha,
    lambda_1se = enet_tuned_full$lambda,
    lambda_min = enet_tuned_full$lambda_min
  )

  hyper_full <- bind_rows(hyper_full, new_ridge_row)
  readr::write_csv(hyper_full, hyper_csv_path)
  log_progress("hyper_updated", paste("updated ridge row in", hyper_csv_path))
} else {
  warning("model_hyperparameters_full.csv not found at ", hyper_csv_path,
          " — skipping hyperparameter update. The driver script will need manual alpha/lambda values.")
}



thresh_csv_path <- file.path(results_path, "model_thresholds_deployment.csv")
if (file.exists(thresh_csv_path)) {
  thresh_deploy <- readr::read_csv(thresh_csv_path, show_col_types = FALSE)
  thresh_deploy <- thresh_deploy %>% filter(model != "ridge")
  thresh_deploy <- bind_rows(thresh_deploy, tibble(
    model             = "ridge",
    threshold_youden  = enet_deploy_th_y,
    threshold_f1      = enet_deploy_th_f,
    threshold_scope   = "full_labeled_cross_fitted_calibrated"
  ))
  readr::write_csv(thresh_deploy, thresh_csv_path)
  log_progress("thresh_updated", paste("updated ridge row in", thresh_csv_path))
}




save_rds_atomic(enet_fit_full,    file.path(enet_root, "enet_model_full.rds"))
save_rds_atomic(enet_tuned_full,  file.path(enet_root, "enet_tuning_full.rds"))


oof_eval_enet <- tibble(
  id             = lab$id,
  row_id         = lab$row_id,
  manual_label   = lab$manual_label,
  prob_ridge     = enet_oof,
  score_ridge    = enet_score_oof,
  label_youden   = enet_pred_youden_oof,
  label_f1       = enet_pred_f1_oof
)
readr::write_csv(oof_eval_enet, file.path(enet_root, "enet_oof_predictions.csv"))


readr::write_csv(
  tibble(
    model = "ridge",
    method = "elastic_net",
    alpha = enet_tuned_full$alpha,
    lambda_1se = enet_tuned_full$lambda,
    lambda_min = enet_tuned_full$lambda_min,
    threshold_youden = enet_deploy_th_y,
    threshold_f1 = enet_deploy_th_f,
    threshold_scope = "full_labeled_cross_fitted_calibrated"
  ),
  file.path(enet_root, "enet_thresholds.csv")
)


readr::write_csv(oof_metrics_enet, file.path(enet_root, "enet_oof_metrics.csv"))


coef_mat <- as.matrix(coef(enet_fit_full, s = enet_tuned_full$lambda))
coef_df  <- tibble(
  feature     = rownames(coef_mat),
  coefficient = as.numeric(coef_mat[, 1])
) %>%
  filter(feature != "(Intercept)", coefficient != 0) %>%
  arrange(desc(abs(coefficient)))

cat("\n===== Top 30 elastic net features (non-zero coefficients) =====\n")
print(head(coef_df, 30), n = 30)

n_nonzero <- sum(coef_df$coefficient != 0)
n_total   <- nrow(coef_mat) - 1
cat(sprintf("\nFeatures with non-zero coefficients: %d / %d (%.1f%%)\n",
            n_nonzero, n_total, 100 * n_nonzero / n_total))

readr::write_csv(coef_df, file.path(enet_root, "enet_coefficients.csv"))

log_progress("done", "elastic net replacement complete")

cat("\n===== SUMMARY =====\n")
cat(sprintf("Selected alpha:   %s\n", enet_tuned_full$alpha))
cat(sprintf("Selected lambda:  %s\n", round(enet_tuned_full$lambda, 6)))
cat(sprintf("Non-zero features: %d / %d\n", n_nonzero, n_total))
cat(sprintf("Deploy threshold (youden): %.4f\n", enet_deploy_th_y))
cat(sprintf("Deploy threshold (f1):     %.4f\n", enet_deploy_th_f))
cat(sprintf("OOF AUC:      %.4f\n", oof_metrics_enet$auc))
cat(sprintf("OOF Log-loss: %.4f\n", oof_metrics_enet$log_loss))
cat(sprintf("OOF F1 (youden threshold): %.4f\n", oof_metrics_enet$f1_youden))
cat(sprintf("OOF MCC (youden threshold): %.4f\n", oof_metrics_enet$mcc_youden))
cat(sprintf("\nUpdated: %s\n", output_csv_path))






replace_ridge_rows <- function(csv_path, new_rows, key_cols = c("model")) {
  if (!file.exists(csv_path)) {
    warning("File not found, skipping: ", csv_path)
    return(invisible(NULL))
  }
  old <- readr::read_csv(csv_path, show_col_types = FALSE)

  keep <- old %>% filter(model != "ridge")

  updated <- bind_rows(keep, new_rows)
  readr::write_csv(updated, csv_path)
  log_progress("file_updated", csv_path)
}


replace_ridge_rows(
  file.path(results_path, "model_thresholds_oof.csv"),
  tibble(
    model            = "ridge",
    threshold_youden = enet_deploy_th_y,
    threshold_f1     = enet_deploy_th_f,
    threshold_scope  = "full_labeled_cross_fitted_calibrated"
  )
)


replace_ridge_rows(
  file.path(results_path, "model_metrics_oof.csv"),
  oof_metrics_enet
)


replace_ridge_rows(
  file.path(results_path, "model_metrics_oof_by_fold.csv"),
  bind_rows(fold_metric_rows),
  key_cols = c("outer_fold", "model")
)


replace_ridge_rows(
  file.path(results_path, "model_thresholds_outer.csv"),
  bind_rows(fold_threshold_rows),
  key_cols = c("outer_fold", "model")
)


replace_ridge_rows(
  file.path(results_path, "model_hyperparameters_outer.csv"),
  bind_rows(fold_hyper_rows),
  key_cols = c("outer_fold", "model")
)


oof_confusion_new <- bind_rows(
  hard_metrics(lab$manual_label, enet_pred_youden_oof) %>%
    mutate(model = "ridge", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, enet_pred_f1_oof) %>%
    mutate(model = "ridge", threshold_kind = "f1")
)
replace_ridge_rows(
  file.path(results_path, "oof_confusion_counts.csv"),
  oof_confusion_new,
  key_cols = c("model", "threshold_kind")
)


oof_eval_path <- file.path(results_path, "oof_eval_predictions.csv")
if (file.exists(oof_eval_path)) {
  oof_eval <- readr::read_csv(oof_eval_path, show_col_types = FALSE)

  oof_eval$prob_ridge         <- enet_oof
  oof_eval$score_ridge_logit  <- enet_score_oof
  oof_eval$label_ridge_youden <- enet_pred_youden_oof
  oof_eval$label_ridge_f1     <- enet_pred_f1_oof
  readr::write_csv(oof_eval, oof_eval_path)
  log_progress("file_updated", oof_eval_path)
}



tuning_full_path <- file.path(results_path, "model_tuning_full.csv")
if (file.exists(tuning_full_path)) {
  tuning_full <- readr::read_csv(tuning_full_path, show_col_types = FALSE)
  tuning_full <- tuning_full %>% filter(model != "ridge")
  new_tuning_row <- tibble(
    model      = "ridge",
    alpha      = enet_tuned_full$alpha,
    lambda_1se = enet_tuned_full$lambda,
    lambda_min = enet_tuned_full$lambda_min
  )
  tuning_full <- bind_rows(tuning_full, new_tuning_row)
  readr::write_csv(tuning_full, tuning_full_path)
  log_progress("file_updated", tuning_full_path)
}

log_progress("all_files_done", "all downstream CSVs updated")
```

