---
title: "Latin text classification (600 AD)"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of Christian vs. non-Christian documents

This pipeline uses **elastic net logistic regression** (plus NB, SVM, XGBoost, LightGBM for
comparison) on TF-IDF features of lemmatized Latin text. A few changes relative to the older
version of this script:

1. **No ridge regression anywhere.** The regularized logistic model is elastic net throughout,
   with `alpha` tuned over a grid (so it can recover ridge-like behaviour if that is what the
   data wants, but it is not forced).
2. **Stopword removal is now case-insensitive.** Previously `tokens_remove()` matched
   case-sensitively, which meant capitalized proper nouns (author names, saint names, emperor
   names, place names) were sailing past the filter and getting included as features at DFM
   time after lowercasing. This was the main source of "high AUC for the wrong reasons":
   the classifier was learning *who wrote the text* or *who it was dedicated to* rather than
   anything about Christian vs. non-Christian language.
3. **Much more aggressive proper-noun / author / dedicatee filtering.** The previous
   `custom_additions` list contained a handful of names (`fridericus`, `heinricus`,
   `conradus`). We replace this with a full blocklist of Christian/non-Christian author names,
   saint names, emperor names, biblical figures, place names, and liturgical markers that
   trivially reveal the class label.
4. Consistent variable names throughout — no more `prob_ridge` being used to hold elastic-net
   probabilities.

## Data loading

```{r setup-paths}
library(tidyverse)
library(quanteda)
library(quanteda.textmodels)
library(glmnet)
library(LiblineaR)
library(xgboost)
library(lightgbm)
library(Matrix)
library(pROC)

set.seed(123)

data_path       <- "../../data/processed_texts/600ad"
results_path    <- "../../results/predictions/non_transformer_lemmatized/600ad_no_name"
checkpoint_path <- results_path
checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
output_dir      <- file.path(checkpoint_path, "model_outputs_checkpointed")
progress_path   <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir,      recursive = TRUE, showWarnings = FALSE)

k_folds_outer     <- 5L
k_folds_inner     <- 5L
xgb_max_rounds    <- 4000L
lgb_max_rounds    <- 4000L
early_stop_rounds <- 100L
min_docfreq       <- 3L
```

```{r load-data}
df <- readr::read_csv(file.path(data_path, "corpus_600ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

dup_corpus <- df %>% count(id, name = "n") %>% filter(n > 1)
if (nrow(dup_corpus) > 0) {
  stop(paste0("Duplicate ids in corpus_600ad.csv: ",
              paste(head(dup_corpus$id, 20), collapse = ", ")))
}

labels_raw <- readr::read_csv(file.path(data_path, "corpus_600ad_labeled.csv"),
                              show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  select(id, label)

label_conflicts <- labels_raw %>%
  distinct(id, label) %>%
  count(id, name = "n_labels") %>%
  filter(n_labels > 1)
if (nrow(label_conflicts) > 0) {
  stop(paste0("Conflicting labels: ",
              paste(head(label_conflicts$id, 20), collapse = ", ")))
}

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
```

## Stopword & proper-noun filter

```{r stopwords}
# -- 1. Generic Latin stopwords (function words, pronouns, auxiliaries) ---------------

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt",
                 what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt",
                 what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

# -- 2. Corpus / OCR artifacts --------------------------------------------------------
# These are single letters, HTML escape fragments, and scan tags that leak through
# OCR and XML stripping. They are not real Latin tokens.

corpus_artifacts <- c(
  "m", "c", "l", "d",       # stray single letters (numerals, abbreviations)
  "gt", "lt", "amp", "quot", # HTML entity fragments
  "br", "href", "http", "https", "www"
)


proper_nouns_christian <- c(
  # Divinity & core figures
  "christus", "iesus", "jesus", "messias",
  # Apostles & evangelists
  "paulus", "petrus", "iohannes", "johannes", "matthaeus", "marcus", "lucas",
  "andreas", "iacobus", "jacobus", "thomas", "philippus", "bartholomaeus",
  "simon", "thaddaeus", "matthias", "barnabas",
  # Old Testament figures
  "moyses", "moses", "abraham", "isaac", "iacob", "jacob", "david", "salomon",
  "isaias", "hieremias", "ezechiel", "daniel", "noe", "adam", "eva",
  # Marian / Holy Family
  "maria",  "joseph", "ioseph",
  # Patristic & early-medieval Christian authors frequently cited or named
  "augustinus", "hieronymus", "ambrosius", "gregorius", "benedictus",
  "cassiodorus", "boethius", "cyprianus",
  "tertullianus", "origenes", "athanasius", "basilius", "chrysostomus",
  "leo", "hilarius", "martinus",
  # Saints, martyrs, popes commonly named
  "stephanus", "laurentius", "vincentius", "agnes", "caecilia", "agatha",
  "sebastianus", "georgius", "nicolaus", "silvester", "perpetua", "felicitas",
  # Liturgical / cult markers (not ordinary theological vocabulary)
  "amen", "alleluia", "halleluia", "kyrie", "eleison",
  # Places with strong Christian association in this period
  "ierusalem", "jerusalem", "bethleem", "nazareth", "galilea",
  "roma", "constantinopolis", "antiochia"
)

proper_nouns_non_christian <- c(
  # Roman / pagan deities
  "iuppiter", "jupiter", "iovis", "juno", "iuno", "minerva", "mars", "martis",
  "venus", "veneris", "mercurius", "apollo", "apollinis", "diana", "neptunus",
  "vulcanus", "ceres", "bacchus", "saturnus", "pluto", "vesta", "ianus", "janus",
  "hercules",
  # Greek deities / mythological figures that recur
  "zeus", "hera", "athena", "hermes", "aphrodite", "poseidon",
  "achilles", "ulixes", "ulysses", "aeneas", "hector", "priamus",
  # Divine epithets
  "phoebus", "lyaeus", "gradivus",
  # Roman emperors / generals / statesmen (pre-Christian and the few anachronistic)
  "augustus", "caesar", "nero", "traianus", "hadrianus", "marcus",
  "cicero", "tullius", "pompeius", "crassus", "brutus", "antonius",
  "fridericus", "heinricus", "henricus", "conradus", "otto", "ottonis",
  "carolus",
  # Classical authors frequently named
  "vergilius", "virgilius", "horatius", "ovidius", "livius", "tacitus",
  "sallustius", "lucretius", "plinius", "seneca"
)


latin_stops <- sort(unique(tolower(c(
  stops_v1,
  stops_v2,
  corpus_artifacts,
  proper_nouns_christian,
  proper_nouns_non_christian
))))
latin_stops <- latin_stops[latin_stops != "*" & nzchar(latin_stops)]

length(latin_stops)
head(latin_stops, 50)
```

## Shared utilities (atomic I/O, logging, metrics, folds)

```{r utilities}
`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {
  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_",
                format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  if (!file.rename(tmp, path)) {
    file.copy(tmp, path, overwrite = TRUE); unlink(tmp)
  }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_",
                format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  if (!file.rename(tmp, path)) {
    file.copy(tmp, path, overwrite = TRUE); unlink(tmp)
  }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}
```

```{r}
append_progress_row <- function(row, path) {
  ensure_dir(dirname(path))
  readr::write_csv(
    row,
    path,
    append    = file.exists(path),
    col_names = !file.exists(path)
  )
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time   = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step   = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step,
                 if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

safe_prob  <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("christian", "non_christian") %in% names(tb))) {
    stop("Both classes required",
         if (nzchar(where)) paste0(" in ", where) else "", ".")
  }
  if (any(tb < 2)) {
    stop("Each class needs >= 2 observations",
         if (nzchar(where)) paste0(" in ", where) else "", ".")
  }
}

make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for CV.")
  idx_pos <- sample(which(y == "christian"))
  idx_neg <- sample(which(y == "non_christian"))
  folds <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos))
    folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg))
    folds[[((i - 1) %% k_use) + 1]] <- c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = p,
      levels = c("non_christian", "christian"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = s,
      levels = c("non_christian", "christian"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p <- safe_prob(p)
  y01 <- as.integer(y == "christian")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "christian")
  p01 <- as.integer(pred_label == "christian")
  tp <- sum(y01 == 1 & p01 == 1); tn <- sum(y01 == 0 & p01 == 0)
  fp <- sum(y01 == 0 & p01 == 1); fn <- sum(y01 == 1 & p01 == 0)
  precision   <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall      <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1  <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0)
           NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(
    accuracy = (tp + tn) / length(y), balanced_accuracy = bal_acc,
    precision = precision, recall = recall, specificity = specificity,
    f1 = f1, mcc = mcc, tp = tp, tn = tn, fp = fp, fn = fn
  )
}

choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(response = y, predictor = p,
                  levels = c("non_christian", "christian"),
                  direction = "<", quiet = TRUE),
        x = "best", input = "threshold", best.method = "youden",
        ret = "threshold", transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01  <- as.integer(y == "christian")
    cand <- sort(unique(c(0, p, 0.5, 1)))
    f1s  <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp <- sum(pred01 == 1 & y01 == 1)
      fp <- sum(pred01 == 1 & y01 == 0)
      fn <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p), brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    tp_youden = hy$tp, tn_youden = hy$tn, fp_youden = hy$fp, fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc,
    tp_f1 = hf$tp, tn_f1 = hf$tn, fp_f1 = hf$fp, fn_f1 = hf$fn
  )
}

summarize_model_fold <- function(y, p, pred_youden, pred_f1, model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold, model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p), brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc
  )
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]), by = keys)
  bind_rows(keep, new_rows)
}

fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores), y = as.integer(y == "christian")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 || length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(),
                 control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores)); z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores),
                      type = "response"))
  }
}
```

## Tokenization & TF-IDF

```{r tokenizer}
# Key fix vs. previous version: tokens_tolower() BEFORE tokens_remove(), so that
# proper nouns like "Fridericus", "Augustinus", "Iesus" are normalised to their
# lowercase form first and then stripped by the stopword / blocklist. Previously
# the filter ran case-sensitively and these tokens survived into the DFM.
make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_tolower() %>%
    quanteda::tokens_remove(pattern = stop_list, valuetype = "fixed") %>%
    quanteda::dfm(tolower = FALSE)  # already lowercased above
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L) {
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  }
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N   <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat    <- as(x_counts, "dgCMatrix")
  # Sublinear TF on non-zero entries: 1 + log(tf)
  x_mat@x  <- 1 + log(x_mat@x)
  idf_vec  <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf  <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}
```

## Model wrappers

### Naive Bayes

```{r nb-wrappers}
nb_prob_christian <- function(pred_obj) {
  if ("christian" %in% colnames(pred_obj)) safe_prob(pred_obj[, "christian"])
  else safe_prob(pred_obj[, ncol(pred_obj)])
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}

predict_nb_score <- function(fit, newx_dfm) {
  safe_logit(nb_prob_christian(predict(fit, newdata = newx_dfm, type = "probability")))
}
```

### Elastic net (replaces ridge everywhere)

```{r enet-wrappers}
# Alpha = 0 would be ridge, alpha = 1 lasso. We tune over values in between; the
# grid does NOT include 0 so ridge is explicitly excluded per request.
enet_alpha_grid <- c(0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.7, 0.9)

fit_enet_fixed <- function(x, y, alpha, lambda) {
  suppressWarnings(
    glmnet::glmnet(
      x = x, y = as.integer(y == "christian"),
      family = "binomial",
      alpha = alpha, lambda = lambda,
      standardize = TRUE, maxit = 1e6
    )
  )
}

enet_has_nonzero <- function(fit) {
  beta <- as.numeric(coef(fit)[-1, 1])
  sum(beta != 0) > 0
}

predict_enet_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_enet_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

enet_lambda_grid <- function(x, y, alpha, nlambda = 60, lambda.min.ratio = 1e-3) {
  base_fit <- suppressWarnings(glmnet::glmnet(
    x = x, y = as.integer(y == "christian"),
    family = "binomial", alpha = alpha,
    standardize = TRUE,
    nlambda = nlambda, lambda.min.ratio = lambda.min.ratio,
    maxit = 1e6
  ))
  lams <- as.numeric(base_fit$lambda)
  if (length(lams) == 0) {
    warning("glmnet returned no lambda path for alpha = ", alpha,
            "; using fallback grid.")
    lams <- 10^seq(log10(0.001), log10(100), length.out = nlambda)
  }
  lams
}

tune_enet_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "",
                           ckpt_dir, nlambda = 60, lambda.min.ratio = 1e-3) {
  ensure_dir(ckpt_dir)
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path) %||%
    list(done = rep(FALSE, length(enet_alpha_grid)))

  for (a_idx in seq_along(enet_alpha_grid)) {
    if (isTRUE(summary_state$done[a_idx])) next
    alpha <- enet_alpha_grid[a_idx]
    log_progress("enet_tuning_alpha", paste(prefix, "alpha =", alpha))

    alpha_dir  <- file.path(ckpt_dir, sprintf("alpha_%03d", a_idx))
    ensure_dir(alpha_dir)
    lambda_grid <- enet_lambda_grid(x, y, alpha = alpha,
                                    nlambda = nlambda,
                                    lambda.min.ratio = lambda.min.ratio)
    save_rds_atomic(lambda_grid, file.path(alpha_dir, "lambda_grid.rds"))

    for (g in seq_along(lambda_grid)) {
      cand_path <- file.path(alpha_dir, sprintf("lambda_%03d.rds", g))
      cand_state <- load_rds_or_null(cand_path) %||%
        list(fold_losses = rep(NA_real_, length(folds)),
             done = rep(FALSE, length(folds)),
             any_zero = FALSE)
      lam <- lambda_grid[g]
      for (j in seq_along(folds)) {
        if (isTRUE(cand_state$done[j])) next
        va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
        fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr],
                              alpha = alpha, lambda = lam)
        if (!enet_has_nonzero(fit)) {
          cand_state$any_zero    <- TRUE
          cand_state$fold_losses <- rep(NA_real_, length(folds))
          cand_state$done[]      <- TRUE
          save_rds_atomic(cand_state, cand_path)
          break
        }
        p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
        cand_state$fold_losses[j] <- log_loss_safe(y[va], p_va)
        cand_state$done[j] <- TRUE
        save_rds_atomic(cand_state, cand_path)
      }
    }
    summary_state$done[a_idx] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  # Assemble tuning table
  rows <- list()
  for (a_idx in seq_along(enet_alpha_grid)) {
    alpha <- enet_alpha_grid[a_idx]
    alpha_dir   <- file.path(ckpt_dir, sprintf("alpha_%03d", a_idx))
    lambda_grid <- readRDS(file.path(alpha_dir, "lambda_grid.rds"))
    for (g in seq_along(lambda_grid)) {
      cs <- readRDS(file.path(alpha_dir, sprintf("lambda_%03d.rds", g)))
      if (isTRUE(cs$any_zero)) {
        rows[[length(rows) + 1]] <-
          tibble(alpha = alpha, lambda = lambda_grid[g],
                 cvm = Inf, cvsd = NA_real_)
      } else {
        cvm  <- mean(cs$fold_losses, na.rm = TRUE)
        cvsd <- stats::sd(cs$fold_losses, na.rm = TRUE) /
                sqrt(sum(is.finite(cs$fold_losses)))
        rows[[length(rows) + 1]] <-
          tibble(alpha = alpha, lambda = lambda_grid[g],
                 cvm = cvm, cvsd = cvsd)
      }
    }
  }
  tuning <- bind_rows(rows) %>% filter(is.finite(cvm))
  if (nrow(tuning) == 0) {
    stop("All alpha/lambda combinations failed to converge.")
  }

  best_per_alpha <- tuning %>%
    group_by(alpha) %>%
    summarise(
      idx_min    = which.min(cvm),
      lambda_min = lambda[idx_min],
      cvm_min    = cvm[idx_min],
      cvsd_min   = cvsd[idx_min],
      .groups    = "drop"
    ) %>%
    mutate(cutoff = cvm_min + cvsd_min)

  best_1se <- tuning %>%
    inner_join(best_per_alpha %>% select(alpha, cutoff), by = "alpha") %>%
    filter(cvm <= cutoff) %>%
    group_by(alpha) %>%
    summarise(lambda_1se = max(lambda), .groups = "drop")

  global_best <- best_per_alpha %>%
    inner_join(best_1se, by = "alpha") %>%
    arrange(cvm_min) %>%
    slice(1)

  list(
    alpha      = global_best$alpha,
    lambda     = global_best$lambda_1se,
    lambda_min = global_best$lambda_min,
    tuning     = tuning
  )
}

cv_enet_scores_ckpt <- function(x, y, alpha, lambda, k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("enet_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr],
                          alpha = alpha, lambda = lambda)
    state$scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

### Linear SVM, XGBoost, LightGBM (unchanged in spirit, kept for benchmarking)

```{r svm-wrappers}
svm_to_matrix <- function(x) {
  if (inherits(x, c("dgCMatrix", "dsCMatrix", "dgRMatrix", "matrix"))) return(x)
  as.matrix(x)
}

fit_svm_fixed <- function(x, y, cost) {
  x_use <- svm_to_matrix(x)
  fit <- LiblineaR::LiblineaR(data = x_use, target = y, type = 7, cost = cost)
  pred_tr <- predict(fit, x_use, decisionValues = TRUE)
  raw <- as.numeric(pred_tr$decisionValues[, 1])
  sgn <- if (mean(raw[y == "christian"]) >= mean(raw[y == "non_christian"])) 1 else -1
  list(model = fit, sign = sgn, cost = cost)
}

predict_svm_score <- function(obj, x) {
  pred <- predict(obj$model, svm_to_matrix(x), decisionValues = TRUE)
  obj$sign * as.numeric(pred$decisionValues[, 1])
}

tune_svm_cost_ckpt <- function(x, y,
                               cost_grid = c(0.01, 0.05, 0.1, 0.25, 0.5, 1, 2, 4, 8, 16),
                               inner_k = 5, seed = 123, prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  rows_path <- file.path(ckpt_dir, "summary.rds")
  state <- load_rds_or_null(rows_path) %||%
    list(done = rep(FALSE, length(cost_grid)),
         rows = vector("list", length(cost_grid)))
  for (i in seq_along(cost_grid)) {
    if (isTRUE(state$done[i])) next
    scores_i <- cv_svm_scores_ckpt(
      x, y, cost = cost_grid[i], k = inner_k, seed = seed + i,
      prefix = prefix, ckpt_dir = file.path(ckpt_dir, sprintf("cost_%03d", i))
    )
    state$rows[[i]] <- tibble(cost = cost_grid[i],
                              auc_margin = auc_margin_safe(y, scores_i))
    state$done[i] <- TRUE
    save_rds_atomic(state, rows_path)
  }
  tuning <- bind_rows(state$rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}

cv_svm_scores_ckpt <- function(x, y, cost, k = 5, seed = 123,
                               prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("svm_calibration_fold",
                 paste(prefix, "cost", cost, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    state$scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

```{r xgb-wrappers}
xgb_grid <- function() {
  tidyr::crossing(
    eta = c(0.01, 0.03, 0.05),
    max_depth = c(3L, 4L, 6L),
    min_child_weight = c(1, 3, 5),
    subsample = c(0.7, 0.85),
    colsample_bytree = c(0.6, 0.8),
    gamma = c(0, 0.1, 0.5)
  ) %>% mutate(candidate_id = row_number()) %>% select(candidate_id, everything())
}

best_iter_xgb <- function(fit, fallback) {
  bi <- fit$best_iteration
  if (is.null(bi) || length(bi) == 0 || !is.finite(bi))
    as.integer(fallback) else as.integer(bi)
}

make_xgb_params <- function(row, scale_pos_weight = 1) {
  list(
    booster = "gbtree", objective = "binary:logistic",
    eval_metric = "logloss", tree_method = "hist",
    eta = as.numeric(row$eta),
    max_depth = as.integer(row$max_depth),
    min_child_weight = as.numeric(row$min_child_weight),
    subsample = as.numeric(row$subsample),
    colsample_bytree = as.numeric(row$colsample_bytree),
    gamma = as.numeric(row$gamma),
    scale_pos_weight = scale_pos_weight
  )
}

fit_xgb_fixed <- function(x, y, params, nrounds, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  xgboost::xgb.train(
    params = params,
    data = xgboost::xgb.DMatrix(x, label = as.integer(y == "christian")),
    nrounds = as.integer(nrounds), verbose = 0
  )
}

predict_xgb_score <- function(fit, x) {
  as.numeric(predict(fit, xgboost::xgb.DMatrix(x), outputmargin = TRUE))
}

tune_xgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- xgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path) %||%
    list(done = rep(FALSE, nrow(grid)))
  n_pos <- sum(y == "christian"); n_neg <- sum(y == "non_christian")
  spw <- n_neg / max(n_pos, 1)
  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path) %||%
      list(oof_prob = rep(NA_real_, length(y)),
           best_iters = rep(NA_integer_, length(folds)),
           done = rep(FALSE, length(folds)))
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("xgb_tuning_fold",
                   paste(prefix, "candidate", g, "of", nrow(grid),
                         "fold", j, "of", length(folds)))
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "christian")
      n_neg_tr <- sum(y[tr] == "non_christian")
      spw_tr <- n_neg_tr / max(n_pos_tr, 1)
      dtr <- xgboost::xgb.DMatrix(x[tr, , drop = FALSE],
                                   label = as.integer(y[tr] == "christian"))
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE],
                                   label = as.integer(y[va] == "christian"))
      fit <- xgboost::xgb.train(
        params = make_xgb_params(pars_row, scale_pos_weight = spw_tr),
        data = dtr, nrounds = xgb_max_rounds,
        evals = list(valid = dva),
        early_stopping_rounds = early_stop_rounds, verbose = 0
      )
      cand_state$oof_prob[va]  <- safe_prob(as.numeric(predict(fit, dva)))
      cand_state$best_iters[j] <- best_iter_xgb(fit, xgb_max_rounds)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }
  tuning_rows <- lapply(seq_len(nrow(grid)), function(g) {
    pars_row <- grid[g, , drop = FALSE]
    cs <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    pars_row %>% mutate(
      log_loss = log_loss_safe(y, cs$oof_prob),
      auc      = auc_safe(y, cs$oof_prob),
      best_nrounds = max(1L, as.integer(round(median(cs$best_iters, na.rm = TRUE))))
    )
  })
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best,
       params = make_xgb_params(best, scale_pos_weight = spw),
       tuning = tuning)
}

cv_xgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123,
                               prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("xgb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "christian"); n_neg_tr <- sum(y[tr] == "non_christian")
    fold_params <- modifyList(params,
                              list(scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr],
                         params = fold_params, nrounds = nrounds, seed = seed + j)
    state$scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

```{r lgb-wrappers}
lgb_grid <- function() {
  tidyr::crossing(
    learning_rate = c(0.01, 0.03, 0.05),
    num_leaves = c(15L, 31L, 63L),
    min_data_in_leaf = c(5L, 10L, 20L),
    feature_fraction = c(0.6, 0.8),
    bagging_fraction = c(0.7, 0.85)
  ) %>% mutate(candidate_id = row_number()) %>% select(candidate_id, everything())
}

make_lgb_params <- function(row, seed = 123) {
  list(
    objective = "binary", metric = "binary_logloss", boosting = "gbdt",
    learning_rate = as.numeric(row$learning_rate),
    num_leaves = as.integer(row$num_leaves),
    min_data_in_leaf = as.integer(row$min_data_in_leaf),
    feature_fraction = as.numeric(row$feature_fraction),
    bagging_fraction = as.numeric(row$bagging_fraction),
    bagging_freq = 1L, is_unbalance = TRUE,
    first_metric_only = TRUE,
    seed = seed, feature_fraction_seed = seed, bagging_seed = seed,
    verbosity = -1
  )
}

make_lgb_dataset_params <- function() list(feature_pre_filter = FALSE)

predict_lgb_prob <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "response"))),
    error = function(e) tryCatch(do.call(predict, args),
                                 error = function(e2) rep(0.5, nrow(x)))
  )
  safe_prob(as.numeric(out))
}

predict_lgb_score <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "raw"))),
    error = function(e) tryCatch(
      do.call(predict, c(args, list(rawscore = TRUE))),
      error = function(e2) qlogis(safe_prob(predict_lgb_prob(model, x, num_iteration)))
    )
  )
  as.numeric(out)
}

fit_lgb_fixed <- function(x, y, params, nrounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x, label = as.integer(y == "christian"),
    params = make_lgb_dataset_params(), free_raw_data = TRUE
  )
  lightgbm::lgb.train(params = params, data = dtrain,
                      nrounds = as.integer(nrounds), verbose = -1)
}

fit_lgb_early_stop <- function(x_tr, y_tr, x_va, y_va, params,
                               max_rounds, early_rounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x_tr, label = as.integer(y_tr == "christian"),
    params = make_lgb_dataset_params(), free_raw_data = FALSE
  )
  dval <- lightgbm::lgb.Dataset(
    data = x_va, label = as.integer(y_va == "christian"),
    params = make_lgb_dataset_params(),
    reference = dtrain, free_raw_data = FALSE
  )
  fit <- lightgbm::lgb.train(
    params = params, data = dtrain, nrounds = as.integer(max_rounds),
    valids = list(valid = dval),
    early_stopping_rounds = as.integer(early_rounds), verbose = -1
  )
  best_iter <- fit$best_iter
  if (is.null(best_iter) || !is.finite(best_iter) || best_iter < 1)
    best_iter <- as.integer(max_rounds)
  list(model = fit, best_iter = as.integer(best_iter))
}

tune_lgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- lgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path) %||%
    list(done = rep(FALSE, nrow(grid)))
  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path) %||%
      list(oof_prob = rep(NA_real_, length(y)),
           best_iters = rep(NA_integer_, length(folds)),
           done = rep(FALSE, length(folds)))
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("lgb_tuning_fold",
                   paste(prefix, "candidate", g, "of", nrow(grid),
                         "fold", j, "of", length(folds)))
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      lgb_params_fold <- make_lgb_params(pars_row, seed = seed + 1000L * g + j)
      result <- fit_lgb_early_stop(
        x_tr = x[tr, , drop = FALSE], y_tr = y[tr],
        x_va = x[va, , drop = FALSE], y_va = y[va],
        params = lgb_params_fold,
        max_rounds = lgb_max_rounds, early_rounds = early_stop_rounds
      )
      cand_state$oof_prob[va] <- predict_lgb_prob(result$model,
                                                  x[va, , drop = FALSE],
                                                  num_iteration = result$best_iter)
      cand_state$best_iters[j] <- result$best_iter
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }
  tuning_rows <- lapply(seq_len(nrow(grid)), function(g) {
    pars_row <- grid[g, , drop = FALSE]
    cs <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    pars_row %>% mutate(
      log_loss = log_loss_safe(y, cs$oof_prob),
      auc = auc_safe(y, cs$oof_prob),
      best_nrounds = max(1L, as.integer(round(median(cs$best_iters, na.rm = TRUE))))
    )
  })
  tuning <- bind_rows(tuning_rows) %>% arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_lgb_params(best, seed = seed + 9999L),
       tuning = tuning)
}

cv_lgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123,
                               prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("lgb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE], y = y[tr],
      params = modifyList(params,
                          list(seed = seed + j,
                               feature_fraction_seed = seed + j,
                               bagging_seed = seed + j)),
      nrounds = nrounds
    )
    state$scores[va] <- predict_lgb_score(fit, x[va, , drop = FALSE],
                                           num_iteration = nrounds)
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

cv_nb_scores_ckpt <- function(x_dfm, y, k = 5, seed = 123,
                              prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("nb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_nb_model(x_dfm[tr, ], y[tr])
    state$scores[va] <- predict_nb_score(fit, x_dfm[va, ])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

## Nested cross-validation (outer loop)

```{r setup-run}
check_binary_y(df_labeled$manual_label, "labeled data")

lab <- df_labeled %>% arrange(row_id)
out <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos     <- match(lab$row_id, out$row_id)
n_lab       <- nrow(lab)
outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)

init_run_state <- function(n_lab) {
  list(
    outer_fold_id = integer(n_lab),

    nb_oof        = rep(NA_real_, n_lab),
    enet_oof      = rep(NA_real_, n_lab),
    svm_oof       = rep(NA_real_, n_lab),
    xgb_oof       = rep(NA_real_, n_lab),
    lgb_oof       = rep(NA_real_, n_lab),
    tfidf_mlp_oof = rep(NA_real_, n_lab),
    word_cnn_oof  = rep(NA_real_, n_lab),

    nb_score_oof        = rep(NA_real_, n_lab),
    enet_score_oof      = rep(NA_real_, n_lab),
    svm_score_oof       = rep(NA_real_, n_lab),
    xgb_score_oof       = rep(NA_real_, n_lab),
    lgb_score_oof       = rep(NA_real_, n_lab),
    tfidf_mlp_score_oof = rep(NA_real_, n_lab),
    word_cnn_score_oof  = rep(NA_real_, n_lab),

    nb_pred_youden_oof        = rep(NA_character_, n_lab),
    enet_pred_youden_oof      = rep(NA_character_, n_lab),
    svm_pred_youden_oof       = rep(NA_character_, n_lab),
    xgb_pred_youden_oof       = rep(NA_character_, n_lab),
    lgb_pred_youden_oof       = rep(NA_character_, n_lab),
    tfidf_mlp_pred_youden_oof = rep(NA_character_, n_lab),
    word_cnn_pred_youden_oof  = rep(NA_character_, n_lab),

    nb_pred_f1_oof        = rep(NA_character_, n_lab),
    enet_pred_f1_oof      = rep(NA_character_, n_lab),
    svm_pred_f1_oof       = rep(NA_character_, n_lab),
    xgb_pred_f1_oof       = rep(NA_character_, n_lab),
    lgb_pred_f1_oof       = rep(NA_character_, n_lab),
    tfidf_mlp_pred_f1_oof = rep(NA_character_, n_lab),
    word_cnn_pred_f1_oof  = rep(NA_character_, n_lab),

    outer_threshold_df = tibble::tibble(),
    fold_metric_df     = tibble::tibble(),
    selected_outer_df  = tibble::tibble(),
    model_done         = list(),
    full_done          = list()
  )
}

mark_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  state[[bucket]][[key]] <- TRUE
  state
}
is_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  isTRUE(state[[bucket]][[key]])
}

state_path <- file.path(checkpoint_root, "run_state.rds")
state <- load_rds_or_null(state_path)
if (is.null(state) || length(state$nb_oof) != n_lab) {
  state <- init_run_state(n_lab)
  save_rds_atomic(state, state_path)
}
```

```{r outer-loop}
for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start", paste("fold", i, "of", length(outer_folds)))
  va   <- outer_folds[[i]]
  tr   <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]

  outer_dir <- file.path(checkpoint_root, sprintf("outer_fold_%02d", i))
  ensure_dir(outer_dir)

  # Re-fit the vectorizer on the OUTER TRAINING SET ONLY — no test leakage.
  vec_fold  <- fit_vectorizer(lab$text[tr], latin_stops)
  x_tr_dfm  <- transform_counts_dfm(lab$text[tr], vec_fold, latin_stops)
  x_va_dfm  <- transform_counts_dfm(lab$text[va], vec_fold, latin_stops)
  x_tr      <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
  x_va      <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)

  state$outer_fold_id[va] <- i
  save_rds_atomic(state, state_path)

  # -- Naive Bayes --------------------------------------------------------------------
  key <- paste0("outer_", i, "_naive_bayes")
  if (!is_done(state, key, "model_done")) {
    model_dir <- file.path(outer_dir, "naive_bayes")
    nb_train_scores <- cv_nb_scores_ckpt(
      x_tr_dfm, y_tr, k = k_folds_inner, seed = 1000 + i,
      prefix = paste("outer", i), ckpt_dir = file.path(model_dir, "calibration")
    )
    nb_platt <- fit_platt(nb_train_scores, y_tr)
    nb_fit   <- fit_nb_model(x_tr_dfm, y_tr)
    nb_va_scores   <- predict_nb_score(nb_fit, x_va_dfm)
    nb_train_prob  <- apply_platt(nb_platt, nb_train_scores)
    nb_va_prob     <- apply_platt(nb_platt, nb_va_scores)
    nb_th_y <- choose_threshold(y_tr, nb_train_prob, "youden")
    nb_th_f <- choose_threshold(y_tr, nb_train_prob, "f1")

    state$nb_oof[va]              <- nb_va_prob
    state$nb_score_oof[va]        <- nb_va_scores
    state$nb_pred_youden_oof[va]  <- ifelse(nb_va_prob >= nb_th_y,
                                             "christian", "non_christian")
    state$nb_pred_f1_oof[va]      <- ifelse(nb_va_prob >= nb_th_f,
                                             "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "naive_bayes",
                            threshold_youden = nb_th_y, threshold_f1 = nb_th_f)
    metric_row <- summarize_model_fold(y_va, nb_va_prob,
                                        state$nb_pred_youden_oof[va],
                                        state$nb_pred_f1_oof[va],
                                        "naive_bayes", i)
    selected_row <- tibble(outer_fold = i, model = "naive_bayes",
                            threshold_youden = nb_th_y, threshold_f1 = nb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row, c("outer_fold", "model"))
    state$fold_metric_df     <- upsert_rows(state$fold_metric_df,
                                             metric_row, c("outer_fold", "model"))
    state$selected_outer_df  <- upsert_rows(state$selected_outer_df,
                                             selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # -- Elastic net (replaces ridge) ----------------------------------------------------
  key <- paste0("outer_", i, "_elastic_net")
  if (!is_done(state, key, "model_done")) {
    model_dir <- file.path(outer_dir, "elastic_net")
    enet_tuned <- tune_enet_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 2000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    enet_train_scores <- cv_enet_scores_ckpt(
      x_tr, y_tr, alpha = enet_tuned$alpha, lambda = enet_tuned$lambda,
      k = k_folds_inner, seed = 2100 + i, prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    enet_platt <- fit_platt(enet_train_scores, y_tr)
    enet_fit   <- fit_enet_fixed(x_tr, y_tr,
                                  alpha = enet_tuned$alpha,
                                  lambda = enet_tuned$lambda)
    enet_va_scores  <- predict_enet_score(enet_fit, x_va)
    enet_train_prob <- apply_platt(enet_platt, enet_train_scores)
    enet_va_prob    <- apply_platt(enet_platt, enet_va_scores)
    enet_th_y <- choose_threshold(y_tr, enet_train_prob, "youden")
    enet_th_f <- choose_threshold(y_tr, enet_train_prob, "f1")

    state$enet_oof[va]             <- enet_va_prob
    state$enet_score_oof[va]       <- enet_va_scores
    state$enet_pred_youden_oof[va] <- ifelse(enet_va_prob >= enet_th_y,
                                              "christian", "non_christian")
    state$enet_pred_f1_oof[va]     <- ifelse(enet_va_prob >= enet_th_f,
                                              "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "elastic_net",
                            threshold_youden = enet_th_y,
                            threshold_f1 = enet_th_f)
    metric_row <- summarize_model_fold(y_va, enet_va_prob,
                                        state$enet_pred_youden_oof[va],
                                        state$enet_pred_f1_oof[va],
                                        "elastic_net", i)
    selected_row <- tibble(
      outer_fold = i, model = "elastic_net",
      alpha      = enet_tuned$alpha,
      lambda_1se = enet_tuned$lambda,
      lambda_min = enet_tuned$lambda_min,
      threshold_youden = enet_th_y, threshold_f1 = enet_th_f
    )

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row, c("outer_fold", "model"))
    state$fold_metric_df     <- upsert_rows(state$fold_metric_df,
                                             metric_row, c("outer_fold", "model"))
    state$selected_outer_df  <- upsert_rows(state$selected_outer_df,
                                             selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
    save_rds_atomic(list(tuning = enet_tuned$tuning,
                          thresholds = threshold_row,
                          metrics = metric_row),
                     file.path(model_dir, "result.rds"))
  }

  # -- Linear SVM ----------------------------------------------------------------------
  key <- paste0("outer_", i, "_svm_platt")
  if (!is_done(state, key, "model_done")) {
    model_dir <- file.path(outer_dir, "svm_platt")
    svm_tuned <- tune_svm_cost_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 3000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    svm_train_scores <- cv_svm_scores_ckpt(
      x_tr, y_tr, cost = svm_tuned$cost, k = k_folds_inner,
      seed = 3100 + i, prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    svm_platt <- fit_platt(svm_train_scores, y_tr)
    svm_fit   <- fit_svm_fixed(x_tr, y_tr, cost = svm_tuned$cost)
    svm_va_scores  <- predict_svm_score(svm_fit, x_va)
    svm_train_prob <- apply_platt(svm_platt, svm_train_scores)
    svm_va_prob    <- apply_platt(svm_platt, svm_va_scores)
    svm_th_y <- choose_threshold(y_tr, svm_train_prob, "youden")
    svm_th_f <- choose_threshold(y_tr, svm_train_prob, "f1")

    state$svm_oof[va]             <- svm_va_prob
    state$svm_score_oof[va]       <- svm_va_scores
    state$svm_pred_youden_oof[va] <- ifelse(svm_va_prob >= svm_th_y,
                                             "christian", "non_christian")
    state$svm_pred_f1_oof[va]     <- ifelse(svm_va_prob >= svm_th_f,
                                             "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "svm_platt",
                            threshold_youden = svm_th_y,
                            threshold_f1 = svm_th_f)
    metric_row <- summarize_model_fold(y_va, svm_va_prob,
                                        state$svm_pred_youden_oof[va],
                                        state$svm_pred_f1_oof[va],
                                        "svm_platt", i)
    selected_row <- tibble(outer_fold = i, model = "svm_platt",
                            cost = svm_tuned$cost,
                            threshold_youden = svm_th_y,
                            threshold_f1 = svm_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row, c("outer_fold", "model"))
    state$fold_metric_df     <- upsert_rows(state$fold_metric_df,
                                             metric_row, c("outer_fold", "model"))
    state$selected_outer_df  <- upsert_rows(state$selected_outer_df,
                                             selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # -- XGBoost -------------------------------------------------------------------------
  key <- paste0("outer_", i, "_xgboost")
  if (!is_done(state, key, "model_done")) {
    model_dir <- file.path(outer_dir, "xgboost")
    xgb_tuned <- tune_xgb_model_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 4000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    xgb_nrounds <- as.integer(xgb_tuned$best$best_nrounds[1])
    xgb_train_scores <- cv_xgb_scores_ckpt(
      x_tr, y_tr, params = xgb_tuned$params, nrounds = xgb_nrounds,
      k = k_folds_inner, seed = 4100 + i, prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    xgb_platt <- fit_platt(xgb_train_scores, y_tr)
    xgb_fit   <- fit_xgb_fixed(x_tr, y_tr, params = xgb_tuned$params,
                                nrounds = xgb_nrounds, seed = 4200 + i)
    xgb_va_scores  <- predict_xgb_score(xgb_fit, x_va)
    xgb_train_prob <- apply_platt(xgb_platt, xgb_train_scores)
    xgb_va_prob    <- apply_platt(xgb_platt, xgb_va_scores)
    xgb_th_y <- choose_threshold(y_tr, xgb_train_prob, "youden")
    xgb_th_f <- choose_threshold(y_tr, xgb_train_prob, "f1")

    state$xgb_oof[va]             <- xgb_va_prob
    state$xgb_score_oof[va]       <- xgb_va_scores
    state$xgb_pred_youden_oof[va] <- ifelse(xgb_va_prob >= xgb_th_y,
                                             "christian", "non_christian")
    state$xgb_pred_f1_oof[va]     <- ifelse(xgb_va_prob >= xgb_th_f,
                                             "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "xgboost",
                            threshold_youden = xgb_th_y,
                            threshold_f1 = xgb_th_f)
    metric_row <- summarize_model_fold(y_va, xgb_va_prob,
                                        state$xgb_pred_youden_oof[va],
                                        state$xgb_pred_f1_oof[va],
                                        "xgboost", i)
    selected_row <- xgb_tuned$best %>%
      mutate(outer_fold = i, model = "xgboost",
             threshold_youden = xgb_th_y, threshold_f1 = xgb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row, c("outer_fold", "model"))
    state$fold_metric_df     <- upsert_rows(state$fold_metric_df,
                                             metric_row, c("outer_fold", "model"))
    state$selected_outer_df  <- upsert_rows(state$selected_outer_df,
                                             selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # -- LightGBM ------------------------------------------------------------------------
  key <- paste0("outer_", i, "_lightgbm")
  if (!is_done(state, key, "model_done")) {
    model_dir <- file.path(outer_dir, "lightgbm")
    lgb_tuned <- tune_lgb_model_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 5000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning")
    )
    lgb_nrounds <- as.integer(lgb_tuned$best$best_nrounds[1])
    lgb_train_scores <- cv_lgb_scores_ckpt(
      x_tr, y_tr, params = lgb_tuned$params, nrounds = lgb_nrounds,
      k = k_folds_inner, seed = 5100 + i, prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration")
    )
    lgb_platt <- fit_platt(lgb_train_scores, y_tr)
    lgb_fit   <- fit_lgb_fixed(x_tr, y_tr, params = lgb_tuned$params,
                                nrounds = lgb_nrounds)
    lgb_va_scores  <- predict_lgb_score(lgb_fit, x_va, num_iteration = lgb_nrounds)
    lgb_train_prob <- apply_platt(lgb_platt, lgb_train_scores)
    lgb_va_prob    <- apply_platt(lgb_platt, lgb_va_scores)
    lgb_th_y <- choose_threshold(y_tr, lgb_train_prob, "youden")
    lgb_th_f <- choose_threshold(y_tr, lgb_train_prob, "f1")

    state$lgb_oof[va]             <- lgb_va_prob
    state$lgb_score_oof[va]       <- lgb_va_scores
    state$lgb_pred_youden_oof[va] <- ifelse(lgb_va_prob >= lgb_th_y,
                                             "christian", "non_christian")
    state$lgb_pred_f1_oof[va]     <- ifelse(lgb_va_prob >= lgb_th_f,
                                             "christian", "non_christian")

    threshold_row <- tibble(outer_fold = i, model = "lightgbm",
                            threshold_youden = lgb_th_y,
                            threshold_f1 = lgb_th_f)
    metric_row <- summarize_model_fold(y_va, lgb_va_prob,
                                        state$lgb_pred_youden_oof[va],
                                        state$lgb_pred_f1_oof[va],
                                        "lightgbm", i)
    selected_row <- lgb_tuned$best %>%
      mutate(outer_fold = i, model = "lightgbm",
             threshold_youden = lgb_th_y, threshold_f1 = lgb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row, c("outer_fold", "model"))
    state$fold_metric_df     <- upsert_rows(state$fold_metric_df,
                                             metric_row, c("outer_fold", "model"))
    state$selected_outer_df  <- upsert_rows(state$selected_outer_df,
                                             selected_row, c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }
}

log_progress("outer_loop_done", "all outer folds complete")
```

## Full-data deployment fits

```{r full-fits}
full_dir <- file.path(checkpoint_root, "full_fit")
ensure_dir(full_dir)

vec_full       <- fit_vectorizer(lab$text, latin_stops)
x_lab_dfm_full <- transform_counts_dfm(lab$text, vec_full, latin_stops)
x_all_dfm_full <- transform_counts_dfm(out$text, vec_full, latin_stops)
x_lab_full     <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all_full     <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

full_predictions <- tibble(row_id = out$row_id, id = out$id,
                           manual_label = out$manual_label)

# -- Naive Bayes --------------------------------------------------------------------
nb_cv_scores  <- cv_nb_scores_ckpt(
  x_lab_dfm_full, lab$manual_label, k = k_folds_inner, seed = 8001,
  prefix = "full", ckpt_dir = file.path(full_dir, "naive_bayes", "calibration")
)
nb_platt_full <- fit_platt(nb_cv_scores, lab$manual_label)
nb_lab_prob_cf <- apply_platt(nb_platt_full, nb_cv_scores)
nb_fit_full   <- fit_nb_model(x_lab_dfm_full, lab$manual_label)
nb_all_scores <- predict_nb_score(nb_fit_full, x_all_dfm_full)
nb_all_prob   <- apply_platt(nb_platt_full, nb_all_scores)
nb_deploy_th_y <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "youden")
nb_deploy_th_f <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "f1")

# -- Elastic net --------------------------------------------------------------------
enet_tuned_full <- tune_enet_ckpt(
  x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8002,
  prefix = "full", ckpt_dir = file.path(full_dir, "elastic_net", "tuning")
)
enet_cv_scores <- cv_enet_scores_ckpt(
  x_lab_full, lab$manual_label,
  alpha = enet_tuned_full$alpha, lambda = enet_tuned_full$lambda,
  k = k_folds_inner, seed = 8102, prefix = "full",
  ckpt_dir = file.path(full_dir, "elastic_net", "calibration")
)
enet_platt_full <- fit_platt(enet_cv_scores, lab$manual_label)
enet_lab_score_cf <- enet_cv_scores
enet_lab_prob_cf  <- apply_platt(enet_platt_full, enet_cv_scores)
enet_fit_full <- fit_enet_fixed(x_lab_full, lab$manual_label,
                                 alpha = enet_tuned_full$alpha,
                                 lambda = enet_tuned_full$lambda)
enet_all_scores <- predict_enet_score(enet_fit_full, x_all_full)
enet_all_prob   <- apply_platt(enet_platt_full, enet_all_scores)
enet_deploy_th_y <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "youden")
enet_deploy_th_f <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "f1")

# -- SVM ---------------------------------------------------------------------------
svm_tuned_full <- tune_svm_cost_ckpt(
  x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8003,
  prefix = "full", ckpt_dir = file.path(full_dir, "svm_platt", "tuning")
)
svm_cv_scores <- cv_svm_scores_ckpt(
  x_lab_full, lab$manual_label, cost = svm_tuned_full$cost,
  k = k_folds_inner, seed = 8103, prefix = "full",
  ckpt_dir = file.path(full_dir, "svm_platt", "calibration")
)
svm_platt_full <- fit_platt(svm_cv_scores, lab$manual_label)
svm_lab_prob_cf <- apply_platt(svm_platt_full, svm_cv_scores)
svm_fit_full  <- fit_svm_fixed(x_lab_full, lab$manual_label, cost = svm_tuned_full$cost)
svm_all_scores <- predict_svm_score(svm_fit_full, x_all_full)
svm_all_prob   <- apply_platt(svm_platt_full, svm_all_scores)
svm_deploy_th_y <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "youden")
svm_deploy_th_f <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "f1")

# -- XGBoost -----------------------------------------------------------------------
xgb_tuned_full <- tune_xgb_model_ckpt(
  x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8004,
  prefix = "full", ckpt_dir = file.path(full_dir, "xgboost", "tuning")
)
xgb_nrounds_full <- as.integer(xgb_tuned_full$best$best_nrounds[1])
xgb_cv_scores <- cv_xgb_scores_ckpt(
  x_lab_full, lab$manual_label, params = xgb_tuned_full$params,
  nrounds = xgb_nrounds_full, k = k_folds_inner, seed = 8104,
  prefix = "full", ckpt_dir = file.path(full_dir, "xgboost", "calibration")
)
xgb_platt_full <- fit_platt(xgb_cv_scores, lab$manual_label)
xgb_lab_prob_cf <- apply_platt(xgb_platt_full, xgb_cv_scores)
xgb_fit_full  <- fit_xgb_fixed(x_lab_full, lab$manual_label,
                                params = xgb_tuned_full$params,
                                nrounds = xgb_nrounds_full, seed = 8204)
xgb_all_scores <- predict_xgb_score(xgb_fit_full, x_all_full)
xgb_all_prob   <- apply_platt(xgb_platt_full, xgb_all_scores)
xgb_deploy_th_y <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "youden")
xgb_deploy_th_f <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "f1")

# -- LightGBM ----------------------------------------------------------------------
lgb_tuned_full <- tune_lgb_model_ckpt(
  x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8005,
  prefix = "full", ckpt_dir = file.path(full_dir, "lightgbm", "tuning")
)
lgb_nrounds_full <- as.integer(lgb_tuned_full$best$best_nrounds[1])
lgb_cv_scores <- cv_lgb_scores_ckpt(
  x_lab_full, lab$manual_label, params = lgb_tuned_full$params,
  nrounds = lgb_nrounds_full, k = k_folds_inner, seed = 8105,
  prefix = "full", ckpt_dir = file.path(full_dir, "lightgbm", "calibration")
)
lgb_platt_full <- fit_platt(lgb_cv_scores, lab$manual_label)
lgb_lab_prob_cf <- apply_platt(lgb_platt_full, lgb_cv_scores)
lgb_fit_full  <- fit_lgb_fixed(x_lab_full, lab$manual_label,
                                params = lgb_tuned_full$params,
                                nrounds = lgb_nrounds_full)
lgb_all_scores <- predict_lgb_score(lgb_fit_full, x_all_full,
                                     num_iteration = lgb_nrounds_full)
lgb_all_prob   <- apply_platt(lgb_platt_full, lgb_all_scores)
lgb_deploy_th_y <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "youden")
lgb_deploy_th_f <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "f1")

log_progress("full_fits_done", "all deployment models trained")
```

## Assemble per-document predictions

```{r assemble-predictions}
# For labeled documents, overwrite the full-fit predictions with the cross-fitted
# ones so no document is evaluated on a model that has already seen it.
prob_nb   <- nb_all_prob;   score_nb_logit   <- nb_all_scores
prob_enet <- enet_all_prob; score_enet_logit <- enet_all_scores
prob_svm  <- svm_all_prob;  score_svm_margin <- svm_all_scores
prob_xgb  <- xgb_all_prob;  score_xgb_margin <- xgb_all_scores
prob_lgb  <- lgb_all_prob;  score_lgb_margin <- lgb_all_scores

prob_nb[lab_pos]    <- nb_lab_prob_cf
prob_enet[lab_pos]  <- enet_lab_prob_cf
prob_svm[lab_pos]   <- svm_lab_prob_cf
prob_xgb[lab_pos]   <- xgb_lab_prob_cf
prob_lgb[lab_pos]   <- lgb_lab_prob_cf

score_enet_logit[lab_pos] <- enet_lab_score_cf

deployment_thresholds <- tibble(
  model = c("naive_bayes", "elastic_net", "svm_platt", "xgboost", "lightgbm"),
  threshold_youden = c(nb_deploy_th_y, enet_deploy_th_y, svm_deploy_th_y,
                       xgb_deploy_th_y, lgb_deploy_th_y),
  threshold_f1     = c(nb_deploy_th_f, enet_deploy_th_f, svm_deploy_th_f,
                       xgb_deploy_th_f, lgb_deploy_th_f),
  threshold_scope  = "full_labeled_cross_fitted_calibrated"
)

get_threshold <- function(model_name, kind = c("youden", "f1")) {
  kind <- match.arg(kind)
  deployment_thresholds %>%
    filter(model == model_name) %>%
    pull(if (kind == "youden") threshold_youden else threshold_f1) %>%
    `[[`(1)
}

out <- out %>%
  mutate(
    prediction_source = if_else(!is.na(manual_label),
                                 "cross_fitted_labeled",
                                 "full_fit_unlabeled"),

    prob_nb = safe_prob(prob_nb), score_nb_logit = score_nb_logit,
    label_nb_youden = if_else(prob_nb >= get_threshold("naive_bayes", "youden"),
                               "christian", "non_christian"),
    label_nb_f1     = if_else(prob_nb >= get_threshold("naive_bayes", "f1"),
                               "christian", "non_christian"),

    prob_enet = safe_prob(prob_enet), score_enet_logit = score_enet_logit,
    label_enet_youden = if_else(prob_enet >= get_threshold("elastic_net", "youden"),
                                 "christian", "non_christian"),
    label_enet_f1     = if_else(prob_enet >= get_threshold("elastic_net", "f1"),
                                 "christian", "non_christian"),

    prob_svm = safe_prob(prob_svm), score_svm_margin = score_svm_margin,
    score_svm_logit = safe_logit(prob_svm),
    label_svm_youden = if_else(prob_svm >= get_threshold("svm_platt", "youden"),
                                "christian", "non_christian"),
    label_svm_f1     = if_else(prob_svm >= get_threshold("svm_platt", "f1"),
                                "christian", "non_christian"),

    prob_xgb = safe_prob(prob_xgb), score_xgb_margin = score_xgb_margin,
    score_xgb_logit = safe_logit(prob_xgb),
    label_xgb_youden = if_else(prob_xgb >= get_threshold("xgboost", "youden"),
                                "christian", "non_christian"),
    label_xgb_f1     = if_else(prob_xgb >= get_threshold("xgboost", "f1"),
                                "christian", "non_christian"),

    prob_lgb = safe_prob(prob_lgb), score_lgb_margin = score_lgb_margin,
    score_lgb_logit = safe_logit(prob_lgb),
    label_lgb_youden = if_else(prob_lgb >= get_threshold("lightgbm", "youden"),
                                "christian", "non_christian"),
    label_lgb_f1     = if_else(prob_lgb >= get_threshold("lightgbm", "f1"),
                                "christian", "non_christian")
  )
```

## OOF metrics, outputs, and elastic-net feature report

```{r metrics-and-outputs}
model_metrics_oof <- bind_rows(
  summarize_model_oof(lab$manual_label, state$nb_oof,
                      state$nb_pred_youden_oof, state$nb_pred_f1_oof,
                      "naive_bayes"),
  summarize_model_oof(lab$manual_label, state$enet_oof,
                      state$enet_pred_youden_oof, state$enet_pred_f1_oof,
                      "elastic_net"),
  summarize_model_oof(lab$manual_label, state$svm_oof,
                      state$svm_pred_youden_oof, state$svm_pred_f1_oof,
                      "svm_platt"),
  summarize_model_oof(lab$manual_label, state$xgb_oof,
                      state$xgb_pred_youden_oof, state$xgb_pred_f1_oof,
                      "xgboost"),
  summarize_model_oof(lab$manual_label, state$lgb_oof,
                      state$lgb_pred_youden_oof, state$lgb_pred_f1_oof,
                      "lightgbm")
)

oof_eval_predictions <- tibble(
  id = lab$id, row_id = lab$row_id, manual_label = lab$manual_label,
  outer_fold = state$outer_fold_id,

  prob_nb   = state$nb_oof,   score_nb_logit   = state$nb_score_oof,
  label_nb_youden   = state$nb_pred_youden_oof,
  label_nb_f1       = state$nb_pred_f1_oof,

  prob_enet = state$enet_oof, score_enet_logit = state$enet_score_oof,
  label_enet_youden = state$enet_pred_youden_oof,
  label_enet_f1     = state$enet_pred_f1_oof,

  prob_svm  = state$svm_oof,  score_svm_margin = state$svm_score_oof,
  label_svm_youden  = state$svm_pred_youden_oof,
  label_svm_f1      = state$svm_pred_f1_oof,

  prob_xgb  = state$xgb_oof,  score_xgb_margin = state$xgb_score_oof,
  label_xgb_youden  = state$xgb_pred_youden_oof,
  label_xgb_f1      = state$xgb_pred_f1_oof,

  prob_lgb  = state$lgb_oof,  score_lgb_margin = state$lgb_score_oof,
  label_lgb_youden  = state$lgb_pred_youden_oof,
  label_lgb_f1      = state$lgb_pred_f1_oof
)

selected_full <- bind_rows(
  tibble(model = "naive_bayes"),
  tibble(model = "elastic_net",
         alpha = enet_tuned_full$alpha,
         lambda_1se = enet_tuned_full$lambda,
         lambda_min = enet_tuned_full$lambda_min),
  tibble(model = "svm_platt", cost = svm_tuned_full$cost),
  xgb_tuned_full$best %>% mutate(model = "xgboost"),
  lgb_tuned_full$best %>% mutate(model = "lightgbm")
)

dir.create(results_path, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(out,                 file.path(results_path,
                                                 "corpus_600ad_model_outputs_oof.csv"))
readr::write_csv(deployment_thresholds, file.path(results_path,
                                                   "model_thresholds_deployment.csv"))
readr::write_csv(state$outer_threshold_df,
                 file.path(results_path, "model_thresholds_outer.csv"))
readr::write_csv(model_metrics_oof,
                 file.path(results_path, "model_metrics_oof.csv"))
readr::write_csv(state$fold_metric_df,
                 file.path(results_path, "model_metrics_oof_by_fold.csv"))
readr::write_csv(oof_eval_predictions,
                 file.path(results_path, "oof_eval_predictions.csv"))
readr::write_csv(state$selected_outer_df,
                 file.path(results_path, "model_hyperparameters_outer.csv"))
readr::write_csv(selected_full,
                 file.path(results_path, "model_hyperparameters_full.csv"))
```

```{r enet-feature-report}
# Interpretability check: which tokens does the elastic net actually rely on? If the
# top features are function words, genuine theological/philosophical vocabulary, or
# discourse markers, interpretation is sound. If they are still proper nouns or
# cult-specific markers, either the blocklist needs more entries or the corpus is
# contaminated with metadata tokens.

coef_mat <- as.matrix(coef(enet_fit_full, s = enet_tuned_full$lambda))
coef_df  <- tibble(
  feature     = rownames(coef_mat),
  coefficient = as.numeric(coef_mat[, 1])
) %>%
  filter(feature != "(Intercept)", coefficient != 0) %>%
  arrange(desc(abs(coefficient))) %>%
  mutate(direction = if_else(coefficient > 0, "christian", "non_christian"))

n_nonzero <- nrow(coef_df)
n_total   <- nrow(coef_mat) - 1

cat(sprintf("\nElastic net: alpha = %s, lambda = %.6f\n",
            enet_tuned_full$alpha, enet_tuned_full$lambda))
cat(sprintf("Non-zero features: %d / %d (%.1f%%)\n\n",
            n_nonzero, n_total, 100 * n_nonzero / n_total))

cat("Top 30 features pushing toward 'christian':\n")
print(coef_df %>% filter(direction == "christian") %>% head(30), n = 30)

cat("\nTop 30 features pushing toward 'non_christian':\n")
print(coef_df %>% filter(direction == "non_christian") %>% head(30), n = 30)

readr::write_csv(coef_df, file.path(results_path, "enet_coefficients.csv"))

cat("\nOOF metrics:\n")
print(model_metrics_oof %>%
        select(model, auc, log_loss, brier, f1_youden, mcc_youden))

log_progress("done", "pipeline complete")
```




# Patch blocks to add non-transformer deep models in R

These are the two R additions I would wire in first:

- `tfidf_mlp`: a dense neural net on TF-IDF built inside `keras3::layer_text_vectorization(output_mode = "tf_idf")`
- `word_cnn`: a token-sequence CNN built in `keras3`

I am **not** adding char-CNN here. It is possible, but it is the first one most likely to become slow and annoying in your current checkpointed nested-CV setup.

---

## 1) Add to the setup chunk

Paste these lines into `setup-paths` right after your existing libraries.

```{r}
library(keras3)
library(tensorflow)

keras3::use_backend("tensorflow")
keras3::config_disable_interactive_logging()
```

```{r}

library(dplyr)
library(tidyr)
library(purrr)
library(rsample)
library(yardstick)
library(keras3)
library(tensorflow)

keras3::clear_session()
gc()
```

```{r}
run_tfidf_mlp <- TRUE
run_word_cnn  <- TRUE

keras_batch_size <- 32L
keras_patience   <- 3L
keras_val_frac   <- 0.1

# Keep these modest at first. Once the code runs end-to-end, widen the grids.
tfidf_mlp_grid <- tidyr::crossing(
  max_tokens   = c(5000L, 10000L),
  hidden_units = c(128L, 256L),
  dropout      = c(0.2, 0.4),
  lr           = c(1e-3, 5e-4),
  epochs       = c(10L, 15L)
) %>%
  dplyr::mutate(candidate_id = dplyr::row_number()) %>%
  dplyr::select(candidate_id, dplyr::everything())

word_cnn_grid <- tidyr::crossing(
  max_tokens  = c(10000L, 20000L),
  seq_length  = c(750L, 1500L),
  embed_dim   = c(100L, 200L),
  filters     = c(128L),
  kernel_size = c(5L, 7L),
  dense_units = c(128L),
  dropout     = c(0.3, 0.5),
  lr          = c(1e-3),
  epochs      = c(10L, 15L)
) %>%
  dplyr::mutate(candidate_id = dplyr::row_number()) %>%
  dplyr::select(candidate_id, dplyr::everything())

```



## 2) Add helper functions

Paste this as a new section after your current model-wrapper chunks.

```{r}
# -----------------------------------------------------------------------------
# Neural text helpers
# -----------------------------------------------------------------------------

as_text_input <- function(text) {
  matrix(as.character(text), ncol = 1)
}

prep_text_for_neural <- function(text_data, stop_list) {
  toks <- quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_tolower() %>%
    quanteda::tokens_remove(pattern = stop_list, valuetype = "fixed")

  vapply(
    seq_along(toks),
    function(i) paste(as.character(toks[[i]]), collapse = " "),
    FUN.VALUE = character(1)
  )
}

build_tfidf_mlp_model <- function(train_text, cfg) {
  text_in <- keras3::layer_input(shape = c(1), dtype = "string", name = "text")

  vec <- keras3::layer_text_vectorization(
    max_tokens   = as.integer(cfg$max_tokens),
    standardize  = "lower_and_strip_punctuation",
    split        = "whitespace",
    output_mode  = "tf_idf",
    sparse       = FALSE,
    pad_to_max_tokens = TRUE
  )
  keras3::adapt(vec, as_text_input(train_text))

  x <- vec(text_in) %>%
    keras3::layer_dense(units = as.integer(cfg$hidden_units), activation = "relu") %>%
    keras3::layer_dropout(rate = as.numeric(cfg$dropout)) %>%
    keras3::layer_dense(units = 1)

  model <- keras3::keras_model(text_in, x)
  model %>% keras3::compile(
    optimizer = keras3::optimizer_adam(learning_rate = as.numeric(cfg$lr)),
    loss = keras3::loss_binary_crossentropy(from_logits = TRUE)
  )
  model
}

build_word_cnn_model <- function(train_text, cfg) {
  text_in <- keras3::layer_input(shape = c(1), dtype = "string", name = "text")

  vec <- keras3::layer_text_vectorization(
    max_tokens             = as.integer(cfg$max_tokens),
    standardize            = "lower_and_strip_punctuation",
    split                  = "whitespace",
    output_mode            = "int",
    output_sequence_length = as.integer(cfg$seq_length)
  )
  keras3::adapt(vec, as_text_input(train_text))

  x <- vec(text_in) %>%
    keras3::layer_embedding(
      input_dim  = as.integer(cfg$max_tokens + 2L),
      output_dim = as.integer(cfg$embed_dim)
    ) %>%
    keras3::layer_conv_1d(
      filters     = as.integer(cfg$filters),
      kernel_size = as.integer(cfg$kernel_size),
      activation  = "relu"
    ) %>%
    keras3::layer_global_max_pooling_1d() %>%
    keras3::layer_dense(units = as.integer(cfg$dense_units), activation = "relu") %>%
    keras3::layer_dropout(rate = as.numeric(cfg$dropout)) %>%
    keras3::layer_dense(units = 1)

  model <- keras3::keras_model(text_in, x)
  model %>% keras3::compile(
    optimizer = keras3::optimizer_adam(learning_rate = as.numeric(cfg$lr)),
    loss = keras3::loss_binary_crossentropy(from_logits = TRUE)
  )
  model
}

fit_predict_keras_text_scores <- function(train_text, y_train, test_text,
                                          build_fun, cfg, seed = 123) {
  keras3::clear_session()
  keras3::set_random_seed(as.integer(seed))

  model <- build_fun(train_text, cfg)

  model %>% keras3::fit(
    x = as_text_input(train_text),
    y = as.numeric(y_train == "christian"),
    validation_split = keras_val_frac,
    epochs = as.integer(cfg$epochs),
    batch_size = as.integer(cfg$batch_size %||% keras_batch_size),
    callbacks = list(
      keras3::callback_early_stopping(
        monitor = "val_loss",
        patience = as.integer(keras_patience),
        restore_best_weights = TRUE
      )
    ),
    verbose = 0
  )

  out <- as.numeric(model %>% keras3::predict(as_text_input(test_text), verbose = 0))
  keras3::clear_session()
  out
}

cv_keras_text_scores_ckpt <- function(text, y, build_fun, cfg,
                                      k = 5, seed = 123,
                                      prefix = "", ckpt_dir,
                                      model_tag = "keras_text") {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path) %||%
    list(scores = rep(NA_real_, length(y)),
         done   = rep(FALSE, length(folds)))

  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress(
      paste0(model_tag, "_calibration_fold"),
      paste(prefix, "fold", j, "of", length(folds))
    )
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    state$scores[va] <- fit_predict_keras_text_scores(
      train_text = text[tr],
      y_train    = y[tr],
      test_text  = text[va],
      build_fun  = build_fun,
      cfg        = cfg,
      seed       = seed + j
    )
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }

  state$scores
}

tune_keras_text_model_ckpt <- function(text, y, grid, build_fun,
                                       inner_k = 5, seed = 123,
                                       prefix = "", ckpt_dir,
                                       model_tag = "keras_text") {
  ensure_dir(ckpt_dir)
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path) %||%
    list(done = rep(FALSE, nrow(grid)))

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path) %||%
      list(scores = rep(NA_real_, length(y)),
           done   = rep(FALSE, length(folds)))
    cfg <- as.list(grid[g, , drop = FALSE])

    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress(
        paste0(model_tag, "_tuning_fold"),
        paste(prefix, "candidate", g, "of", nrow(grid),
              "fold", j, "of", length(folds))
      )
      va <- folds[[j]]
      tr <- setdiff(seq_along(y), va)
      cand_state$scores[va] <- fit_predict_keras_text_scores(
        train_text = text[tr],
        y_train    = y[tr],
        test_text  = text[va],
        build_fun  = build_fun,
        cfg        = cfg,
        seed       = seed + 1000L * g + j
      )
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }

    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- lapply(seq_len(nrow(grid)), function(g) {
    cs <- readRDS(file.path(ckpt_dir, sprintf("candidate_%03d.rds", g)))
    bind_cols(
      grid[g, , drop = FALSE],
      tibble(
        log_loss   = log_loss_safe(y, plogis(cs$scores)),
        auc_margin = auc_margin_safe(y, cs$scores)
      )
    )
  })

  tuning <- bind_rows(tuning_rows) %>%
    arrange(log_loss, desc(auc_margin))

  list(best = tuning[1, , drop = FALSE], tuning = tuning)
}
```


## 3) Extend `init_run_state()`

Inside `init_run_state <- function(n_lab) { ... }`, add these fields.

```{r}
    tfidf_mlp_oof = rep(NA_real_, n_lab),
    word_cnn_oof  = rep(NA_real_, n_lab),

    tfidf_mlp_score_oof = rep(NA_real_, n_lab),
    word_cnn_score_oof  = rep(NA_real_, n_lab),

    tfidf_mlp_pred_youden_oof = rep(NA_character_, n_lab),
    word_cnn_pred_youden_oof  = rep(NA_character_, n_lab),

    tfidf_mlp_pred_f1_oof = rep(NA_character_, n_lab),
    word_cnn_pred_f1_oof  = rep(NA_character_, n_lab),
```

---

## 4) Add neural-preprocessed text inside the outer loop

Inside the outer loop, right after you create `x_tr` / `x_va`, add:

```{r}
  text_tr_neural <- prep_text_for_neural(lab$text[tr], latin_stops)
  text_va_neural <- prep_text_for_neural(lab$text[va], latin_stops)
```


## 5) Add the TF-IDF MLP outer-fold block

Paste this inside the outer loop after your LightGBM block.

```r
  # -- TF-IDF MLP ----------------------------------------------------------------------
  if (isTRUE(run_tfidf_mlp)) {
    key <- paste0("outer_", i, "_tfidf_mlp")
    if (!is_done(state, key, "model_done")) {
      model_dir <- file.path(outer_dir, "tfidf_mlp")

      tfidf_tuned <- tune_keras_text_model_ckpt(
        text = text_tr_neural,
        y = y_tr,
        grid = tfidf_mlp_grid,
        build_fun = build_tfidf_mlp_model,
        inner_k = k_folds_inner,
        seed = 6000 + i,
        prefix = paste("outer", i),
        ckpt_dir = file.path(model_dir, "tuning"),
        model_tag = "tfidf_mlp"
      )

      tfidf_cfg <- as.list(tfidf_tuned$best[1, , drop = FALSE])
      tfidf_train_scores <- cv_keras_text_scores_ckpt(
        text = text_tr_neural,
        y = y_tr,
        build_fun = build_tfidf_mlp_model,
        cfg = tfidf_cfg,
        k = k_folds_inner,
        seed = 6100 + i,
        prefix = paste("outer", i),
        ckpt_dir = file.path(model_dir, "calibration"),
        model_tag = "tfidf_mlp"
      )

      tfidf_platt <- fit_platt(tfidf_train_scores, y_tr)
      tfidf_va_scores <- fit_predict_keras_text_scores(
        train_text = text_tr_neural,
        y_train = y_tr,
        test_text = text_va_neural,
        build_fun = build_tfidf_mlp_model,
        cfg = tfidf_cfg,
        seed = 6200 + i
      )
      tfidf_train_prob <- apply_platt(tfidf_platt, tfidf_train_scores)
      tfidf_va_prob    <- apply_platt(tfidf_platt, tfidf_va_scores)
      tfidf_th_y <- choose_threshold(y_tr, tfidf_train_prob, "youden")
      tfidf_th_f <- choose_threshold(y_tr, tfidf_train_prob, "f1")

      state$tfidf_mlp_oof[va]             <- tfidf_va_prob
      state$tfidf_mlp_score_oof[va]       <- tfidf_va_scores
      state$tfidf_mlp_pred_youden_oof[va] <- ifelse(tfidf_va_prob >= tfidf_th_y,
                                                    "christian", "non_christian")
      state$tfidf_mlp_pred_f1_oof[va]     <- ifelse(tfidf_va_prob >= tfidf_th_f,
                                                    "christian", "non_christian")

      threshold_row <- tibble(
        outer_fold = i,
        model = "tfidf_mlp",
        threshold_youden = tfidf_th_y,
        threshold_f1 = tfidf_th_f
      )
      metric_row <- summarize_model_fold(
        y_va,
        tfidf_va_prob,
        state$tfidf_mlp_pred_youden_oof[va],
        state$tfidf_mlp_pred_f1_oof[va],
        "tfidf_mlp",
        i
      )
      selected_row <- tfidf_tuned$best %>%
        mutate(
          outer_fold = i,
          model = "tfidf_mlp",
          threshold_youden = tfidf_th_y,
          threshold_f1 = tfidf_th_f
        )

      state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                              threshold_row,
                                              c("outer_fold", "model"))
      state$fold_metric_df <- upsert_rows(state$fold_metric_df,
                                          metric_row,
                                          c("outer_fold", "model"))
      state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                             selected_row,
                                             c("outer_fold", "model"))
      state <- mark_done(state, key, "model_done")
      save_rds_atomic(state, state_path)
    }
  }
```

---

## 6) Add the word-CNN outer-fold block

Paste this immediately after the TF-IDF MLP block.

```r
  # -- Word CNN ------------------------------------------------------------------------
  if (isTRUE(run_word_cnn)) {
    key <- paste0("outer_", i, "_word_cnn")
    if (!is_done(state, key, "model_done")) {
      model_dir <- file.path(outer_dir, "word_cnn")

      cnn_tuned <- tune_keras_text_model_ckpt(
        text = text_tr_neural,
        y = y_tr,
        grid = word_cnn_grid,
        build_fun = build_word_cnn_model,
        inner_k = k_folds_inner,
        seed = 7000 + i,
        prefix = paste("outer", i),
        ckpt_dir = file.path(model_dir, "tuning"),
        model_tag = "word_cnn"
      )

      cnn_cfg <- as.list(cnn_tuned$best[1, , drop = FALSE])
      cnn_train_scores <- cv_keras_text_scores_ckpt(
        text = text_tr_neural,
        y = y_tr,
        build_fun = build_word_cnn_model,
        cfg = cnn_cfg,
        k = k_folds_inner,
        seed = 7100 + i,
        prefix = paste("outer", i),
        ckpt_dir = file.path(model_dir, "calibration"),
        model_tag = "word_cnn"
      )

      cnn_platt <- fit_platt(cnn_train_scores, y_tr)
      cnn_va_scores <- fit_predict_keras_text_scores(
        train_text = text_tr_neural,
        y_train = y_tr,
        test_text = text_va_neural,
        build_fun = build_word_cnn_model,
        cfg = cnn_cfg,
        seed = 7200 + i
      )
      cnn_train_prob <- apply_platt(cnn_platt, cnn_train_scores)
      cnn_va_prob    <- apply_platt(cnn_platt, cnn_va_scores)
      cnn_th_y <- choose_threshold(y_tr, cnn_train_prob, "youden")
      cnn_th_f <- choose_threshold(y_tr, cnn_train_prob, "f1")

      state$word_cnn_oof[va]             <- cnn_va_prob
      state$word_cnn_score_oof[va]       <- cnn_va_scores
      state$word_cnn_pred_youden_oof[va] <- ifelse(cnn_va_prob >= cnn_th_y,
                                                   "christian", "non_christian")
      state$word_cnn_pred_f1_oof[va]     <- ifelse(cnn_va_prob >= cnn_th_f,
                                                   "christian", "non_christian")

      threshold_row <- tibble(
        outer_fold = i,
        model = "word_cnn",
        threshold_youden = cnn_th_y,
        threshold_f1 = cnn_th_f
      )
      metric_row <- summarize_model_fold(
        y_va,
        cnn_va_prob,
        state$word_cnn_pred_youden_oof[va],
        state$word_cnn_pred_f1_oof[va],
        "word_cnn",
        i
      )
      selected_row <- cnn_tuned$best %>%
        mutate(
          outer_fold = i,
          model = "word_cnn",
          threshold_youden = cnn_th_y,
          threshold_f1 = cnn_th_f
        )

      state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                              threshold_row,
                                              c("outer_fold", "model"))
      state$fold_metric_df <- upsert_rows(state$fold_metric_df,
                                          metric_row,
                                          c("outer_fold", "model"))
      state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                             selected_row,
                                             c("outer_fold", "model"))
      state <- mark_done(state, key, "model_done")
      save_rds_atomic(state, state_path)
    }
  }
```

---

## 7) Add neural text for the full-data section

Inside `full-fits`, right after `x_all_full`, add:

```r
text_lab_neural_full <- prep_text_for_neural(lab$text, latin_stops)
text_all_neural_full <- prep_text_for_neural(out$text, latin_stops)
```

---

## 8) Add full-data TF-IDF MLP fit

Paste this after your LightGBM full fit.

```r
# -- TF-IDF MLP ----------------------------------------------------------------------
if (isTRUE(run_tfidf_mlp)) {
  tfidf_tuned_full <- tune_keras_text_model_ckpt(
    text = text_lab_neural_full,
    y = lab$manual_label,
    grid = tfidf_mlp_grid,
    build_fun = build_tfidf_mlp_model,
    inner_k = k_folds_inner,
    seed = 9001,
    prefix = "full",
    ckpt_dir = file.path(full_dir, "tfidf_mlp", "tuning"),
    model_tag = "tfidf_mlp"
  )
  tfidf_cfg_full <- as.list(tfidf_tuned_full$best[1, , drop = FALSE])
  tfidf_cv_scores <- cv_keras_text_scores_ckpt(
    text = text_lab_neural_full,
    y = lab$manual_label,
    build_fun = build_tfidf_mlp_model,
    cfg = tfidf_cfg_full,
    k = k_folds_inner,
    seed = 9101,
    prefix = "full",
    ckpt_dir = file.path(full_dir, "tfidf_mlp", "calibration"),
    model_tag = "tfidf_mlp"
  )
  tfidf_platt_full <- fit_platt(tfidf_cv_scores, lab$manual_label)
  tfidf_lab_prob_cf <- apply_platt(tfidf_platt_full, tfidf_cv_scores)
  tfidf_all_scores <- fit_predict_keras_text_scores(
    train_text = text_lab_neural_full,
    y_train = lab$manual_label,
    test_text = text_all_neural_full,
    build_fun = build_tfidf_mlp_model,
    cfg = tfidf_cfg_full,
    seed = 9201
  )
  tfidf_all_prob <- apply_platt(tfidf_platt_full, tfidf_all_scores)
  tfidf_deploy_th_y <- choose_threshold(lab$manual_label, tfidf_lab_prob_cf, "youden")
  tfidf_deploy_th_f <- choose_threshold(lab$manual_label, tfidf_lab_prob_cf, "f1")
}
```

---

## 9) Add full-data word-CNN fit

Paste this after the TF-IDF MLP full-fit block.

```r
# -- Word CNN ------------------------------------------------------------------------
if (isTRUE(run_word_cnn)) {
  cnn_tuned_full <- tune_keras_text_model_ckpt(
    text = text_lab_neural_full,
    y = lab$manual_label,
    grid = word_cnn_grid,
    build_fun = build_word_cnn_model,
    inner_k = k_folds_inner,
    seed = 9002,
    prefix = "full",
    ckpt_dir = file.path(full_dir, "word_cnn", "tuning"),
    model_tag = "word_cnn"
  )
  cnn_cfg_full <- as.list(cnn_tuned_full$best[1, , drop = FALSE])
  cnn_cv_scores <- cv_keras_text_scores_ckpt(
    text = text_lab_neural_full,
    y = lab$manual_label,
    build_fun = build_word_cnn_model,
    cfg = cnn_cfg_full,
    k = k_folds_inner,
    seed = 9102,
    prefix = "full",
    ckpt_dir = file.path(full_dir, "word_cnn", "calibration"),
    model_tag = "word_cnn"
  )
  cnn_platt_full <- fit_platt(cnn_cv_scores, lab$manual_label)
  cnn_lab_prob_cf <- apply_platt(cnn_platt_full, cnn_cv_scores)
  cnn_all_scores <- fit_predict_keras_text_scores(
    train_text = text_lab_neural_full,
    y_train = lab$manual_label,
    test_text = text_all_neural_full,
    build_fun = build_word_cnn_model,
    cfg = cnn_cfg_full,
    seed = 9202
  )
  cnn_all_prob <- apply_platt(cnn_platt_full, cnn_all_scores)
  cnn_deploy_th_y <- choose_threshold(lab$manual_label, cnn_lab_prob_cf, "youden")
  cnn_deploy_th_f <- choose_threshold(lab$manual_label, cnn_lab_prob_cf, "f1")
}
```

---

## 10) Extend `deployment_thresholds`

Add these rows to the `deployment_thresholds <- tibble(...)` block.

```r
  model = c("naive_bayes", "elastic_net", "svm_platt", "xgboost", "lightgbm",
            if (isTRUE(run_tfidf_mlp)) "tfidf_mlp",
            if (isTRUE(run_word_cnn))  "word_cnn"),
  threshold_youden = c(nb_deploy_th_y, enet_deploy_th_y, svm_deploy_th_y,
                       xgb_deploy_th_y, lgb_deploy_th_y,
                       if (isTRUE(run_tfidf_mlp)) tfidf_deploy_th_y,
                       if (isTRUE(run_word_cnn))  cnn_deploy_th_y),
  threshold_f1 = c(nb_deploy_th_f, enet_deploy_th_f, svm_deploy_th_f,
                   xgb_deploy_th_f, lgb_deploy_th_f,
                   if (isTRUE(run_tfidf_mlp)) tfidf_deploy_th_f,
                   if (isTRUE(run_word_cnn))  cnn_deploy_th_f),
```

If you want to avoid messing with the existing vectorized construction, it is cleaner to just rebuild `deployment_thresholds` like this:

```r
deployment_thresholds <- bind_rows(
  tibble(model = "naive_bayes", threshold_youden = nb_deploy_th_y, threshold_f1 = nb_deploy_th_f),
  tibble(model = "elastic_net", threshold_youden = enet_deploy_th_y, threshold_f1 = enet_deploy_th_f),
  tibble(model = "svm_platt", threshold_youden = svm_deploy_th_y, threshold_f1 = svm_deploy_th_f),
  tibble(model = "xgboost", threshold_youden = xgb_deploy_th_y, threshold_f1 = xgb_deploy_th_f),
  tibble(model = "lightgbm", threshold_youden = lgb_deploy_th_y, threshold_f1 = lgb_deploy_th_f),
  if (isTRUE(run_tfidf_mlp)) tibble(model = "tfidf_mlp", threshold_youden = tfidf_deploy_th_y, threshold_f1 = tfidf_deploy_th_f),
  if (isTRUE(run_word_cnn)) tibble(model = "word_cnn", threshold_youden = cnn_deploy_th_y, threshold_f1 = cnn_deploy_th_f)
) %>%
  mutate(threshold_scope = "full_labeled_cross_fitted_calibrated")
```

---

## 11) Extend the assembled per-document predictions

Inside `assemble-predictions`, add these objects before `out <- out %>% mutate(...)`:

```r
if (isTRUE(run_tfidf_mlp)) {
  prob_tfidf_mlp  <- tfidf_all_prob
  score_tfidf_mlp <- tfidf_all_scores
  prob_tfidf_mlp[lab_pos] <- tfidf_lab_prob_cf
}

if (isTRUE(run_word_cnn)) {
  prob_word_cnn  <- cnn_all_prob
  score_word_cnn <- cnn_all_scores
  prob_word_cnn[lab_pos] <- cnn_lab_prob_cf
}
```

Then add these fields inside the `mutate(...)` call:

```r
    prob_tfidf_mlp = if (isTRUE(run_tfidf_mlp)) safe_prob(prob_tfidf_mlp) else NA_real_,
    score_tfidf_mlp_logit = if (isTRUE(run_tfidf_mlp)) score_tfidf_mlp else NA_real_,
    label_tfidf_mlp_youden = if (isTRUE(run_tfidf_mlp)) if_else(
      prob_tfidf_mlp >= get_threshold("tfidf_mlp", "youden"),
      "christian", "non_christian"
    ) else NA_character_,
    label_tfidf_mlp_f1 = if (isTRUE(run_tfidf_mlp)) if_else(
      prob_tfidf_mlp >= get_threshold("tfidf_mlp", "f1"),
      "christian", "non_christian"
    ) else NA_character_,

    prob_word_cnn = if (isTRUE(run_word_cnn)) safe_prob(prob_word_cnn) else NA_real_,
    score_word_cnn_logit = if (isTRUE(run_word_cnn)) score_word_cnn else NA_real_,
    label_word_cnn_youden = if (isTRUE(run_word_cnn)) if_else(
      prob_word_cnn >= get_threshold("word_cnn", "youden"),
      "christian", "non_christian"
    ) else NA_character_,
    label_word_cnn_f1 = if (isTRUE(run_word_cnn)) if_else(
      prob_word_cnn >= get_threshold("word_cnn", "f1"),
      "christian", "non_christian"
    ) else NA_character_,
```

---

## 12) Extend the OOF metrics block

In `model_metrics_oof <- bind_rows(...)`, add:

```r
  if (isTRUE(run_tfidf_mlp)) summarize_model_oof(
    lab$manual_label,
    state$tfidf_mlp_oof,
    state$tfidf_mlp_pred_youden_oof,
    state$tfidf_mlp_pred_f1_oof,
    "tfidf_mlp"
  ),
  if (isTRUE(run_word_cnn)) summarize_model_oof(
    lab$manual_label,
    state$word_cnn_oof,
    state$word_cnn_pred_youden_oof,
    state$word_cnn_pred_f1_oof,
    "word_cnn"
  )
```

In `oof_eval_predictions <- tibble(...)`, add:

```r
  prob_tfidf_mlp = if (isTRUE(run_tfidf_mlp)) state$tfidf_mlp_oof else NA_real_,
  score_tfidf_mlp_logit = if (isTRUE(run_tfidf_mlp)) state$tfidf_mlp_score_oof else NA_real_,
  label_tfidf_mlp_youden = if (isTRUE(run_tfidf_mlp)) state$tfidf_mlp_pred_youden_oof else NA_character_,
  label_tfidf_mlp_f1 = if (isTRUE(run_tfidf_mlp)) state$tfidf_mlp_pred_f1_oof else NA_character_,

  prob_word_cnn = if (isTRUE(run_word_cnn)) state$word_cnn_oof else NA_real_,
  score_word_cnn_logit = if (isTRUE(run_word_cnn)) state$word_cnn_score_oof else NA_real_,
  label_word_cnn_youden = if (isTRUE(run_word_cnn)) state$word_cnn_pred_youden_oof else NA_character_,
  label_word_cnn_f1 = if (isTRUE(run_word_cnn)) state$word_cnn_pred_f1_oof else NA_character_,
```

In `selected_full <- bind_rows(...)`, add:

```r
  if (isTRUE(run_tfidf_mlp)) tfidf_tuned_full$best %>% mutate(model = "tfidf_mlp"),
  if (isTRUE(run_word_cnn)) cnn_tuned_full$best %>% mutate(model = "word_cnn")
```

---

## 13) One-time install block if keras is not set up yet

Run this once outside the pipeline, not every time.

```r
install.packages("keras3")
keras3::install_keras(backend = "tensorflow")
```

---

## 14) What I would actually run first

Start with this:

```r
run_tfidf_mlp <- TRUE
run_word_cnn  <- FALSE
```

Once that works end-to-end and your checkpointing is behaving, turn on `word_cnn`.

If you try both at once before the first one runs cleanly, debugging will be annoying.


```



