---
title: "Latin word interactions (200ad)"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">
    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>
    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Latin word interactions in the 200ad corpus

## Data loading and cleaning

```{r}
library(tidyverse)
library(stringi)
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textstats)
library(Matrix)
library(ComplexHeatmap)
library(circlize)
library(tidygraph)
library(ggraph)
library(igraph)
library(glasso)
library(irlba)
library(ggrepel)
library(infotheo)
library(purrr)
library(tidyr)
library(dplyr)

data_path <- "../../data/processed_texts/200ad"
results_dir <- "../../results/word_interactions/200ad"

dir.create(results_dir, recursive = TRUE, showWarnings = FALSE)

set.seed(1234)

theme_pub <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      text             = element_text(family = "serif", colour = "#1a1a2e"),
      plot.title       = element_text(size = rel(1.25), face = "bold",
                                      hjust = 0, margin = margin(b = 10)),
      plot.subtitle    = element_text(size = rel(0.9), colour = "#555555",
                                      hjust = 0, margin = margin(b = 12)),
      plot.caption     = element_text(size = rel(0.7), colour = "#888888",
                                      hjust = 1, margin = margin(t = 10)),
      axis.title       = element_text(size = rel(0.95), face = "bold"),
      axis.title.x     = element_text(margin = margin(t = 8)),
      axis.title.y     = element_text(margin = margin(r = 8)),
      axis.text        = element_text(size = rel(0.85), colour = "#333333"),
      panel.grid.major = element_line(colour = "#e8e8e8", linewidth = 0.35),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      legend.position  = "bottom",
      legend.title     = element_text(face = "bold", size = rel(0.85)),
      legend.text      = element_text(size = rel(0.8)),
      plot.margin      = margin(15, 15, 15, 15),
      plot.background  = element_rect(fill = "white", colour = NA),
      panel.background = element_rect(fill = "white", colour = NA)
    )
}


pal_blue   <- c("#F7FBFF", "#DEEBF7", "#C6DBEF", "#9ECAE1",
                "#6BAED6", "#4292C6", "#2171B5", "#08519C", "#08306B")
col_accent <- "#08519C"
col_neg    <- "#C0392B"
col_pos    <- "#08519C"
```

```{r}
df <- readr::read_csv(file.path(data_path, "corpus_200ad.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

labels <- readr::read_csv(file.path(data_path, "corpus_200ad_labeled.csv"), show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("christian", "non_christian")) %>%
  transmute(id, manual_label = label) %>%
  distinct(id, .keep_all = TRUE)

df_labeled <- df %>%
  inner_join(labels, by = "id")
```

## Stopwords

```{r}
custom_additions <- c(
  "m", "c", "l", "d", "gt", "lt", "br", "href",
  "fridericus", "heinricus", "conradus", "meus", "tuus"
)

stops_v1 <- scan("../../data/stopwords/stopwords_latin1.txt", what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan("../../data/stopwords/stopwords_latin2.txt", what = "character", comment.char = "#", quiet = TRUE, encoding = "UTF-8")

latin_stops <- c(stops_v1, stops_v2, custom_additions)
latin_stops <- latin_stops[latin_stops != "*"]
latin_stops <- unique(latin_stops)
latin_stops <- sort(latin_stops)

head(latin_stops, 200)
```

## Tokenization

```{r}
#| label: tokenization

get_standard_dfm <- function(text_data, stop_list) {
  toks <- tokens(
    text_data,
    remove_punct   = TRUE,
    remove_symbols = TRUE,
    remove_numbers = TRUE,
    remove_separators = TRUE,
    split_hyphens  = TRUE
  ) %>%
    tokens_tolower() %>%

    tokens_remove(pattern = stop_list, valuetype = "fixed") %>%

    tokens_remove(pattern = "^\\d+$",          valuetype = "regex") %>%
    tokens_remove(pattern = "^[[:punct:]]+$",  valuetype = "regex")

  dfm(toks)
}

dfm_train <- get_standard_dfm(df_labeled$text, latin_stops)
dim(dfm_train)
```

## Corpus matrix preparation

```{r}
dfm_all <- get_standard_dfm(df_labeled$text, latin_stops)

feat_stats <- textstat_frequency(dfm_all)

feature_keep <- feat_stats %>%
  filter(frequency >= 10, docfreq >= 5) %>%
  arrange(desc(frequency)) %>%
  slice_head(n = 400) %>%
  pull(feature)

dfm_small <- dfm_select(dfm_all, pattern = feature_keep)

X_count <- as(dfm_small, "dgCMatrix")
X_bin <- X_count
X_bin@x[X_bin@x > 0] <- 1

doc_n <- nrow(X_count)
feat_n <- ncol(X_count)
feature_names <- colnames(X_count)

doc_n
feat_n
```

## Basic prevalence and frequency tables

```{r}
feature_summary <- tibble(
  feature    = feature_names,
  term_count = Matrix::colSums(X_count),
  doc_freq   = Matrix::colSums(X_bin),
  prevalence = doc_freq / nrow(X_bin)
) %>%
  arrange(desc(term_count))

feature_summary %>% slice_head(n = 30)
```

## Pairwise co-occurrence counts

```{r}
cooc <- Matrix::crossprod(X_bin)
diag(cooc) <- 0

freq_vec <- Matrix::colSums(X_bin)
names(freq_vec) <- feature_names
N_docs <- nrow(X_bin)
```

## PMI and PPMI

```{r}
p_doc <- freq_vec / N_docs
p_ij  <- as.matrix(cooc) / N_docs

pmi_mat <- matrix(NA_real_, nrow = feat_n, ncol = feat_n,
                  dimnames = list(feature_names, feature_names))

for (i in seq_len(feat_n)) {
  for (j in seq_len(feat_n)) {
    if (i != j && p_ij[i, j] > 0) {
      pmi_mat[i, j] <- log(p_ij[i, j] / (p_doc[i] * p_doc[j]))
    }
  }
}

ppmi_mat <- pmi_mat
ppmi_mat[is.na(ppmi_mat)] <- 0
ppmi_mat[ppmi_mat < 0] <- 0
```

## Top marginal associations

```{r}
top_ppmi <- expand.grid(
  word1 = rownames(ppmi_mat),
  word2 = colnames(ppmi_mat),
  stringsAsFactors = FALSE
) %>%
  mutate(
    ppmi   = as.vector(ppmi_mat),
    i      = match(word1, feature_names),
    j      = match(word2, feature_names),
    cooc_n = as.vector(cooc[cbind(i, j)])
  ) %>%
  select(-i, -j) %>%
  filter(word1 < word2, ppmi > 0) %>%
  arrange(desc(ppmi), desc(cooc_n))

top_ppmi %>% slice_head(n = 50)
```

## Permutation test for co-occurrence enrichment

```{r}
permute_columns_within_docs <- function(Xb) {
  out <- Xb
  for (j in seq_len(ncol(out))) {
    out[, j] <- sample(out[, j])
  }
  out
}

B_perm <- 100

obs_ppmi <- ppmi_mat

perm_max_ppmi  <- numeric(B_perm)
perm_mean_ppmi <- numeric(B_perm)

for (b in seq_len(B_perm)) {
  Xp <- permute_columns_within_docs(as.matrix(X_bin))
  cooc_p <- crossprod(Xp)
  diag(cooc_p) <- 0

  p_doc_p <- colSums(Xp) / nrow(Xp)
  p_ij_p  <- cooc_p / nrow(Xp)

  pmi_p <- matrix(0, ncol(Xp), ncol(Xp))
  for (i in seq_len(ncol(Xp))) {
    for (j in seq_len(ncol(Xp))) {
      if (i != j && p_ij_p[i, j] > 0) {
        pmi_p[i, j] <- max(0, log(p_ij_p[i, j] / (p_doc_p[i] * p_doc_p[j])))
      }
    }
  }

  perm_max_ppmi[b]  <- max(pmi_p)
  perm_mean_ppmi[b] <- mean(pmi_p[pmi_p > 0])
}

perm_test_summary <- tibble(
  observed_max_ppmi                = max(obs_ppmi),
  permuted_max_ppmi_mean           = mean(perm_max_ppmi),
  permuted_max_ppmi_q95            = quantile(perm_max_ppmi, 0.95),
  observed_mean_positive_ppmi      = mean(obs_ppmi[obs_ppmi > 0]),
  permuted_mean_positive_ppmi_mean = mean(perm_mean_ppmi),
  permuted_mean_positive_ppmi_q95  = quantile(perm_mean_ppmi, 0.95)
)

perm_test_summary
```

## PPMI heatmap

```{r}
#| fig-width: 10
#| fig-height: 9

top_heat_words <- feature_summary %>%
  slice_head(n = 60) %>%
  pull(feature)

ppmi_heat <- ppmi_mat[top_heat_words, top_heat_words]

ht_ppmi <- Heatmap(
  ppmi_heat,
  name = "PPMI",
  cluster_rows    = TRUE,
  cluster_columns = TRUE,
  row_names_gp    = grid::gpar(fontsize = 7, fontfamily = "serif"),
  column_names_gp = grid::gpar(fontsize = 7, fontfamily = "serif"),
  rect_gp         = grid::gpar(col = "white", lwd = 0.3),
  col = colorRamp2(
    c(0, quantile(ppmi_heat, 0.5), quantile(ppmi_heat, 0.8), max(ppmi_heat)),
    c("#F7FBFF", "#C6DBEF", "#4292C6", "#08306B")
  ),
  column_title      = "PPMI — Top 60 Features",
  column_title_gp   = grid::gpar(fontsize = 14, fontface = "bold", fontfamily = "serif"),
  heatmap_legend_param = list(
    title_gp  = grid::gpar(fontsize = 10, fontface = "bold", fontfamily = "serif"),
    labels_gp = grid::gpar(fontsize = 8, fontfamily = "serif")
  ),
  border = TRUE,
  border_gp = grid::gpar(col = "#cccccc", lwd = 0.5)
)

draw(ht_ppmi)
```

## Word association network

```{r}
#| fig-width: 12
#| fig-height: 10

net_edges <- top_ppmi %>%
  filter(cooc_n >= 5) %>%
  slice_head(n = 250) %>%
  transmute(from = word1, to = word2, weight = ppmi)

net_nodes <- tibble(name = unique(c(net_edges$from, net_edges$to))) %>%
  left_join(feature_summary %>% transmute(name = feature, doc_freq, term_count), by = "name")

g_ppmi <- tbl_graph(nodes = net_nodes, edges = net_edges, directed = FALSE)

p_ppmi_net <- ggraph(g_ppmi, layout = "fr") +
  geom_edge_link(
    aes(width = weight, alpha = weight),
    colour = "#6BAED6",
    show.legend = FALSE
  ) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_edge_alpha(range = c(0.15, 0.7)) +
  geom_node_point(
    aes(size = doc_freq),
    colour = col_accent,
    alpha = 0.85,
    show.legend = TRUE
  ) +
  scale_size_continuous(range = c(1.5, 10), name = "Document\nfrequency") +
  geom_node_text(
    aes(label = name),
    repel = TRUE, size = 3.2,
    colour = "#1a1a2e", family = "serif",
    max.overlaps = 30,
    segment.colour = "#aaaaaa",
    segment.size = 0.25
  ) +
  labs(
    title    = "Word Association Network (PPMI)",
    subtitle = "Top 250 co-occurrence edges, minimum 5 shared documents",
    caption  = "600 AD Latin corpus"
  ) +
  theme_void(base_size = 13) +
  theme(
    text           = element_text(family = "serif"),
    plot.title     = element_text(face = "bold", size = 16, hjust = 0,
                                  margin = margin(b = 4)),
    plot.subtitle  = element_text(size = 11, colour = "#555555", hjust = 0,
                                  margin = margin(b = 8)),
    plot.caption   = element_text(size = 9, colour = "#888888", hjust = 1),
    legend.position = "right",
    plot.margin    = margin(15, 15, 15, 15),
    plot.background = element_rect(fill = "white", colour = NA)
  )

p_ppmi_net
```

## Conditional dependence network via sparse Gaussian graphical model

```{r}
X_tfidf <- as(dfm_tfidf(dfm_small), "dgCMatrix")
X_dense <- as.matrix(X_tfidf)

valid_cols <- apply(X_dense, 2, function(x) stats::sd(x) > 0 & all(is.finite(x)))
X_dense <- X_dense[, valid_cols, drop = FALSE]

feature_names_glasso <- colnames(X_dense)

X_scaled <- scale(X_dense)
X_scaled <- as.matrix(X_scaled)

S <- cov(X_scaled)
dimnames(S) <- list(feature_names_glasso, feature_names_glasso)

lambda_grid <- seq(0.05, 0.30, by = 0.01)

ebic_score <- function(Theta, S, n, gamma = 0.5) {
  p  <- ncol(S)
  E  <- sum(Theta[upper.tri(Theta)] != 0)
  ll <- (n / 2) * (determinant(Theta, logarithm = TRUE)$modulus[1] - sum(S * Theta))
  as.numeric(-2 * ll + E * log(n) + 4 * E * gamma * log(p))
}

glasso_fits <- vector("list", length(lambda_grid))
ebic_vals   <- numeric(length(lambda_grid))

for (k in seq_along(lambda_grid)) {
  fit_k           <- glasso(S, rho = lambda_grid[k])
  glasso_fits[[k]] <- fit_k
  ebic_vals[k]    <- ebic_score(fit_k$wi, S, nrow(X_scaled), gamma = 0.5)
}

best_k      <- which.min(ebic_vals)
glasso_best <- glasso_fits[[best_k]]
best_lambda <- lambda_grid[best_k]

Theta <- glasso_best$wi
dimnames(Theta) <- list(feature_names_glasso, feature_names_glasso)

Pcor <- -cov2cor(Theta)
diag(Pcor) <- 0
dimnames(Pcor) <- list(feature_names_glasso, feature_names_glasso)

best_lambda
dim(Pcor)
length(feature_names_glasso)
```

## Top conditional dependencies

```{r}
top_pcor <- expand.grid(
  word1 = rownames(Pcor),
  word2 = colnames(Pcor),
  stringsAsFactors = FALSE
) %>%
  mutate(
    partial_cor     = as.vector(Pcor),
    abs_partial_cor = abs(partial_cor)
  ) %>%
  filter(word1 < word2, partial_cor != 0) %>%
  arrange(desc(abs_partial_cor))

top_pcor %>% slice_head(n = 50)
```

## Conditional dependence network plot

```{r}
#| fig-width: 12
#| fig-height: 10

pcor_edges <- top_pcor %>%
  slice_head(n = 200) %>%
  transmute(
    from   = word1,
    to     = word2,
    weight = partial_cor,
    sign   = ifelse(partial_cor > 0, "positive", "negative")
  )

pcor_nodes <- tibble(name = unique(c(pcor_edges$from, pcor_edges$to))) %>%
  left_join(feature_summary %>% transmute(name = feature, doc_freq, term_count), by = "name")

g_pcor <- tbl_graph(nodes = pcor_nodes, edges = pcor_edges, directed = FALSE)

p_pcor_net <- ggraph(g_pcor, layout = "fr") +
  geom_edge_link(
    aes(width = abs(weight), alpha = abs(weight), colour = sign),
    show.legend = FALSE
  ) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_edge_alpha(range = c(0.15, 0.7)) +
  scale_edge_colour_manual(values = c(positive = col_pos, negative = col_neg)) +
  geom_node_point(
    aes(size = doc_freq),
    colour = "#1a1a2e",
    alpha = 0.8,
    show.legend = TRUE
  ) +
  scale_size_continuous(range = c(1.5, 9), name = "Document\nfrequency") +
  geom_node_text(
    aes(label = name),
    repel = TRUE, size = 3.2,
    family = "serif",
    max.overlaps = 30,
    segment.colour = "#aaaaaa",
    segment.size = 0.25
  ) +
  labs(
    title    = "Conditional Dependence Network (Graphical Lasso)",
    subtitle = paste0("Top 200 edges — blue = positive, red = negative  (λ = ",
                      round(best_lambda, 3), ")"),
    caption  = "600 AD Latin corpus"
  ) +
  theme_void(base_size = 13) +
  theme(
    text            = element_text(family = "serif"),
    plot.title      = element_text(face = "bold", size = 16, hjust = 0,
                                   margin = margin(b = 4)),
    plot.subtitle   = element_text(size = 11, colour = "#555555", hjust = 0,
                                   margin = margin(b = 8)),
    plot.caption    = element_text(size = 9, colour = "#888888", hjust = 1),
    legend.position = "right",
    plot.margin     = margin(15, 15, 15, 15),
    plot.background = element_rect(fill = "white", colour = NA)
  )

p_pcor_net
```

## Stability analysis for the conditional network

```{r}
B_boot    <- 50
edge_keep <- vector("list", B_boot)

for (b in seq_len(B_boot)) {
  idx <- sample(seq_len(nrow(X_scaled)), replace = TRUE)
  Sb  <- cov(X_scaled[idx, , drop = FALSE])
  dimnames(Sb) <- list(feature_names_glasso, feature_names_glasso)

  fit_b <- glasso(Sb, rho = best_lambda)
  Pb    <- -cov2cor(fit_b$wi)
  diag(Pb) <- 0
  dimnames(Pb) <- list(feature_names_glasso, feature_names_glasso)

  eb <- expand.grid(
    word1 = rownames(Pb),
    word2 = colnames(Pb),
    stringsAsFactors = FALSE
  ) %>%
    mutate(partial_cor = as.vector(Pb)) %>%
    filter(word1 < word2, partial_cor != 0) %>%
    transmute(edge = paste(word1, word2, sep = "~~"))

  edge_keep[[b]] <- eb$edge
}

edge_stability <- bind_rows(lapply(edge_keep, function(x) tibble(edge = x)), .id = "boot") %>%
  count(edge, name = "times_selected") %>%
  mutate(selection_rate = times_selected / B_boot) %>%
  tidyr::separate(edge, into = c("word1", "word2"), sep = "~~") %>%
  arrange(desc(selection_rate))

edge_stability %>% slice_head(n = 50)
```

## Low-rank geometry

```{r}
k_svd <- min(50, ncol(X_tfidf) - 1, nrow(X_tfidf) - 1)
stopifnot(k_svd >= 2)

svd_fit <- irlba(X_tfidf, nv = k_svd, nu = k_svd)

sv_tbl <- tibble(
  component         = seq_along(svd_fit$d),
  singular_value    = svd_fit$d,
  variance_explained = svd_fit$d^2 / sum(svd_fit$d^2),
  cumulative_variance = cumsum(variance_explained)
)

sv_tbl %>% slice_head(n = 20)
```

## Permutation baseline for singular values

```{r}
B_svd      <- 50
n_sv_plot  <- min(20, length(svd_fit$d))
perm_sv    <- matrix(NA_real_, nrow = B_svd, ncol = n_sv_plot)

X_tfidf_dense <- as.matrix(X_tfidf)

for (b in seq_len(B_svd)) {
  Xp <- X_tfidf_dense
  for (j in seq_len(ncol(Xp))) {
    Xp[, j] <- sample(Xp[, j])
  }
  sv_b <- irlba(Xp, nv = n_sv_plot, nu = n_sv_plot)$d
  perm_sv[b, ] <- sv_b
}

sv_plot_tbl <- tibble(
  component = seq_len(n_sv_plot),
  observed  = svd_fit$d[seq_len(n_sv_plot)],
  perm_mean = colMeans(perm_sv, na.rm = TRUE),
  perm_q95  = apply(perm_sv, 2, quantile, 0.95, na.rm = TRUE)
)

sv_plot_tbl
```

## Singular value plot

```{r}
#| fig-width: 8
#| fig-height: 5.5

p_svd <- ggplot(sv_plot_tbl, aes(component)) +
  geom_ribbon(aes(ymin = 0, ymax = perm_q95),
              fill = "#D9EAF7", alpha = 0.6) +
  geom_line(aes(y = perm_mean), linewidth = 0.9, colour = "#9ECAE1",
            linetype = "dashed") +
  geom_line(aes(y = observed), linewidth = 1.2, colour = col_accent) +
  geom_point(aes(y = observed), size = 2.5, colour = col_accent,
             fill = "white", shape = 21, stroke = 1.1) +
  annotate("text", x = n_sv_plot * 0.75, y = max(sv_plot_tbl$perm_q95) * 1.15,
           label = "95th percentile\n(permuted)", colour = "#6BAED6",
           size = 3.5, family = "serif", fontface = "italic") +
  scale_x_continuous(breaks = seq_len(n_sv_plot)) +
  labs(
    title    = "Singular Value Spectrum",
    subtitle = "Observed vs. column-permuted null distribution",
    x = "Component", y = "Singular value",
    caption = "Shaded region: null 95th percentile envelope"
  ) +
  theme_pub() +
  theme(panel.grid.major.x = element_blank())

p_svd
```

## Word coordinates from low-rank structure

```{r}
#| fig-width: 10
#| fig-height: 8

word_coords <- as_tibble(svd_fit$v[, 1:2, drop = FALSE]) %>%
  setNames(c("dim1", "dim2")) %>%
  mutate(feature = colnames(X_tfidf)) %>%
  left_join(feature_summary, by = "feature")

plot_words <- word_coords %>%
  slice_max(order_by = term_count, n = min(120, nrow(word_coords)))

p_embed <- ggplot(plot_words, aes(dim1, dim2)) +
  geom_hline(yintercept = 0, linewidth = 0.3, colour = "grey80", linetype = "dotted") +
  geom_vline(xintercept = 0, linewidth = 0.3, colour = "grey80", linetype = "dotted") +
  geom_point(
    aes(size = doc_freq),
    colour = col_accent, alpha = 0.6,
    show.legend = TRUE
  ) +
  scale_size_continuous(range = c(1.5, 9), name = "Document frequency") +
  ggrepel::geom_text_repel(
    aes(label = feature), size = 3.2,
    family = "serif",
    max.overlaps = 200,
    segment.colour = "#bbbbbb",
    segment.size = 0.2,
    colour = "#1a1a2e"
  ) +
  labs(
    title    = "Word Embedding — First Two SVD Dimensions",
    subtitle = "TF-IDF weighted, top 120 terms by frequency",
    x = "Dimension 1", y = "Dimension 2",
    caption = "600 AD Latin corpus"
  ) +
  theme_pub() +
  theme(panel.grid.major.x = element_line(colour = "#e8e8e8", linewidth = 0.35))

p_embed
```

## Pairwise dependence beyond correlation: mutual information

```{r}
top_mi_words <- feature_summary %>%
  slice_head(n = min(80, nrow(feature_summary))) %>%
  pull(feature)

X_mi <- as.matrix(X_bin[, top_mi_words, drop = FALSE])

mi_tbl <- expand_grid(word1 = top_mi_words, word2 = top_mi_words) %>%
  filter(word1 < word2) %>%
  mutate(
    i  = match(word1, colnames(X_mi)),
    j  = match(word2, colnames(X_mi)),
    mi = map2_dbl(i, j, ~ mutinformation(X_mi[, .x], X_mi[, .y]))
  ) %>%
  select(word1, word2, mi) %>%
  arrange(desc(mi))

mi_tbl %>% slice_head(n = 50)
```

## Compare PPMI to mutual information

```{r}
#| fig-width: 7.5
#| fig-height: 5.5

compare_dep <- top_ppmi %>%
  select(word1, word2, ppmi) %>%
  inner_join(mi_tbl, by = c("word1", "word2"))

p_ppmi_mi <- ggplot(compare_dep, aes(ppmi, mi)) +
  geom_point(alpha = 0.4, colour = col_accent, size = 1.8) +
  geom_smooth(method = "lm", se = TRUE, colour = col_neg,
              linewidth = 0.9, fill = "#f4cccc", alpha = 0.3) +
  labs(
    title    = "PPMI vs. Mutual Information",
    subtitle = "Each point is a word pair",
    x = "Positive PMI", y = "Mutual information (bits)",
    caption  = "600 AD Latin corpus"
  ) +
  theme_pub() +
  theme(panel.grid.major.x = element_line(colour = "#e8e8e8", linewidth = 0.35))

p_ppmi_mi
```

## Triplewise interaction screening

```{r}
top_trip_words <- feature_summary %>%
  slice_head(n = min(30, nrow(feature_summary))) %>%
  pull(feature)

X_trip <- as.matrix(X_bin[, top_trip_words, drop = FALSE])

triples <- combn(top_trip_words, 3, simplify = FALSE)

safe_mi <- function(x, y) {
  if (length(unique(x)) < 2 || length(unique(y)) < 2) return(0)
  mutinformation(x, y)
}

interaction_information <- function(a, b, c) {
  I_ab <- safe_mi(a, b)

  idx1 <- which(c == 1)
  idx0 <- which(c == 0)

  I_ab_c1 <- if (length(idx1) > 1) safe_mi(a[idx1], b[idx1]) else 0
  I_ab_c0 <- if (length(idx0) > 1) safe_mi(a[idx0], b[idx0]) else 0

  I_ab - (mean(c == 1) * I_ab_c1 + mean(c == 0) * I_ab_c0)
}

triple_tbl <- purrr::map_dfr(triples, function(tri) {
  a <- X_trip[, tri[1]]
  b <- X_trip[, tri[2]]
  c <- X_trip[, tri[3]]

  tibble(
    word1            = tri[1],
    word2            = tri[2],
    word3            = tri[3],
    interaction_info = interaction_information(a, b, c)
  )
}) %>%
  mutate(abs_interaction_info = abs(interaction_info)) %>%
  arrange(desc(abs_interaction_info))

triple_tbl %>% slice_head(n = 40)
```

## Plot higher-order interaction strengths

```{r}
#| fig-width: 9
#| fig-height: 8

triple_plot <- triple_tbl %>%
  slice_head(n = min(30, nrow(triple_tbl))) %>%
  mutate(
    triple = paste(word1, word2, word3, sep = "  ·  "),
    triple = forcats::fct_reorder(triple, interaction_info),
    direction = ifelse(interaction_info >= 0, "synergistic", "redundant")
  )

p_triple <- ggplot(triple_plot, aes(interaction_info, triple, fill = direction)) +
  geom_col(alpha = 0.85, width = 0.7) +
  geom_vline(xintercept = 0, linewidth = 0.4, colour = "#333333") +
  scale_fill_manual(
    values = c(synergistic = col_pos, redundant = col_neg),
    name   = "Interaction type"
  ) +
  labs(
    title    = "Three-Way Interaction Information",
    subtitle = "Top 30 triples by absolute interaction strength",
    x = "Interaction information",
    y = NULL,
    caption = "Positive = synergistic; negative = redundant"
  ) +
  theme_pub() +
  theme(
    axis.text.y     = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = "#e8e8e8", linewidth = 0.35)
  )

p_triple
```

## Summary diagnostics table

```{r}
summary_diagnostics <- tibble(
  quantity = c(
    "mean positive PPMI",
    "max PPMI",
    "mean abs partial correlation",
    "number of nonzero conditional edges",
    "top 5 singular values cumulative variance",
    "top 10 singular values cumulative variance",
    "mean pairwise mutual information",
    "mean abs triple interaction information"
  ),
  value = c(
    if (any(ppmi_mat > 0, na.rm = TRUE)) mean(ppmi_mat[ppmi_mat > 0], na.rm = TRUE) else NA_real_,
    max(ppmi_mat, na.rm = TRUE),
    mean(abs(Pcor[Pcor != 0]), na.rm = TRUE),
    sum(Pcor[upper.tri(Pcor)] != 0, na.rm = TRUE),
    if (nrow(sv_tbl) >= 5) sv_tbl$cumulative_variance[5] else NA_real_,
    if (nrow(sv_tbl) >= 10) sv_tbl$cumulative_variance[10] else NA_real_,
    mean(mi_tbl$mi, na.rm = TRUE),
    mean(abs(triple_tbl$interaction_info), na.rm = TRUE)
  )
)

summary_diagnostics
```

## MI heatmap

```{r}
#| fig-width: 11
#| fig-height: 9.5

top_mi_heat_words <- mi_tbl %>%
  slice_head(n = 150) %>%
  select(word1, word2) %>%
  unlist(use.names = FALSE) %>%
  unique() %>%
  head(40)

mi_heat_tbl <- expand_grid(
  word1 = top_mi_heat_words,
  word2 = top_mi_heat_words
) %>%
  mutate(
    i  = match(word1, colnames(X_mi)),
    j  = match(word2, colnames(X_mi)),
    mi = map2_dbl(i, j, ~ if (is.na(.x) || is.na(.y)) NA_real_
                          else mutinformation(X_mi[, .x], X_mi[, .y]))
  )

p_mi_heat <- ggplot(mi_heat_tbl, aes(word1, word2, fill = mi)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  scale_fill_gradientn(
    colours  = c("#F7FBFF", "#C6DBEF", "#6BAED6", "#2171B5", "#08306B"),
    na.value = "grey95",
    name     = "MI (bits)"
  ) +
  labs(
    title    = "Pairwise Mutual Information",
    subtitle = "Top 40 most-associated terms",
    caption  = "600 AD Latin corpus"
  ) +
  theme_pub() +
  theme(
    axis.title       = element_blank(),
    axis.text.x      = element_text(angle = 50, hjust = 1, size = 9),
    axis.text.y      = element_text(size = 9),
    panel.grid       = element_blank(),
    legend.position  = "right"
  ) +
  coord_fixed()

p_mi_heat
```

## Save tables

```{r}
readr::write_csv(feature_summary, file.path(results_dir, "feature_summary.csv"))
readr::write_csv(top_ppmi, file.path(results_dir, "top_ppmi.csv"))
readr::write_csv(perm_test_summary, file.path(results_dir, "ppmi_permutation_summary.csv"))
readr::write_csv(top_pcor, file.path(results_dir, "top_partial_correlations.csv"))
readr::write_csv(edge_stability, file.path(results_dir, "edge_stability_bootstrap.csv"))
readr::write_csv(sv_tbl, file.path(results_dir, "svd_spectrum.csv"))
readr::write_csv(sv_plot_tbl, file.path(results_dir, "svd_permutation_comparison.csv"))
readr::write_csv(word_coords, file.path(results_dir, "word_embeddings_2d.csv"))
readr::write_csv(mi_tbl, file.path(results_dir, "pairwise_mutual_information.csv"))
readr::write_csv(compare_dep, file.path(results_dir, "ppmi_vs_mi.csv"))
readr::write_csv(triple_tbl, file.path(results_dir, "triple_interactions.csv"))
readr::write_csv(summary_diagnostics, file.path(results_dir, "summary_diagnostics.csv"))
readr::write_csv(mi_heat_tbl, file.path(results_dir, "mi_heatmap_table.csv"))
readr::write_csv(net_edges, file.path(results_dir, "ppmi_network_edges.csv"))
readr::write_csv(net_nodes, file.path(results_dir, "ppmi_network_nodes.csv"))
readr::write_csv(pcor_edges, file.path(results_dir, "partial_correlation_network_edges.csv"))
readr::write_csv(pcor_nodes, file.path(results_dir, "partial_correlation_network_nodes.csv"))
```

## Save matrices and objects

```{r}
saveRDS(ppmi_mat, file.path(results_dir, "ppmi_matrix.rds"))
saveRDS(cooc, file.path(results_dir, "cooccurrence_matrix.rds"))
saveRDS(Pcor, file.path(results_dir, "partial_correlation_matrix.rds"))
saveRDS(X_tfidf, file.path(results_dir, "tfidf_matrix.rds"))
saveRDS(X_bin, file.path(results_dir, "binary_matrix.rds"))
saveRDS(g_ppmi, file.path(results_dir, "ppmi_network.rds"))
saveRDS(g_pcor, file.path(results_dir, "partial_correlation_network.rds"))
saveRDS(svd_fit, file.path(results_dir, "svd_fit.rds"))
saveRDS(glasso_best, file.path(results_dir, "glasso_fit.rds"))
```

## Save figures

```{r}
ggsave(
  filename = file.path(results_dir, "svd_spectrum_plot.png"),
  plot = p_svd,
  width = 8, height = 5.5, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "word_embedding_plot.png"),
  plot = p_embed,
  width = 10, height = 8, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "ppmi_vs_mi_plot.png"),
  plot = p_ppmi_mi,
  width = 7.5, height = 5.5, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "triple_interactions_plot.png"),
  plot = p_triple,
  width = 9, height = 8, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "ppmi_network_plot.png"),
  plot = p_ppmi_net,
  width = 12, height = 10, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "partial_correlation_network_plot.png"),
  plot = p_pcor_net,
  width = 12, height = 10, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(results_dir, "mi_heatmap_plot.png"),
  plot = p_mi_heat,
  width = 11, height = 9.5, dpi = 300, bg = "white"
)

png(file.path(results_dir, "ppmi_heatmap.png"), width = 2800, height = 2400, res = 300)
draw(ht_ppmi)
dev.off()
```

## Final check

```{r}
saved_files <- list.files(results_dir, full.names = TRUE)
print(saved_files)
length(saved_files)
```
