---
title: "Latin text classification"
author: "Muhammad Rehan"
date: 2026-01-10
format:
  html:
    theme:
      light: [lux, assets/theme.scss]
      dark: [darkly, assets/theme-dark.scss]
    toc: true
    toc-depth: 3
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-overflow: wrap
    code-fold: show
    code-tools: true
    code-link: true
    highlight-style: nord
    lightbox: auto
    citations-hover: true
    crossrefs-hover: true
    fig-align: center
    fig-width: 8
    fig-asp: 0.62
    fig-dpi: 300
    fig-format: svg
    grid:
      body-width: 980px
      margin-width: 280px
      gutter-width: 1.2rem
    css: [assets/custom.css]

include-in-header:
  text: |
    <!-- Standard meta -->
    <meta name="viewport" content="width=device-width, initial-scale=1" />
    <meta name="theme-color" content="#0b5ed7" />

    <!-- KaTeX for LaTeX math -->
    <link rel="stylesheet"
          href="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.css"
          integrity="sha384-/7mQh8gJtX+NEwGDbVS9/6IKxhpcJn/qdNqxabWWMwBLT/gR59OA8sLu7saEh1N5"
          crossorigin="anonymous">

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/katex.min.js"
            integrity="sha384-G/jPD7WvyaManIeZDV4SSQdlqzTeWY5Avzkdxl3pNGdisz8Iky3Uczdlz7YTpcvA"
            crossorigin="anonymous"></script>

    <script defer
            src="https://cdn.jsdelivr.net/npm/katex@0.16.8/dist/contrib/auto-render.min.js"
            integrity="sha384-mll67QQj1YOD9OFwP8CMLJrjCwXTlzAIXazhbugzuDUFcPRl1BDpRP70dNFu7Bbf"
            crossorigin="anonymous"
            onload="renderMathInElement(document.body);"></script>

execute:
  echo: true
  warning: false
  message: false
  freeze: auto
  cache: true
  dpi: 600

math: katex
---

# Classification of Christian vs. Non-Christian Documents

## Setup and Configuration

### Paths and packages

```{r setup}
suppressPackageStartupMessages({
  library(tidyverse)
  library(stringi)
  library(quanteda)
  library(quanteda.textmodels)
  library(glmnet)
  library(LiblineaR)
  library(xgboost)
  library(lightgbm)
  library(Matrix)
  library(pROC)
  library(knitr)
})

set.seed(123)

# ── Paths ──────────────────────────────────────────────────────────────────────
# Adjust these three paths to match your project layout.
data_path    <- "../../data/processed_texts/non_late_late"
results_path <- "../../results/predictions/non_transformer_lemmatized/non_late_late"
checkpoint_path <- results_path

# Derived paths
checkpoint_root <- file.path(checkpoint_path, "checkpoints_nested_text_models")
output_dir      <- file.path(checkpoint_path, "model_outputs_checkpointed")
progress_path   <- file.path(checkpoint_path, "model_progress_log.csv")

dir.create(checkpoint_root, recursive = TRUE, showWarnings = FALSE)
dir.create(output_dir,      recursive = TRUE, showWarnings = FALSE)

# ── Hyper-parameters ──────────────────────────────────────────────────────────
k_folds_outer    <- 5L
k_folds_inner    <- 5L
xgb_max_rounds   <- 4000L
lgb_max_rounds   <- 4000L
early_stop_rounds <- 100L
min_docfreq      <- 3L

# Elastic net alpha grid (replaces pure ridge α=0)
enet_alpha_grid <- c(0.05, 0.1, 0.2, 0.3, 0.4, 0.5)
```

## Utility Functions

```{r utilities}
# ── Filesystem helpers ─────────────────────────────────────────────────────────
`%||%` <- function(x, y) if (is.null(x)) y else x

ensure_dir <- function(path) {

  dir.create(path, recursive = TRUE, showWarnings = FALSE)
  invisible(path)
}

save_rds_atomic <- function(object, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_",
                format(Sys.time(), "%Y%m%d%H%M%S"))
  saveRDS(object, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) { file.copy(tmp, path, overwrite = TRUE); unlink(tmp) }
  invisible(path)
}

write_csv_atomic <- function(x, path) {
  ensure_dir(dirname(path))
  tmp <- paste0(path, ".tmp_", Sys.getpid(), "_",
                format(Sys.time(), "%Y%m%d%H%M%S"), ".csv")
  readr::write_csv(x, tmp)
  ok <- file.rename(tmp, path)
  if (!ok) { file.copy(tmp, path, overwrite = TRUE); unlink(tmp) }
  invisible(path)
}

load_rds_or_null <- function(path) {
  if (!file.exists(path)) return(NULL)
  readRDS(path)
}

# ── Progress logging ──────────────────────────────────────────────────────────
append_progress_row <- function(row, path) {
  if (!file.exists(path)) {
    write_csv_atomic(row, path)
  } else {
    old <- readr::read_csv(path, show_col_types = FALSE) %>%
      mutate(across(everything(), as.character))
    row <- row %>% mutate(across(everything(), as.character))
    write_csv_atomic(bind_rows(old, row), path)
  }
}

log_progress <- function(step, detail = "") {
  row <- tibble(
    time   = format(Sys.time(), "%Y-%m-%d %H:%M:%S"),
    step   = step,
    detail = detail
  )
  message(paste0("[", row$time, "] ", step,
                 if (nzchar(detail)) paste0(" | ", detail) else ""))
  append_progress_row(row, progress_path)
}

# ── Probability / score helpers ───────────────────────────────────────────────
safe_prob  <- function(p, eps = 1e-15) pmin(pmax(as.numeric(p), eps), 1 - eps)
safe_logit <- function(p, eps = 1e-15) qlogis(safe_prob(p, eps))

# ── Label checks ──────────────────────────────────────────────────────────────
check_binary_y <- function(y, where = "") {
  tb <- table(y)
  if (!all(c("late", "non_late") %in% names(tb)))
    stop(paste0("Both classes required",
                if (nzchar(where)) paste0(" in ", where) else "", "."))
  if (any(tb < 2))
    stop(paste0("Each class needs >=2 obs",
                if (nzchar(where)) paste0(" in ", where) else "", "."))
}

# ── Fold creation ─────────────────────────────────────────────────────────────
make_folds <- function(y, k, seed = 123) {
  check_binary_y(y, "make_folds")
  set.seed(seed)
  k_use <- min(as.integer(k), min(table(y)))
  if (k_use < 2) stop("Not enough observations per class for CV.")
  idx_pos <- sample(which(y == "late"))
  idx_neg <- sample(which(y == "non_late"))
  folds   <- vector("list", k_use)
  for (i in seq_len(k_use)) folds[[i]] <- integer(0)
  for (i in seq_along(idx_pos))
    folds[[((i - 1) %% k_use) + 1]] <-
      c(folds[[((i - 1) %% k_use) + 1]], idx_pos[i])
  for (i in seq_along(idx_neg))
    folds[[((i - 1) %% k_use) + 1]] <-
      c(folds[[((i - 1) %% k_use) + 1]], idx_neg[i])
  folds
}

# ── Evaluation metrics ────────────────────────────────────────────────────────
auc_safe <- function(y, p) {
  p <- safe_prob(p)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = p,
      levels = c("non_late", "late"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

auc_margin_safe <- function(y, s) {
  s <- as.numeric(s)
  tryCatch(
    as.numeric(pROC::auc(pROC::roc(
      response = y, predictor = s,
      levels = c("non_late", "late"),
      direction = "<", quiet = TRUE
    ))),
    error = function(e) NA_real_
  )
}

log_loss_safe <- function(y, p) {
  p   <- safe_prob(p)
  y01 <- as.integer(y == "late")
  -mean(y01 * log(p) + (1 - y01) * log(1 - p))
}

brier_safe <- function(y, p) {
  p   <- safe_prob(p)
  y01 <- as.integer(y == "late")
  mean((y01 - p)^2)
}

hard_metrics <- function(y, pred_label) {
  y01 <- as.integer(y == "late")
  p01 <- as.integer(pred_label == "late")
  tp  <- sum(y01 == 1 & p01 == 1)
  tn  <- sum(y01 == 0 & p01 == 0)
  fp  <- sum(y01 == 0 & p01 == 1)
  fn  <- sum(y01 == 1 & p01 == 0)
  precision   <- if ((tp + fp) == 0) NA_real_ else tp / (tp + fp)
  recall      <- if ((tp + fn) == 0) NA_real_ else tp / (tp + fn)
  specificity <- if ((tn + fp) == 0) NA_real_ else tn / (tn + fp)
  f1  <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0)
    NA_real_ else 2 * precision * recall / (precision + recall)
  bal_acc <- mean(c(recall, specificity), na.rm = TRUE)
  mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
  mcc <- if (mcc_den == 0) NA_real_ else ((tp * tn) - (fp * fn)) / mcc_den
  tibble(accuracy = (tp + tn) / length(y), balanced_accuracy = bal_acc,
         precision = precision, recall = recall, specificity = specificity,
         f1 = f1, mcc = mcc, tp = tp, tn = tn, fp = fp, fn = fn)
}

# ── Threshold selection ───────────────────────────────────────────────────────
choose_threshold <- function(y, p, metric = c("youden", "f1")) {
  metric <- match.arg(metric)
  p <- safe_prob(p)
  if (metric == "youden") {
    th <- tryCatch(
      pROC::coords(
        pROC::roc(response = y, predictor = p,
                  levels = c("non_late", "late"),
                  direction = "<", quiet = TRUE),
        x = "best", input = "threshold", best.method = "youden",
        ret = "threshold", transpose = FALSE
      )[[1]],
      error = function(e) NA_real_
    )
    if (!is.finite(th)) 0.5 else as.numeric(th)
  } else {
    y01  <- as.integer(y == "late")
    cand <- sort(unique(c(0, p, 0.5, 1)))
    f1s  <- vapply(cand, function(th) {
      pred01 <- as.integer(p >= th)
      tp  <- sum(pred01 == 1 & y01 == 1)
      fp  <- sum(pred01 == 1 & y01 == 0)
      fn  <- sum(pred01 == 0 & y01 == 1)
      den <- 2 * tp + fp + fn
      if (den == 0) 0 else 2 * tp / den
    }, numeric(1))
    cand[which.max(f1s)]
  }
}

# ── Summary helpers ───────────────────────────────────────────────────────────
summarize_model_oof <- function(y, p, pred_youden, pred_f1, model_name) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    threshold_policy_youden = "nested_outer_train_only",
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    tp_youden = hy$tp, tn_youden = hy$tn, fp_youden = hy$fp, fn_youden = hy$fn,
    threshold_policy_f1 = "nested_outer_train_only",
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc,
    tp_f1 = hf$tp, tn_f1 = hf$tn, fp_f1 = hf$fp, fn_f1 = hf$fn
  )
}

summarize_model_fold <- function(y, p, pred_youden, pred_f1,
                                  model_name, outer_fold) {
  hy <- hard_metrics(y, pred_youden)
  hf <- hard_metrics(y, pred_f1)
  tibble(
    outer_fold = outer_fold, model = model_name,
    auc = auc_safe(y, p), log_loss = log_loss_safe(y, p),
    brier = brier_safe(y, p),
    accuracy_youden = hy$accuracy, balanced_accuracy_youden = hy$balanced_accuracy,
    precision_youden = hy$precision, recall_youden = hy$recall,
    specificity_youden = hy$specificity, f1_youden = hy$f1, mcc_youden = hy$mcc,
    accuracy_f1 = hf$accuracy, balanced_accuracy_f1 = hf$balanced_accuracy,
    precision_f1 = hf$precision, recall_f1 = hf$recall,
    specificity_f1 = hf$specificity, f1_f1 = hf$f1, mcc_f1 = hf$mcc
  )
}

upsert_rows <- function(df, new_rows, keys) {
  if (is.null(df) || nrow(df) == 0) return(new_rows)
  keep <- dplyr::anti_join(df, dplyr::distinct(new_rows[, keys, drop = FALSE]),
                           by = keys)
  bind_rows(keep, new_rows)
}

# ── Platt scaling ─────────────────────────────────────────────────────────────
fit_platt <- function(scores, y) {
  z <- tibble(score = as.numeric(scores),
              y = as.integer(y == "late")) %>%
    filter(is.finite(score))
  if (nrow(z) < 10 || length(unique(z$score)) < 2 ||
      length(unique(z$y)) < 2) return(NULL)
  suppressWarnings(
    tryCatch({
      fit <- glm(y ~ score, data = z, family = binomial(),
                 control = glm.control(maxit = 200))
      coef_score <- coef(fit)[["score"]]
      if (!is.finite(coef_score) || coef_score < 0) return(NULL)
      if (abs(coef_score) > 50) return(NULL)
      fit
    }, error = function(e) NULL)
  )
}

apply_platt <- function(platt, scores) {
  scores <- as.numeric(scores)
  if (is.null(platt)) {
    z <- as.numeric(scale(scores))
    z[!is.finite(z)] <- 0
    safe_prob(plogis(z))
  } else {
    safe_prob(predict(platt, newdata = data.frame(score = scores),
                      type = "response"))
  }
}
```

## Text Vectorization Functions

```{r vectorization}
make_counts_dfm <- function(text_data, stop_list) {
  quanteda::tokens(text_data, remove_punct = TRUE, remove_numbers = TRUE) %>%
    quanteda::tokens_remove(pattern = stop_list) %>%
    quanteda::dfm()
}

fit_vectorizer <- function(text_data, stop_list, min_df = min_docfreq) {
  x <- make_counts_dfm(text_data, stop_list)
  if (min_df > 1L)
    x <- quanteda::dfm_trim(x, min_docfreq = min_df, docfreq_type = "count")
  if (quanteda::nfeat(x) == 0) stop("No features remain after filtering.")
  N   <- quanteda::ndoc(x)
  n_k <- quanteda::docfreq(x, scheme = "count")
  idf <- log((N + 1) / (n_k + 1)) + 1
  list(features = quanteda::featnames(x), idf = idf, min_df = min_df)
}

transform_counts_dfm <- function(text_data, vec, stop_list) {
  x <- make_counts_dfm(text_data, stop_list)
  quanteda::dfm_match(x, features = vec$features)
}

transform_tfidf_sparse <- function(text_data, vec, stop_list) {
  x_counts <- transform_counts_dfm(text_data, vec, stop_list)
  x_mat    <- as(x_counts, "dgCMatrix")
  x_mat@x  <- 1 + log(x_mat@x)
  idf_vec  <- vec$idf[match(colnames(x_mat), names(vec$idf))]
  idf_vec[is.na(idf_vec)] <- 0
  x_tfidf  <- x_mat %*% Matrix::Diagonal(x = idf_vec)
  colnames(x_tfidf) <- colnames(x_mat)
  row_norms <- sqrt(Matrix::rowSums(x_tfidf^2))
  row_norms[row_norms == 0] <- 1
  x_tfidf <- x_tfidf / row_norms
  Matrix::drop0(x_tfidf)
}
```

## Model-Specific Functions

### Naive Bayes

```{r nb_functions}
nb_prob_late <- function(pred_obj) {
  if ("late" %in% colnames(pred_obj))
    safe_prob(pred_obj[, "late"])
  else
    safe_prob(pred_obj[, ncol(pred_obj)])
}

fit_nb_model <- function(x_dfm, y) {
  quanteda.textmodels::textmodel_nb(x_dfm, y)
}

predict_nb_score <- function(fit, newx_dfm) {
  safe_logit(nb_prob_late(
    predict(fit, newdata = newx_dfm, type = "probability")
  ))
}

cv_nb_scores_ckpt <- function(x_dfm, y, k = 5, seed = 123,
                               prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)),
                  done   = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("nb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]
    tr <- setdiff(seq_along(y), va)
    fit <- fit_nb_model(x_dfm[tr, ], y[tr])
    state$scores[va] <- predict_nb_score(fit, x_dfm[va, ])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

### Elastic Net (replaces Ridge)

```{r enet_functions}
fit_enet_fixed <- function(x, y, alpha, lambda) {
  suppressWarnings(
    glmnet::glmnet(
      x = x, y = as.integer(y == "late"),
      family = "binomial", alpha = alpha, lambda = lambda,
      standardize = TRUE, maxit = 1e6
    )
  )
}

enet_has_nonzero <- function(fit) {
  beta <- as.numeric(coef(fit)[-1, 1])
  sum(beta != 0) > 0
}

predict_enet_score <- function(fit, newx) {
  as.numeric(predict(fit, newx = newx, type = "link"))
}

predict_enet_prob <- function(fit, newx) {
  safe_prob(as.numeric(predict(fit, newx = newx, type = "response")))
}

enet_lambda_grid <- function(x, y, alpha, nlambda = 60,
                              lambda.min.ratio = 1e-3) {
  base_fit <- suppressWarnings(glmnet::glmnet(
    x = x, y = as.integer(y == "late"),
    family = "binomial", alpha = alpha, standardize = TRUE,
    nlambda = nlambda, lambda.min.ratio = lambda.min.ratio, maxit = 1e6
  ))
  lams <- as.numeric(base_fit$lambda)
  if (length(lams) == 0) {
    warning("glmnet returned no lambda path for alpha = ", alpha,
            "; using fallback grid.")
    lams <- 10^seq(log10(0.001), log10(100), length.out = nlambda)
  }
  lams
}

tune_enet_ckpt <- function(x, y, inner_k = 5, seed = 123, prefix = "",
                            ckpt_dir,
                            nlambda = 60, lambda.min.ratio = 1e-3) {
  ensure_dir(ckpt_dir)
  result_path <- file.path(ckpt_dir, "result.rds")
  cached <- load_rds_or_null(result_path)
  if (!is.null(cached)) return(cached)

  folds <- make_folds(y, inner_k, seed)
  all_results <- list()

  for (a_idx in seq_along(enet_alpha_grid)) {
    alpha <- enet_alpha_grid[a_idx]
    log_progress("enet_tuning_alpha",
                 paste(prefix, "alpha =", alpha))
    lambda_grid <- enet_lambda_grid(x, y, alpha = alpha,
                                     nlambda = nlambda,
                                     lambda.min.ratio = lambda.min.ratio)
    lambda_results <- vector("list", length(lambda_grid))
    for (g in seq_along(lambda_grid)) {
      lam     <- lambda_grid[g]
      fold_losses <- rep(NA_real_, length(folds))
      fold_ok <- TRUE
      for (j in seq_along(folds)) {
        va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
        fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr],
                              alpha = alpha, lambda = lam)
        if (!enet_has_nonzero(fit)) { fold_ok <- FALSE; break }
        p_va <- predict_enet_prob(fit, x[va, , drop = FALSE])
        fold_losses[j] <- log_loss_safe(y[va], p_va)
      }
      if (!fold_ok) {
        lambda_results[[g]] <- tibble(alpha = alpha, lambda = lam,
                                      cvm = Inf, cvsd = NA_real_)
      } else {
        cvm  <- mean(fold_losses, na.rm = TRUE)
        cvsd <- sd(fold_losses, na.rm = TRUE) /
                sqrt(sum(is.finite(fold_losses)))
        lambda_results[[g]] <- tibble(alpha = alpha, lambda = lam,
                                      cvm = cvm, cvsd = cvsd)
      }
    }
    all_results[[a_idx]] <- bind_rows(lambda_results)
  }

  tuning <- bind_rows(all_results) %>% filter(is.finite(cvm))
  if (nrow(tuning) == 0)
    stop("All alpha/lambda combinations failed. Check your data.")

  best_per_alpha <- tuning %>%
    group_by(alpha) %>%
    summarise(idx_min = which.min(cvm), lambda_min = lambda[idx_min],
              cvm_min = cvm[idx_min], cvsd_min = cvsd[idx_min],
              .groups = "drop") %>%
    mutate(cutoff = cvm_min + cvsd_min)

  best_1se <- tuning %>%
    inner_join(best_per_alpha %>% select(alpha, cutoff), by = "alpha") %>%
    filter(cvm <= cutoff) %>%
    group_by(alpha) %>%
    summarise(lambda_1se = max(lambda), .groups = "drop")

  global_best <- best_per_alpha %>%
    inner_join(best_1se, by = "alpha") %>%
    arrange(cvm_min) %>%
    slice(1)

  out <- list(
    alpha      = global_best$alpha,
    lambda     = global_best$lambda_1se,
    lambda_min = global_best$lambda_min,
    tuning     = tuning
  )
  save_rds_atomic(out, result_path)
  out
}

cv_enet_scores_ckpt <- function(x, y, alpha, lambda, k = 5, seed = 123,
                                 prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)),
                  done   = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("enet_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr],
                          alpha = alpha, lambda = lambda)
    state$scores[va] <- predict_enet_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

cv_enet_probs_ckpt <- function(x, y, alpha, lambda, k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state_prob.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$probs) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(probs = rep(NA_real_, length(y)),
                  done  = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("enet_prob_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_enet_fixed(x[tr, , drop = FALSE], y[tr],
                          alpha = alpha, lambda = lambda)
    state$probs[va] <- predict_enet_prob(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$probs
}
```

### SVM (Platt-calibrated)

```{r svm_functions}
svm_to_matrix <- function(x) {
  if (inherits(x, "dgCMatrix") || inherits(x, "dsCMatrix") ||
      inherits(x, "dgRMatrix")) return(x)
  if (inherits(x, "matrix")) return(x)
  as.matrix(x)
}

fit_svm_fixed <- function(x, y, cost) {
  x_use <- svm_to_matrix(x)
  fit   <- LiblineaR::LiblineaR(data = x_use, target = y,
                                 type = 7, cost = cost)
  pred_tr <- predict(fit, x_use, decisionValues = TRUE)
  raw <- as.numeric(pred_tr$decisionValues[, 1])
  sgn <- if (mean(raw[y == "late"]) >=
             mean(raw[y == "non_late"])) 1 else -1
  list(model = fit, sign = sgn, cost = cost)
}

predict_svm_score <- function(obj, x) {
  pred <- predict(obj$model, svm_to_matrix(x), decisionValues = TRUE)
  obj$sign * as.numeric(pred$decisionValues[, 1])
}

cv_svm_scores_ckpt <- function(x, y, cost, k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)),
                  done   = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("svm_calibration_fold",
                 paste(prefix, "cost", cost, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_svm_fixed(x[tr, , drop = FALSE], y[tr], cost = cost)
    state$scores[va] <- predict_svm_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}

tune_svm_cost_ckpt <- function(x, y,
                                cost_grid = c(0.01, 0.05, 0.1, 0.25,
                                              0.5, 1, 2, 4, 8, 16),
                                inner_k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  rows_path <- file.path(ckpt_dir, "summary.rds")
  state <- load_rds_or_null(rows_path)
  if (is.null(state) || length(state$done) != length(cost_grid)) {
    state <- list(done = rep(FALSE, length(cost_grid)),
                  rows = vector("list", length(cost_grid)))
  }
  for (i in seq_along(cost_grid)) {
    if (isTRUE(state$done[i])) next
    scores_i <- cv_svm_scores_ckpt(
      x = x, y = y, cost = cost_grid[i], k = inner_k,
      seed = seed + i, prefix = prefix,
      ckpt_dir = file.path(ckpt_dir, sprintf("cost_%03d", i))
    )
    state$rows[[i]] <- tibble(cost = cost_grid[i],
                              auc_margin = auc_margin_safe(y, scores_i))
    state$done[i] <- TRUE
    save_rds_atomic(state, rows_path)
  }
  tuning <- bind_rows(state$rows) %>% arrange(desc(auc_margin), cost)
  list(cost = tuning$cost[1], tuning = tuning)
}
```

### XGBoost

```{r xgb_functions}
xgb_grid <- function() {
  tidyr::crossing(
    eta = c(0.01, 0.03, 0.05),
    max_depth = c(3L, 4L, 6L),
    min_child_weight = c(1, 3, 5),
    subsample = c(0.7, 0.85),
    colsample_bytree = c(0.6, 0.8),
    gamma = c(0, 0.1, 0.5)
  ) %>%
    mutate(candidate_id = row_number()) %>%
    select(candidate_id, everything())
}

best_iter_xgb <- function(fit, fallback) {
  bi <- fit$best_iteration
  if (is.null(bi) || length(bi) == 0 || !is.finite(bi))
    as.integer(fallback)
  else
    as.integer(bi)
}

make_xgb_params <- function(row, scale_pos_weight = 1) {
  list(
    booster = "gbtree", objective = "binary:logistic",
    eval_metric = "logloss", tree_method = "hist",
    eta = as.numeric(row$eta),
    max_depth = as.integer(row$max_depth),
    min_child_weight = as.numeric(row$min_child_weight),
    subsample = as.numeric(row$subsample),
    colsample_bytree = as.numeric(row$colsample_bytree),
    gamma = as.numeric(row$gamma),
    scale_pos_weight = scale_pos_weight
  )
}

fit_xgb_fixed <- function(x, y, params, nrounds, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  xgboost::xgb.train(
    params = params,
    data   = xgboost::xgb.DMatrix(x, label = as.integer(y == "late")),
    nrounds = as.integer(nrounds), verbose = 0
  )
}

predict_xgb_score <- function(fit, x) {
  as.numeric(predict(fit, xgboost::xgb.DMatrix(x), outputmargin = TRUE))
}

tune_xgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123,
                                  prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- xgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }
  n_pos <- sum(y == "late"); n_neg <- sum(y == "non_late")
  spw   <- n_neg / max(n_pos, 1)

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row  <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) ||
        length(cand_state$done) != length(folds)) {
      cand_state <- list(oof_prob   = rep(NA_real_, length(y)),
                         best_iters = rep(NA_integer_, length(folds)),
                         done       = rep(FALSE, length(folds)))
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("xgb_tuning_fold",
                   paste(prefix, "candidate", g, "of", nrow(grid),
                         "fold", j, "of", length(folds)))
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      n_pos_tr <- sum(y[tr] == "late")
      n_neg_tr <- sum(y[tr] == "non_late")
      spw_tr   <- n_neg_tr / max(n_pos_tr, 1)
      dtr <- xgboost::xgb.DMatrix(x[tr, , drop = FALSE],
                                    label = as.integer(y[tr] == "late"))
      dva <- xgboost::xgb.DMatrix(x[va, , drop = FALSE],
                                    label = as.integer(y[va] == "late"))
      fit <- xgboost::xgb.train(
        params = make_xgb_params(pars_row, scale_pos_weight = spw_tr),
        data = dtr, nrounds = xgb_max_rounds,
        evals = list(valid = dva),
        early_stopping_rounds = early_stop_rounds, verbose = 0
      )
      cand_state$oof_prob[va]  <- safe_prob(as.numeric(predict(fit, dva)))
      cand_state$best_iters[j] <- best_iter_xgb(fit, xgb_max_rounds)
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir,
                                     sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss     = log_loss_safe(y, cand_state$oof_prob),
        auc          = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(
          median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>%
    arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_xgb_params(best, scale_pos_weight = spw),
       tuning = tuning)
}

cv_xgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)),
                  done   = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("xgb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    n_pos_tr <- sum(y[tr] == "late")
    n_neg_tr <- sum(y[tr] == "non_late")
    fold_params <- modifyList(params, list(
      scale_pos_weight = n_neg_tr / max(n_pos_tr, 1)))
    fit <- fit_xgb_fixed(x[tr, , drop = FALSE], y[tr],
                         params = fold_params, nrounds = nrounds,
                         seed = seed + j)
    state$scores[va] <- predict_xgb_score(fit, x[va, , drop = FALSE])
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

### LightGBM

```{r lgb_functions}
lgb_grid <- function() {
  tidyr::crossing(
    learning_rate    = c(0.01, 0.03, 0.05),
    num_leaves       = c(15L, 31L, 63L),
    min_data_in_leaf = c(5L, 10L, 20L),
    feature_fraction = c(0.6, 0.8),
    bagging_fraction = c(0.7, 0.85)
  ) %>%
    mutate(candidate_id = row_number()) %>%
    select(candidate_id, everything())
}

make_lgb_params <- function(row, seed = 123) {
  list(
    objective = "binary", metric = "binary_logloss", boosting = "gbdt",
    learning_rate    = as.numeric(row$learning_rate),
    num_leaves       = as.integer(row$num_leaves),
    min_data_in_leaf = as.integer(row$min_data_in_leaf),
    feature_fraction = as.numeric(row$feature_fraction),
    bagging_fraction = as.numeric(row$bagging_fraction),
    bagging_freq = 1L, is_unbalance = TRUE, first_metric_only = TRUE,
    seed = seed, feature_fraction_seed = seed, bagging_seed = seed,
    verbosity = -1
  )
}

make_lgb_dataset_params <- function() list(feature_pre_filter = FALSE)

predict_lgb_prob <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "response"))),
    error = function(e) tryCatch(
      do.call(predict, args),
      error = function(e2) rep(0.5, nrow(x))
    )
  )
  safe_prob(as.numeric(out))
}

predict_lgb_score <- function(model, x, num_iteration = NULL) {
  args <- list(model, x)
  if (!is.null(num_iteration)) args$num_iteration <- as.integer(num_iteration)
  out <- tryCatch(
    do.call(predict, c(args, list(type = "raw"))),
    error = function(e) tryCatch(
      do.call(predict, c(args, list(rawscore = TRUE))),
      error = function(e2) {
        p <- predict_lgb_prob(model, x, num_iteration)
        qlogis(safe_prob(p))
      }
    )
  )
  as.numeric(out)
}

fit_lgb_fixed <- function(x, y, params, nrounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x, label = as.integer(y == "late"),
    params = make_lgb_dataset_params(), free_raw_data = TRUE
  )
  lightgbm::lgb.train(
    params = params, data = dtrain,
    nrounds = as.integer(nrounds), verbose = -1
  )
}

fit_lgb_early_stop <- function(x_tr, y_tr, x_va, y_va, params,
                                max_rounds, early_rounds) {
  dtrain <- lightgbm::lgb.Dataset(
    data = x_tr, label = as.integer(y_tr == "late"),
    params = make_lgb_dataset_params(), free_raw_data = FALSE
  )
  dval <- lightgbm::lgb.Dataset(
    data = x_va, label = as.integer(y_va == "late"),
    params = make_lgb_dataset_params(), reference = dtrain,
    free_raw_data = FALSE
  )
  fit <- lightgbm::lgb.train(
    params = params, data = dtrain,
    nrounds = as.integer(max_rounds),
    valids = list(valid = dval),
    early_stopping_rounds = as.integer(early_rounds), verbose = -1
  )
  best_iter <- fit$best_iter
  if (is.null(best_iter) || !is.finite(best_iter) || best_iter < 1)
    best_iter <- as.integer(max_rounds)
  list(model = fit, best_iter = as.integer(best_iter))
}

run_lgb_preflight <- function(text_vec, y, stop_list, seed = 999L) {
  folds <- make_folds(y, 2L, seed)
  va <- folds[[1]]; tr <- setdiff(seq_along(y), va)
  vec  <- fit_vectorizer(text_vec[tr], stop_list)
  x_tr <- transform_tfidf_sparse(text_vec[tr], vec, stop_list)
  x_va <- transform_tfidf_sparse(text_vec[va], vec, stop_list)
  pars_row <- lgb_grid()[1, , drop = FALSE]
  pars     <- make_lgb_params(pars_row, seed = seed)
  result   <- fit_lgb_early_stop(
    x_tr = x_tr, y_tr = y[tr], x_va = x_va, y_va = y[va],
    params = pars, max_rounds = 100L, early_rounds = 20L
  )
  if (!is.finite(result$best_iter))
    stop("LightGBM preflight failed.")
  invisible(result)
}

tune_lgb_model_ckpt <- function(x, y, inner_k = 5, seed = 123,
                                 prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  grid <- lgb_grid()
  folds <- make_folds(y, inner_k, seed)
  summary_path <- file.path(ckpt_dir, "summary.rds")
  summary_state <- load_rds_or_null(summary_path)
  if (is.null(summary_state) || length(summary_state$done) != nrow(grid)) {
    summary_state <- list(done = rep(FALSE, nrow(grid)))
  }

  for (g in seq_len(nrow(grid))) {
    if (isTRUE(summary_state$done[g])) next
    pars_row  <- grid[g, , drop = FALSE]
    cand_path <- file.path(ckpt_dir, sprintf("candidate_%03d.rds", g))
    cand_state <- load_rds_or_null(cand_path)
    if (is.null(cand_state) || length(cand_state$oof_prob) != length(y) ||
        length(cand_state$done) != length(folds)) {
      cand_state <- list(oof_prob   = rep(NA_real_, length(y)),
                         best_iters = rep(NA_integer_, length(folds)),
                         done       = rep(FALSE, length(folds)))
    }
    for (j in seq_along(folds)) {
      if (isTRUE(cand_state$done[j])) next
      log_progress("lgb_tuning_fold",
                   paste(prefix, "candidate", g, "of", nrow(grid),
                         "fold", j, "of", length(folds)))
      va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
      lgb_params_fold <- make_lgb_params(pars_row,
                                          seed = seed + 1000L * g + j)
      result <- fit_lgb_early_stop(
        x_tr = x[tr, , drop = FALSE], y_tr = y[tr],
        x_va = x[va, , drop = FALSE], y_va = y[va],
        params = lgb_params_fold,
        max_rounds = lgb_max_rounds, early_rounds = early_stop_rounds
      )
      cand_state$oof_prob[va] <- predict_lgb_prob(
        result$model, x[va, , drop = FALSE],
        num_iteration = result$best_iter)
      cand_state$best_iters[j] <- result$best_iter
      cand_state$done[j] <- TRUE
      save_rds_atomic(cand_state, cand_path)
    }
    summary_state$done[g] <- TRUE
    save_rds_atomic(summary_state, summary_path)
  }

  tuning_rows <- vector("list", nrow(grid))
  for (g in seq_len(nrow(grid))) {
    pars_row   <- grid[g, , drop = FALSE]
    cand_state <- readRDS(file.path(ckpt_dir,
                                     sprintf("candidate_%03d.rds", g)))
    tuning_rows[[g]] <- pars_row %>%
      mutate(
        log_loss     = log_loss_safe(y, cand_state$oof_prob),
        auc          = auc_safe(y, cand_state$oof_prob),
        best_nrounds = max(1L, as.integer(round(
          median(cand_state$best_iters, na.rm = TRUE))))
      )
  }
  tuning <- bind_rows(tuning_rows) %>%
    arrange(log_loss, desc(auc), best_nrounds)
  best <- tuning[1, , drop = FALSE]
  list(best = best, params = make_lgb_params(best, seed = seed + 9999L),
       tuning = tuning)
}

cv_lgb_scores_ckpt <- function(x, y, params, nrounds, k = 5, seed = 123,
                                prefix = "", ckpt_dir) {
  ensure_dir(ckpt_dir)
  state_path <- file.path(ckpt_dir, "state.rds")
  folds <- make_folds(y, k, seed)
  state <- load_rds_or_null(state_path)
  if (is.null(state) || length(state$scores) != length(y) ||
      length(state$done) != length(folds)) {
    state <- list(scores = rep(NA_real_, length(y)),
                  done   = rep(FALSE, length(folds)))
  }
  for (j in seq_along(folds)) {
    if (isTRUE(state$done[j])) next
    log_progress("lgb_calibration_fold",
                 paste(prefix, "fold", j, "of", length(folds)))
    va <- folds[[j]]; tr <- setdiff(seq_along(y), va)
    fit <- fit_lgb_fixed(
      x = x[tr, , drop = FALSE], y = y[tr],
      params = modifyList(params, list(
        seed = seed + j, feature_fraction_seed = seed + j,
        bagging_seed = seed + j)),
      nrounds = nrounds
    )
    state$scores[va] <- predict_lgb_score(
      fit, x[va, , drop = FALSE], num_iteration = nrounds)
    state$done[j] <- TRUE
    save_rds_atomic(state, state_path)
  }
  state$scores
}
```

### Model saving helper

```{r save_model}
save_model_object <- function(model, path, engine) {
  ensure_dir(dirname(path))
  if (engine == "xgboost") {
    xgboost::xgb.save(model, path)
  } else if (engine == "lightgbm") {
    lightgbm::lgb.save(model, path)
  } else {
    save_rds_atomic(model, path)
  }
  invisible(path)
}
```

## Data Loading and Cleaning

```{r load_data}
log_progress("load_data", "reading corpus and labels")

df <- readr::read_csv(file.path(data_path, "corpus_non_late_late.csv"),
                       show_col_types = FALSE) %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(text) & text != "") %>%
  mutate(row_id = row_number())

dup_corpus <- df %>% count(id, name = "n") %>% filter(n > 1)
if (nrow(dup_corpus) > 0)
  stop(paste0("Duplicate ids in corpus_non_late_late.csv: ",
              paste(head(dup_corpus$id, 20), collapse = ", ")))

labels_raw <- readr::read_csv(
  file.path(data_path, "corpus_non_late_late_labeled.csv"),
  show_col_types = FALSE
) %>%
  mutate(id = as.character(id)) %>%
  filter(label %in% c("late", "non_late")) %>%
  select(id, label)

label_conflicts <- labels_raw %>%
  distinct(id, label) %>%
  count(id, name = "n_labels") %>%
  filter(n_labels > 1)

if (nrow(label_conflicts) > 0)
  stop(paste0("Conflicting labels: ",
              paste(head(label_conflicts$id, 20), collapse = ", ")))

labels <- labels_raw %>%
  distinct(id, label) %>%
  transmute(id, manual_label = label)

df_labeled <- df %>% inner_join(labels, by = "id")
check_binary_y(df_labeled$manual_label, "labeled data")

cat(sprintf("Corpus: %d documents, %d labeled (%d late, %d non_late)\n",
            nrow(df), nrow(df_labeled),
            sum(df_labeled$manual_label == "late"),
            sum(df_labeled$manual_label == "non_late")))
```

### Stopwords

```{r stopwords}
custom_additions <- c("m", "c", "l", "d", "gt", "lt", "br", "href",
                       "fridericus", "heinricus", "conradus", "meus", "tuus")
stops_v1 <- scan(file.path(data_path, "../../stopwords/stopwords_latin1.txt"),
                 what = "character", quiet = TRUE, encoding = "UTF-8")
stops_v2 <- scan(file.path(data_path, "../../stopwords/stopwords_latin2.txt"),
                 what = "character", comment.char = "#",
                 quiet = TRUE, encoding = "UTF-8")
latin_stops <- sort(unique(c(stops_v1, stops_v2, custom_additions)))
latin_stops <- latin_stops[latin_stops != "*"]

cat(sprintf("Stopword list: %d terms\n", length(latin_stops)))
```

### Prepare working data

```{r prepare_data}
lab     <- df_labeled %>% arrange(row_id)
out     <- df %>% left_join(lab %>% select(row_id, manual_label), by = "row_id")
lab_pos <- match(lab$row_id, out$row_id)

outer_folds <- make_folds(lab$manual_label, k_folds_outer, seed = 123)
n_lab       <- nrow(lab)
```

## Run State Initialization

```{r run_state}
init_run_state <- function(n_lab) {
  list(
    preflight_done = FALSE,
    outer_fold_id  = integer(n_lab),

    # OOF probabilities (Platt-calibrated)
    nb_oof    = rep(NA_real_, n_lab),
    enet_oof  = rep(NA_real_, n_lab),
    svm_oof   = rep(NA_real_, n_lab),
    xgb_oof   = rep(NA_real_, n_lab),
    lgb_oof   = rep(NA_real_, n_lab),

    # OOF raw scores
    nb_score_oof    = rep(NA_real_, n_lab),
    enet_score_oof  = rep(NA_real_, n_lab),
    svm_score_oof   = rep(NA_real_, n_lab),
    xgb_score_oof   = rep(NA_real_, n_lab),
    lgb_score_oof   = rep(NA_real_, n_lab),

    # OOF predictions (Youden)
    nb_pred_youden_oof    = rep(NA_character_, n_lab),
    enet_pred_youden_oof  = rep(NA_character_, n_lab),
    svm_pred_youden_oof   = rep(NA_character_, n_lab),
    xgb_pred_youden_oof   = rep(NA_character_, n_lab),
    lgb_pred_youden_oof   = rep(NA_character_, n_lab),

    # OOF predictions (F1)
    nb_pred_f1_oof    = rep(NA_character_, n_lab),
    enet_pred_f1_oof  = rep(NA_character_, n_lab),
    svm_pred_f1_oof   = rep(NA_character_, n_lab),
    xgb_pred_f1_oof   = rep(NA_character_, n_lab),
    lgb_pred_f1_oof   = rep(NA_character_, n_lab),

    outer_threshold_df = tibble(),
    fold_metric_df     = tibble(),
    selected_outer_df  = tibble(),
    model_done = list(),
    full_done  = list()
  )
}

mark_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  state[[bucket]][[key]] <- TRUE
  state
}

is_done <- function(state, key, bucket = c("model_done", "full_done")) {
  bucket <- match.arg(bucket)
  isTRUE(state[[bucket]][[key]])
}

state_path <- file.path(checkpoint_root, "run_state.rds")

state <- load_rds_or_null(state_path)
if (is.null(state) || length(state$nb_oof) != n_lab) {
  state <- init_run_state(n_lab)
  save_rds_atomic(state, state_path)
}
```

## LightGBM Preflight

```{r preflight}
if (!isTRUE(state$preflight_done)) {
  log_progress("lgb_preflight_start", "testing LightGBM")
  invisible(run_lgb_preflight(lab$text, lab$manual_label,
                               latin_stops, seed = 999L))
  state$preflight_done <- TRUE
  save_rds_atomic(state, state_path)
  log_progress("lgb_preflight_done", "LightGBM preflight passed")
}
```

## Nested Cross-Validation (5 models)

The five models are: **Naive Bayes**, **Elastic Net** (replacing pure ridge), **SVM (Platt)**, **XGBoost**, **LightGBM**.

```{r nested_cv}
for (i in seq_along(outer_folds)) {
  log_progress("outer_fold_start",
               paste("fold", i, "of", length(outer_folds)))
  va <- outer_folds[[i]]
  tr <- setdiff(seq_len(n_lab), va)
  y_tr <- lab$manual_label[tr]
  y_va <- lab$manual_label[va]

  outer_dir <- file.path(checkpoint_root, sprintf("outer_fold_%02d", i))
  ensure_dir(outer_dir)

  # Vectorize once per outer fold

  vec_fold   <- fit_vectorizer(lab$text[tr], latin_stops)
  x_tr_dfm   <- transform_counts_dfm(lab$text[tr], vec_fold, latin_stops)
  x_va_dfm   <- transform_counts_dfm(lab$text[va], vec_fold, latin_stops)
  x_tr       <- transform_tfidf_sparse(lab$text[tr], vec_fold, latin_stops)
  x_va       <- transform_tfidf_sparse(lab$text[va], vec_fold, latin_stops)

  state$outer_fold_id[va] <- i
  save_rds_atomic(state, state_path)

  # ── 1. LightGBM ────────────────────────────────────────────────────────────
  key <- paste0("outer_", i, "_lightgbm")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_lgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "lightgbm")
    lgb_tuned <- tune_lgb_model_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 5000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning"))
    lgb_train_scores <- cv_lgb_scores_ckpt(
      x_tr, y_tr, params = lgb_tuned$params,
      nrounds = as.integer(lgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner, seed = 5100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration"))
    lgb_platt     <- fit_platt(lgb_train_scores, y_tr)
    lgb_fit       <- fit_lgb_fixed(x_tr, y_tr, params = lgb_tuned$params,
                                    nrounds = as.integer(
                                      lgb_tuned$best$best_nrounds[1]))
    lgb_va_scores <- predict_lgb_score(
      lgb_fit, x_va,
      num_iteration = as.integer(lgb_tuned$best$best_nrounds[1]))
    lgb_train_prob <- apply_platt(lgb_platt, lgb_train_scores)
    lgb_va_prob    <- apply_platt(lgb_platt, lgb_va_scores)
    lgb_th_y <- choose_threshold(y_tr, lgb_train_prob, "youden")
    lgb_th_f <- choose_threshold(y_tr, lgb_train_prob, "f1")

    state$lgb_oof[va]             <- lgb_va_prob
    state$lgb_score_oof[va]       <- lgb_va_scores
    state$lgb_pred_youden_oof[va] <- ifelse(lgb_va_prob >= lgb_th_y,
                                             "late", "non_late")
    state$lgb_pred_f1_oof[va]     <- ifelse(lgb_va_prob >= lgb_th_f,
                                             "late", "non_late")

    threshold_row <- tibble(outer_fold = i, model = "lightgbm",
                            threshold_youden = lgb_th_y,
                            threshold_f1 = lgb_th_f)
    metric_row <- summarize_model_fold(y_va, lgb_va_prob,
                                        state$lgb_pred_youden_oof[va],
                                        state$lgb_pred_f1_oof[va],
                                        "lightgbm", i)
    selected_row <- lgb_tuned$best %>%
      mutate(outer_fold = i, model = "lightgbm",
             threshold_youden = lgb_th_y, threshold_f1 = lgb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row,
                                             c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row,
                                         c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                            selected_row,
                                            c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # ── 2. Naive Bayes ─────────────────────────────────────────────────────────
  key <- paste0("outer_", i, "_naive_bayes")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_nb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "naive_bayes")
    nb_train_scores <- cv_nb_scores_ckpt(
      x_tr_dfm, y_tr, k = k_folds_inner, seed = 1000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration"))
    nb_platt     <- fit_platt(nb_train_scores, y_tr)
    nb_fit       <- fit_nb_model(x_tr_dfm, y_tr)
    nb_va_scores <- predict_nb_score(nb_fit, x_va_dfm)
    nb_train_prob <- apply_platt(nb_platt, nb_train_scores)
    nb_va_prob    <- apply_platt(nb_platt, nb_va_scores)
    nb_th_y <- choose_threshold(y_tr, nb_train_prob, "youden")
    nb_th_f <- choose_threshold(y_tr, nb_train_prob, "f1")

    state$nb_oof[va]             <- nb_va_prob
    state$nb_score_oof[va]       <- nb_va_scores
    state$nb_pred_youden_oof[va] <- ifelse(nb_va_prob >= nb_th_y,
                                            "late", "non_late")
    state$nb_pred_f1_oof[va]     <- ifelse(nb_va_prob >= nb_th_f,
                                            "late", "non_late")

    threshold_row <- tibble(outer_fold = i, model = "naive_bayes",
                            threshold_youden = nb_th_y,
                            threshold_f1 = nb_th_f)
    metric_row <- summarize_model_fold(y_va, nb_va_prob,
                                        state$nb_pred_youden_oof[va],
                                        state$nb_pred_f1_oof[va],
                                        "naive_bayes", i)
    selected_row <- tibble(outer_fold = i, model = "naive_bayes",
                           threshold_youden = nb_th_y,
                           threshold_f1 = nb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row,
                                             c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row,
                                         c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                            selected_row,
                                            c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # ── 3. Elastic Net (replaces ridge) ─────────────────────────────────────────
  key <- paste0("outer_", i, "_elastic_net")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_enet_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "elastic_net")

    enet_tuned <- tune_enet_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 2000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning"))

    log_progress("enet_outer_tuned",
                 paste("outer", i, "| alpha =", enet_tuned$alpha,
                       "| lambda =", round(enet_tuned$lambda, 6)))

    # Cross-fitted training probabilities for threshold selection
    enet_train_prob <- cv_enet_probs_ckpt(
      x_tr, y_tr, alpha = enet_tuned$alpha, lambda = enet_tuned$lambda,
      k = k_folds_inner, seed = 2100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration_prob"))

    # Fit on full training fold, predict on validation
    enet_fit       <- fit_enet_fixed(x_tr, y_tr,
                                     alpha = enet_tuned$alpha,
                                     lambda = enet_tuned$lambda)
    enet_va_scores <- predict_enet_score(enet_fit, x_va)
    enet_va_prob   <- predict_enet_prob(enet_fit, x_va)

    enet_th_y <- choose_threshold(y_tr, enet_train_prob, "youden")
    enet_th_f <- choose_threshold(y_tr, enet_train_prob, "f1")

    state$enet_oof[va]             <- enet_va_prob
    state$enet_score_oof[va]       <- enet_va_scores
    state$enet_pred_youden_oof[va] <- ifelse(enet_va_prob >= enet_th_y,
                                              "late", "non_late")
    state$enet_pred_f1_oof[va]     <- ifelse(enet_va_prob >= enet_th_f,
                                              "late", "non_late")

    threshold_row <- tibble(outer_fold = i, model = "elastic_net",
                            threshold_youden = enet_th_y,
                            threshold_f1 = enet_th_f)
    metric_row <- summarize_model_fold(y_va, enet_va_prob,
                                        state$enet_pred_youden_oof[va],
                                        state$enet_pred_f1_oof[va],
                                        "elastic_net", i)
    selected_row <- tibble(outer_fold = i, model = "elastic_net",
                           alpha = enet_tuned$alpha,
                           lambda_1se = enet_tuned$lambda,
                           lambda_min = enet_tuned$lambda_min,
                           threshold_youden = enet_th_y,
                           threshold_f1 = enet_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row,
                                             c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row,
                                         c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                            selected_row,
                                            c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # ── 4. SVM (Platt) ─────────────────────────────────────────────────────────
  key <- paste0("outer_", i, "_svm_platt")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_svm_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "svm_platt")
    svm_tuned <- tune_svm_cost_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 3000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning"))
    svm_train_scores <- cv_svm_scores_ckpt(
      x_tr, y_tr, cost = svm_tuned$cost, k = k_folds_inner,
      seed = 3100 + i, prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration"))
    svm_platt     <- fit_platt(svm_train_scores, y_tr)
    svm_fit       <- fit_svm_fixed(x_tr, y_tr, cost = svm_tuned$cost)
    svm_va_scores <- predict_svm_score(svm_fit, x_va)
    svm_train_prob <- apply_platt(svm_platt, svm_train_scores)
    svm_va_prob    <- apply_platt(svm_platt, svm_va_scores)
    svm_th_y <- choose_threshold(y_tr, svm_train_prob, "youden")
    svm_th_f <- choose_threshold(y_tr, svm_train_prob, "f1")

    state$svm_oof[va]             <- svm_va_prob
    state$svm_score_oof[va]       <- svm_va_scores
    state$svm_pred_youden_oof[va] <- ifelse(svm_va_prob >= svm_th_y,
                                             "late", "non_late")
    state$svm_pred_f1_oof[va]     <- ifelse(svm_va_prob >= svm_th_f,
                                             "late", "non_late")

    threshold_row <- tibble(outer_fold = i, model = "svm_platt",
                            threshold_youden = svm_th_y,
                            threshold_f1 = svm_th_f)
    metric_row <- summarize_model_fold(y_va, svm_va_prob,
                                        state$svm_pred_youden_oof[va],
                                        state$svm_pred_f1_oof[va],
                                        "svm_platt", i)
    selected_row <- tibble(outer_fold = i, model = "svm_platt",
                           cost = svm_tuned$cost,
                           threshold_youden = svm_th_y,
                           threshold_f1 = svm_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row,
                                             c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row,
                                         c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                            selected_row,
                                            c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  # ── 5. XGBoost ─────────────────────────────────────────────────────────────
  key <- paste0("outer_", i, "_xgboost")
  if (!is_done(state, key, "model_done")) {
    log_progress("fit_xgb_outer", paste("outer fold", i))
    model_dir <- file.path(outer_dir, "xgboost")
    xgb_tuned <- tune_xgb_model_ckpt(
      x_tr, y_tr, inner_k = k_folds_inner, seed = 4000 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "tuning"))
    xgb_train_scores <- cv_xgb_scores_ckpt(
      x_tr, y_tr, params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      k = k_folds_inner, seed = 4100 + i,
      prefix = paste("outer", i),
      ckpt_dir = file.path(model_dir, "calibration"))
    xgb_platt    <- fit_platt(xgb_train_scores, y_tr)
    xgb_fit      <- fit_xgb_fixed(
      x_tr, y_tr, params = xgb_tuned$params,
      nrounds = as.integer(xgb_tuned$best$best_nrounds[1]),
      seed = 4200 + i)
    xgb_va_scores <- predict_xgb_score(xgb_fit, x_va)
    xgb_train_prob <- apply_platt(xgb_platt, xgb_train_scores)
    xgb_va_prob    <- apply_platt(xgb_platt, xgb_va_scores)
    xgb_th_y <- choose_threshold(y_tr, xgb_train_prob, "youden")
    xgb_th_f <- choose_threshold(y_tr, xgb_train_prob, "f1")

    state$xgb_oof[va]             <- xgb_va_prob
    state$xgb_score_oof[va]       <- xgb_va_scores
    state$xgb_pred_youden_oof[va] <- ifelse(xgb_va_prob >= xgb_th_y,
                                             "late", "non_late")
    state$xgb_pred_f1_oof[va]     <- ifelse(xgb_va_prob >= xgb_th_f,
                                             "late", "non_late")

    threshold_row <- tibble(outer_fold = i, model = "xgboost",
                            threshold_youden = xgb_th_y,
                            threshold_f1 = xgb_th_f)
    metric_row <- summarize_model_fold(y_va, xgb_va_prob,
                                        state$xgb_pred_youden_oof[va],
                                        state$xgb_pred_f1_oof[va],
                                        "xgboost", i)
    selected_row <- xgb_tuned$best %>%
      mutate(outer_fold = i, model = "xgboost",
             threshold_youden = xgb_th_y, threshold_f1 = xgb_th_f)

    state$outer_threshold_df <- upsert_rows(state$outer_threshold_df,
                                             threshold_row,
                                             c("outer_fold", "model"))
    state$fold_metric_df <- upsert_rows(state$fold_metric_df, metric_row,
                                         c("outer_fold", "model"))
    state$selected_outer_df <- upsert_rows(state$selected_outer_df,
                                            selected_row,
                                            c("outer_fold", "model"))
    state <- mark_done(state, key, "model_done")
    save_rds_atomic(state, state_path)
  }

  log_progress("outer_fold_done",
               paste("fold", i, "of", length(outer_folds)))
}
```

## OOF Evaluation Metrics

```{r oof_metrics}
oof_metrics <- bind_rows(
  summarize_model_oof(lab$manual_label, state$nb_oof,
                       state$nb_pred_youden_oof, state$nb_pred_f1_oof,
                       "naive_bayes"),
  summarize_model_oof(lab$manual_label, state$enet_oof,
                       state$enet_pred_youden_oof, state$enet_pred_f1_oof,
                       "elastic_net"),
  summarize_model_oof(lab$manual_label, state$svm_oof,
                       state$svm_pred_youden_oof, state$svm_pred_f1_oof,
                       "svm_platt"),
  summarize_model_oof(lab$manual_label, state$xgb_oof,
                       state$xgb_pred_youden_oof, state$xgb_pred_f1_oof,
                       "xgboost"),
  summarize_model_oof(lab$manual_label, state$lgb_oof,
                       state$lgb_pred_youden_oof, state$lgb_pred_f1_oof,
                       "lightgbm")
)

cat("\n===== OOF Metrics Summary =====\n")
oof_metrics %>%
  select(model, auc, log_loss, brier,
         f1_youden, mcc_youden, f1_f1, mcc_f1) %>%
  as.data.frame() %>%
  print()

oof_predictions <- lab %>%
  transmute(
    row_id, id, manual_label,
    outer_fold         = state$outer_fold_id,
    nb_oof             = state$nb_oof,
    nb_score_oof       = state$nb_score_oof,
    nb_pred_youden_oof = state$nb_pred_youden_oof,
    nb_pred_f1_oof     = state$nb_pred_f1_oof,
    enet_oof             = state$enet_oof,
    enet_score_oof       = state$enet_score_oof,
    enet_pred_youden_oof = state$enet_pred_youden_oof,
    enet_pred_f1_oof     = state$enet_pred_f1_oof,
    svm_oof             = state$svm_oof,
    svm_score_oof       = state$svm_score_oof,
    svm_pred_youden_oof = state$svm_pred_youden_oof,
    svm_pred_f1_oof     = state$svm_pred_f1_oof,
    xgb_oof             = state$xgb_oof,
    xgb_score_oof       = state$xgb_score_oof,
    xgb_pred_youden_oof = state$xgb_pred_youden_oof,
    xgb_pred_f1_oof     = state$xgb_pred_f1_oof,
    lgb_oof             = state$lgb_oof,
    lgb_score_oof       = state$lgb_score_oof,
    lgb_pred_youden_oof = state$lgb_pred_youden_oof,
    lgb_pred_f1_oof     = state$lgb_pred_f1_oof
  )

write_csv_atomic(arrange(state$outer_threshold_df, outer_fold, model),
                 file.path(output_dir, "outer_thresholds.csv"))
write_csv_atomic(arrange(state$fold_metric_df, outer_fold, model),
                 file.path(output_dir, "outer_fold_metrics.csv"))
write_csv_atomic(arrange(state$selected_outer_df, outer_fold, model),
                 file.path(output_dir, "selected_outer_hyperparameters.csv"))
write_csv_atomic(oof_metrics,
                 file.path(output_dir, "oof_model_metrics.csv"))
write_csv_atomic(oof_predictions,
                 file.path(output_dir, "oof_predictions_labeled.csv"))
```

## Full-Data Deployment Models

```{r full_fit}
full_dir <- file.path(checkpoint_root, "full_fit")
ensure_dir(full_dir)

# Vectorize on all labeled data
vec_full_path <- file.path(full_dir, "vectorizer_full.rds")
vec_full <- load_rds_or_null(vec_full_path)
if (is.null(vec_full)) {
  vec_full <- fit_vectorizer(lab$text, latin_stops)
  save_rds_atomic(vec_full, vec_full_path)
}

x_lab_dfm_full <- transform_counts_dfm(lab$text, vec_full, latin_stops)
x_all_dfm_full <- transform_counts_dfm(out$text, vec_full, latin_stops)
x_lab_full     <- transform_tfidf_sparse(lab$text, vec_full, latin_stops)
x_all_full     <- transform_tfidf_sparse(out$text, vec_full, latin_stops)

full_predictions <- tibble(row_id = out$row_id, id = out$id,
                            manual_label = out$manual_label)

# ── Naive Bayes ───────────────────────────────────────────────────────────────
key <- "full_naive_bayes"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_nb_full", "full-data fit")
  model_dir <- file.path(full_dir, "naive_bayes")
  nb_cv_scores <- cv_nb_scores_ckpt(
    x_lab_dfm_full, lab$manual_label, k = k_folds_inner, seed = 8001,
    prefix = "full", ckpt_dir = file.path(model_dir, "calibration"))
  nb_platt <- fit_platt(nb_cv_scores, lab$manual_label)
  nb_fit   <- fit_nb_model(x_lab_dfm_full, lab$manual_label)
  save_model_object(nb_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(nb_platt, file.path(model_dir, "platt.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

nb_fit   <- readRDS(file.path(full_dir, "naive_bayes", "model.rds"))
nb_platt <- load_rds_or_null(file.path(full_dir, "naive_bayes", "platt.rds"))
nb_all_scores <- predict_nb_score(nb_fit, x_all_dfm_full)
full_predictions <- full_predictions %>%
  mutate(nb_score = nb_all_scores,
         nb_prob  = apply_platt(nb_platt, nb_all_scores))

# ── Elastic Net (replaces ridge) ─────────────────────────────────────────────
key <- "full_elastic_net"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_enet_full", "full-data fit")
  model_dir <- file.path(full_dir, "elastic_net")
  enet_tuned_full <- tune_enet_ckpt(
    x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8002,
    prefix = "full", ckpt_dir = file.path(model_dir, "tuning"))
  enet_cv_scores <- cv_enet_scores_ckpt(
    x_lab_full, lab$manual_label,
    alpha = enet_tuned_full$alpha, lambda = enet_tuned_full$lambda,
    k = k_folds_inner, seed = 8102, prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration"))
  enet_platt <- fit_platt(enet_cv_scores, lab$manual_label)
  enet_fit   <- fit_enet_fixed(x_lab_full, lab$manual_label,
                                alpha = enet_tuned_full$alpha,
                                lambda = enet_tuned_full$lambda)
  save_model_object(enet_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(enet_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(enet_tuned_full, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

enet_fit          <- readRDS(file.path(full_dir, "elastic_net", "model.rds"))
enet_platt        <- load_rds_or_null(file.path(full_dir, "elastic_net",
                                                 "platt.rds"))
enet_tuned_full   <- readRDS(file.path(full_dir, "elastic_net",
                                        "tuning_result.rds"))
enet_all_scores   <- predict_enet_score(enet_fit, x_all_full)
full_predictions  <- full_predictions %>%
  mutate(enet_score = enet_all_scores,
         enet_prob  = apply_platt(enet_platt, enet_all_scores))

# ── SVM (Platt) ──────────────────────────────────────────────────────────────
key <- "full_svm_platt"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_svm_full", "full-data fit")
  model_dir <- file.path(full_dir, "svm_platt")
  svm_tuned_full <- tune_svm_cost_ckpt(
    x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8003,
    prefix = "full", ckpt_dir = file.path(model_dir, "tuning"))
  svm_cv_scores <- cv_svm_scores_ckpt(
    x_lab_full, lab$manual_label, cost = svm_tuned_full$cost,
    k = k_folds_inner, seed = 8103, prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration"))
  svm_platt <- fit_platt(svm_cv_scores, lab$manual_label)
  svm_fit   <- fit_svm_fixed(x_lab_full, lab$manual_label,
                              cost = svm_tuned_full$cost)
  save_model_object(svm_fit, file.path(model_dir, "model.rds"), "rds")
  save_rds_atomic(svm_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(svm_tuned_full, file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

svm_fit        <- readRDS(file.path(full_dir, "svm_platt", "model.rds"))
svm_platt      <- load_rds_or_null(file.path(full_dir, "svm_platt",
                                              "platt.rds"))
svm_all_scores <- predict_svm_score(svm_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(svm_score = svm_all_scores,
         svm_prob  = apply_platt(svm_platt, svm_all_scores))

# ── XGBoost ──────────────────────────────────────────────────────────────────
key <- "full_xgboost"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_xgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "xgboost")
  xgb_tuned_full <- tune_xgb_model_ckpt(
    x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8004,
    prefix = "full", ckpt_dir = file.path(model_dir, "tuning"))
  xgb_cv_scores <- cv_xgb_scores_ckpt(
    x_lab_full, lab$manual_label, params = xgb_tuned_full$params,
    nrounds = as.integer(xgb_tuned_full$best$best_nrounds[1]),
    k = k_folds_inner, seed = 8104, prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration"))
  xgb_platt <- fit_platt(xgb_cv_scores, lab$manual_label)
  xgb_fit   <- fit_xgb_fixed(
    x_lab_full, lab$manual_label, params = xgb_tuned_full$params,
    nrounds = as.integer(xgb_tuned_full$best$best_nrounds[1]),
    seed = 8204)
  save_model_object(xgb_fit, file.path(model_dir, "model.xgb"), "xgboost")
  save_rds_atomic(xgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = xgb_tuned_full$best,
                       tuning = xgb_tuned_full$tuning),
                  file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

xgb_fit        <- xgboost::xgb.load(file.path(full_dir, "xgboost",
                                                "model.xgb"))
xgb_platt      <- load_rds_or_null(file.path(full_dir, "xgboost",
                                              "platt.rds"))
xgb_all_scores <- predict_xgb_score(xgb_fit, x_all_full)
full_predictions <- full_predictions %>%
  mutate(xgb_score = xgb_all_scores,
         xgb_prob  = apply_platt(xgb_platt, xgb_all_scores))

# ── LightGBM ────────────────────────────────────────────────────────────────
key <- "full_lightgbm"
if (!is_done(state, key, "full_done")) {
  log_progress("fit_lgb_full", "full-data fit")
  model_dir <- file.path(full_dir, "lightgbm")
  lgb_tuned_full <- tune_lgb_model_ckpt(
    x_lab_full, lab$manual_label, inner_k = k_folds_inner, seed = 8005,
    prefix = "full", ckpt_dir = file.path(model_dir, "tuning"))
  lgb_cv_scores <- cv_lgb_scores_ckpt(
    x_lab_full, lab$manual_label, params = lgb_tuned_full$params,
    nrounds = as.integer(lgb_tuned_full$best$best_nrounds[1]),
    k = k_folds_inner, seed = 8105, prefix = "full",
    ckpt_dir = file.path(model_dir, "calibration"))
  lgb_platt <- fit_platt(lgb_cv_scores, lab$manual_label)
  lgb_fit   <- fit_lgb_fixed(
    x = x_lab_full, y = lab$manual_label,
    params = lgb_tuned_full$params,
    nrounds = as.integer(lgb_tuned_full$best$best_nrounds[1]))
  save_model_object(lgb_fit, file.path(model_dir, "model.txt"), "lightgbm")
  save_rds_atomic(lgb_platt, file.path(model_dir, "platt.rds"))
  save_rds_atomic(list(best = lgb_tuned_full$best,
                       tuning = lgb_tuned_full$tuning),
                  file.path(model_dir, "tuning_result.rds"))
  state <- mark_done(state, key, "full_done")
  save_rds_atomic(state, state_path)
}

lgb_fit <- lightgbm::lgb.load(file.path(full_dir, "lightgbm", "model.txt"))
lgb_platt <- load_rds_or_null(file.path(full_dir, "lightgbm", "platt.rds"))
lgb_tuning_result <- readRDS(file.path(full_dir, "lightgbm",
                                        "tuning_result.rds"))
lgb_all_scores <- predict_lgb_score(
  lgb_fit, x_all_full,
  num_iteration = as.integer(lgb_tuning_result$best$best_nrounds[1]))
full_predictions <- full_predictions %>%
  mutate(lgb_score = lgb_all_scores,
         lgb_prob  = apply_platt(lgb_platt, lgb_all_scores))

write_csv_atomic(full_predictions,
                 file.path(output_dir, "full_corpus_model_scores.csv"))
```

## Deployment Thresholds and Final Corpus Scoring

```{r deployment}
log_progress("deployment", "computing deployment thresholds and final labels")

# Cross-fitted calibration probabilities on labeled data for each model
# (already computed during full-fit; re-derive from saved calibration scores)

# NB
nb_cv_scores_full <- cv_nb_scores_ckpt(
  x_lab_dfm_full, lab$manual_label, k = k_folds_inner, seed = 8001,
  prefix = "full_deploy",
  ckpt_dir = file.path(full_dir, "naive_bayes", "calibration"))
nb_platt_full <- load_rds_or_null(file.path(full_dir, "naive_bayes",
                                             "platt.rds"))
nb_lab_prob_cf <- apply_platt(nb_platt_full, nb_cv_scores_full)

# Elastic net
enet_cv_scores_full <- cv_enet_scores_ckpt(
  x_lab_full, lab$manual_label,
  alpha = enet_tuned_full$alpha, lambda = enet_tuned_full$lambda,
  k = k_folds_inner, seed = 8102, prefix = "full_deploy",
  ckpt_dir = file.path(full_dir, "elastic_net", "calibration"))
enet_platt_full <- load_rds_or_null(file.path(full_dir, "elastic_net",
                                               "platt.rds"))
enet_lab_prob_cf <- apply_platt(enet_platt_full, enet_cv_scores_full)

# SVM
svm_tuned_full <- readRDS(file.path(full_dir, "svm_platt",
                                     "tuning_result.rds"))
svm_cv_scores_full <- cv_svm_scores_ckpt(
  x_lab_full, lab$manual_label, cost = svm_tuned_full$cost,
  k = k_folds_inner, seed = 8103, prefix = "full_deploy",
  ckpt_dir = file.path(full_dir, "svm_platt", "calibration"))
svm_platt_full <- load_rds_or_null(file.path(full_dir, "svm_platt",
                                              "platt.rds"))
svm_lab_prob_cf <- apply_platt(svm_platt_full, svm_cv_scores_full)

# XGBoost
xgb_tuning_result <- readRDS(file.path(full_dir, "xgboost",
                                        "tuning_result.rds"))
xgb_cv_scores_full <- cv_xgb_scores_ckpt(
  x_lab_full, lab$manual_label, params = xgb_tuning_result$best %>%
    { make_xgb_params(., scale_pos_weight =
        sum(lab$manual_label == "non_late") /
        max(sum(lab$manual_label == "late"), 1)) },
  nrounds = as.integer(xgb_tuning_result$best$best_nrounds[1]),
  k = k_folds_inner, seed = 8104, prefix = "full_deploy",
  ckpt_dir = file.path(full_dir, "xgboost", "calibration"))
xgb_platt_full <- load_rds_or_null(file.path(full_dir, "xgboost",
                                              "platt.rds"))
xgb_lab_prob_cf <- apply_platt(xgb_platt_full, xgb_cv_scores_full)

# LightGBM
lgb_cv_scores_full <- cv_lgb_scores_ckpt(
  x_lab_full, lab$manual_label, params = lgb_tuning_result$best %>%
    { make_lgb_params(., seed = 8005 + 9999L) },
  nrounds = as.integer(lgb_tuning_result$best$best_nrounds[1]),
  k = k_folds_inner, seed = 8105, prefix = "full_deploy",
  ckpt_dir = file.path(full_dir, "lightgbm", "calibration"))
lgb_platt_full <- load_rds_or_null(file.path(full_dir, "lightgbm",
                                              "platt.rds"))
lgb_lab_prob_cf <- apply_platt(lgb_platt_full, lgb_cv_scores_full)

# Deployment thresholds
nb_deploy_th_y   <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "youden")
nb_deploy_th_f   <- choose_threshold(lab$manual_label, nb_lab_prob_cf, "f1")
enet_deploy_th_y <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "youden")
enet_deploy_th_f <- choose_threshold(lab$manual_label, enet_lab_prob_cf, "f1")
svm_deploy_th_y  <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "youden")
svm_deploy_th_f  <- choose_threshold(lab$manual_label, svm_lab_prob_cf, "f1")
xgb_deploy_th_y  <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "youden")
xgb_deploy_th_f  <- choose_threshold(lab$manual_label, xgb_lab_prob_cf, "f1")
lgb_deploy_th_y  <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "youden")
lgb_deploy_th_f  <- choose_threshold(lab$manual_label, lgb_lab_prob_cf, "f1")

deployment_thresholds <- tibble(
  model = c("naive_bayes", "elastic_net", "svm_platt", "xgboost", "lightgbm"),
  threshold_youden = c(nb_deploy_th_y, enet_deploy_th_y, svm_deploy_th_y,
                       xgb_deploy_th_y, lgb_deploy_th_y),
  threshold_f1 = c(nb_deploy_th_f, enet_deploy_th_f, svm_deploy_th_f,
                   xgb_deploy_th_f, lgb_deploy_th_f),
  threshold_scope = "full_labeled_cross_fitted_calibrated"
)

get_threshold <- function(model_name, kind = c("youden", "f1")) {
  kind <- match.arg(kind)
  deployment_thresholds %>%
    filter(model == model_name) %>%
    pull(if (kind == "youden") threshold_youden else threshold_f1) %>%
    .[[1]]
}

# Build full-corpus predictions with cross-fitted probs for labeled rows
prob_nb   <- full_predictions$nb_prob
prob_enet <- full_predictions$enet_prob
prob_svm  <- full_predictions$svm_prob
prob_xgb  <- full_predictions$xgb_prob
prob_lgb  <- full_predictions$lgb_prob

score_nb   <- full_predictions$nb_score
score_enet <- full_predictions$enet_score
score_svm  <- full_predictions$svm_score
score_xgb  <- full_predictions$xgb_score
score_lgb  <- full_predictions$lgb_score

# Replace labeled rows with cross-fitted values
prob_nb[lab_pos]   <- nb_lab_prob_cf
prob_enet[lab_pos] <- enet_lab_prob_cf
prob_svm[lab_pos]  <- svm_lab_prob_cf
prob_xgb[lab_pos]  <- xgb_lab_prob_cf
prob_lgb[lab_pos]  <- lgb_lab_prob_cf

score_nb[lab_pos]   <- nb_cv_scores_full
score_enet[lab_pos] <- enet_cv_scores_full
score_svm[lab_pos]  <- svm_cv_scores_full
score_xgb[lab_pos]  <- xgb_cv_scores_full
score_lgb[lab_pos]  <- lgb_cv_scores_full

out <- out %>%
  mutate(
    prediction_source = if_else(!is.na(manual_label),
                                "cross_fitted_labeled",
                                "full_fit_unlabeled"),

    prob_nb = safe_prob(prob_nb),
    score_nb_logit = score_nb,
    label_nb_youden = if_else(prob_nb >= get_threshold("naive_bayes", "youden"),
                              "late", "non_late"),
    label_nb_f1 = if_else(prob_nb >= get_threshold("naive_bayes", "f1"),
                          "late", "non_late"),

    prob_enet = safe_prob(prob_enet),
    score_enet_logit = score_enet,
    label_enet_youden = if_else(prob_enet >= get_threshold("elastic_net", "youden"),
                                "late", "non_late"),
    label_enet_f1 = if_else(prob_enet >= get_threshold("elastic_net", "f1"),
                            "late", "non_late"),

    prob_svm = safe_prob(prob_svm),
    score_svm_margin = score_svm,
    score_svm_logit  = safe_logit(prob_svm),
    label_svm_youden = if_else(prob_svm >= get_threshold("svm_platt", "youden"),
                               "late", "non_late"),
    label_svm_f1 = if_else(prob_svm >= get_threshold("svm_platt", "f1"),
                           "late", "non_late"),

    prob_xgb = safe_prob(prob_xgb),
    score_xgb_margin = score_xgb,
    score_xgb_logit  = safe_logit(prob_xgb),
    label_xgb_youden = if_else(prob_xgb >= get_threshold("xgboost", "youden"),
                               "late", "non_late"),
    label_xgb_f1 = if_else(prob_xgb >= get_threshold("xgboost", "f1"),
                           "late", "non_late"),

    prob_lgb = safe_prob(prob_lgb),
    score_lgb_margin = score_lgb,
    score_lgb_logit  = safe_logit(prob_lgb),
    label_lgb_youden = if_else(prob_lgb >= get_threshold("lightgbm", "youden"),
                               "late", "non_late"),
    label_lgb_f1 = if_else(prob_lgb >= get_threshold("lightgbm", "f1"),
                           "late", "non_late")
  )
```

## OOF Confusion Counts and Eval Predictions

```{r oof_eval}
oof_eval_predictions <- tibble(
  id = lab$id, row_id = lab$row_id,
  outer_fold   = state$outer_fold_id,
  manual_label = lab$manual_label,

  prob_nb = state$nb_oof, score_nb_logit = state$nb_score_oof,
  label_nb_youden = state$nb_pred_youden_oof,
  label_nb_f1     = state$nb_pred_f1_oof,

  prob_enet = state$enet_oof, score_enet_logit = state$enet_score_oof,
  label_enet_youden = state$enet_pred_youden_oof,
  label_enet_f1     = state$enet_pred_f1_oof,

  prob_svm = state$svm_oof, score_svm_margin = state$svm_score_oof,
  label_svm_youden = state$svm_pred_youden_oof,
  label_svm_f1     = state$svm_pred_f1_oof,

  prob_xgb = state$xgb_oof, score_xgb_margin = state$xgb_score_oof,
  label_xgb_youden = state$xgb_pred_youden_oof,
  label_xgb_f1     = state$xgb_pred_f1_oof,

  prob_lgb = state$lgb_oof, score_lgb_margin = state$lgb_score_oof,
  label_lgb_youden = state$lgb_pred_youden_oof,
  label_lgb_f1     = state$lgb_pred_f1_oof
)

oof_confusion_counts <- bind_rows(
  hard_metrics(lab$manual_label, state$nb_pred_youden_oof) %>%
    mutate(model = "naive_bayes", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, state$enet_pred_youden_oof) %>%
    mutate(model = "elastic_net", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, state$svm_pred_youden_oof) %>%
    mutate(model = "svm_platt", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, state$xgb_pred_youden_oof) %>%
    mutate(model = "xgboost", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, state$lgb_pred_youden_oof) %>%
    mutate(model = "lightgbm", threshold_kind = "youden"),
  hard_metrics(lab$manual_label, state$nb_pred_f1_oof) %>%
    mutate(model = "naive_bayes", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, state$enet_pred_f1_oof) %>%
    mutate(model = "elastic_net", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, state$svm_pred_f1_oof) %>%
    mutate(model = "svm_platt", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, state$xgb_pred_f1_oof) %>%
    mutate(model = "xgboost", threshold_kind = "f1"),
  hard_metrics(lab$manual_label, state$lgb_pred_f1_oof) %>%
    mutate(model = "lightgbm", threshold_kind = "f1")
)
```

## Elastic Net Coefficients

```{r enet_coefs}
coef_mat <- as.matrix(coef(enet_fit, s = enet_tuned_full$lambda))
coef_df  <- tibble(
  feature     = rownames(coef_mat),
  coefficient = as.numeric(coef_mat[, 1])
) %>%
  filter(feature != "(Intercept)", coefficient != 0) %>%
  arrange(desc(abs(coefficient)))

n_nonzero <- nrow(coef_df)
n_total   <- nrow(coef_mat) - 1L

cat(sprintf("\nElastic net: alpha = %s, lambda = %s\n",
            enet_tuned_full$alpha, round(enet_tuned_full$lambda, 6)))
cat(sprintf("Non-zero features: %d / %d (%.1f%%)\n",
            n_nonzero, n_total, 100 * n_nonzero / n_total))
cat("\nTop 30 features by |coefficient|:\n")
print(head(coef_df, 30), n = 30)
```

## Data Audit Summary

```{r audit}
audit_summary <- tibble(
  corpus_rows               = nrow(df),
  labeled_rows              = nrow(df_labeled),
  labels_raw_rows           = nrow(labels_raw),
  duplicate_corpus_ids      = nrow(dup_corpus),
  conflicting_label_ids     = nrow(label_conflicts),
  feature_weighting         = "sublinear_tfidf_l2norm",
  min_docfreq_filter        = min_docfreq,
  probability_policy        = "all_models_platt_calibrated",
  oof_probability_policy    = "outer_validation_predictions_with_training_only_calibration",
  oof_threshold_policy      = "nested_outer_fold_train_only",
  deployment_threshold_policy = "full_labeled_cross_fitted_calibrated"
)
```

## Write All Outputs

```{r write_outputs}
# Full selected hyperparameters
selected_full <- bind_rows(
  tibble(model = "naive_bayes"),
  tibble(model = "elastic_net",
         alpha = enet_tuned_full$alpha,
         lambda_1se = enet_tuned_full$lambda,
         lambda_min = enet_tuned_full$lambda_min),
  tibble(model = "svm_platt",
         cost = svm_tuned_full$cost),
  xgb_tuning_result$best %>% mutate(model = "xgboost"),
  lgb_tuning_result$best %>% mutate(model = "lightgbm")
)

dir.create(results_path, recursive = TRUE, showWarnings = FALSE)

readr::write_csv(
  out %>% select(-any_of(c("text", "text_raw"))),
  file.path(results_path, "corpus_non_late_late_model_outputs_oof.csv"))

readr::write_csv(deployment_thresholds,
  file.path(results_path, "model_thresholds_deployment.csv"))

readr::write_csv(
  arrange(state$outer_threshold_df, outer_fold, model),
  file.path(results_path, "model_thresholds_outer.csv"))

readr::write_csv(oof_metrics,
  file.path(results_path, "model_metrics_oof.csv"))

readr::write_csv(
  arrange(state$fold_metric_df, outer_fold, model),
  file.path(results_path, "model_metrics_oof_by_fold.csv"))

readr::write_csv(oof_eval_predictions,
  file.path(results_path, "oof_eval_predictions.csv"))

readr::write_csv(oof_confusion_counts,
  file.path(results_path, "oof_confusion_counts.csv"))

readr::write_csv(
  arrange(state$selected_outer_df, outer_fold, model),
  file.path(results_path, "model_hyperparameters_outer.csv"))

readr::write_csv(selected_full,
  file.path(results_path, "model_hyperparameters_full.csv"))

readr::write_csv(coef_df,
  file.path(results_path, "enet_coefficients.csv"))

readr::write_csv(audit_summary,
  file.path(results_path, "data_audit_summary.csv"))

log_progress("done", "all outputs written")

cat("\n===== FINAL SUMMARY =====\n")
cat(sprintf("Models: NB, Elastic Net (α=%s, λ=%s), SVM, XGB, LGB\n",
            enet_tuned_full$alpha, round(enet_tuned_full$lambda, 6)))
cat(sprintf("Non-zero elastic net features: %d / %d\n", n_nonzero, n_total))
cat("\nOOF metrics:\n")
oof_metrics %>%
  select(model, auc, log_loss, brier, f1_youden, mcc_youden) %>%
  as.data.frame() %>%
  print()
```
